\section{Related Work}
\label{ch01:sec:related_work}

This section reviews the literature most relevant to the themes developed in this
monograph. The discussion begins with formal methods for task and motion
synthesis, then turns to planning under incomplete models and infeasible
specifications, multi-robot coordination under temporal tasks, and execution-time
adaptation under interaction constraints. The final part focuses on contingent
collaboration and human supervision, which have become increasingly important in
large-scale and long-horizon robotic systems.

%==================================================
%==================================================
\subsection{Task Planning and Allocation in Multi-Robot Systems}
\label{sec:ch01:rw:allocation}

Task planning and allocation in multi-robot systems concerns how a mission is
decomposed into executable subtasks, how these subtasks are assigned to
individual robots or subteams, and how the resulting activities are coordinated
over time and space. In this literature, the dominant abstractions are jobs,
visits, service requests, coalitions, and resource claims, rather than formal
behavioral specifications. Classical reviews have established the main taxonomies
of multi-robot task-allocation problems and cooperative planning architectures,
with particular emphasis on the distinction between centralized and distributed
coordination \cite{khamis2015multi}. A related survey literature in cooperative
multi-agent planning has further clarified how decomposition, coupling, and
information sharing shape the complexity of team planning
\cite{torreno2017cooperative}. These works provide the algorithmic foundation for
fleet-level coordination, but they typically treat task structure at a more
abstract level than the temporally rich mission models considered later in this
monograph.

\subsubsection{Task decomposition and centralized fleet planning}
\label{sec:ch01:rw:allocation:centralized}

A large portion of the literature adopts a centralized viewpoint, in which a
global planner reasons over the full task set and computes assignments for the
team as a whole. This formulation is natural when the planner has access to task
descriptions, robot capabilities, and resource constraints in a common model.
It is also consistent with classical formulations from routing and scheduling,
where a mission is encoded through precedence relations, timing windows, service
costs, and assignment variables. In multi-robot settings, such centralized
formulations are attractive precisely because they can model global coupling
explicitly, including heterogeneous robot capabilities, ordered tasks, and
shared resources \cite{gini2017multi,ji2021multi}.

This advantage is particularly clear in mixed-integer and related combinatorial
formulations. Such methods can represent rich coupling patterns with high
precision, and they often yield assignments that are globally consistent and
near-optimal with respect to well-defined cost criteria. Their strength,
however, is inseparable from their computational burden. As the number of
robots, subtasks, and inter-task constraints increases, the search or
optimization problem quickly becomes difficult to solve and even harder to
re-solve after mission updates. Branch-and-bound and hybrid exact--heuristic
approaches alleviate this burden in some regimes, but the underlying scaling
problem remains substantial \cite{martin2021mrta}. Centralized fleet planning is
therefore most effective when the task set is moderately sized, the coupling is
structured, and replanning is relatively infrequent.

\subsubsection{Distributed allocation and market-based coordination}
\label{sec:ch01:rw:allocation:distributed}

To improve scalability and robustness, many works distribute the allocation
process across the team. Market-based coordination is one of the most influential
examples. In auction-based schemes, robots evaluate candidate tasks using local
cost or utility estimates and then compete or negotiate for assignment. This
reduces the need for solving a monolithic team-level optimization problem after
every task update. It also fits heterogeneous fleets well, because each robot can
incorporate its own capabilities, motion costs, or uncertainty estimates into
its bid \cite{zhao2021market}. As a result, market-based methods often provide a
practical compromise between allocation quality and computational tractability.

A broader distributed perspective appears in negotiation-based and
constraint-based coordination. In these formulations, the allocation is shaped by
local message exchange rather than by a single global optimizer, and consistency
is achieved through iterative agreement among agents. Distributed constraint
optimization has become a particularly important abstraction in this context,
because it offers a general language for utility, coupling, and decentralized
decision making \cite{fioretto2018distributed}. Approximate and metaheuristic
methods, including evolutionary search, further enlarge the range of tractable
problem sizes \cite{liu2019multi}. The corresponding trade-off is clear. These
approaches scale better and are often more robust to local disturbances, but
their guarantees of global optimality are weaker, and their behavior becomes
harder to analyze when feasibility depends on tightly coupled timing, ordering,
or collaborative-action constraints. They are therefore most effective when task
interactions remain relatively local or can be evaluated through local
performance surrogates.

\subsubsection{Hierarchical and continual fleet coordination}
\label{sec:ch01:rw:allocation:hierarchical}

More recent work has attempted to bridge the gap between globally informed
planning and scalable execution through hierarchical coordination. In this family
of approaches, the mission is decomposed across multiple abstraction levels,
such as strategic mission partitioning, fleet-level allocation, and local
execution planning. This architecture reduces the dimensionality of each planning
layer while preserving enough global structure to maintain coherent team
behavior. It is therefore especially attractive for large fleets, where a fully
flat optimization becomes difficult to maintain. Hierarchical planning has also
been combined with richer symbolic or temporal mission models, which allows task
assignment and high-level task structure to be addressed within a common
pipeline \cite{leahy2021scalable,luo2024decomposition}.

A closely related trend concerns continual or reactive fleet coordination, in
which new tasks may arise during execution and assignments must be revised
online. This shift reflects the practical observation that real robotic fleets
rarely operate over a fixed set of tasks known completely in advance. Recent
methods on dynamic coalition formation and reactive task allocation represent
important progress in this direction \cite{dai2024dynamic,chen2025real}. Even
so, many such approaches still assume relatively structured task models,
moderate update rates, or simplified interactions between allocation and
low-level execution. In particular, the task is often treated as an abstract job
to be assigned, whereas in more demanding settings the feasibility of a task may
depend on motion constraints, action preconditions, temporal ordering, and
execution-time interaction among robots.

Taken together, the literature on task planning and allocation provides strong
algorithmic tools for mission decomposition, assignment, and scheduling in
multi-robot systems. Its primary concern is how to distribute workload
efficiently across a fleet under computational and operational constraints.
However, many formulations abstract away the richer semantics that arise when
missions involve temporal ordering, recurrence, safety, and contingent
collaboration. That limitation motivates the next subsection, which turns to
formal methods for task and motion synthesis and to logic-based models that can
represent and reason about complex robotic behavior with stronger semantic
precision.

%==================================================
%==================================================
\subsection{Formal Methods for Task and Motion Synthesis}
\label{sec:ch01:rw:formal}

Formal methods provide a principled framework for specifying, verifying, and
synthesizing system behavior from high-level requirements. Their importance in
robotics increased once missions could no longer be described adequately as
simple goal-reaching or waypoint-following problems. Temporal logic made it
possible to express ordering, recurrence, safety, and response within a single
semantic language, while automata-based reasoning connected these requirements
to algorithmic synthesis procedures. As a result, this literature is relevant
not only because it offers expressive task descriptions, but also because it
seeks correctness guarantees that are difficult to obtain from purely heuristic
planning pipelines.

\subsubsection{Symbolic abstraction and automata-based synthesis}
\label{sec:ch01:rw:formal:abstraction}

A foundational idea in this area is to replace a continuous or hybrid dynamical
system by a finite symbolic model that preserves the properties relevant to a
given specification. Once such a model is available, model checking and
automata-theoretic synthesis can be applied to construct system behaviors whose
traces satisfy temporal formulas. The algorithmic basis for this viewpoint is
developed systematically by Baier and Katoen in
\cite{baier2008principles}. In the context of hybrid and continuous systems,
early formal abstraction results such as \cite{alur2000discrete} showed how
continuous dynamics can be related to finite-state models, while approximate
bisimulation and related abstractions, as in
\cite{girard2007approximately}, clarified how finite symbolic models can
preserve behavioral properties with controlled approximation error.

Building on this formal backbone, temporal-logic control synthesis has been
developed for several important model classes. For discrete-time linear
systems, Tabuada and Pappas in \cite{tabuada2006linear} and Kloetzer and Belta
in \cite{kloetzer2008fully} showed how LTL specifications can be enforced
through finite abstractions and automata-based constructions. For piecewise
affine systems, Yordanov and Belta in \cite{yordanov2010formal} and later
Yordanov et al. in \cite{yordanov2012temporal} extended this approach to richer
hybrid dynamics. For stochastic models, Ding et al. in \cite{ding2011mdp}
illustrated how temporal-logic constraints can be incorporated into controlled
Markov decision processes. Across these settings, the key conceptual shift is
the same: task satisfaction is no longer treated as an external check on a
candidate plan, but as the object of synthesis itself.

\subsubsection{Temporal-logic planning for robotic motion}
\label{sec:ch01:rw:formal:motion}

Within robotics, these developments led to a now standard architecture for
planning under temporal tasks. A continuous robot model is first abstracted into
a finite transition system, typically by partitioning the workspace into regions
and encoding the feasible motion among them. A temporal formula is then
translated into an automaton, a discrete plan is synthesized over the product
between the robot abstraction and the task automaton, and the resulting plan is
finally refined into executable motion by a lower-level controller. The survey
by Belta et al. in \cite{belta2007symbolic} helped establish this architecture
as a common bridge between formal verification and robotic planning.

This abstraction-to-synthesis pipeline substantially broadened the class of
robotic missions that could be handled in a systematic way. Rather than
reasoning only about destination points, the literature could now address
surveillance, sequencing, persistent visitation, and conditional response
within a unified framework. Early robotic formulations of temporal-logic motion
planning include dynamic-robot synthesis in \cite{fainekos2009temporal} and
sampling-based planning under temporal goals in \cite{bhatia2010sampling}.
Related work also explored interfaces and software layers that make temporal
specifications more accessible in robotic practice. For example,
\cite{srinivas2013graphical} proposed a graphical language for motion and
mission planning, while \cite{he2015towards} connected temporal specifications
to manipulation-oriented planning. The central achievement of this literature is
that it transformed high-level mission requirements into motion-planning
problems with explicit logical semantics.

\subsubsection{Integrated task and motion synthesis with actions}
\label{sec:ch01:rw:formal:action}

The classical motion-planning pipeline becomes less adequate when the mission
requires not only movement among regions, but also actions whose availability
depends on local conditions, internal robot state, or discrete choices among
alternatives. In planning theory, such action structure is naturally described
through preconditions and effects, as in the action-language perspective of
\cite{gelfond1998action} and the planning formalisms surveyed by
\cite{pednault1989adl}. In robotics, however, integrating this type of action
semantics with motion planning under temporal logic is substantially more
difficult. The feasibility of an action may depend on where the robot is, what
has already occurred, and which local resources are currently available.

Several works moved formal robotic synthesis in this direction by enriching the
symbolic state or label structure used for planning. The manipulation-oriented
framework in \cite{he2015towards} already reflects the need to reason about
actions more explicitly than in pure region-visitation tasks. Yet many
classical formulations remain restrictive in one of two ways. Some attach
actions permanently to locations, which simplifies planning but cannot represent
changing local conditions or alternative action choices. Others treat action
propositions as freely switchable symbolic labels, which weakens the connection
between the task specification and the actual feasibility of execution. This
limitation becomes visible when comparing surveillance-style formulations such
as \cite{smith2011optimal} and reactive mission planning such as
\cite{kress2009temporal} with motion--action coupling approaches such as
\cite{guo2013motion}, where navigation and action-description models are
combined more directly. Once local actions, service behaviors, or collaborative
operations become part of the mission, the planning model must represent more
than region transitions alone.

\subsubsection{Expressiveness, robustness, and scalability limits}
\label{sec:ch01:rw:formal:limits}

As the expressiveness of the task language increases, the synthesis problem
also becomes more demanding. One line of work introduced quantitative or
robustness-based extensions in order to move beyond purely Boolean task
satisfaction. For example, the robustness semantics developed by Fainekos and
Pappas in \cite{fainekos2009robustness} quantify how strongly a trajectory
satisfies or violates a temporal requirement, while probabilistic and
falsification-oriented analyses such as \cite{abbas2013probabilistic} connect
formal semantics to testing and optimization. These developments are important
because they make formal specifications more useful under uncertainty, model
mismatch, and optimization-driven planning. At the same time, they also enlarge
the computational burden and complicate the relation between symbolic synthesis
and executable control.

A more persistent limitation concerns scalability. The practical success of
formal task and motion synthesis depends critically on the quality of the
abstraction and on the size of the automata and product structures generated
during planning. Even when the continuous model is approximated well, automaton
construction and product-based search may grow rapidly with the specification
length and the size of the workspace abstraction, as emphasized in
\cite{baier2008principles}. For this reason, many classical methods remain most
effective when the workspace is fixed, the relevant semantic labels are known in
advance, and synthesis can be performed largely offline. These assumptions are
restrictive in the presence of incomplete knowledge, changing workspace
information, or nominal tasks that turn out to be infeasible. This observation
motivates the next subsection, which turns to planning under incomplete models
and infeasible specifications.


%==================================================
%==================================================
\subsection{Planning under Incomplete Models and Infeasible Specifications}
\label{sec:ch01:rw:incomplete}

The synthesis frameworks reviewed above are most effective when the workspace,
its semantic labels, and the task model are all available in a reliable form
before planning begins. In that setting, a symbolic abstraction can be built
once, a temporal objective can be compiled into an automaton, and execution can
proceed against a fixed model. Robotic deployment is often less orderly. The
planner may rely on only partial knowledge of traversability, obstacles, or
task-relevant properties, and this knowledge may continue to evolve after
execution has started. A second difficulty appears when the nominal
specification itself cannot be realized over the available model. The relevant
literature therefore departs from the classical setting in two connected ways:
it either weakens the assumption of complete prior knowledge, or it weakens the
expectation that the original task can be satisfied exactly.

\subsubsection{Uncertain and partially known workspace models}
\label{sec:ch01:rw:incomplete:workspace}

One response to incomplete knowledge is to encode uncertainty directly in the
planning model. In stochastic formulations, temporal-task synthesis is combined
with Markov decision processes so that uncertain transitions or observations are
treated probabilistically rather than ignored. For example, \cite{ding2011mdp}
shows how temporal-logic specifications can be synthesized over controlled
stochastic models, while \cite{wolff2012robust} studies robust satisfaction of
temporal objectives under uncertainty in transition probabilities. These works
are important because they preserve the formal link between high-level task
semantics and quantitative uncertainty, allowing planning objectives to be
expressed in terms of satisfaction probability or robust performance rather than
only exact correctness over a deterministic abstraction.

A different route treats the environment as reactive or only partially observed
during planning. Receding-horizon synthesis in \cite{wongpiromsarn2010receding}
illustrates how temporal-logic tasks can be enforced through repeated online
optimization, so that new environmental information can influence the evolving
control policy. More recently, perception-aware planning has incorporated
uncertain semantic information into the formal synthesis loop, as in
\cite{kantaros2022perception}, where temporal planning is coupled to uncertain
maps and perception outcomes. What these methods share is an attempt to weaken
the assumption of complete prior knowledge without discarding formal task
semantics. Their common limitation is that uncertainty is usually represented as
part of the model class from the outset. In many robotic settings, however, the
difficulty is not only stochasticity or partial observation in a static sense,
but the progressive correction of a symbolic model that was incomplete or
partially inaccurate at the moment of planning.

\subsubsection{Execution-time plan repair and revision}
\label{sec:ch01:rw:incomplete:repair}

Once execution begins, the practical question often changes from how to plan
under uncertainty to how to respond when the current symbolic plan is no longer
consistent with newly acquired information. This leads to a complementary body
of work on plan repair and revision. Rather than resynthesizing from scratch
after every inconsistency, these methods attempt to preserve as much of the
existing plan structure as possible while restoring feasibility or task
correctness. An early example is the backtracking framework of
\cite{livingston2012backtracking}, where execution retreats to a previously
consistent point and replanning proceeds from there. The appeal of this
perspective is computational as well as semantic: it tries to exploit the
structure already present in the nominal plan instead of discarding it
completely.

This idea becomes especially meaningful in temporally constrained motion
planning, where the effect of a model update is not limited to local geometry.
If a transition becomes blocked or a proposition label changes, the question is
not merely whether another route exists, but whether the remaining behavior
still satisfies the intended temporal formula. Iterative planning for partially
unknown environments in \cite{maly2013iterative} already reflects this
execution-centered viewpoint, while the revision framework in
\cite{guo2013revising} addresses the semantic consequences of model updates more
directly by revising LTL motion plans in partially known workspaces. The
importance of this literature lies precisely in that shift: plan adaptation is
treated as a logic-sensitive operation, not simply as local geometric
replanning.

Even so, many plan-revision methods remain limited in scope. They are often
developed for single-robot settings, or for environments in which the main
uncertainty concerns blocked transitions and updated labels rather than tightly
coupled interaction with other agents. Moreover, local repair does not always
make clear how far the revised behavior may drift from the original
specification when multiple updates accumulate over time. These questions
become more pronounced once task dependencies, collaboration, and shared
resources enter the planning loop.

\subsubsection{Task relaxation and least-violating synthesis}
\label{sec:ch01:rw:incomplete:relaxation}

Model incompleteness is not the only reason that a nominal synthesis problem can
fail. In some instances, the symbolic model is sufficiently accurate, but the
specification itself is unrealizable or infeasible. Under such conditions, a
binary report of failure is rarely satisfactory. One natural response is to ask
how the specification should be changed so that the revised task becomes
realizable with minimal deviation from the original intent. This question is
studied explicitly in \cite{fainekos2011revising}, where temporal-logic
specifications are revised in response to synthesis failure, and in
\cite{kim2012approximate}, where minimal revision of specification automata is
treated as an optimization problem. Such works are valuable because they place
the burden of repair on the specification itself, rather than only on the model
or the plan.

A related stream of research focuses first on diagnosis. Instead of immediately
modifying the task, unrealizability analysis seeks to identify which assumptions
or constraints are responsible for failure. The analysis in
\cite{raman2011analyzing} is representative of this perspective. By isolating
incompatible requirements, the method provides useful information to a designer
or supervisory layer and clarifies whether failure is caused by the environment
assumptions, the system model, or the task structure. Diagnosis, however, does
not yet determine what behavior should be generated once exact realization is
known to be impossible. That question leads to least-violating and partially
satisfying synthesis.

In least-violating planning, the objective is no longer exact task satisfaction
but the best admissible compromise relative to the original specification. The
minimum-violation formulations in \cite{tumova2013least} and
\cite{tumova2013minimum} treat selected rules as violable and quantify the cost
of doing so. Time-sensitive partial satisfaction is considered in
\cite{lahijanian2015time}, where degradation is measured more explicitly against
the structure of the task. Closely related ideas appear in maximally satisfying
planning, such as \cite{tumova2014maximally}, where the goal is to preserve as
much of the original mission as possible under feasibility limitations. This
body of work is especially relevant when specifications contain requirements of
unequal importance. In that case, synthesis becomes a problem of prioritization
as much as one of feasibility, and the distinction between hard and soft
constraints becomes conceptually unavoidable.

Taken together, these studies show that formal planning can be extended beyond
the idealized setting of complete knowledge and exactly realizable tasks.
Uncertainty can be embedded into the model, nominal plans can be repaired as the
workspace abstraction changes, and infeasible specifications can be revised or
relaxed in a principled manner. The remaining difficulty is that much of this
literature still operates in single-robot or weakly coupled settings. Once
multiple robots interact under shared temporal requirements, model updates and
specification compromises no longer remain local. This is the point at which the
review must turn from adaptive single-system planning to temporal-logic
coordination in multi-robot systems.


%==================================================
%==================================================
\subsection{Temporal-Logic Coordination under Coupled Constraints}
\label{sec:ch01:rw:coordination}

Extending temporal-logic synthesis from a single robot to a multi-robot team
changes the nature of the planning problem. The central issue is no longer only
how to construct one behavior that satisfies a temporal specification, but how
to coordinate several behaviors whose feasibility and correctness are coupled
through shared motion space, timing requirements, and interaction constraints.
This shift has produced two main strands in the literature. One strand adopts a
team-level viewpoint and synthesizes joint behavior from a global temporal
mission. The other starts from locally assigned specifications and treats
coordination as an emergent consequence of task dependencies, resource
contention, and intermittent collaboration. Both inherit the automata-based
machinery reviewed in Section~\ref{sec:ch01:rw:formal}, but they differ
substantially in how coordination is modeled and where the computational burden
is placed.

\subsubsection{Global temporal-task coordination}
\label{sec:ch01:rw:coordination:global}

The earliest and most direct extensions of temporal-logic planning to
multi-robot systems followed a top-down paradigm. In these formulations, the
team is viewed as a single composite system, and a global temporal task is
specified over the collective behavior of all robots. Synthesis then proceeds by
constructing coordinated joint executions whose projection onto the individual
robots satisfies the team-level mission. Representative examples include
\cite{kloetzer2010automatic}, where distributed robotic teams are deployed from
temporal-logic specifications, and \cite{ding2011automatic}, where global task
requirements are translated into coordinated team behaviors through formal
synthesis. In this class of methods, inter-robot dependence is handled directly
within the symbolic planning object rather than deferred to execution time.

Such formulations are appealing because they provide a strong notion of
correctness. Since the mission is specified globally, logical dependencies among
robots are represented explicitly, and synchronization requirements can be
verified at the team level. This makes the framework well suited to tasks that
involve shared temporal objectives, coordinated visitation patterns, or
mission-level ordering constraints. It also explains why early work in this area
was able to provide strong guarantees about joint task satisfaction despite the
complexity of the underlying system.

As the literature evolved, the team-level paradigm was extended in order to
address robustness and richer coordination structure. The effect of uncertain
travel times on temporal satisfaction was studied in \cite{ulusoy2012robust} and
\cite{ulusoy2013optimality}, where the emphasis lies on preserving correctness
despite bounded execution mismatch. Related optimization-based formulations made
synchronization and routing constraints more explicit, which improved the
representation of heterogeneous capabilities and coupled motion schedules.
Recent developments have pushed this direction further by integrating allocation,
sampling, and scalable decomposition into the synthesis loop.
Simultaneous task allocation and planning under temporal goals were considered
in \cite{schillinger2018simultaneous}. Sampling-based optimal synthesis was
developed in \cite{kantaros2020stylus}. Abstraction-free formulations such as
\cite{luo2021abstractionfree} and temporal-task allocation methods such as
\cite{luo2022temporal} expanded the size and variety of missions that can be
handled. Capability-aware and counting-style extensions, represented by
\cite{sahin2019multirobot} and \cite{cardona2024planning}, broadened the
expressive scope of global team specifications.

Even with these advances, the main limitation of the top-down literature remains
structural rather than conceptual. The team-level viewpoint is powerful because
it captures coupling directly, but the same feature causes the planning object
to grow rapidly with the number of robots, synchronization points, and temporal
dependencies. Hierarchical remedies alleviate this burden in some cases, yet the
global paradigm is still most effective when the coordination pattern can be
encoded within a centralized or hierarchically centralized synthesis pipeline.

\subsubsection{Locally assigned tasks and loosely coupled collaboration}
\label{sec:ch01:rw:coordination:local}

Another branch of the literature begins from a different modeling choice. Rather
than specifying one temporal task for the entire team, it assigns local
specifications to individual robots and studies how these locally meaningful
behaviors can coexist or cooperate. This viewpoint is particularly natural in
heterogeneous fleets, where each robot may have a distinct role, sensing
profile, or action capability, and where the full coordination pattern need not
be fixed in advance. Decentralized multi-agent control from local temporal
specifications was studied in \cite{filippidis2012decentralized}, and receding-
horizon coordination from local LTL tasks was developed in \cite{tumova2014receding}.
These works showed that formal coordination need not start from a single
team-wide formula in order to retain a meaningful logical structure.

The attraction of the locally specified viewpoint is not only computational.
It also matches many practical deployments more closely. In warehouse systems,
inspection scenarios, and service settings, robots often receive task
descriptions that are already local in content, while collaboration occurs only
when these local activities intersect through shared resources, timing
dependencies, or temporary service requirements. Under such conditions, a
globally synchronized team model may be unnecessarily rigid. Local
specifications make it possible to exploit heterogeneity and asynchronous
execution more naturally, and they reduce the need to maintain a fully coupled
global planning object throughout the mission.

This line of work becomes especially relevant when inter-robot dependence is
intermittent rather than persistent. Distributed plan reconfiguration through
knowledge transfer in \cite{guo2014distributed}, bottom-up coordination for
dependent local tasks in \cite{guo2015multi}, and task-and-motion coordination
for heterogeneous teams with loosely coupled local specifications in
\cite{guo2016task} all reflect this logic. In such formulations, collaboration
is not imposed by a permanent team-level objective. Instead, it arises when one
robot's local task becomes feasible only through assistance, information
exchange, or conflict resolution with others. This shift greatly improves
scalability and modeling realism, but it also changes the main technical burden.
Correctness is no longer guaranteed by one global synthesis step. It must be
maintained across distributed decisions, evolving dependencies, and execution-
time modifications of local plans.

Large systems have motivated further refinement of this perspective. Work such as
\cite{leahy2021scalable}, \cite{luo2024decomposition}, and \cite{luo2025hiera}
shows that hierarchical decomposition can partially recover global structure
without returning fully to flat centralized synthesis. Even so, the conceptual
shift remains: once temporal coordination is organized around local
specifications, attention moves from global satisfiability alone to consistency
among partial plans, dependency management, and the timing of collaboration.
That shift is closely aligned with the perspective developed later in this
monograph.

\subsubsection{Interaction constraints and relative-motion safety}
\label{sec:ch01:rw:coordination:interaction}

Temporal-task satisfaction does not by itself guarantee that a multi-robot team
can operate safely and coherently in shared space. Even if each robot follows a
correct symbolic plan, the team must still avoid collisions, preserve necessary
communication structure, and respect relative-motion constraints induced by the
physical environment or the task itself. These requirements are often treated as
background assumptions in high-level planning, but in practice they are central
components of coordination. Classical stabilization and collision-avoidance
schemes for multiple robots were developed in \cite{dimarogonas2006feedback},
while connectivity-preserving motion control was studied in \cite{zavlanos2011graph}.
Taken together, these works show that safe team behavior depends on persistent
interaction constraints that remain active throughout execution, not only at
isolated synchronization points.

This observation led to more explicit efforts to combine formal task planning
with relative-motion and communication constraints. Distributed connectedness
maintenance in \cite{ji2007distributed} and the control of relative motion in
\cite{guo2014controlling} illustrate how local interaction requirements can be
embedded into continuous coordination laws. Temporal-task variants appear in
\cite{guo2014cooperative}, \cite{guo2015hybrid}, and
\cite{guo2015communicationfree}, where local temporal objectives are coupled to
distance maintenance, connectivity, or communication-free coordination rules.
Related deployment problems with reduced communication were also considered in
\cite{kloetzer2011multi}. What makes this literature especially relevant is that
it refuses to separate task satisfaction from team integrity. The planning
problem is no longer only to achieve a sequence of symbolic objectives, but to
do so while continuously preserving the geometric and communication relations
that keep the team operational.

Another useful perspective comes from graph-based local interaction models.
Embedded graph grammars, introduced in \cite{mcnew2006locally} and developed
further in \cite{smith2009multi}, \cite{mcnew2007solving}, and
\cite{pickem2013complete}, provide a compact way to describe local mode
transitions, neighbor-dependent behaviors, and structural reconfiguration in
large robotic teams. Their appeal lies in locality. Coordination rules are
expressed through neighboring relations and graph rewrites, which makes them
well suited to settings in which only local interaction is available. At the
same time, this literature often leaves the connection to rich temporal-task
semantics only partially developed. The hybrid graph-grammar viewpoint in
\cite{guo2016graphgrammar} is representative of attempts to close that gap by
linking local interaction rules to more expressive task-level coordination
objectives.

The overall picture is therefore less a neat progression than a persistent
tension. Team-level temporal synthesis captures coordination explicitly, but
faces a steep growth in complexity. Locally specified planning improves
flexibility and tractability, but shifts the burden toward distributed
consistency and execution-time coordination. Interaction constraints make the
problem sharper still, because high-level correctness must coexist with
collision avoidance, connectivity preservation, and relative-motion safety. The
next subsection examines what happens once these already difficult coordination
problems are further complicated by online task arrivals, contingent requests,
limited communication, and human involvement.


%==================================================
%==================================================
\subsection{Online, Communication-Constrained, and Human-Supervised Coordination}
\label{sec:ch01:rw:online}

The coordination models reviewed so far still rely, to different degrees, on the
assumption that the relevant task structure can be determined before execution
and that enough information is available to preserve consistency among the
robots. Practical deployment is often more fluid. New service demands may arise
after the mission has already begun, temporary collaborations may be requested
by robots that were previously independent, information may propagate only
intermittently, and human operators may need to redirect or validate decisions
without taking over low-level control. Under these conditions, coordination is
not only a matter of satisfying temporal requirements over a fixed team model.
It also becomes a matter of managing evolving commitments, delayed information,
and selective intervention while the mission is still unfolding.

\subsubsection{Online and contingent collaboration}
\label{sec:ch01:rw:online:contingent}

A substantial recent literature addresses the shift from static mission
synthesis to continual and reactive coordination. In these formulations, the
task set is no longer assumed to be fixed at the outset. New jobs, requests, or
priorities may appear during execution, and the team must update its decisions
while preserving enough structure to maintain progress toward the original
mission. Reactive task allocation under temporal constraints, studied in
\cite{chen2025real}, illustrates this shift clearly. Hierarchical coordination
under continually arriving task streams, as in \cite{luo2025hulk}, makes the
same point at a larger scale: the problem is not merely how to solve a larger
planning instance, but how to preserve mission coherence while the planning
instance itself continues to change.

This online viewpoint becomes more demanding when collaboration is contingent
rather than persistent. A robot may require assistance only after reaching a
particular local state, may issue a request only when a resource becomes
available or unavailable, or may need another robot to satisfy a transient
formation or service condition. Such events are not well captured by the
classical notion of reactive synthesis against a predefined class of
environmental behaviors. In reactive temporal-logic formulations such as
\cite{kress2009temporal}, the environment is modeled in advance and the
controller is synthesized against that known interaction structure. In
contingent collaboration, by contrast, important coordination events are often
generated by the evolving internal state of the robotic team itself, and they
need not belong to a finite class that can be exhaustively encoded before
deployment.

This distinction has motivated work that treats online requests as explicit
coordination objects rather than as incidental disturbances. Event-based
collaborative mechanisms involving service and formation requests were studied
in \cite{guo2017event}, where coordination is triggered by the evolving status
of local tasks. What makes this line of work particularly relevant is that it
places execution-time negotiation inside the planning problem instead of outside
it. The team must not only respond to change, but also reason about when a new
request should be generated, how it should be propagated, and how it should
alter the surrounding local plans. Existing methods have improved responsiveness
substantially, but many still rely on repeated replanning of expensive global
coordination objects or on update structures simpler than those encountered in
long-horizon, heterogeneous missions. The perspective adopted later in this
monograph is especially close to the contingent case, where collaboration is
triggered locally and must be coordinated without assuming that all interaction
patterns are known in advance.

\subsubsection{Coordination under intermittent or limited communication}
\label{sec:ch01:rw:online:communication}

The difficulty of online coordination increases sharply once communication can no
longer be treated as effectively unlimited. A large part of the multi-robot
planning literature assumes communication rich enough to support timely sharing
of plans, state estimates, and coordination decisions. That assumption is often
reasonable in structured indoor domains or simulations, but it is difficult to
maintain in subterranean exploration, disaster response, large indoor
facilities, or infrastructure inspection, where range limits, occlusion, and
network instability severely restrict information flow. In such settings,
communication becomes a resource that must itself be planned, preserved, or
recovered.

One response has been to schedule communication explicitly rather than requiring
it continuously. Intermittent communication protocols, developed in works such
as \cite{stump2011connectivity}, \cite{hollinger2013sampling}, and
\cite{nikou2016cooperative}, allow robots to operate independently for periods
of time and then re-establish contact through planned rendezvous. Different
variants make different assumptions about whether communication times are fixed
or optimized, whether meetings are pairwise or team-wide, and whether the
rendezvous structure is imposed centrally or emerges through distributed
coordination. The appeal of these methods lies in the freedom they return to the
robots between communication events. Their cost is that information no longer
propagates immediately, which complicates temporal consistency, dependency
management, and the validity of local decisions made on stale data.

These issues are especially visible in collaborative exploration. Frontier-based
and information-gain approaches, including \cite{yamauchi1997frontier},
\cite{burgard2000collaborative}, and \cite{sheng2006distributed}, benefit
strongly from shared maps and timely mutual awareness. Once that assumption is
weakened, the team must decide not only where to explore, but also when to
reconnect, whether to deploy relays, and how to balance map growth against
information latency. Relay-based and rendezvous-oriented exploration strategies,
as in \cite{mosteo2013framework}, \cite{dehoog2010role}, and related work,
demonstrate that exploration and communication are tightly coupled decision
problems rather than separable modules.

Recent studies have begun to connect communication constraints to richer
high-level mission models. Team-wise intermittent communication in unknown
environments was investigated in \cite{guo2024teamwise}, while communication-
aware fleet coordination with online requests appears in \cite{guo2025moroco}.
These efforts are important because they treat communication limits not merely
as performance degradation, but as structural constraints that reshape the
planning problem. Task assignment, temporal consistency, and situational
awareness all become functions of when information can be exchanged and how much
of it can be moved through the network. Once communication is treated in this
way, the distinction between planning and networking becomes much less rigid.

\subsubsection{Human-fleet interaction and mixed-initiative supervision}
\label{sec:ch01:rw:online:human}

Limited communication also changes the role of the human operator. In many field
deployments, the human is not simply an external observer but an active
participant who monitors the fleet, approves risky actions, inserts new goals,
or resolves ambiguities that are difficult to settle autonomously. At the same
time, direct teleoperation scales poorly with the number of robots and becomes
even less practical when communication is delayed or intermittent. This tension
has motivated a substantial literature on human--swarm and human--multi-robot
interaction. Surveys such as \cite{crandall2017human} and \cite{nam2018survey}
emphasize that scalable supervision depends on abstraction, situational
awareness, and control modes that allow operators to influence the mission
without micromanaging individual robots.

Mixed-initiative coordination provides one influential answer to this problem.
Instead of assigning all authority to either the autonomy stack or the operator,
mixed-initiative systems allow decision authority to shift according to
uncertainty, task difficulty, and operator intent. This perspective is evident
in \cite{jiang2019mixed}, where human input and robotic planning are integrated
more directly, and it is closely related to scaled-autonomy ideas such as
\cite{kortenkamp2016integrating}, where the level of human involvement depends
on the state of the mission. Such approaches are especially relevant for
temporally structured tasks, because intervention cannot be treated as an
isolated override. It must be incorporated into a coordination architecture that
still preserves task progress, inter-robot consistency, and mission-level
safety.

The challenge becomes richer when the operator is embedded in a
communication-constrained fleet. Then the design problem is not only how to
present the state of the team, but how to preserve timely and selective human
intervention when the robots themselves may not remain continuously connected.
Interactive exploration frameworks such as \cite{jhunjhunwala2021ihero} already
illustrate the value of online operator input in dynamic missions. Work on
multimodal and immersive interfaces, including \cite{orton2020madcat} and
\cite{walker2023ar}, shows how richer communication channels can improve
situational awareness and command efficiency. Yet much of this literature still
assumes communication support more reliable than what is available in many
real-world deployments. The unresolved issue is therefore deeper than adding a
human to the loop. The real question is how to preserve scalable autonomy,
temporal consistency, and meaningful oversight at the same time.

What emerges from this body of work is not a simple extension of classical
multi-robot planning, but a noticeable change in emphasis. Coordination must now
accommodate evolving commitments, delayed information, and selective human
intervention as part of the mission logic itself. This shift explains why later
strands of the literature turn increasingly toward abstraction, uncertainty, and
dynamic task models rather than treating them as secondary complications.

%========================================
%========================================
\subsection{Uncertainty, Abstraction, and Dynamic Task Models}
\label{sec:ch01:rw:extensions}

As temporal-task planning moved beyond static deterministic workspaces, several
adjacent strands of the literature became increasingly important. One strand
replaces deterministic transition models by stochastic decision processes in
order to represent uncertain outcomes more faithfully. Another revisits the
meaning of safety once long-horizon execution, recovery, or partial task
violation must be taken into account. A third addresses the computational burden
of formal synthesis through symbolic and temporal abstraction. More recent work
has also expanded the task language itself, moving from propositional formulas
over fixed labels toward signal-based specifications and dynamically changing
task models. These developments do not form a single unified school, yet they
collectively mark a shift in emphasis: uncertainty, abstraction, and evolving
task semantics are no longer secondary complications, but part of the core
planning problem.

\subsubsection{Probabilistic temporal-task planning}
\label{sec:ch01:rw:extensions:mdp}

Markov decision processes provide one of the main formal models for robotic
planning under uncertainty. In this setting, stochasticity may arise from
actuation noise, imperfect sensing, or uncertain environmental response, and the
planner seeks a policy rather than a single deterministic path. Classical MDP
planning typically focuses on reachability or on the optimization of expected
cumulative reward. Temporal-task formulations alter that objective by requiring
the policy-induced behavior to satisfy a high-level logical specification.
Representative work in this direction includes \cite{ding2011mdp}, where LTL
constraints are incorporated into MDP control, and \cite{wolff2012robust}, where
task satisfaction is studied under uncertainty in transition probabilities.
These formulations are important because they preserve formal task semantics
while moving from deterministic execution traces to probabilistic policies.

Once the planning object becomes a policy, several new questions emerge. The
objective may be to maximize the probability of satisfying a temporal formula,
to optimize cost subject to a minimum satisfaction threshold, or to compare
different operating points across multiple verification criteria. Quantitative
multi-objective verification in \cite{forejt2011quantitative} and Pareto-style
analysis in \cite{forejt2012pareto} make these trade-offs explicit, while tool
support such as \cite{kwiatkowska2011prism} has helped establish a practical
computational workflow for probabilistic verification and synthesis. The
conceptual gain is substantial. Temporal planning is no longer limited to the
question of whether a task can be satisfied, but can instead reason about how
well, how reliably, and at what cost a task can be achieved under uncertainty.

This same line of thought has also influenced learning-based control. Rather
than assuming a fully known stochastic model, recent work has combined temporal
logic with reinforcement learning, as in \cite{hasanbeig2019reinforcement},
\cite{bozkurt2020control}, and \cite{wang2015temporal}. These approaches are
appealing because they broaden the reach of formal planning to settings where
transition structure must be estimated from experience. Their limitations are
equally instructive. Temporal reasoning does not disappear when learning is
introduced. It is relocated into reward shaping, policy optimization, or
probabilistic verification of the learned behavior. In that sense, learning and
formal methods are not substitutes here, but overlapping responses to the same
difficulty of planning under uncertainty.

\subsubsection{Safety requirements and constrained satisfaction}
\label{sec:ch01:rw:extensions:safety}

Safety is often introduced into temporal planning as one requirement among many,
for example by encoding obstacle avoidance, separation, or return conditions
directly in the task formula. Under this interpretation, safety is part of the
same logical object that also captures reachability, surveillance, or service
requirements. This view remains natural and powerful, especially when safety can
be expressed as an invariant or as a temporal exclusion rule. It also underlies
much of the literature on minimum-violation planning, where some requirements
are treated as violable but costly rules rather than as absolutely rigid
constraints. Representative examples include
\cite{tumova2013least,tumova2013minimum,vasile2017minimum}. These methods are
especially useful when the system must continue to operate despite conflicting,
partially infeasible, or degraded mission requirements.

A more structured notion of safety appears when execution unfolds over long
horizons or under substantial uncertainty. In such settings, safety may depend
not only on avoiding a bad state at the present instant, but on maintaining the
ability to recover, return, or hand control to a safe policy later on. This
shift from state avoidance to recoverability is evident in safe exploration for
MDPs, studied in \cite{moldovan2012safe} and \cite{turchetta2016safe}. Formal
synthesis for stochastic hybrid processes, as in \cite{laurenti2020formal},
makes the same point in a different model class: the interpretation of safety
depends on the probabilistic and dynamical structure of the system, not merely
on a static forbidden set. Once safety is viewed in this way, it becomes tightly
connected to model uncertainty, task persistence, and the availability of
future corrective action.

This distinction matters for the broader literature reviewed in this monograph.
If safety is treated only as a propositional side condition, then its relation
to adaptation, partial knowledge, or online coordination may remain unclear. If
it is treated as a constrained-satisfaction problem with its own semantic
structure, then safety becomes a central component of how plans are compared,
revised, or relaxed. That perspective is particularly important in settings
where exact task realization may fail and where different parts of the mission
carry different levels of criticality.

\subsubsection{Hierarchical and symbolic abstraction}
\label{sec:ch01:rw:extensions:abstraction}

The computational cost of formal synthesis depends heavily on how the system is
abstracted. A naive discretization of the full continuous state space quickly
becomes intractable as dimension grows, and this remains one of the main
practical bottlenecks in automata-based planning. Symbolic abstraction is a
response to this difficulty. Rather than enumerating all possible states at a
uniform resolution, it seeks finite representations that preserve exactly those
behavioral distinctions relevant to the task. The general symbolic viewpoint is
developed in \cite{tabuada2009verification}, where verification and control are
organized around finite abstractions of hybrid systems. Related work on
approximate similarity and refinement, such as \cite{haesaert2018verification},
clarifies how abstract policies can be related back to the original stochastic
or continuous model with controlled loss of fidelity.

A complementary response is temporal abstraction. The options framework in
\cite{sutton1999between} introduced the idea that temporally extended behaviors
can be treated as higher-level decision units. This insight has had lasting
influence because it reduces the effective planning horizon and allows the
planner to reason over behaviors rather than only over primitive transitions.
The same impulse appears in formal planning through hierarchical decomposition
and abstraction refinement. The hierarchical LTL synthesis framework in
\cite{meyer2019hierarchical} and incremental synthesis by abstraction
refinement in \cite{nilsson2014incremental} both reflect the view that the
structure of the task should shape the structure of the abstraction. This is not
merely a computational convenience. It changes the modeling philosophy. The
planner is no longer asked to solve a monolithic problem over the finest
available state space, but to exploit the layered structure of the mission and
the dynamics.

Feature-based abstraction becomes especially attractive when only part of the
system state is relevant to the temporal objective. In such cases, the most
useful representation is often not the most detailed one, but the one that
retains the distinctions that matter for satisfaction, refinement, and
execution. This observation connects directly to later parts of this monograph,
where abstraction is not treated as a preprocessing step alone, but as an
integral design choice shaped by task structure, scalability demands, and the
need for online adaptation.

\subsubsection{Signal temporal logic and dynamic task models}
\label{sec:ch01:rw:extensions:stl}

Much of the classical temporal-planning literature is built on propositional
labels attached to regions or abstract states. This representation is natural
for many motion-planning problems, but it becomes restrictive when the mission
depends directly on continuous quantities such as relative distance, velocity,
formation error, or time-varying state predicates. Signal Temporal Logic offers
a richer language for such settings because its semantics are defined directly
over real-valued trajectories. This makes it especially attractive for tasks
that involve transient formation constraints, continuous safety margins, or
specifications that depend on evolving physical signals rather than on static
state labels alone.

This increased expressiveness has encouraged optimization-based formulations for
continuous multi-robot planning. Coupled multi-robot motion under temporal goals
was studied in \cite{lindemann2019coupled}, while recent work such as
\cite{liu2022stlq} and \cite{kurtz2022mixedinteger} illustrates how STL
constraints can be integrated into trajectory optimization and mixed-integer
planning pipelines. The resulting methods narrow the gap between high-level task
specification and continuous-team behavior. At the same time, they inherit the
usual trade-off between expressiveness and tractability. As the task language
moves closer to the signal level, the synthesis problem often moves further away
from the discrete combinatorial structure that made classical automata-based
methods attractive.

Dynamic task models push this evolution one step further. In these settings, the
relevant propositions, semantic features, or spatial targets may change during
execution, so that the task is not evaluated over a fixed symbolic environment.
Reactive planning for heterogeneous teams under dynamic collaborative temporal
missions was studied in \cite{zhang2024reactive}, while dynamic semantic
environments were considered in \cite{kalluraya2023multirobot}. Online learning
and control for STL tasks, as in \cite{aksaray2022online}, points in a similar
direction by coupling adaptation to richer continuous-time specifications.
What these works reveal is a broader transition in the field. Temporal planning
is no longer confined to static propositions over a fixed abstraction. The task
itself may now depend on evolving signals, moving semantic features, and
interaction constraints whose meaning changes as execution proceeds.

Taken together, the lines of work reviewed in this subsection extend formal
planning along several fronts at once. They incorporate uncertainty through
probabilistic decision models, reinterpret safety as a structured planning
constraint, use abstraction to control computational growth, and broaden the
task language beyond propositional formulas over static labels. These
developments help explain why the remaining chapters of this monograph combine
formal task semantics with abstraction, adaptation, and richer coordination
models instead of treating those ingredients separately.
