%==============================
\section{Probabilistic Temporal Tasks and Safe-Return Constraints}
\label{ch06:sec:ch06_safe_plan}

This section considers motion planning under uncertainty for a robot whose
environment is modeled by a labeled Markov decision process, abbreviated as
MDP. The nominal mission is specified by an LTL formula, while safety is
captured by a separate safe-return requirement. This separation is important,
because productive operation and permanent return to a designated safe set are
generally incompatible on the same infinite execution. Related temporal
logic background is given in Chapter~\ref{chap:preliminaries}; see in
particular Sec.~\ref{ch02:sec:LTL} and
Secs.~\ref{ch02:subsec:NBA}--\ref{ch02:subsec:product}.
The present section retains only the material needed for the synthesis
problem, while omitting repeated preliminaries and auxiliary discussion.
The resulting model remains consistent with prior work on
probabilistically labeled MDPs and temporal-logic planning
\cite{guo2018probabilistic,belta2017formal,ding2014ltl}.

%==============================================================
\subsection{Uncertain system and probabilistic temporal task}
\label{ch06:sec:problem-formulate}

This subsection establishes the stochastic task-planning model used throughout the section. It first defines the labeled MDP and the probabilistic temporal-task specification, and then states the safe-return requirement that couples mission fulfillment with recoverability.

\subsubsection{Uncertain system as labeled MDP}
\label{ch06:subsec:probMDP}

The robot and its environment are modeled by a finite labeled MDP. This model
captures the stochastic state evolution together with the propositions needed
for temporal-task evaluation. The formal definition is given by:
\begin{equation}
\label{ch06:eq:mdp}
\mathcal{M} \triangleq
\bigl(X, U, D, p_D, AP, L, c_D, x_0\bigr),
\end{equation}
where $X$ is a finite state space, $U(x)$ is the set of admissible control
actions at state $x \in X$, and
$D \triangleq \{(x,u) \mid x \in X,\; u \in U(x)\}$ is the admissible
state-action set. The transition probability function
$p_D \colon X \times U \times X \rightarrow [0,1]$ satisfies
$\sum_{\check{x} \in X} p_D(x,u,\check{x}) = 1$ for every $(x,u) \in D$.
Moreover, $AP$ is the set of atomic propositions, the labeling map
$L \colon X \rightarrow 2^{AP}$ assigns the propositions satisfied at each
state, $c_D \colon D \rightarrow \mathbb{R}_{>0}$ is the stage-cost function,
and $x_0 \in X$ is the initial state. Equation~\eqref{ch06:eq:mdp} follows the
standard MDP model in \cite{puterman2014markov}, augmented with a deterministic
labeling map.
The deterministic labeling in \eqref{ch06:eq:mdp} is sufficient for the
development below. When the workspace features are uncertain, the same approach
can be extended to probabilistically labeled MDPs by augmenting the state or by
using the construction in \cite{guo2018probabilistic}. In the present section,
the deterministic form is retained to keep the exposition compact.
In a search-and-rescue mission, a state may encode the current region together
with local conditions relevant to motion and task execution. Admissible actions
represent motion primitives or local maneuvers, transition probabilities model
uncertainty due to debris or slippage, and labels record task-relevant features
such as exits, victims, charging stations, or hazards.


\subsubsection{Task specification}
\label{ch06:subsec:LTL}


The nominal mission is specified by an LTL formula $\varphi$ over the
proposition set $AP$ from \eqref{ch06:eq:mdp}. The syntax and semantics of LTL
are recalled in Sec.~\ref{ch02:sec:LTL}. The automata-based synthesis used
later follows the same overall paradigm as in
Secs.~\ref{ch02:subsec:NBA}--\ref{ch02:subsec:product}, with deterministic
automata used here for probabilistic planning
\cite{baier2008principles,belta2017formal}.
Typical tasks include persistent surveillance,
$\varphi = (\Box \Diamond r_1) \wedge (\Box \Diamond r_2)
\wedge (\Box \Diamond r_3)$, ordered service,
$\varphi = \Diamond \bigl((\texttt{pick} \wedge r_1) \wedge
\Diamond (\texttt{drop} \wedge r_2)\bigr)$, and response requirements,
$\varphi = \Box \bigl(\texttt{low} \rightarrow
\Diamond (\texttt{supply} \wedge r_1)\bigr)$.


For each stage $T \geq 0$, let
$R_T \triangleq x_0 l_0 u_0 \cdots x_T l_T$ denote a finite run, where
$l_t = L(x_t)$ and $p_D(x_t,u_t,x_{t+1}) > 0$ for every valid transition.
Let $\mathbf{R}_T$ be the set of all runs of length $T+1$, and let
$\mathbf{R}_{\infty}$ be the set of all infinite runs. A finite-memory policy
is written as $\boldsymbol{\mu} \triangleq \mu_0 \mu_1 \cdots$, where each
$\mu_t \colon \mathbf{R}_t \times U \rightarrow [0,1]$ assigns a distribution
over admissible actions based on the observed history. The set of such policies
is denoted by $\overline{\boldsymbol{\mu}}$.
Given $\boldsymbol{\mu} \in \overline{\boldsymbol{\mu}}$, the MDP in
\eqref{ch06:eq:mdp} induces a probability measure over
$\mathbf{R}_{\infty}$ \cite{baier2008principles}. The corresponding task
satisfaction probability is
\begin{equation}
\label{ch06:eq:satisfy}
\textbf{Sat}_{\mathcal{M}}^{\boldsymbol{\mu}} \triangleq
Pr_{\mathcal{M}}^{\boldsymbol{\mu}}
\bigl(
\{R_{\infty} \in \mathbf{R}_{\mathcal{M}}^{\boldsymbol{\mu}} \mid
R_{\infty}|_{L} \models \varphi\}
\bigr),
\end{equation}
where $\mathbf{R}_{\mathcal{M}}^{\boldsymbol{\mu}} \subseteq
\mathbf{R}_{\infty}$ is the set of infinite runs generated by
$\mathcal{M}$ under $\boldsymbol{\mu}$. Equation~\eqref{ch06:eq:satisfy}
therefore measures the probability that the label sequence induced by an
infinite run satisfies the task formula.
The outbound performance criterion is the expected mean cost
\begin{equation}
\label{ch06:eq:total-cost}
\textbf{Cost}_{\mathcal{M}}^{\boldsymbol{\mu}} \triangleq
\mathbb{E}_{\mathcal{M}}^{\boldsymbol{\mu}}
\left[
\limsup_{T \rightarrow \infty}
\frac{1}{T} \sum_{t=0}^{T-1} c_D(x_t,u_t)
\right],
\end{equation}
which is well suited to persistent tasks over infinite executions. In
particular, an average-cost criterion is more appropriate here than a
stochastic-shortest-path cost, because the latter is tailored to terminating
tasks rather than recurring temporal objectives
\cite{puterman2014markov,bertsekas1995neuro,polychronopoulos1996stochastic}.

\subsubsection{Safe-return constraints}
\label{ch06:subsec:safety}

Safe return is described by a separate LTL formula. This specification captures
the requirement that, after a return request, the robot reaches a designated
safe set and remains there. The requirement is written as:
\begin{equation}
\label{ch06:eq:safe-task}
\varphi_{\texttt{r}} \triangleq
\varphi_{\texttt{r}}^{1} \wedge \Diamond \Box \,\varphi_{\texttt{r}}^{2},
\end{equation}
where $\varphi_{\texttt{r}}^{1}$ is an sc-LTL formula describing the transient
return phase, and $\varphi_{\texttt{r}}^{2}$ identifies the terminal safe set.
More precisely,
$\varphi_{\texttt{r}}^{2} \triangleq \bigvee_{l \in L_s} l$, where
$L_s \subseteq AP$ is the set of labels associated with the safe states.
Equation~\eqref{ch06:eq:safe-task} requires eventual completion of the transient
phase and, afterwards, permanent residence in the safe set
\cite{baier2008principles,pnueli1981temporal}.

Under a return policy
$\boldsymbol{\mu}_{\texttt{r}} \in \overline{\boldsymbol{\mu}}$, the
probability of satisfying the safe-return specification in
\eqref{ch06:eq:safe-task} is given by:
\begin{equation}
\label{ch06:eq:sat-return}
\textbf{Sat}_{\mathcal{M}}^{\boldsymbol{\mu}_{\texttt{r}}} \triangleq
Pr_{\mathcal{M}}^{\boldsymbol{\mu}_{\texttt{r}}}
\bigl(
\{R_{\infty} \in \mathbf{R}_{\mathcal{M}}^{\boldsymbol{\mu}_{\texttt{r}}}
\mid R_{\infty}|_{L} \models \varphi_{\texttt{r}}\}
\bigr),
\end{equation}
where $\mathbf{R}_{\mathcal{M}}^{\boldsymbol{\mu}_{\texttt{r}}}$ denotes the
set of infinite runs generated under the return policy.
For a search-and-rescue mission, a representative safe-return requirement is
$\varphi_{\texttt{r}} = \Diamond \texttt{ex} \wedge
\Diamond \Box \,\texttt{bs}$, which requires the robot to pass through a
designated exit and then remain at the base station.
The role of $\varphi_{\texttt{r}}$ is different from that of the nominal task
$\varphi$. Collision avoidance, emergency response, and similar operational
requirements should normally be encoded directly in $\varphi$. By contrast,
$\varphi_{\texttt{r}}$ expresses recoverability: at any switching time chosen by
an external command, execution should admit a return policy that drives the
system back to safety with sufficiently high probability. This interpretation is
close in spirit to safe exploration and recoverability constraints studied in
\cite{moldovan2012safe,turchetta2016safe}, but it is used here together with a
general temporal mission.
Let $\boldsymbol{\mu}_{\texttt{o}}$ denote the outbound policy for the nominal
task, and let $\boldsymbol{\mu}_{\texttt{r}}$ denote the return policy
activated after a return request. Let $\mathcal{M}'_{\texttt{o}}$ denote the
same MDP as in \eqref{ch06:eq:mdp}, but initialized with the state
distribution induced at the switching instant under
$\boldsymbol{\mu}_{\texttt{o}}$. The safety level of the pair
$(\boldsymbol{\mu}_{\texttt{o}}, \boldsymbol{\mu}_{\texttt{r}})$ is then
defined by:
\begin{equation}
\label{ch06:eq:sat-return-new}
\textbf{Safe}_{\mathcal{M}}^{
\boldsymbol{\mu}_{\texttt{o}},\boldsymbol{\mu}_{\texttt{r}}
}
\triangleq
\textbf{Sat}_{\mathcal{M}'_{\texttt{o}}}^{
\boldsymbol{\mu}_{\texttt{r}}
},
\end{equation}
so \eqref{ch06:eq:sat-return-new} measures the return success probability from
the distribution induced by the outbound execution.

\begin{definition}
\label{ch06:def:safety}
Given the MDP $\mathcal{M}$ and an outbound policy
$\boldsymbol{\mu}_{\texttt{o}}$, the policy
$\boldsymbol{\mu}_{\texttt{o}}$ is called $\chi_{\texttt{r}}$-safe if there
exists a return policy $\boldsymbol{\mu}_{\texttt{r}}$ such that:
\begin{equation}
\label{ch06:eq:safe}
\textbf{Safe}_{\mathcal{M}}^{
\boldsymbol{\mu}_{\texttt{o}},\boldsymbol{\mu}_{\texttt{r}}
}
\geq \chi_{\texttt{r}},
\end{equation}
where $\chi_{\texttt{r}} \in [0,1]$ is the prescribed safety threshold.
\end{definition}

\subsubsection{Problem statement}
\label{ch06:subsec:problem}

\begin{problem}
\label{ch06:prob:main}
Given the labeled MDP $\mathcal{M}$ in \eqref{ch06:eq:mdp}, the task formula
$\varphi$, and the safe-return formula $\varphi_{\texttt{r}}$, synthesize an
outbound policy $\boldsymbol{\mu}_{\texttt{o}}$ and a return policy
$\boldsymbol{\mu}_{\texttt{r}}$ such that
\begin{equation}
\label{ch06:eq:objective}
\begin{aligned}
\min_{
\boldsymbol{\mu}_{\texttt{o}},\boldsymbol{\mu}_{\texttt{r}}
\in \overline{\boldsymbol{\mu}}
}
\quad
&\textbf{Cost}_{\mathcal{M}}^{\boldsymbol{\mu}_{\texttt{o}}}, \\
\text{s.t.}\quad
&\textbf{Sat}_{\mathcal{M}}^{\boldsymbol{\mu}_{\texttt{o}}}
\geq \chi_{\texttt{o}}, \; \textbf{Safe}_{\mathcal{M}}^{
\boldsymbol{\mu}_{\texttt{o}},\boldsymbol{\mu}_{\texttt{r}}
}
\geq \chi_{\texttt{r}};
\end{aligned}
\end{equation}
where $\chi_{\texttt{o}}, \chi_{\texttt{r}} \in [0,1]$ are prescribed lower
bounds on task satisfaction and safe return.
\end{problem}

Problem~\ref{ch06:prob:main} is a two-policy synthesis problem and is
therefore not a standard multi-objective synthesis problem with one policy
that must satisfy all requirements on a single run.
Instead, the outbound policy is admissible only when it preserves the existence
of a return policy satisfying \eqref{ch06:eq:safe}. This distinction is central
throughout the remainder of Sec.~\ref{ch06:sec:ch06_safe_plan}.




%=========================================================
%=========================================================
%=========================================================
\subsection{Product-based reformulation and sequential synthesis}
\label{ch06:sec:analysis}

This subsection reformulates the two constraints in
Problem~\ref{ch06:prob:main} on suitable product automata. The resulting
construction yields a sequential synthesis method: the return policy is
computed first, and the outbound policy is then synthesized under the induced
safe-return values. The product construction follows the standard automata-
based treatment of MDPs under temporal logic specifications
\cite{klein2007ltl2dstar,baier2008principles,belta2017formal,
guo2018probabilistic}.

\subsubsection{Product automata and accepting components}
\label{ch06:subsec:product}

To connect temporal specifications with the MDP model, deterministic Rabin
automata are used. Given an LTL formula \(\psi\) over the proposition set
\(AP\), let \(\mathcal{A}_{\psi}\) denote the associated deterministic Rabin
automaton, written as:
\begin{equation}
\label{ch06:eq:dra}
\mathcal{A}_{\psi} \triangleq
\bigl(
Q_{\psi}, 2^{AP}, \delta_{\psi}, q_{\psi}^{0},
\mathrm{Acc}_{\mathcal{A}_{\psi}}
\bigr),
\end{equation}
where \(Q_{\psi}\) is a finite state set, \(2^{AP}\) is the input alphabet,
\(\delta_{\psi} \colon Q_{\psi} \times 2^{AP} \rightarrow Q_{\psi}\) is the
transition function, \(q_{\psi}^{0} \in Q_{\psi}\) is the initial state, and
\(\mathrm{Acc}_{\mathcal{A}_{\psi}}\) is the Rabin accepting condition.
The accepting condition is given by:
\begin{equation}
\label{ch06:eq:dra-acc}
\mathrm{Acc}_{\mathcal{A}_{\psi}} \triangleq
\left\{
\bigl(H_{\mathcal{A}_{\psi}}^{1}, I_{\mathcal{A}_{\psi}}^{1}\bigr),
\cdots,
\bigl(H_{\mathcal{A}_{\psi}}^{N}, I_{\mathcal{A}_{\psi}}^{N}\bigr)
\right\},
\end{equation}
where
\(H_{\mathcal{A}_{\psi}}^{i}, I_{\mathcal{A}_{\psi}}^{i} \subseteq Q_{\psi}\)
for each \(i = 1,\cdots,N\). Given an infinite word
\(w = l_{0} l_{1} l_{2} \cdots\) over \(2^{AP}\), the corresponding run of
\(\mathcal{A}_{\psi}\) is the sequence
\(q_{0} q_{1} q_{2} \cdots\), where \(q_{0} = q_{\psi}^{0}\) and
\(q_{k+1} = \delta_{\psi}(q_{k}, l_{k})\) for all \(k \geq 0\). This run is
accepting if there exists at least one pair
\(\bigl(H_{\mathcal{A}_{\psi}}^{i}, I_{\mathcal{A}_{\psi}}^{i}\bigr)\) such
that \(H_{\mathcal{A}_{\psi}}^{i}\) is visited only finitely many times and
\(I_{\mathcal{A}_{\psi}}^{i}\) is visited infinitely many times
\cite{baier2008principles}. Translation tools such as \texttt{ltl2dstar} can
be used to construct \(\mathcal{A}_{\psi}\) from \(\psi\)
\cite{klein2007ltl2dstar}.
For the task formula \(\varphi\), let
\begin{equation}
\label{ch06:eq:dra-task}
\mathcal{A}_{\textup{\texttt{o}}} \triangleq
\bigl(
Q_{\textup{\texttt{o}}},
2^{AP},
\delta_{\textup{\texttt{o}}},
q_{\textup{\texttt{o}}}^{0},
\mathrm{Acc}_{\mathcal{A}_{\textup{\texttt{o}}}}
\bigr),
\end{equation}
be the associated deterministic Rabin automaton. The product automaton between
\(\mathcal{M}\) in \eqref{ch06:eq:mdp} and
\(\mathcal{A}_{\textup{\texttt{o}}}\) is defined as follows.

\begin{definition}
\label{ch06:def:product-o}
The product automaton for the nominal task is
\(\mathcal{P}_{\textup{\texttt{o}}} \triangleq
\mathcal{M} \times \mathcal{A}_{\textup{\texttt{o}}}\), given by:
\begin{equation}
\label{ch06:eq:prod-o}
\mathcal{P}_{\textup{\texttt{o}}} \triangleq
\bigl(
S_{\textup{\texttt{o}}},
U_{\textup{\texttt{o}}},
E_{\textup{\texttt{o}}},
p_{E,\textup{\texttt{o}}},
c_{E,\textup{\texttt{o}}},
s^{0}_{\textup{\texttt{o}}},
\mathrm{Acc}_{\mathcal{P}_{\textup{\texttt{o}}}}
\bigr),
\end{equation}
where \(S_{\textup{\texttt{o}}} \subseteq
X \times Q_{\textup{\texttt{o}}}\) is the reachable product-state set,
\(U_{\textup{\texttt{o}}}(\langle x,q\rangle) \triangleq U(x)\), and
\(E_{\textup{\texttt{o}}} \triangleq
\{(s,u) \mid s \in S_{\textup{\texttt{o}}},
u \in U_{\textup{\texttt{o}}}(s)\}\).
The transition probability is given by:
\begin{equation}
\label{ch06:eq:new_pro}
p_{E,\textup{\texttt{o}}}
\bigl(
\langle x,q\rangle, u, \langle \check{x},\check{q}\rangle
\bigr)
\triangleq p_D(x,u,\check{x}),
\end{equation}
where \((x,u) \in D\) and
\(\check{q} = \delta_{\textup{\texttt{o}}}(q,L(x))\).
The stage cost is
\(c_{E,\textup{\texttt{o}}}(\langle x,q\rangle,u)
\triangleq c_D(x,u)\), the initial state
\(s^{0}_{\textup{\texttt{o}}} \triangleq
\langle x_0, q^{0}_{\textup{\texttt{o}}}\rangle\), and the accepting pairs
induced from
\(\mathrm{Acc}_{\mathcal{A}_{\textup{\texttt{o}}}}\) in the standard
manner \cite{baier2008principles}.
\end{definition}

The product automaton \(\mathcal{P}_{\textup{\texttt{o}}}\) represents the
intersection between the traces of \(\mathcal{M}\) and the words accepted by
\(\mathcal{A}_{\textup{\texttt{o}}}\). Its accepting behavior is captured by
accepting maximum end components, abbreviated as AMECs. Let
\(\Xi_{\mathrm{acc}}^{\textup{\texttt{o}}}\) denote the set of AMECs of
\(\mathcal{P}_{\textup{\texttt{o}}}\). If
\((S^{\textup{\texttt{o}}}_{c}, U^{\textup{\texttt{o}}}_{c}) \in
\Xi_{\mathrm{acc}}^{\textup{\texttt{o}}}\), then
\(S^{\textup{\texttt{o}}}_{c}\) is an accepting recurrent region and
\(U^{\textup{\texttt{o}}}_{c}\) specifies the admissible actions that keep
the product inside that region. The union of all accepting components is
\begin{equation}
\label{ch06:eq:AMECs}
S^{\textup{\texttt{o}}}_{\Xi} \triangleq
\bigcup_{
(S^{\textup{\texttt{o}}}_{c},U^{\textup{\texttt{o}}}_{c})
\in \Xi_{\mathrm{acc}}^{\textup{\texttt{o}}}
}
S^{\textup{\texttt{o}}}_{c},
\end{equation}
which is the set of product states from which the accepting condition can be
sustained indefinitely,
see Definitions~10.116, 10.117, 10.124 of~\cite{baier2008principles}.
Let \(\mathcal{A}_{\textup{\texttt{r}}}\) be the deterministic Rabin
automaton associated with the safe-return formula
\(\varphi_{\texttt{r}}\). The corresponding product automaton is defined in
the same way.
The product automaton for the safe-return specification is
\(\mathcal{P}_{\textup{\texttt{r}}} \triangleq
\mathcal{M} \times \mathcal{A}_{\textup{\texttt{r}}}\). Its construction is
the same as in Definition~\ref{ch06:def:product-o}, with
\(\mathcal{A}_{\textup{\texttt{o}}}\) replaced by
\(\mathcal{A}_{\textup{\texttt{r}}}\). Let
\(\Xi_{\mathrm{acc}}^{\textup{\texttt{r}}}\) denote the set of AMECs of
\(\mathcal{P}_{\textup{\texttt{r}}}\), and define
$S^{\textup{\texttt{r}}}_{\Xi} \triangleq
\bigcup_{
(S^{\textup{\texttt{r}}}_{c},U^{\textup{\texttt{r}}}_{c})
\in \Xi_{\mathrm{acc}}^{\textup{\texttt{r}}}
}S^{\textup{\texttt{r}}}_{c}$.
Because both automata are deterministic, each policy on
\(\mathcal{P}_{\textup{\texttt{o}}}\) or
\(\mathcal{P}_{\textup{\texttt{r}}}\) induces a unique policy on
\(\mathcal{M}\), and conversely. The product policies are denoted by
\(\boldsymbol{\pi}_{\textup{\texttt{o}}}\) and
\(\boldsymbol{\pi}_{\textup{\texttt{r}}}\), while the corresponding MDP
policies remain \(\boldsymbol{\mu}_{\textup{\texttt{o}}}\) and
\(\boldsymbol{\mu}_{\textup{\texttt{r}}}\).

\subsubsection{Reformulation of the task and safe-return constraints}
\label{ch06:subsec:task-reformulate}

Task satisfaction can first be expressed as a reachability property on the
product automaton.
Then, the safe-return constraint can be reformulated similarly, but the
associated value is evaluated on the return product and then lifted to the
outbound product through the switching state.
Theorems~\ref{ch06:theorem:task-bound} and
\ref{ch06:theorem:safe-bound} below convert the two constraints in
Problem~\ref{ch06:prob:main} into quantities defined on product automata.
Equation~\eqref{ch06:eq:theo-sat-bound} shows that task satisfaction is a
reachability condition on
\(\widetilde{\mathcal{P}}_{\textup{\texttt{o}}}\), whereas
\eqref{ch06:eq:theo-safe-bound} shows that the safe-return requirement is
evaluated on \(\mathcal{P}_{\textup{\texttt{o}}}\) through the value
function \eqref{ch06:eq:vrstar} computed on
\(\widetilde{\mathcal{P}}_{\textup{\texttt{r}}}\).

\begin{theorem}
\label{ch06:theorem:task-bound}
For any outbound policy
\(\boldsymbol{\mu}_{\textup{\texttt{o}}}\) and its corresponding product
policy \(\boldsymbol{\pi}_{\textup{\texttt{o}}}\), the task-satisfaction
probability in \eqref{ch06:eq:satisfy} satisfies
\begin{equation}
\label{ch06:eq:theo-sat-bound}
\textbf{Sat}_{\mathcal{M}}^{
\boldsymbol{\mu}_{\textup{\texttt{o}}}}
=
\mathbb{E}_{\widetilde{\mathcal{P}}_{\textup{\texttt{o}}}}^{
\boldsymbol{\pi}_{\textup{\texttt{o}}}}
\left[
\sum_{t=0}^{\infty}
b_{s_t,S^{\textup{\texttt{o}}}_{\Xi}}
\right],
\end{equation}
where \(b_{s,S} \triangleq \mathds{1}_{\{s \in S\}}\), and
\(\widetilde{\mathcal{P}}_{\textup{\texttt{o}}}\) is obtained from
\(\mathcal{P}_{\textup{\texttt{o}}}\) by making
\(S^{\textup{\texttt{o}}}_{\Xi}\) absorbing.
\end{theorem}

\begin{proof}
For deterministic Rabin acceptance, a run satisfies \(\varphi\) if and only
if the corresponding run on \(\mathcal{P}_{\textup{\texttt{o}}}\) reaches an
AMEC and then remains inside an accepting component under an admissible
policy \cite{baier2008principles,forejt2011quantitative,
guo2018probabilistic}. Hence, the probability of satisfying \(\varphi\)
equals the probability of reaching
\(S^{\textup{\texttt{o}}}_{\Xi}\) at least once. In the modified product
\(\widetilde{\mathcal{P}}_{\textup{\texttt{o}}}\), the set
\(S^{\textup{\texttt{o}}}_{\Xi}\) is absorbing, so the indicator
\(b_{s_t,S^{\textup{\texttt{o}}}_{\Xi}}\) is counted exactly once along each
accepting trajectory. This yields \eqref{ch06:eq:theo-sat-bound}.
\end{proof}

\begin{theorem}
\label{ch06:theorem:safe-bound}
For any pair of policies
\((\boldsymbol{\mu}_{\textup{\texttt{o}}},
\boldsymbol{\mu}_{\textup{\texttt{r}}})\) and the corresponding product
policies
\((\boldsymbol{\pi}_{\textup{\texttt{o}}},
\boldsymbol{\pi}_{\textup{\texttt{r}}})\), the safe-return quantity in
\eqref{ch06:eq:safe} can be written as
\begin{equation}
\label{ch06:eq:theo-safe-bound}
\textbf{Safe}_{\mathcal{M}}^{
\boldsymbol{\mu}_{\textup{\texttt{o}}},
\boldsymbol{\mu}_{\textup{\texttt{r}}}}
=
\mathbb{E}_{\mathcal{P}_{\textup{\texttt{o}}}}^{
\boldsymbol{\pi}_{\textup{\texttt{o}}}}
\left[
\sum_{t=0}^{\infty}
v^{\star}_{\textup{\texttt{r}}}
\bigl(s^{\textup{\texttt{r}}}_{t}\bigr)
\right],
\end{equation}
where
\(s^{\textup{\texttt{r}}}_{t} \triangleq
\langle x_t, q^{0}_{\textup{\texttt{r}}}\rangle\) is the product state in
\(\mathcal{P}_{\textup{\texttt{r}}}\) associated with the MDP state \(x_t\),
and
\begin{equation}
\label{ch06:eq:vrstar}
v^{\star}_{\textup{\texttt{r}}}(s) \triangleq
\mathbb{E}_{\widetilde{\mathcal{P}}_{\textup{\texttt{r}}}}^{
s,\boldsymbol{\pi}_{\textup{\texttt{r}}}}
\left[
\sum_{\ell=0}^{\infty}
b_{s_{\ell},S^{\textup{\texttt{r}}}_{\Xi}}
\right],
\end{equation}
where \(\widetilde{\mathcal{P}}_{\textup{\texttt{r}}}\) is obtained from
\(\mathcal{P}_{\textup{\texttt{r}}}\) by making
\(S^{\textup{\texttt{r}}}_{\Xi}\) absorbing.
\end{theorem}

\begin{proof}
Consider an arbitrary switching stage \(t \geq 0\). Conditional on the
state \(x_t\) reached under the outbound policy at stage \(t\), the
subsequent evolution of the system is governed by the return policy.
For the modified initial state \(x_t\), the probability of
satisfying \(\varphi_{\texttt{r}}\) under
\(\boldsymbol{\mu}_{\textup{\texttt{r}}}\) is equal to the probability of
reaching \(S^{\textup{\texttt{r}}}_{\Xi}\) in the return product. By the
same argument used in Theorem~\ref{ch06:theorem:task-bound}, this quantity is
given by \(v^{\star}_{\textup{\texttt{r}}}
(s^{\textup{\texttt{r}}}_{t})\) in \eqref{ch06:eq:vrstar}. Averaging these
values over the outbound evolution yields
\eqref{ch06:eq:theo-safe-bound}.
\end{proof}


\subsubsection{Sequential baseline method}
\label{ch06:subsec:complexity}

The reformulation above yields a sequential baseline method. The return
policy is synthesized first, because its value function enters the outbound
optimization through \eqref{ch06:eq:theo-safe-bound}. The return problem is
formulated as:
\begin{equation}
\label{ch06:eq:new-problem-safe}
\max_{
\boldsymbol{\pi}_{\textup{\texttt{r}}}
\in \overline{\boldsymbol{\pi}}
}
\mathbb{E}_{\widetilde{\mathcal{P}}_{\textup{\texttt{r}}}}^{
\boldsymbol{\pi}_{\textup{\texttt{r}}}}
\left[
\sum_{\ell=0}^{\infty}
b_{s_{\ell},S^{\textup{\texttt{r}}}_{\Xi}}
\right],
\end{equation}
which maximizes the reachability of
\(S^{\textup{\texttt{r}}}_{\Xi}\) in
\(\widetilde{\mathcal{P}}_{\textup{\texttt{r}}}\). Once
\(\boldsymbol{\pi}_{\textup{\texttt{r}}}\) is fixed, the value function
\(v^{\star}_{\textup{\texttt{r}}}\) in \eqref{ch06:eq:vrstar} is computed on
\(\widetilde{\mathcal{P}}_{\textup{\texttt{r}}}\).
The outbound policy is then obtained from the constrained problem below:
\begin{equation}
\label{ch06:eq:new-problem}
\begin{aligned}
\min_{
\boldsymbol{\pi}_{\textup{\texttt{o}}}
\in \overline{\boldsymbol{\pi}}
}
\quad
&\textbf{Cost}_{\mathcal{P}_{\textup{\texttt{o}}}}^{
\boldsymbol{\pi}_{\textup{\texttt{o}}}}, \\
\text{s.t.}\quad
&\mathbb{E}_{\widetilde{\mathcal{P}}_{\textup{\texttt{o}}}}^{
\boldsymbol{\pi}_{\textup{\texttt{o}}}}
\left[
\sum_{t=0}^{\infty}
b_{s_t,S^{\textup{\texttt{o}}}_{\Xi}}
\right]
\geq \chi_{\texttt{o}}, \quad
\mathbb{E}_{\mathcal{P}_{\textup{\texttt{o}}}}^{
\boldsymbol{\pi}_{\textup{\texttt{o}}}}
\left[
\sum_{t=0}^{\infty}
v^{\star}_{\textup{\texttt{r}}}
\bigl(s^{\textup{\texttt{r}}}_{t}\bigr)
\right]
\geq \chi_{\texttt{r}};
\end{aligned}
\end{equation}
where the first constraint follows from
\eqref{ch06:eq:theo-sat-bound}, and the second follows from
\eqref{ch06:eq:theo-safe-bound}. Equation~\eqref{ch06:eq:new-problem} is
therefore a constrained MDP problem
\cite{altman1999constrained,bertsekas1991analysis}. A standard occupancy-
measure linear program can be used to synthesize the prefix and suffix parts
of the outbound policy, as in related probabilistic temporal-logic planning
methods \cite{guo2018probabilistic,ding2014optimal}. The mean-cost objective
applies only to the outbound policy, which is consistent with
\eqref{ch06:eq:objective}.
The resulting procedure is summarized in Algorithm~\ref{ch06:alg:direct}.
Although the two linear programs are polynomial in the sizes of the
corresponding product MDPs, the products themselves grow rapidly with the
size of \(\mathcal{M}\) and with the sizes of the automata associated with
\(\varphi\) and \(\varphi_{\texttt{r}}\). In particular,
\(\mathcal{P}_{\textup{\texttt{o}}}\) is typically much larger than
\(\mathcal{P}_{\textup{\texttt{r}}}\), and the constrained optimization in
\eqref{ch06:eq:new-problem} becomes the dominant computational bottleneck.
For this reason, the method above is best viewed as a baseline reference and
as a starting point for the more structured synthesis developed next.

\begin{algorithm}[t]
\caption{Sequential baseline method}
\label{ch06:alg:direct}
\LinesNumbered
\KwIn{\(\mathcal{M}\), \((\varphi,\chi_{\texttt{o}})\), and
\((\varphi_{\texttt{r}},\chi_{\texttt{r}})\).}
\KwOut{\((\mathcal{P}_{\textup{\texttt{r}}},\boldsymbol{\pi}_{\textup{\texttt{r}}},
v^{\star}_{\textup{\texttt{r}}})\), and
\((\mathcal{P}_{\textup{\texttt{o}}},\boldsymbol{\pi}_{\textup{\texttt{o}}})\).}
Construct \(\mathcal{P}_{\textup{\texttt{r}}}\) and compute
\(S^{\textup{\texttt{r}}}_{\Xi}\)\;
Solve \eqref{ch06:eq:new-problem-safe} for
\(\boldsymbol{\pi}_{\textup{\texttt{r}}}\) and
\(v^{\star}_{\textup{\texttt{r}}}\)\;
Construct \(\mathcal{P}_{\textup{\texttt{o}}}\) and compute
\(S^{\textup{\texttt{o}}}_{\Xi}\)\;
Solve \eqref{ch06:eq:new-problem} for
\(\boldsymbol{\pi}_{\textup{\texttt{o}}}\)\;
\end{algorithm}

















%=========================================================
%=========================================================
%=========================================================
\subsection{Hierarchical abstraction and safety-ensured synthesis}
\label{ch06:sec:solution}

The baseline method in Sec.~\ref{ch06:sec:analysis} becomes expensive because
both temporal constraints are enforced directly on product automata built over
the original MDP \(\mathcal{M}\). To reduce this complexity, two feature-based
abstractions are introduced as semi-MDPs. One abstraction is constructed for
the safe-return specification, and the other is constructed for the nominal
task. Planning is then carried out on these abstractions rather than on the
full model.
The construction follows the semi-MDP and option framework in
\cite{sutton1999between}. Each abstract transition represents a motion policy
between two feature states. As in the baseline method, the return policy is
synthesized first, because its value function is required when building the
task-level abstraction. The overall structure is hierarchical: the return policy is computed first, then used to construct the task-level abstraction for outbound planning.


\subsubsection{Hierarchical synthesis of the return policy}
\label{ch06:subsec:safe-return-plan}

The return policy is synthesized first, together with the associated value
function that will later be used in the task abstraction. This requires a
feature-based abstraction tailored to the safe-return specification
\(\varphi_{\texttt{r}}\), followed by policy synthesis on the corresponding
abstract product.

(I) \emph{Return-level abstraction.}
Let
\(\mathcal{A}_{\textup{\texttt{r}}} \triangleq
(Q_{\textup{\texttt{r}}}, 2^{AP}, \delta_{\textup{\texttt{r}}},
q^{0}_{\textup{\texttt{r}}},
\mathrm{Acc}_{\mathcal{A}_{\textup{\texttt{r}}}})\)
be the deterministic Rabin automaton associated with
\(\varphi_{\texttt{r}}\). The effective labels for the return task are defined
by
\begin{equation}
\label{ch06:eq:effectve-safe}
\Theta_{\textup{\texttt{r}}} \triangleq
\left\{
\theta \in 2^{AP}
\;\middle|\;
\exists q \in Q_{\textup{\texttt{r}}}
\text{ such that }
\delta_{\textup{\texttt{r}}}(q,\theta) \neq q
\right\},
\end{equation}
which collects the labels that induce non-self transitions in
\(\mathcal{A}_{\textup{\texttt{r}}}\).
Using \(\Theta_{\textup{\texttt{r}}}\), a feature-based abstraction is
constructed as the labeled semi-MDP
\begin{equation}
\label{ch06:eq:abstract-mdp-safe}
\mathcal{M}'_{\textup{\texttt{r}}} \triangleq
\bigl(
X'_{\textup{\texttt{r}}},
U'_{\textup{\texttt{r}}},
p'_{\textup{\texttt{r}},D},
AP,
L'_{\textup{\texttt{r}}}
\bigr),
\end{equation}
where
$X'_{\textup{\texttt{r}}} \triangleq
\left\{
x \in X
\;\middle|\;
L(x) \in \Theta_{\textup{\texttt{r}}}
\right\}$,
and \(L'_{\textup{\texttt{r}}}\) is the restriction of the labeling map to
\(X'_{\textup{\texttt{r}}}\). Thus,
\(X'_{\textup{\texttt{r}}}\) contains only those states that are relevant to
the progress of \(\mathcal{A}_{\textup{\texttt{r}}}\). The abstract action set
\(U'_{\textup{\texttt{r}}}\) consists of macro actions, each of which
represents an underlying motion policy between two abstract states. The
transition probability
\(p'_{\textup{\texttt{r}},D} \colon
X'_{\textup{\texttt{r}}} \times
U'_{\textup{\texttt{r}}} \times
X'_{\textup{\texttt{r}}} \rightarrow [0,1]\)
is obtained as follows.

For each ordered pair
\((x_f,x_t) \in
X'_{\textup{\texttt{r}}} \times X'_{\textup{\texttt{r}}}\), the associated
macro action is denoted by
\((x_f,x_t) \in U'_{\textup{\texttt{r}}}\). First, a modified MDP
\(\mathcal{M}_{\textup{\texttt{r}}}(x_f,x_t)\) is constructed from
\(\mathcal{M}\), where \(x_f\) is the source state and every state in
\(X'_{\textup{\texttt{r}}} \setminus \{x_f\}\) is treated as a sink state.
Once a sink state is reached, the process remains there through a self-loop
with probability one and zero additional cost. Second, a motion policy
\(\boldsymbol{\mu}'_{\textup{\texttt{r}}}(x_f,x_t)\) is synthesized to
maximize the probability of reaching the target sink \(x_t\). This policy
plays the role of an option in the sense of \cite{sutton1999between}. As in
\eqref{ch06:eq:new-problem-safe}, the synthesis reduces to a maximum-
reachability problem and can be solved by a linear program over state-action
occupancy measures. Third, once
\(\boldsymbol{\mu}'_{\textup{\texttt{r}}}(x_f,x_t)\) is fixed,
\(\mathcal{M}_{\textup{\texttt{r}}}(x_f,x_t)\) becomes a Markov chain, and the
limiting distribution over the sink states defines the abstract transition
probability
\(p'_{\textup{\texttt{r}},D}
(x_f,\boldsymbol{\mu}'_{\textup{\texttt{r}}}(x_f,x_t),x_\ell)\)
for every \(x_\ell \in X'_{\textup{\texttt{r}}}\). If the corresponding linear
program has no feasible solution, then \(x_t\) is not reachable from \(x_f\)
through the abstraction and the macro transition is omitted.
If the automaton
\(\mathcal{A}_{\textup{\texttt{r}}}\) contains \(N\) distinct relevant labels,
and at most \(M\) regions of interest are associated with each label in the
workspace, then \(\mathcal{M}'_{\textup{\texttt{r}}}\) has at most \(MN\)
states and fewer than \(M^2N^2\) abstract transitions. In large workspaces
with relatively few relevant regions, this abstraction can be substantially
smaller than the original MDP.


(II) \emph{Abstract return-policy synthesis.}
Given \(\mathcal{M}'_{\textup{\texttt{r}}}\) and the automaton
\(\mathcal{A}_{\textup{\texttt{r}}}\), their product is given by:
\begin{equation}
\label{ch06:eq:abstract-safe-product}
\mathcal{P}'_{\textup{\texttt{r}}} \triangleq
\mathcal{M}'_{\textup{\texttt{r}}}
\times
\mathcal{A}_{\textup{\texttt{r}}}
=
\bigl(
S'_{\textup{\texttt{r}}},
U'_{\textup{\texttt{r}}},
E'_{\textup{\texttt{r}}},
p'_{\textup{\texttt{r}},E},
s'^{0}_{\textup{\texttt{r}}},
\mathrm{Acc}'_{\textup{\texttt{r}}}
\bigr),
\end{equation}
where the notation is defined in the same way as in
Definition~\ref{ch06:def:product-o}. The transition probability
\(p'_{\textup{\texttt{r}},E}\) is induced from
\(p'_{\textup{\texttt{r}},D}\) in \eqref{ch06:eq:abstract-mdp-safe}.
The optimal abstract return policy
\(\boldsymbol{\pi}'_{\textup{\texttt{r}}}\) and the associated value function
\(v'^{\star}_{\textup{\texttt{r}}}\) are computed from
\(\mathcal{P}'_{\textup{\texttt{r}}}\) in the same manner as
\(\boldsymbol{\pi}_{\textup{\texttt{r}}}\) was computed from
\(\mathcal{P}_{\textup{\texttt{r}}}\) in
Sec.~\ref{ch06:subsec:complexity}. Since the safe-return specification is of
sc-LTL type, satisfaction reduces to reaching the union of AMECs in the
abstract product. Denote this set by \(S'_{\textup{\texttt{r}},\Xi}\). Once
\(S'_{\textup{\texttt{r}},\Xi}\) is reached, the safe-return objective can be
maintained within the accepting component.

\subsubsection{Hierarchical synthesis of the nominal task}
\label{ch06:subsec:task-plan}

After the return policy has been obtained, the nominal task is treated in the
same hierarchical spirit. The task abstraction differs from the return
abstraction in that it already incorporates the safe-return constraint through
the value function \(v'^{\star}_{\textup{\texttt{r}}}\).

(I) \emph{Safety-aware task abstraction.}
Let
\(\mathcal{A}_{\textup{\texttt{o}}}\) be the deterministic Rabin automaton
associated with \(\varphi\). The corresponding set of effective labels is
defined analogously to \eqref{ch06:eq:effectve-safe} and is denoted by
\(\Theta_{\textup{\texttt{o}}}\). The task-level abstraction is then given by
the semi-MDP below:
\begin{equation}
\label{ch06:eq:abstract-mdp-task}
\mathcal{M}'_{\textup{\texttt{o}}} \triangleq
\bigl(
X'_{\textup{\texttt{o}}},
U'_{\textup{\texttt{o}}},
p'_{\textup{\texttt{o}},D},
AP,
L'_{\textup{\texttt{o}}},
c'_{\textup{\texttt{o}},D}
\bigr),
\end{equation}
where the notation is analogous to
\eqref{ch06:eq:abstract-mdp-safe}. The main difference is that the abstract
transitions and abstract costs are computed under the additional safe-return
requirement.
For each ordered pair
\((x_f,x_t) \in
X'_{\textup{\texttt{o}}} \times X'_{\textup{\texttt{o}}}\), the associated
macro action is written as
\((x_f,x_t) \in U'_{\textup{\texttt{o}}}\). As before, a modified MDP
\(\mathcal{M}_{\textup{\texttt{o}}}(x_f,x_t)\) is constructed from the
original model. A motion policy
\(\boldsymbol{\mu}'_{\textup{\texttt{o}}}(x_f,x_t)\) is then synthesized to
maximize the probability of reaching \(x_t\), while also satisfying the
safe-return bound. By Theorem~\ref{ch06:theorem:safe-bound}, the latter can be
imposed through the constraint below:
\begin{equation}
\label{ch06:eq:abstract-task-lp}
\mathbb{E}_{\mathcal{M}_{\textup{\texttt{o}}}(x_f,x_t)}^{
\boldsymbol{\mu}'_{\textup{\texttt{o}}}}
\left[
\sum_{t=0}^{\infty}
v'^{\star}_{\textup{\texttt{r}}}(s^{\textup{\texttt{r}}}_t)
\right]
\geq
\chi_{\texttt{r}},
\end{equation}
where
\(s^{\textup{\texttt{r}}}_t \triangleq
\langle x_t,q^{0}_{\textup{\texttt{r}}}\rangle\)
is the product state in
\(\mathcal{P}'_{\textup{\texttt{r}}}\) associated with the MDP state
encountered at stage \(t\).

Once the option
\(\boldsymbol{\mu}'_{\textup{\texttt{o}}}(x_f,x_t)\) is fixed, the induced
Markov chain yields the abstract transition probabilities
\(p'_{\textup{\texttt{o}},D}\). The corresponding abstract cost is defined by:
\begin{equation}
\label{ch06:eq:task-edge-cost}
c'_{\textup{\texttt{o}},D}(x_f,x_t) \triangleq
\mathbb{E}_{\mathcal{M}_{\textup{\texttt{o}}}(x_f,x_t)}^{
\boldsymbol{\mu}'_{\textup{\texttt{o}}}}
\left[
\sum_{t=0}^{\infty}
c_D(x_t,u_t)
\right],
\end{equation}
that is, the expected cumulative cost of executing the option from \(x_f\)
until termination at one of the sink states. If the constrained linear program
is infeasible, then the transition \((x_f,x_t)\) is removed from
\(U'_{\textup{\texttt{o}}}\).
This local synthesis emphasizes reachability and safety rather than exact cost
optimality for each macro transition. Consequently,
\(c'_{\textup{\texttt{o}},D}\) need not coincide with the minimum possible
cost between the corresponding feature states. This introduces an
approximation with respect to the original objective in
\eqref{ch06:eq:objective}, but avoids the much higher complexity of a global
optimization on the full product automaton.

(II) \emph{Abstract outbound-policy synthesis.}
Given the safety-ensured semi-MDP
\(\mathcal{M}'_{\textup{\texttt{o}}}\), the corresponding product automaton is
given by:
\begin{equation}
\label{ch06:eq:abstract-task-product}
\mathcal{P}'_{\textup{\texttt{o}}} \triangleq
\mathcal{M}'_{\textup{\texttt{o}}}
\times
\mathcal{A}_{\textup{\texttt{o}}},
\end{equation}
where the detailed notation is defined as in
Definition~\ref{ch06:def:product-o}, and the edge cost is induced by
\eqref{ch06:eq:task-edge-cost}.
Unlike the baseline product
\(\mathcal{P}_{\textup{\texttt{o}}}\), the abstract product
\(\mathcal{P}'_{\textup{\texttt{o}}}\) inherits the safe-return guarantee from
Lemma~\ref{ch06:lemma:safe}. Hence, no additional safe-return constraint is
imposed when synthesizing the outbound policy on
\(\mathcal{P}'_{\textup{\texttt{o}}}\). Since \(\varphi\) is a general LTL
formula, the policy consists of a prefix part
\(\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{pre}}\), executed until an
accepting component is reached, and a suffix part
\(\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{suf}}\), executed
thereafter. Following the framework in \cite{guo2018probabilistic}, the two
parts are obtained from the constrained optimization below:
\begin{subequations}
\label{ch06:eq:abstract-prod-task-lp}
\begin{align}
\min_{
\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{pre}},
\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{suf}}
}
\quad
&\textbf{Cost}_{\mathcal{P}'_{\textup{\texttt{o}}}}^{
\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{suf}}},
\label{ch06:eq:abstract-prod-task-lp-objective} \\
\text{s.t.}\quad
&\textbf{Sat}_{\mathcal{P}'_{\textup{\texttt{o}}}}^{
\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{pre}}}
\geq \chi_{\texttt{o}},
\label{ch06:eq:abstract-prod-task-lp-condition}
\end{align}
\end{subequations}
where the prefix policy ensures that the union of AMECs, denoted by
\(S'_{\textup{\texttt{o}},\Xi}\), is reached with probability at least
\(\chi_{\texttt{o}}\), while the suffix policy keeps the execution inside
\(S'_{\textup{\texttt{o}},\Xi}\) and minimizes the corresponding long-term
cost. The detailed linear program is omitted here.

\subsubsection{Algorithmic summary, online execution, and guarantee}
\label{ch06:subsec:summary}

The complete procedure is summarized in
Algorithm~\ref{ch06:alg:all}. The first stage constructs the return-level
abstraction \(\mathcal{M}'_{\textup{\texttt{r}}}\), forms the abstract
product \(\mathcal{P}'_{\textup{\texttt{r}}}\), and synthesizes the return
policy together with the value function
\(v'^{\star}_{\textup{\texttt{r}}}\). The second stage constructs the
safety-aware task abstraction \(\mathcal{M}'_{\textup{\texttt{o}}}\), forms
the task product \(\mathcal{P}'_{\textup{\texttt{o}}}\), and synthesizes the
prefix and suffix parts of the outbound policy.
During online execution, the high-level policy selects a macro action between
two abstract states, and the associated option executes the corresponding
low-level motion policy on the original MDP. For the outbound phase, starting
from the current abstract state \(x_k\), the prefix policy
\(\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{pre}}\) selects the next
target feature state \(x^{\star}_{k+1}\). The option
\(\boldsymbol{\mu}'_{\textup{\texttt{o}}}(x_k,x^{\star}_{k+1})\) is then
executed on \(\mathcal{M}\) until the trajectory reaches some sink state in
\(X'_{\textup{\texttt{o}}}\). Because the option is stochastic, the realized
next state need not equal the intended target. The high-level policy is then
reapplied from the realized state. This process repeats until the execution
reaches \(S'_{\textup{\texttt{o}},\Xi}\), after which the suffix policy
\(\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{suf}}\) is used. The
safe-return policy is invoked only when a return request is issued. Its
execution is analogous, except that it terminates after reaching
\(S'_{\textup{\texttt{r}},\Xi}\). The execution therefore consists of an outbound phase that switches from prefix to suffix control after entering \(S'_{\textup{\texttt{o}},\Xi}\), and a return phase that terminates upon reaching \(S'_{\textup{\texttt{r}},\Xi}\).
The algorithmic difference between
Algorithm~\ref{ch06:alg:direct} and
Algorithm~\ref{ch06:alg:all} is as follows. The baseline method
synthesizes the return and outbound policies directly on product automata built
from the original MDP \(\mathcal{M}\). The hierarchical method first
constructs the abstractions \(\mathcal{M}'_{\textup{\texttt{r}}}\) and
\(\mathcal{M}'_{\textup{\texttt{o}}}\), together with their associated
options, and then synthesizes the high-level policies on the corresponding
abstract products. This separation between low-level motion synthesis and
high-level temporal planning is the main source of the computational savings.

\begin{algorithm}[t]
\caption{Hierarchical synthesis algorithm}
\label{ch06:alg:all}
\LinesNumbered
\KwIn{\(\mathcal{M}\), \((\varphi,\chi_{\texttt{o}})\), and
\((\varphi_{\texttt{r}},\chi_{\texttt{r}})\).}
\KwOut{\((\mathcal{M}'_{\textup{\texttt{r}}},
\boldsymbol{\mu}'_{\textup{\texttt{r}}}),
(\mathcal{P}'_{\textup{\texttt{r}}},
\boldsymbol{\pi}'_{\textup{\texttt{r}}},
v'^{\star}_{\textup{\texttt{r}}})\), and
\((\mathcal{M}'_{\textup{\texttt{o}}},
\boldsymbol{\mu}'_{\textup{\texttt{o}}}),
(\mathcal{P}'_{\textup{\texttt{o}}},
(\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{pre}},
\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{suf}}))\).}
Construct \(\mathcal{A}_{\textup{\texttt{r}}}\)\;
Build \(\mathcal{M}'_{\textup{\texttt{r}}}\) and its options
\(\boldsymbol{\mu}'_{\textup{\texttt{r}}}\)\;
Construct \(\mathcal{P}'_{\textup{\texttt{r}}}\) by
\eqref{ch06:eq:abstract-safe-product}\;
Synthesize \(\boldsymbol{\pi}'_{\textup{\texttt{r}}}\) and
\(v'^{\star}_{\textup{\texttt{r}}}\)\;
Construct \(\mathcal{A}_{\textup{\texttt{o}}}\)\;
Build \(\mathcal{M}'_{\textup{\texttt{o}}}\) and its options
\(\boldsymbol{\mu}'_{\textup{\texttt{o}}}\), using
\(v'^{\star}_{\textup{\texttt{r}}}\)\;
Construct \(\mathcal{P}'_{\textup{\texttt{o}}}\) by
\eqref{ch06:eq:abstract-task-product}\;
Synthesize
\((\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{pre}},
\boldsymbol{\pi}'_{\textup{\texttt{o}},\texttt{suf}})\) from
\eqref{ch06:eq:abstract-prod-task-lp}\;
\end{algorithm}



\begin{lemma}
\label{ch06:lemma:safe}
The semi-MDP \(\mathcal{M}'_{\textup{\texttt{o}}}\) is safety-ensured.
More precisely, for every transition
\((x_f,x_t) \in U'_{\textup{\texttt{o}}}\), the associated motion policy
\(\boldsymbol{\mu}'_{\textup{\texttt{o}}}(x_f,x_t)\), together with the
return policy \(\boldsymbol{\mu}'_{\textup{\texttt{r}}}\), satisfies the
safe-return constraint in \eqref{ch06:eq:safe}.
\end{lemma}

\begin{proof}
For each macro transition \((x_f,x_t)\), the motion policy
\(\boldsymbol{\mu}'_{\textup{\texttt{o}}}(x_f,x_t)\) is synthesized subject
to \eqref{ch06:eq:abstract-task-lp}. By
Theorem~\ref{ch06:theorem:safe-bound}, the left-hand side of
\eqref{ch06:eq:abstract-task-lp} is exactly the safe-return probability
associated with the option execution and the return policy. Hence, every
admissible transition in \(\mathcal{M}'_{\textup{\texttt{o}}}\) preserves the
probability bound \(\chi_{\texttt{r}}\). The entire abstraction is therefore
safety-ensured.
\end{proof}


\begin{theorem}
\label{ch06:thm:safety-guarantee}
The hierarchical policies
\(\boldsymbol{\pi}'_{\textup{\texttt{r}}}\) and
\(\boldsymbol{\pi}'_{\textup{\texttt{o}}}\) satisfy the task and safe-return
constraints in \eqref{ch06:eq:objective}. Moreover, the optimization in
\eqref{ch06:eq:abstract-prod-task-lp} minimizes the cost induced on the
abstract product \(\mathcal{P}'_{\textup{\texttt{o}}}\).
\end{theorem}

\begin{proof}
By Lemma~\ref{ch06:lemma:safe}, every macro transition in
\(\mathcal{M}'_{\textup{\texttt{o}}}\) satisfies the safe-return bound
\(\chi_{\texttt{r}}\) together with the return policy
\(\boldsymbol{\mu}'_{\textup{\texttt{r}}}\). Hence, any execution generated by
the outbound policy on \(\mathcal{P}'_{\textup{\texttt{o}}}\) preserves the
safe-return constraint in \eqref{ch06:eq:objective}.
For the task requirement, the prefix-suffix synthesis on
\(\mathcal{P}'_{\textup{\texttt{o}}}\) follows the same automata-based
construction used in \cite{guo2018probabilistic}. In particular, the prefix
policy ensures that the union of AMECs \(S'_{\textup{\texttt{o}},\Xi}\) is
reached with probability at least \(\chi_{\texttt{o}}\), and the suffix policy
maintains the execution within the accepting component. Therefore, the
task-satisfaction bound in \eqref{ch06:eq:objective} is satisfied.
Finally, \eqref{ch06:eq:abstract-prod-task-lp-objective} minimizes the cost
defined on the abstract product \(\mathcal{P}'_{\textup{\texttt{o}}}\).
Because the local options used to construct
\(\mathcal{M}'_{\textup{\texttt{o}}}\) prioritize reachability and safety
rather than exact transition-cost minimization, this hierarchical cost is an
approximation of the original cost in \eqref{ch06:eq:objective}. The
constraint satisfaction, however, is preserved.
\end{proof}
