%==============================
\section{Collaborative Optimal Transport under Signal Temporal Tasks}
\label{ch06:sec:ch06_stl}


The preceding sections of this chapter address uncertainty and reactivity
primarily at the task-planning and coordination levels.
Section~\ref{ch06:sec:ch06_safe_plan} studies probabilistic temporal tasks
under safe-return requirements, whereas
Section~\ref{ch06:sec:ch06_uncertain} develops an online coordination method
for dynamic temporal tasks with moving targets.
A complementary difficulty arises when the task specification is fixed,
but the collective motion must still be synthesized in a continuous
workspace under nonlinear dynamics, obstacle avoidance, inter-robot
coupling, and high-level signal temporal logic (STL) constraints.
In this setting, the central challenge is not merely to decide
\emph{which} task-relevant behaviors should occur, but to compute
feasible and high-quality trajectories for all robots in a
high-dimensional and strongly non-convex search space.
Existing mixed-integer encodings and analytical planners can model parts
of this problem, but their scalability often deteriorates rapidly with
the planning horizon, the number of robots, and the logical structure of
the task; sampling-based planners, on the other hand, may be less
effective in exploiting the quantitative information carried by STL
robustness.

This section addresses that gap by formulating multi-robot motion
planning under STL tasks as a collaborative trajectory-optimization
problem.
Smooth candidate trajectories are represented through Gaussian-process
priors, while dynamic feasibility, geometric safety requirements, and
task satisfaction are incorporated into a unified cost through control
penalties, collision-avoidance terms, and STL robustness measures.
The resulting solution architecture has three tightly connected parts.
First, Section~\ref{ch06:subsec:batch-stl} develops a batch evaluation rule
for STL robustness under waypoint perturbations, so that task-related
cost variations can be assessed efficiently across many candidate
updates.
Second, Section~\ref{ch06:sec:local} introduces a sequential local optimal
transport procedure that refines robot trajectories through zero-order
updates while accounting for inter-robot couplings.
Third, the hybrid search layer in
Section~\ref{ch06:subsec:search-algorithm} explores alternative planning
orders and partial trajectory sets to improve scalability and solution
quality.
In this way, the section complements the symbolic and reactive methods
developed earlier in the chapter with a continuous optimization
framework for tightly coupled multi-robot trajectory synthesis.







%==============================
%==============================
%==============================
\subsection{Problem formulation}
\label{ch06:subsec:problem_formulation}

This subsection introduces the continuous multi-robot planning model, the STL task semantics, and the collaborative motion-planning problem solved in the remainder of the section.

%==============================
\subsubsection{Multi-robot system}
\label{ch06:sec:multi-robot-system}

Consider a system of $N$ robots indexed by
$\mathcal{N} \triangleq \{1,\cdots,N\}$ and evolving in a workspace
$\mathcal{W}\subset\mathbb{R}^D$, where $D\geq 2$.
Each robot $i\in\mathcal{N}$ follows the discrete-time first-order dynamics
\begin{equation}
\label{ch06:eq:robot-model}
\mathbf{p}_i(t+1)=\mathbf{p}_i(t)+\mathbf{v}_i(t)\Delta t,
\end{equation}
where $\mathbf{p}_i(t),\mathbf{v}_i(t)\in\mathbb{R}^D$ denote,
respectively, the position and velocity of robot $i$ at time
$t\in\{0,\cdots,T\}$, and $\Delta t>0$ is the sampling period.
Each robot occupies a physical region $\mathcal{A}_i(t)$ at time $t$,
for instance a disk centered at $\mathbf{p}_i(t)$.
The full state of robot $i$ is defined by
$\mathbf{q}_i(t)\triangleq (\mathbf{p}_i(t),\mathbf{v}_i(t))$.
For the team-level description, let
$\mathbf{x}(t)\triangleq [\mathbf{p}_1(t),\cdots,\mathbf{p}_N(t)]$ and
$\mathbf{u}(t)\triangleq [\mathbf{v}_1(t),\cdots,\mathbf{v}_N(t)]$
be the stacked position and control vectors, respectively.
The collective state is then
$\mathbf{z}(t)\triangleq (\mathbf{x}(t),\mathbf{u}(t))$.
Over a horizon of length $T$, the system trajectory is defined by
$\mathbf{Z}(T)\triangleq [\mathbf{z}(0),\cdots,\mathbf{z}(T)]^\intercal$,
and $\mathbf{Z}_i(T)$ denotes the corresponding trajectory of robot $i$.
The workspace also contains static obstacles represented by
$\mathcal{O}\subset\mathcal{W}$.

The control effort and trajectory smoothness of robot $i$ are measured by a
stage cost
$h_i:\mathbb{R}^{2D}\times\mathbb{R}^{2D}\rightarrow\mathbb{R}^+$.
Using $h_i$, the cumulative control cost over $\mathbf{Z}_i(T)$ is defined by
\begin{equation}
\label{ch06:eq:control-cost}
c_{\texttt{ctr}}\big(\mathbf{Z}_i(T)\big)
\triangleq \sum_{t=0}^{T-1}
h_i\big(\mathbf{q}_i(t),\mathbf{q}_i(t+1)\big),
\end{equation}
where the function $h_i$ may differ across robots to capture
heterogeneous motion characteristics.
In addition to \eqref{ch06:eq:control-cost}, the collective trajectory is
penalized by state-dependent costs associated with obstacle avoidance,
inter-robot separation, and task satisfaction.
For obstacle avoidance, define
$g_{\texttt{obs}}\big(\mathbf{p}_i(t)\big)\triangleq
\max\{0,r_{\texttt{o}}-\texttt{MSD}(\mathcal{A}_i(t),\mathcal{O})\}$,
where $r_{\texttt{o}}>0$ is a safety margin and
$\texttt{MSD}(\cdot,\cdot)$ is the minimum signed distance between two
manifolds~\cite{schulman2014motion}.
For inter-robot collision avoidance, define
$g_{\texttt{int}}\big(\mathbf{p}_i(t),\mathbf{p}_j(t)\big)\triangleq
\max\{0,r_{\texttt{s}}-
\texttt{MSD}(\mathcal{A}_i(t),\mathcal{A}_j(t))\}$,
where $r_{\texttt{s}}>0$ is the required pairwise safety margin.
Both penalties vanish when the corresponding clearance requirement is met.



%==============================
\subsubsection{Signal temporal logic tasks and robustness}
\label{ch06:subsubsec:stl_robustness}

Let $\mathbf{s}:\mathbb{N}\rightarrow\mathcal{X}$ be a discrete-time signal,
where $\mathcal{X}$ denotes the signal space.
A predicate is an expression of the form $f(\mathbf{s})>0$, where
$f:\mathcal{X}\rightarrow\mathbb{R}$.
An STL formula $\phi$ is generated recursively as
$\phi ::= \mu \mid \lnot \phi \mid \phi_1 \wedge \phi_2
\mid \phi_1 \vee \phi_2 \mid \phi_1 \mathsf{U}_I \phi_2$,
where $\mu$ is a predicate and
$I \triangleq [t_1,t_2]$ is a bounded time interval satisfying
$0 \leq t_1 \leq t_2$.
The operator $\mathsf{U}_I$ requires $\phi_1$ to hold until $\phi_2$
becomes true within $I$.
The operators $\mathsf{F}_I$ and $\mathsf{G}_I$ are derived from
$\mathsf{U}_I$ in the usual way.
Only bounded-time STL formulas are considered here.
The satisfaction of $\phi$ by $\mathbf{s}$ at time $t$ is denoted by
$(\mathbf{s},t)\models \phi$.

To quantify the degree of satisfaction, the space robustness
$\rho^\phi(\mathbf{s},t)\in\mathbb{R}$ is adopted.
Its recursive definition is given in \eqref{ch06:eq:stl-robustness}:
\begin{equation}
\label{ch06:eq:stl-robustness}
\begin{aligned}
\rho^\mu(\mathbf{s},t)
&\triangleq f(\mathbf{s}(t)), \\
\rho^{\lnot\phi}(\mathbf{s},t)
&\triangleq -\rho^\phi(\mathbf{s},t), \\
\rho^{\phi_1\wedge\phi_2}(\mathbf{s},t)
&\triangleq \min\big\{
\rho^{\phi_1}(\mathbf{s},t),\rho^{\phi_2}(\mathbf{s},t)\big\}, \\
\rho^{\phi_1\vee\phi_2}(\mathbf{s},t)
&\triangleq \max\big\{
\rho^{\phi_1}(\mathbf{s},t),\rho^{\phi_2}(\mathbf{s},t)\big\}, \\
\rho^{\mathsf{G}_I\phi}(\mathbf{s},t)
&\triangleq \min_{t'\in t\oplus I}\rho^\phi(\mathbf{s},t'), \\
\rho^{\mathsf{F}_I\phi}(\mathbf{s},t)
&\triangleq \max_{t'\in t\oplus I}\rho^\phi(\mathbf{s},t'), \\
\rho^{\phi_1\mathsf{U}_I\phi_2}(\mathbf{s},t)
&\triangleq \max_{t'\in t\oplus I}
\min\Big\{\rho^{\phi_2}(\mathbf{s},t'),
\min_{t''\in (t,t')}\rho^{\phi_1}(\mathbf{s},t'')\Big\},
\end{aligned}
\end{equation}
where $t\oplus I$ denotes the translation of the interval $I$ by $t$.
A positive value of $\rho^\phi(\mathbf{s},t)$ implies that
$(\mathbf{s},t)\models \phi$ holds
under the standard quantitative semantics of STL
\cite{fainekos2009robustness,deshmukh2017robust,lindemann2025formal}.



The task-related term is defined through STL robustness.
Two classes of atomic predicates are used.
(I) \emph{Regional predicates}.
Given a region $\mathcal{W}_k\subset\mathcal{W}$, let
$\mu_A^k\triangleq
-\max_{i=1,\cdots,N}\big(
\texttt{MSD}(\mathcal{A}_i(t),\mathcal{W}_k)+2r\big)\geq 0$,
where $r$ is the robot radius.
The predicate $\mu_A^k$ is satisfied when all robots lie inside
$\mathcal{W}_k$ at time $t$.
(II) \emph{Collaborative predicates}.
Let $\mu_B^k\triangleq \mathcal{F}(\mathbf{z}(t))\geq 0$, where
$\mathcal{F}$ captures a coupled team-level requirement.
Two representative collaborative constraints are considered.
(I) The first is \emph{relative formation}, described by
$\mathcal{F}_{\texttt{form}}\big(\mathbf{z}(t)\big)\triangleq
\epsilon_{\texttt{f}}-\sum_{(i,j)\in\mathcal{E}}
f_{ij}(\mathbf{q}_i,\mathbf{q}_j)$,
where $\epsilon_{\texttt{f}}>0$ is the maximum admissible formation error.
The term $f_{ij}$ quantifies the formation mismatch between neighboring
robots $(i,j)\in\mathcal{E}$.
Typical choices include
$f_{ij}\triangleq \big|\|\mathbf{p}_i-\mathbf{p}_j\|-d_{ij}\big|$
for distance-based formations~\cite{antonelli2014decentralized} and
$f_{ij}\triangleq
\|(\mathbf{p}_i-\mathbf{p}_j)-\overline{\mathbf{p}}_{ij}\|$
for relative-pose formations~\cite{alonso2017multi}.
(II) The second is \emph{connectivity maintenance}, described by
$\mathcal{F}_{\texttt{con}}\big(\mathbf{z}(t)\big)\triangleq
\lambda_2\big(L(t)\big)$,
where $L(t)\in\mathbb{R}^{N\times N}$ is the Laplacian matrix of the
communication graph $\mathcal{G}_{\texttt{c}}(t)$.
The graph $\mathcal{G}_{\texttt{c}}(t)$ is connected if and only if
$\lambda_2\big(L(t)\big)>0$~\cite{sabattini2013distributed}.
Unlike pairwise spacing constraints, this condition is a system-level
property.

Using the STL syntax introduced above, a collaborative task is specified by
a bounded-time formula $\phi$ over predicates in
$\mu\triangleq \{\mu_A^k,\mu_B^k\}$.
Its trajectory-level penalty is evaluated through the robustness
$\rho^\phi(\mathbf{Z}(T),0)$ together with the geometric costs defined
above.
The resulting state-dependent cost is given by \eqref{ch06:eq:state-cost}:
\begin{equation}
\label{ch06:eq:state-cost}
\begin{aligned}
c_{\texttt{tsk}}\big(\mathbf{Z}(T)\big)\triangleq
&-\rho^\phi\big(\mathbf{Z}(T),0\big) \\
&+\sum_{t=0}^{T}
\Bigg(
\sum_{i\in\mathcal{N}} g_{\texttt{obs}}\big(\mathbf{p}_i(t)\big)
+\sum_{i<j\in\mathcal{N}}
g_{\texttt{int}}\big(\mathbf{p}_i(t),\mathbf{p}_j(t)\big)
\Bigg),
\end{aligned}
\end{equation}
where the first term promotes STL satisfaction, while the remaining terms
penalize obstacle proximity and inter-robot collisions.
Hence, minimizing \eqref{ch06:eq:state-cost} encourages both safety and
task fulfillment.

%==============================
\subsubsection{Collaborative motion planning problem}

Given initial and target states
$\mathbf{x}_{\texttt{s}},\mathbf{x}_{\texttt{g}}\in\mathbb{R}^{DN}$ and an
STL task $\phi$, the collaborative motion planning problem is formulated as
follows.

\begin{problem}
\label{ch06:prob:multi-robot-control}
Find a system trajectory $\mathbf{Z}(T)$ that minimizes the sum of the
task-dependent cost in \eqref{ch06:eq:state-cost} and the individual control
costs in \eqref{ch06:eq:control-cost}, subject to the boundary conditions and
the dynamics in \eqref{ch06:eq:robot-model}:
\begin{equation}
\label{ch06:eq:prob}
\begin{aligned}
\underset{\mathbf{Z}(T)}{\textbf{min}}\quad
& c_{\texttt{tsk}}\big(\mathbf{Z}(T)\big)
+ \sum_{i\in\mathcal{N}} c_{\texttt{ctr}}\big(\mathbf{Z}_i(T)\big), \\
\textbf{s.t.}\quad
& \mathbf{x}(0)=\mathbf{x}_{\texttt{s}}, \mathbf{x}(T)=\mathbf{x}_{\texttt{g}}, \\
& \mathbf{p}_i(t+1)=\mathbf{p}_i(t)+\mathbf{v}_i(t)\Delta t,
\quad \forall i\in\mathcal{N},\ \forall t\in\{0,\cdots,T-1\},
\end{aligned}
\end{equation}
where the optimization variable is the collective trajectory
$\mathbf{Z}(T)$ over the horizon $T$.
\end{problem}

Problem~\ref{ch06:prob:multi-robot-control} is a constrained and highly
non-convex trajectory-optimization problem.
In particular, STL satisfaction introduces time-indexed conjunctions and
disjunctions, which are commonly handled through mixed-integer encodings.
As a consequence, direct gradient-based nonlinear programming is generally
not applicable in a straightforward manner, while MILP and MICP
formulations typically suffer from exponential and combinatorial growth
with respect to the formula size, the planning horizon, and the number of
robots~\cite{fainekos2009robustness,lindemann2025formal,
liu2022stlq,kurtz2022mixed}.
































%==============================
%==============================
%==============================
%==============================
\subsection{Batch computation of STL perturbed robustness}
\label{ch06:subsec:batch-stl}

Efficient batch evaluation of STL robustness is required by the
waypoint-perturbation procedure used later in the Sinkhorn update.
The objective of this subsection is to quantify how the robustness
changes when a single waypoint is perturbed along multiple probing
directions.

%%===================
\subsubsection{STL tree representation}
\label{ch06:subsec:stl-tree}

An STL formula $\phi$ can be decomposed recursively into subformulas and
Boolean operators
\cite{fainekos2009robustness,kurtz2022mixed,leung2023backpropagation}.
This recursive structure is represented by the STL tree
\begin{equation}
\label{ch06:eq:stl-tree}
\mathcal{T}^{\phi}\triangleq
\big(\widehat{\mathcal{T}},\mathcal{I},\circ\big),
\end{equation}
where
$\widehat{\mathcal{T}}\triangleq
[\mathcal{T}^{\phi_1},\cdots,\mathcal{T}^{\phi_K}]$
is the list of $K\geq 1$ child subtrees,
$\mathcal{I}\triangleq [I^{\phi_1},\cdots,I^{\phi_K}]$
collects the associated time intervals, and
$\circ\in\{\wedge,\vee\}$ specifies the Boolean composition at the
current node.
Leaf nodes correspond to predicates evaluated at particular time
instants, whereas internal nodes aggregate the child robustness values.
In particular, the temporal operators $\mathsf{F}_I$ and
$\mathsf{G}_I$ are realized through disjunctive and conjunctive
compositions, respectively, while $\mathsf{U}_I$ induces a recursive
combination of both.
The representation in \eqref{ch06:eq:stl-tree} enables bottom-up
robustness evaluation over the formula tree.


%%===================
\subsubsection{Batched evaluation of perturbed robustness}
\label{ch06:subsec:stl-robustness}

As recalled in Section~\ref{ch06:subsubsec:stl_robustness}, the standard STL robustness is
computed by evaluating the predicates at the leaves of
$\mathcal{T}^{\phi}$ and then applying $\min$ and $\max$ recursively to
conjunction and disjunction nodes.
For the present perturbation-based optimization, a direct application of
this rule may become locally insensitive.
In particular, if the perturbed child of a conjunction node is not the
minimizer, the effect of the perturbation can be completely masked by
another child whose value remains unchanged.
A surrogate perturbed robustness is therefore introduced to retain
sensitivity to the local waypoint variation while preserving the
tree-based structure of the STL formula.
Let $\widehat{\mathbf{Z}}_{t^\star}(T)$ denote a trajectory obtained by
perturbing only the $t^\star$-th waypoint of $\mathbf{Z}(T)$.
For a fixed time index $t^\star$, let
$\widehat{\boldsymbol{\mathcal{Z}}}_{t^\star}\triangleq
[\widehat{\mathbf{Z}}_{t^\star}^{(1)}(T),\cdots,
\widehat{\mathbf{Z}}_{t^\star}^{(M)}(T)]$
be the batch of perturbed trajectories generated along the $M$ probing
directions used by the local transport step.
The predicate values at the leaf nodes are first computed in parallel
for all elements of
$\widehat{\boldsymbol{\mathcal{Z}}}_{t^\star}$.
The internal-node values are then propagated upward by masked reductions
over the intervals in $\mathcal{I}$.

To reduce the masking effect of hard conjunctions, the recursive
evaluation is modified as follows.
At the leaf level, the pointwise predicate robustness values
$f_i^k\big(\widehat{\mathbf{z}}(t)\big)$ are computed in the usual way.
At the first conjunction level above the leaves, the value
$f_i^k\big(\widehat{\mathbf{z}}(t^\star)\big)$ is passed upward whenever
$t^\star\in I^{\phi_i}$.
At higher conjunction levels, the child values are aggregated by their
mean.
Disjunction nodes continue to use the standard maximum operator.
Accordingly, the perturbed robustness of a single perturbed trajectory is
defined in \eqref{ch06:eq:perturbed-robustness}:
\begin{equation}
\label{ch06:eq:perturbed-robustness}
\widehat{\rho}^{\phi}\big(\widehat{\mathbf{Z}}_{t^\star}(T)\big)
\triangleq
\texttt{Eval}_{\mathcal{T}^{\phi}}
\Big(
\big\{f_i^k\big(\widehat{\mathbf{z}}(t)\big)\big\},
\{\max,\min,\operatorname{mean}\}
\Big),
\end{equation}
where $\widehat{\mathbf{z}}(t)$ is the state of the perturbed trajectory
at time $t$, and $f_i^k$ denotes the predicate-robustness function
associated with the corresponding leaf node.
The same recursion is then applied elementwise to the batch
$\widehat{\boldsymbol{\mathcal{Z}}}_{t^\star}$, so that the perturbed
robustness values for all probing directions are obtained in a single
bottom-up traversal.
This construction preserves the original tree-based interpretation of
the formula, improves local sensitivity to waypoint perturbations, and
supports efficient batched computation in the subsequent optimization
steps.
The surrogate robustness in \eqref{ch06:eq:perturbed-robustness} is used
to guide the local update, while the satisfaction of the final
trajectory is verified using the standard STL robustness in
Section~\ref{ch06:subsubsec:stl_robustness}.


























%==============================
%==============================
%==============================
%==============================
%==============================
\subsection{Sequential local optimal transport}
\label{ch06:sec:local}

Given the batched evaluation of perturbed STL robustness in
Section~\ref{ch06:subsec:batch-stl}, the multi-robot planning problem is
decomposed into a sequence of lower-dimensional subproblems.
At each stage, only one robot trajectory is optimized, while the
trajectories already computed for the preceding robots are treated as
time-dependent constraints in the local objective.
This decomposition greatly reduces the dimension of each optimization
step and makes local transport updates computationally tractable.
Its main limitation is that the final collective solution depends on the
chosen planning order, since earlier trajectory decisions restrict the
feasible and low-cost regions available to the later robots.

%==============================
\subsubsection{Design of planning sequence}
\label{ch06:subsec:sequence-design}

To encode the couplings that must be respected during sequential
planning, introduce an undirected dependency graph
$\overline{\mathcal{G}}\triangleq
(\mathcal{N},\overline{\mathcal{E}})$,
where $\overline{\mathcal{E}}$ represents the pairwise constraints that
link robot trajectories.
Global requirements, such as connectivity maintenance, are represented
through the pairwise dependencies induced by the initial communication
graph.
For each robot $i\in\mathcal{N}$, define the neighbor set
$\mathcal{N}_i^{\overline{\mathcal{E}}}\triangleq
\{j\in\mathcal{N}\mid (i,j)\in\overline{\mathcal{E}},\ i\neq j\}$.
The planning sequence is written as below:
\begin{equation}
\label{ch06:eq:order}
\kappa\triangleq [i_1,\cdots,i_N],
\end{equation}
where $i_\ell\in\mathcal{N}$ for each
$\ell\in\{1,\cdots,N\}$.
For every stage $\ell>1$, let
$\mathcal{N}_\ell^-\triangleq \{i_1,\cdots,i_{\ell-1}\}$ denote the set
of robots planned before $i_\ell$.
The sequence $\kappa$ is called valid if every robot $i_\ell$ with
$\ell>1$ is constrained by at least one preceding robot, namely,
\[
i_\ell\in
\bigcup_{j\in\mathcal{N}_\ell^-}\mathcal{N}_j^{\overline{\mathcal{E}}},
\]
which ensures that, when robot $i_\ell$ is optimized, the local
problem already reflects the relevant interactions with the previously
planned trajectories.
Moreover, for time-varying pairwise constraints, let
$\overline{\mathcal{G}}_s\triangleq
(\mathcal{N},\overline{\mathcal{E}}_s)$,
$s\in\{1,\cdots,S\}$, denote the dependency graphs associated with the
different time indices or operating modes.
The validity condition is then strengthened to
\[
i_\ell\in
\bigcap_{s=1}^S
\Big(
\bigcup_{j\in\mathcal{N}_\ell^-}
\mathcal{N}_j^{\overline{\mathcal{E}}_s}
\Big),
\]
which guarantees that robot $i_\ell$ remains coupled to the preceding
set across all graphs $\overline{\mathcal{G}}_s$.
Hence, valid sequences support feasible sequential decompositions,
whereas arbitrary orderings can easily produce incompatible local
subproblems.

%==============================
\subsubsection{Constrained trajectory optimization for each robot}
\label{ch06:subsec:sequential-planning}

Fix a valid planning sequence
$\kappa=[i_1,\cdots,i_N]$.
The sequential procedure solves one local trajectory optimization problem
for each robot in the order prescribed by \eqref{ch06:eq:order}.
For the first robot $i_1$, no previously planned trajectory is
available.
Its local task cost therefore consists only of the obstacle-avoidance
term together with the STL robustness term, and the corresponding
solution is denoted by $\mathbf{Z}_{i_1}^\star(T)$.
For each later stage $\ell\in\{2,\cdots,N\}$, define the previously
planned trajectory set as:
\[
\Gamma_\ell^-\triangleq
\{\mathbf{Z}_{i_1}^\star(T),\cdots,
\mathbf{Z}_{i_{\ell-1}}^\star(T)\},
\]
and let
$\Gamma_\ell\triangleq
\Gamma_\ell^-\cup\{\mathbf{Z}_{i_\ell}(T)\}$
be the collective trajectory obtained after adding the candidate
trajectory of robot $i_\ell$.
The local state-dependent cost for robot $i_\ell$ is then defined by:
\begin{equation}
\label{ch06:eq:state-cost-sequential}
\begin{aligned}
c_{\texttt{tsk}}^{i_\ell}\big(\mathbf{Z}_{i_\ell}(T)\big)
\triangleq
&\sum_{t=0}^{T}
\Bigg(
\sum_{j\in\mathcal{N}_\ell^-}
g_{\texttt{int}}
\big(\mathbf{p}_{i_\ell}(t),\mathbf{p}_{j}(t)\big)
+ g_{\texttt{obs}}\big(\mathbf{p}_{i_\ell}(t)\big)
\Bigg)-\widehat{\rho}^{\phi}(\Gamma_\ell),
\end{aligned}
\end{equation}
where $\widehat{\rho}^{\phi}(\Gamma_\ell)$ denotes the perturbed
robustness surrogate in \eqref{ch06:eq:perturbed-robustness}, evaluated
on the collective trajectory represented by $\Gamma_\ell$.
The first term in \eqref{ch06:eq:state-cost-sequential} penalizes
collisions with the preceding robots and with obstacles, while the last
term promotes collective task satisfaction.
Combining \eqref{ch06:eq:state-cost-sequential} with the control cost in
\eqref{ch06:eq:control-cost}, the local optimization problem for robot
$i_\ell$ is written as
\begin{equation}
\label{ch06:eq:local-problem}
\mathbf{Z}_{i_\ell}^\star(T)\triangleq
\underset{\mathbf{Z}_{i_\ell}(T)}{\textbf{argmin}}
\Big\{
c_{\texttt{tsk}}^{i_\ell}\big(\mathbf{Z}_{i_\ell}(T)\big)
+
c_{\texttt{ctr}}\big(\mathbf{Z}_{i_\ell}(T)\big)
\Big\},
\end{equation}
where the optimization is subject to the robot dynamics in
\eqref{ch06:eq:robot-model} and to the boundary conditions inherited from the
global problem.
After solving \eqref{ch06:eq:local-problem}, the resulting trajectory
$\mathbf{Z}_{i_\ell}^\star(T)$ is appended to the collective set and is
used in the optimization of the subsequent robots.

%==============================
\begin{algorithm}[t!]
  \caption{Local Sinkhorn update for a single robot.}
  \label{ch06:alg:sinkhorn}
  \SetAlgoLined
  \KwIn{State cost $c_{\texttt{tsk}}(\cdot)$ and control cost
  $c_{\texttt{ctr}}(\cdot)$.}
  \KwOut{Optimized trajectory $\mathbf{Z}_{i_\ell}^\star(T)$.}
  Initialize $\alpha_0$, $\lambda$, and $\mathbf{Z}_{i_\ell}^{0}(T)$\;
  Set histograms
  $\boldsymbol{n}=\boldsymbol{1}_{T}/T$ and
  $\boldsymbol{m}=\boldsymbol{1}_{M}/M$\;
  \While{($k\leq K$) and
    ($\|\mathbf{Z}_{i_\ell}^{k+1}(T)-\mathbf{Z}_{i_\ell}^{k}(T)\|
    > \epsilon$)}{
    Build $\boldsymbol{D}_{\mathcal{P}}$ with rotation\;
    Compute $\boldsymbol{C}$ from
    $(\mathbf{Z}_{i_\ell}^{k}(T),\boldsymbol{D}_{\mathcal{P}})$\;
    Compute $\boldsymbol{W}_\lambda^\star$ from
    \eqref{ch06:eq:OT}\;
    Update $\mathbf{Z}_{i_\ell}^{k+1}(T)$ by
    \eqref{ch06:eq:sinkhorn}\;
    Update $\alpha_k$ and set $k\leftarrow k+1$\;
  }
  \Return $\mathbf{Z}_{i_\ell}^{k}(T)$\;
\end{algorithm}
%==============================


%%===================
\subsubsection{Local optimal transport via Sinkhorn step}
\label{ch06:subsec:sinkhorn}

The local problem in \eqref{ch06:eq:local-problem} remains nonlinear and
non-convex.
To compute approximate solutions efficiently, each single-robot update is
recast as an entropy-regularized optimal transport problem and solved by
a Sinkhorn-type iteration
\cite{cuturi2013sinkhorn,peyre2019computational,le2023accelerating}.
The main idea is to probe a set of local perturbation directions at
every waypoint, evaluate their costs in parallel, and then transport the
trajectory toward lower-cost regions by a weighted combination of these
directions.
The procedure summarized in Algorithm~\ref{ch06:alg:sinkhorn} consists of four
stages.
(I) \emph{Initialization}.
At iteration $k$, the current trajectory of robot $i_\ell$ is denoted by
$\mathbf{Z}_{i_\ell}^{k}(T)$ and is initialized from the
Gaussian-process prior.
(II) \emph{Probing}.
Each waypoint is perturbed along $M$ directions collected in
$\boldsymbol{D}_{\mathcal{P}}\in\mathbb{R}^{M\times 2D}$, where the
directions are generated from a polytope
$\mathcal{P}\subset\mathbb{R}^{2D}$.
Evaluating the local objective along these perturbations produces the
cost matrix $\boldsymbol{C}\in\mathbb{R}^{T\times M}$.
(III) \emph{Weight optimization}.
The transport weights are obtained by solving the regularized \emph{optimal
transport problem} below:
\begin{equation}
\label{ch06:eq:OT}
OT_{\lambda}(\boldsymbol{n},\boldsymbol{m})\triangleq
\underset{\boldsymbol{W}\in U(\boldsymbol{n},\boldsymbol{m})}
{\textbf{min}}
\Big\{
\langle \boldsymbol{W},\boldsymbol{C}\rangle
-\lambda H(\boldsymbol{W})
\Big\},
\end{equation}
where
$\boldsymbol{n}\in\Sigma_T$ and $\boldsymbol{m}\in\Sigma_M$ are the
histograms over time indices and probing directions, respectively,
$\Sigma_q\triangleq
\{\mathbf{a}\in\mathbb{R}_{\geq 0}^{q}\mid
\sum_{j=1}^{q} a_j=1\}$ is the probability simplex,
$U(\boldsymbol{n},\boldsymbol{m})$ is the set of transport plans with
marginals $\boldsymbol{n}$ and $\boldsymbol{m}$,
$H(\boldsymbol{W})$ is the entropy of the plan, and $\lambda>0$
controls the regularization strength.
(IV) \emph{Trajectory update}.
Using the optimizer $\boldsymbol{W}_\lambda^\star$ of
\eqref{ch06:eq:OT}, the trajectory is updated by:
\begin{equation}
\label{ch06:eq:sinkhorn}
\mathbf{Z}_{i_\ell}^{k+1}(T)
=
\mathbf{Z}_{i_\ell}^{k}(T)
+
\alpha_k\,
\big(\texttt{diag}(\boldsymbol{n})^{-1}\big)
\boldsymbol{W}_\lambda^\star
\boldsymbol{D}_{\mathcal{P}},
\end{equation}
where $\alpha_k>0$ is the step size.
The iteration continues until convergence or until the maximum number of
iterations is reached.
After the Sinkhorn update has converged, the optimized trajectory
$\mathbf{Z}_{i_\ell}^\star(T)$ is added to the current collective set,
and the next robot in the sequence is planned using the updated
preceding set.
In this way, the sequential procedure yields the collective trajectory
\[
\Gamma_\kappa\triangleq
\{\mathbf{Z}_{i_1}^\star(T),\cdots,\mathbf{Z}_{i_N}^\star(T)\}.
\]
The overall solution quality depends on the planning sequence
$\kappa$, because early decisions influence the feasible and
cost-effective regions available to later robots.
This interaction between sequence selection and sequential transport motivates the hybrid search
scheme introduced in the next subsection.

Several implementation heuristics can further improve the numerical
behavior of Algorithm~\ref{ch06:alg:sinkhorn}.
(I) \emph{Randomized probing directions}.
To reduce correlation among the waypoint updates, the probing directions
in $\boldsymbol{D}_{\mathcal{P}}$ can be randomly rotated at each
iteration and at each trajectory point.
(II) \emph{Directional cost averaging}.
Instead of evaluating only one perturbed point along each probing
direction, let $N_{\texttt{s}}>0$ denote the number of auxiliary samples
taken along the same direction.
The corresponding costs are then averaged to obtain a more stable
estimate of the directional cost.
This leads to an extended probing array
$\widetilde{\boldsymbol{D}}_{\mathcal{P}}
\in \mathbb{R}^{T\times M\times N_{\texttt{s}}\times 2D}$.
Such averaging reduces variance in the cost evaluation and, in practice,
often improves convergence.
(III) \emph{Annealing of step size and exploration radius}.
The step size $\alpha_k$ and the exploration radius $\beta_k$ can be
decreased gradually, for instance through exponential decay schedules,
so that early iterations emphasize exploration, whereas later
iterations focus on local refinement and stabilization
\cite{altschuler2017near,le2023accelerating}.































%==============================
%==============================
%==============================
%==============================
\subsection{Hybrid search for sequence and collective trajectory}
\label{ch06:sec:hybrid}

The sequential local transport scheme in Section~\ref{ch06:sec:local}
reduces the dimension of the original multi-robot planning problem, but
its performance depends strongly on the chosen planning sequence.
Early local decisions restrict the feasible and low-cost regions
available to the remaining robots, so the final collective solution is
determined jointly by sequence selection and trajectory refinement.
To address this coupling, a hybrid search layer is introduced that
combines discrete exploration over partial planning sequences with
continuous local optimization through the Sinkhorn-based updates. The two components are tightly coupled, since each partial sequence restricts the feasible local transports and each local transport evaluation influences subsequent sequence expansion.

%==============================
\subsubsection{Hybrid search framework}
\label{ch06:subsec:search-algorithm}

The search is organized as a rooted tree
$\mathfrak{T}\triangleq (\mathcal{V},\rightarrow)$.
Each node is denoted by
$\nu\triangleq (\mathcal{N}_\nu,\Gamma_\nu)$, where
$\mathcal{N}_\nu\subset\mathcal{N}$ is the set of robots that have
already been planned, and
$\Gamma_\nu\triangleq
\{\mathbf{Z}_{i_1},\cdots,\mathbf{Z}_{i_\ell}\}$ is the associated
partial collective trajectory.
The root node is
$\nu_0\triangleq (\emptyset,\emptyset)$.
A child node is created by appending one additional robot to the current
partial plan and solving the corresponding local trajectory optimization
problem.
As summarized in Algorithm~\ref{ch06:alg:ss-MPOT}, the search proceeds
through the following four stages.

(I) \emph{Selection}.
At each search round, a subset of $H$ nodes is selected for parallel
expansion, denoted by
$\mathcal{V}_H\triangleq \{\nu_1^\star,\cdots,\nu_H^\star\}$.
These nodes are chosen according to the value function
\begin{equation}
\label{ch06:eq:value}
\chi(\nu)\triangleq
\frac{|\mathcal{N}_\nu|}{N}
-\eta\,\xi(\Gamma_\nu)
+\psi(\nu),
\end{equation}
where the first term measures planning progress, the second term
penalizes the estimated accumulated trajectory cost, and the third term
carries feedback from previous expansions.
Here $\eta>0$ is a weighting parameter,
$\xi(\Gamma_\nu)$ is a surrogate of the partial trajectory cost based on
the same local cost model used in the sequential stage, and
$\psi(\nu)$ is initialized at zero and updated during the search.
Nodes are therefore favored when they correspond to deeper progress,
lower estimated cost, and stronger positive evidence from their
descendants.

(II) \emph{Expansion}.
For each selected node $\nu_h^\star$, the robots that are eligible to be
planned next are collected in
$\mathcal{N}_{\nu_h^\star}^{\texttt{ext}}\triangleq
\Big(
\bigcup_{i\in\mathcal{N}_{\nu_h^\star}}
\mathcal{N}_i^{\overline{\mathcal{E}}}
\Big)
\setminus \mathcal{N}_{\nu_h^\star}$,
which restricts the search to robots that remain compatible with the
current partial plan.
At the root node, the initialization
$\mathcal{N}_{\nu_0}^{\texttt{ext}}\triangleq \mathcal{N}$ is used so
that the search can start from any robot.
Once a robot
$i^\star\in\mathcal{N}_{\nu_h^\star}^{\texttt{ext}}$ is chosen, the
local Sinkhorn step in Algorithm~\ref{ch06:alg:sinkhorn} generates a set of
candidate trajectories denoted by:
\[
\Pi_{i^\star}\triangleq
\{\boldsymbol{\tau}_{i^\star}^1,\cdots,
\boldsymbol{\tau}_{i^\star}^K\},
\]
where the notation $\Pi_{i^\star}$ is used to avoid confusion with a
single trajectory variable.
Each candidate defines a child node.
This expansion is written below as:
\begin{equation}
\label{ch06:eq:expand}
\nu_k^+\triangleq
\big(
\mathcal{N}_{\nu_h^\star}\cup\{i^\star\},
\Gamma_{\nu_h^\star}\cup\{\boldsymbol{\tau}_{i^\star}^k\}
\big),
\end{equation}
where an edge
$(\nu_h^\star,\nu_k^+)\in\rightarrow$
is added for each admissible candidate.
If no feasible candidate trajectory is produced, then the corresponding
branch is not expanded further.

(III) \emph{Back-propagation}.
After expansion, the feedback term in \eqref{ch06:eq:value} is updated
along the ancestor chain of each new node.
For a newly created node $\nu^+$ and any ancestor
$\nu\in\mathcal{V}_{\nu^+}^-$, the update rule is given by:
\begin{equation}
\label{ch06:eq:propagation}
\psi(\nu)\leftarrow
\psi(\nu)+\psi(\nu^+)\gamma^{d(\nu,\nu^+)},
\end{equation}
where $\mathcal{V}_{\nu^+}^-\subset\mathcal{V}$ is the set of ancestors
of $\nu^+$, $d(\nu,\nu^+)$ is the path length between $\nu$ and $\nu^+$
in the search tree, and $\gamma\in(0,1)$ is a decay factor.
The value $\psi(\nu^+)$ of each newly accepted node is initialized by a
positive score, so feasible expansions reinforce the priorities of the
ancestor nodes from which they originate.
This mechanism biases later selections toward branches that repeatedly
lead to successful continuations.

(IV) \emph{Termination}.
The cycle of selection, expansion, and back-propagation is repeated
until the prescribed search budget is exhausted.
Complete solutions correspond to leaf nodes satisfying
$\overline{\mathcal{V}}^\star\triangleq
\{\nu\in\mathcal{V}\mid \mathcal{N}_\nu=\mathcal{N}\}$,
and the final solution is selected by:
\[
\overline{\nu}^\star\triangleq
\operatorname*{argmax}_{\nu\in\overline{\mathcal{V}}^\star}\chi(\nu),
\]
with associated collective trajectory
$\Gamma_{\overline{\nu}^\star}$.
In this way, the search layer evaluates multiple candidate planning
orders while preserving the continuous local refinement of the
trajectories at each expansion step.

%==============================
\subsubsection{Parallelization and practical improvements}

The hybrid search is naturally amenable to parallel computation.
During expansion, the local trajectory optimization problems associated
with all nodes in $\mathcal{V}_H$ can be solved simultaneously, and the
resulting child nodes can also be generated in parallel.
Likewise, the updates in \eqref{ch06:eq:propagation} can be aggregated
collectively after each expansion round rather than processed one node at
a time.
These features substantially reduce runtime and are especially useful
when the probing-based local updates are executed on modern parallel
hardware.
Several practical enhancements further improve scalability.
First, the number of expanded nodes and the number of retained
trajectory candidates can be adjusted adaptively during the search.
A broader expansion is beneficial in the early stages to explore diverse
planning orders, whereas later stages benefit more from a narrower but
deeper refinement of the most promising branches.
Second, when the expansion of a node $\nu$ repeatedly fails, a
similarity-based pruning rule can be used to remove nearby redundant
nodes,
$\widetilde{\mathcal{V}}_\nu\triangleq
\Big\{
\nu'\in\mathcal{V}\ \Big|\
\|\Gamma_\nu-\Gamma_{\nu'}\|<\delta,\
\mathcal{N}_\nu=\mathcal{N}_{\nu'}
\Big\}$,
where $\|\Gamma_\nu-\Gamma_{\nu'}\|$ denotes a suitable distance
between the two partial collective trajectories and $\delta>0$ is a
pruning threshold.
This rule helps avoid repeated exploration of infeasible or nearly
duplicate branches.
Together, these implementation choices improve scalability, reduce
cascading failures, and accelerate convergence.














































%==========
\begin{algorithm}[t!]
  \caption{Hybrid search with collaborative optimal transport under STL tasks}
  \label{ch06:alg:ss-MPOT}
  \SetAlgoLined
  \KwIn{
    Initial state $\mathbf{x}_{\texttt{s}}$,
    goal state $\mathbf{x}_{\texttt{g}}$,
    cost functions $c_{\texttt{ctr}}(\cdot)$ and
    $c_{\texttt{tsk}}(\cdot)$.
  }
  \KwOut{Feasible collective trajectory $\Gamma_{\overline{\nu}^\star}$.}

  Initialize $\mathfrak{T}=(\{\nu_0\},\emptyset)$\;

  \While{termination condition is not met \textup{(in parallel)}}{
    \tcc{Selection}
    Select $\mathcal{V}_H$ using \eqref{ch06:eq:value}\;

    \tcc{Expansion}
    \ForEach{$\nu_h^\star\in\mathcal{V}_H$}{
      \eIf{$\mathcal{N}_{\nu_h^\star}=\emptyset$}{
        $\mathcal{N}_{\nu_h^\star}^{\texttt{ext}}\gets \mathcal{N}$\;
      }{
        $\mathcal{N}_{\nu_h^\star}^{\texttt{ext}}\gets
        \big(
        \bigcup_{i\in\mathcal{N}_{\nu_h^\star}}
        \mathcal{N}_i^{\overline{\mathcal{E}}}
        \big)
        \setminus
        \mathcal{N}_{\nu_h^\star}$\;
      }

      \If{$\mathcal{N}_{\nu_h^\star}^{\texttt{ext}}\neq\emptyset$}{
        Choose $i^\star\in
        \mathcal{N}_{\nu_h^\star}^{\texttt{ext}}$\;
        $\Pi_{i^\star}\gets \texttt{SinkhornStep}(\cdot)$ by
        Algorithm~\ref{ch06:alg:sinkhorn}\;
        \ForEach{$\boldsymbol{\tau}_{i^\star}^k\in \Pi_{i^\star}$}{
          Generate $\nu_k^+$ using \eqref{ch06:eq:expand}\;
          Update $\mathcal{V}$ and $\rightarrow$\;
        }
      }
    }

    \tcc{Back-propagation}
    Update $\psi(\nu)$ using \eqref{ch06:eq:propagation}\;
    $\overline{\mathcal{V}}^\star\gets
    \{\nu\in\mathcal{V}\mid \mathcal{N}_\nu=\mathcal{N}\}$\;
  }

  \tcc{Termination}
  $\overline{\nu}^\star\gets
  \operatorname*{argmax}_{\nu\in\overline{\mathcal{V}}^\star}\chi(\nu)$\;
  \Return $\Gamma_{\overline{\nu}^\star}$\;
\end{algorithm}
%==========


%==============================
%==============================
%==============================
%==============================
\subsection{Overall analyses}
\label{ch06:subsec:overall}

The construction above yields a complete pipeline from surrogate robustness computation to hybrid search over planning orders. The two parts below summarize how candidate trajectories are verified for execution and how the dominant computational costs scale.

%==============================
\subsubsection{Verification and execution}
\label{ch06:subsubsec:execute}

After each local optimization step, a candidate trajectory is retained
only after explicit verification of the hard constraints.
This verification stage checks collision avoidance, inter-robot
constraints, and positive robustness under the standard STL semantics,
rather than under the surrogate perturbed robustness used during the
local search.
Consequently, only trajectories that satisfy the required feasibility
and safety conditions are inserted into the search tree.
This separation is important: the surrogate robustness is used to guide
the optimization efficiently, whereas the standard robustness is used to
certify correctness of the final accepted trajectory.
For execution, the optimized trajectories
$\{\mathbf{Z}_i^\star(T)\}_{i\in\mathcal{N}}$ can be refined further
into denser waypoint sequences with a smaller sampling step.
This post-processing improves tracking quality, reduces abrupt changes
between consecutive waypoints, and yields trajectories that are more
suitable for closed-loop execution.
If needed, the refined trajectories can be rechecked under the same hard
constraints before deployment.
They can then be tracked either by onboard low-level controllers or by a
model predictive control scheme, depending on the sensing, actuation,
and computation capabilities of the robotic platform.

%==============================
\subsubsection{Complexity analysis}
\label{ch06:subsubsec:complexity}

The dominant computational cost of
Algorithm~\ref{ch06:alg:ss-MPOT} arises from the repeated Sinkhorn updates
performed during node expansion.
Using the complexity estimate for the entropy-regularized optimal
transport step, the cost of one such update is
$\widetilde{O}\!\left(
\tfrac{2D}{\epsilon^2}\cdot
\tfrac{(HN_pT)^2}{\varepsilon^3}
\right)$,
where $D$ is the spatial dimension, $\epsilon>0$ is the convergence
precision of the local transport update, $H$ is the number of parallel
expansions, $N_p$ is the number of trajectory samples, $T$ is the
trajectory length, and $\varepsilon>0$ is the tolerance used in the
Sinkhorn--Knopp scaling
\cite{cuturi2013sinkhorn,peyre2019computational,le2023accelerating}.
In addition, evaluating the task-related constraints through the STL
tree contributes a cost of
$O\big(M|\mathcal{T}^\phi|\big)$ per robot, where $M$ is the number of
probing directions and $|\mathcal{T}^\phi|$ is the number of nodes in
the STL tree.
The remaining bookkeeping operations of the search tree are of lower
order compared with the repeated transport updates and robustness
evaluations.
With $N$ robots, the search tree depth is at most $N$.
In the worst case, the number of explored nodes can still grow rapidly.
In practice, however, pruning and heuristic guidance substantially limit
the effective search.
Empirically, the search often remains close to about $CN$ expansions,
with $C\in[1,3]$ depending on the problem structure and on the degree of
branch pruning.
Thus, although the worst-case complexity remains high, the combination
of parallel expansion, local transport updates, and pruning keeps the
method computationally practical for the scenarios considered here.

%==============================
\subsubsection{Generalization}
\label{ch06:subsubsec:generalization}

The framework extends naturally to several more complex scenarios.
(I) \emph{Split-and-merge coordination for large fleets}.
A large fleet can be partitioned into subgroups
$\{\mathcal{N}_1,\cdots,\mathcal{N}_G\}$, each equipped with its own
local formation constraints.
A higher-level STL formula can then be imposed over these groups rather
than over individual robots.
This allows the subgroups to separate, navigate independently, and later
merge again while preserving the required collective task logic.
The same hybrid search principle can then be applied at the group level,
while the local transport step is retained within each subgroup.
(II) \emph{Dynamic obstacles}.
When obstacle trajectories are known in advance or can be predicted with
sufficient accuracy, they can be incorporated directly into the
state-dependent cost in \eqref{ch06:eq:state-cost-sequential}.
In this case, the same planning framework continues to enforce collision
avoidance and STL robustness in a time-varying environment.
When obstacle motion is uncertain, conservative reachable sets or safety
buffers can be used instead so that the resulting plan remains robust to
prediction errors.
This extension preserves the overall structure of the method, while
modifying only the underlying cost evaluation.
