%==============================
%==============================
\section{Reactive Coordination under Dynamic Temporal Tasks}
\label{ch06:sec:ch06_uncertain}

The previous section addressed uncertainty at the motion-planning level through
probabilistic safety and return guarantees. The emphasis in this section is on a
different source of difficulty. Here, the task-relevant targets are dynamic, their
future motion is only partially predictable, and additional tasks may be released
online in response to newly observed events. Consequently, a task assignment that
is feasible and efficient at one time instant may soon become inadequate as the
target configuration evolves. The coordination problem must therefore be updated
online while preserving the temporal and collaborative structure of the mission.
To address this difficulty, the section develops an online coordination method
that combines target-trajectory estimation, uncertainty quantification, compact
task decomposition, and uncertainty-aware task assignment. In particular,
target-motion predictions and their uncertainty sets are constructed in
Section~\ref{ch06:subsec:trajectory}, the temporal structure of the task is encoded
through R-posets in Section~\ref{ch06:subsec:poset}, and these components are used
in Section~\ref{ch06:subsec:assignment} to guide the assignment process.
Section~\ref{ch06:subsec:execution} then embeds this assignment scheme into an
online replanning mechanism so that the team can react to updated target
information and newly released tasks during execution.



%==============================
%==============================
%==============================
%==============================
\subsection{Task coordination over dynamic temporal tasks}
\label{ch06:sec:problem}

This subsection introduces the dynamic task-coordination model used throughout the section. It specifies the robots, targets, temporal-task language, and problem statement before the predictive and assignment layers are developed.

%==============================
\subsubsection{Robots and targets}
\label{ch06:subsec:robots}

Consider a team of $N$ autonomous robots operating in a workspace
$\mathcal{W} \subset \mathbb{R}^{3}$. Let
$\mathcal{N} \triangleq \{1,\cdots,N\}$ denote the robot index set. The
state of robot $n \in \mathcal{N}$ at time $t \geq 0$ is denoted by
$x_{t,n} \in \mathcal{W}$, and the stacked team state is
$x_t \triangleq (x_{t,1},\cdots,x_{t,N})$. The maximum speed of robot
$n$ is denoted by $v_n$. Each robot $n$ can execute actions from the set
$\mathcal{A}_n$. Let
$\mathcal{A} \triangleq \bigcup_{n \in \mathcal{N}} \mathcal{A}_n$.
Let $\mathcal{C}$ denote the set of pre-programmed collaborations. Each
collaboration $C_k \in \mathcal{C}$ consists of a list of actions,
written as $C_k \triangleq [a_1,\cdots,a_{\ell_k}]$, where $\ell_k > 0$
is the number of required actions and $a_{\ell} \in \mathcal{A}$ for
every $\ell \in \{1,\cdots,\ell_k\}$. Each action $a_{\ell} \in C_k$
must be assigned to a capable robot $n$ satisfying
$a_{\ell} \in \mathcal{A}_n$. Each collaboration has a fixed duration
given by the map $\rho : \mathcal{C} \rightarrow \mathbb{R}_{+}$.

The workspace also contains $M$ dynamic targets with unknown
trajectories. Let $\mathcal{M} \triangleq \{1,\cdots,M\}$ denote the
target index set. The position of target $m \in \mathcal{M}$ at time
$t \geq 0$ is modeled as a random variable
$Y_{t,m} \in \mathbb{R}^{2}$, and the joint target state is
$Y_t \triangleq (Y_{t,1},\cdots,Y_{t,M}) \in \mathbb{R}^{2M}$. The
target trajectories are assumed to follow an unknown distribution
$\mathcal{D}$, namely $(Y_0,Y_1,\cdots) \sim \mathcal{D}$. This
distribution is assumed to be independent of the robot-team state $x_t$.
A total of $\bar{K}$ independent trajectory samples
$Y^{(i)} \triangleq (Y_0^{(i)},Y_1^{(i)},\cdots)$ are available. These
samples are partitioned into the calibration set
$D_{\texttt{cal}} \triangleq \{Y^{(1)},\cdots,Y^{(K)}\}$ and the
training set
$D_{\texttt{tra}} \triangleq \{Y^{(K+1)},\cdots,Y^{(\bar{K})}\}$. The
datasets are independently drawn from $\mathcal{D}$
\cite{Lars2023CP,Matthew2024CP}. Perfect observations of target
positions and velocities are assumed to be available throughout
execution. They are denoted by
$Y_{0:t} \triangleq (Y_0,\cdots,Y_t)$ and
$V_{0:t} \triangleq (V_0,\cdots,V_t)$, where
$V_t \triangleq (V_{t,1},\cdots,V_{t,M})$. A centralized estimation
scheme aggregates these measurements and produces real-time estimates of
the target states. For convenience, static task regions and obstacles
are modeled as targets with zero velocity.

%==============================
\subsubsection{Task specification}
\label{ch06:subsec:task-specification}

The task formulas are defined over two classes of atomic propositions.
For each robot $n \in \mathcal{N}$ and target $m \in \mathcal{M}$, the
proposition $p_n^m$ is \textit{true} when the distance between robot $n$
and target $m$ is below a prescribed threshold. Let
$\textbf{p} \triangleq
\{p_n^m \mid m \in \mathcal{M},\, n \in \mathcal{N}\}$. For each
collaboration $C_k \in \mathcal{C}$ and target $m \in \mathcal{M}$, the
proposition $c_k^m$ is \textit{true} when collaboration $C_k$ is
executed on target $m$. Let
$\textbf{c} \triangleq
\{c_k^m \mid m \in \mathcal{M},\, C_k \in \mathcal{C}\}$. The complete
set of atomic propositions is
$AP \triangleq \textbf{p} \cup \textbf{c}$.
Each team-level task is represented as a syntactically co-safe LTL
sc-LTL formula over $AP$. The overall task specification is defined by
\eqref{ch06:eq:reactive-task}:
\begin{equation}
\label{ch06:eq:reactive-task}
\varphi \triangleq \varphi_{\texttt{static}}
\wedge
\bigwedge_{\bar{e} \in \bar{E}}
\Box\bigl(
\varphi_{\texttt{obs}}^{\bar{e}}
\rightarrow
\Diamond \varphi_{\texttt{rep}}^{\bar{e}}
\bigr),
\end{equation}
where $\varphi_{\texttt{static}}$ denotes the offline-specified
sc-LTL task component, $\bar{E}$ is the set of events triggered by
online observations, $\varphi_{\texttt{obs}}^{\bar{e}}$ is the
propositional condition associated with event $\bar{e} \in \bar{E}$,
and $\varphi_{\texttt{rep}}^{\bar{e}}$ is the corresponding response
task, also expressed as an sc-LTL formula. According to
\eqref{ch06:eq:reactive-task}, whenever
$\varphi_{\texttt{obs}}^{\bar{e}}$ holds, the response task
$\varphi_{\texttt{rep}}^{\bar{e}}$ must eventually be satisfied.
The standard LTL syntax and semantics introduced in
Chapter~\ref{chap:preliminaries} are adopted here; see also
\cite{baier2008principles}. The fragment used in this section is sc-LTL,
which admits satisfaction by finite traces and is written in positive
normal form using the temporal operators $\bigcirc$, $\textsf{U}$, and
$\Diamond$ \cite{belta2017formal}.

Thus, given the robot model and the task specification,
the task plan for the robot team is defined by:
\begin{equation}
\label{ch06:plan}
\boldsymbol{\tau}_{\mathcal{N}} \triangleq (\pi_1,\pi_2,\cdots,\tau_n),
\end{equation}
where
$\tau_n \triangleq
(t_n^1,a_n^1,m_n^1)(t_n^2,a_n^2,m_n^2)\cdots
(t_n^{K_n},a_n^{K_n},m_n^{K_n})$
is the timed sequence of actions and corresponding targets for robot
$n \in \mathcal{N}$. Here, each action $a_n^k \in \mathcal{A}_n$ is
executed on target $m_n^k \in \mathcal{M}$ at time $t_n^k > 0$, for
$k \in \mathcal{K}_n \triangleq \{1,\cdots,K_n\}$. The plan
\eqref{ch06:plan} induces the trace
$w_{\tau} \triangleq \sigma_1 \sigma_2 \cdots \sigma_L$, where
$\sigma_{\ell} \in 2^{AP}$ for every $\ell \in \{1,\cdots,L\}$. The
language of $\varphi$ is
$\mathcal{L}_{\varphi} \triangleq \{w \mid w \models \varphi\}$, where
$\models$ is the satisfaction relation. Since sc-LTL formulas admit
satisfaction by finite traces
\cite{belta2017formal,baier2008principles}, the plan satisfies
$\varphi$ if $w_{\tau} \in \mathcal{L}_{\varphi}$, which is denoted by
$\boldsymbol{\tau}_{\mathcal{N}} \models \varphi$.
The makespan $T_{\varphi}$ is the minimum time required to generate a
trace satisfying $\varphi$. Let
$\varphi(t) \triangleq \{\varphi_1,\cdots,\varphi_L\}$ denote the set of
tasks released up to time $t$. The average makespan is then defined by:
\begin{equation}
\label{ch06:eq:avg-makespan}
\overline{T}_{\varphi} \triangleq
\frac{1}{L}\sum_{\ell=1}^{L} T_{\varphi_{\ell}},
\end{equation}
which is the performance metric used in the sequel.

%==============================
\subsubsection{Problem statement}
\label{ch06:subsec:prob-statement}

Given the task formulas $\varphi(t)$ and the observations
$\{Y_{0:t},V_{0:t}\}$, the objective is to synthesize the team plan
$\boldsymbol{\tau}_{\mathcal{N}}$ such that $\boldsymbol{\tau}_{\mathcal{N}} \models \varphi(t)$
and the average makespan in \eqref{ch06:eq:avg-makespan} is minimized.

%==============================
\begin{example}
\label{ch06:exp:task}

Consider a fleet of UAVs and ground robots deployed for wildlife
monitoring in a workspace labeled by \texttt{poacher}, \texttt{trap},
and \texttt{animal}. An offline collaborative task is specified by
\eqref{ch06:eq:task}:
\begin{equation}
\label{ch06:eq:task}
\begin{aligned}
\varphi_{\texttt{1}}
&\triangleq
\varphi_{\texttt{p}-\texttt{s}_1}
\wedge
\varphi_{\texttt{mf}-\texttt{a}}, \\
\varphi_{\texttt{p}-\texttt{s}_1}
&\triangleq
\Diamond \texttt{patrol}_{\texttt{s}_1}, \\
\varphi_{\texttt{mf}-\texttt{a}}
&\triangleq
\Diamond\bigl(
\texttt{monitor}_{\texttt{a}}
\wedge
\neg \texttt{film}_{\texttt{a}}
\wedge
\Diamond \texttt{film}_{\texttt{a}}
\bigr),
\end{aligned}
\end{equation}
which requires patrol of region $\texttt{s}_1$ and sequential
monitoring and filming of the antelope. Ground robots must also avoid
obstacles in the workspace. In addition, the reactive rule
$\texttt{poacher} \rightarrow \Diamond \texttt{arrest}$ releases an
arrest task when a poacher is detected, whereas the rule
$\texttt{trap} \wedge \texttt{animal} \rightarrow
\Diamond \texttt{rescue}$ releases a rescue task when a trapped animal
is observed.
\end{example}









































%==============================
%==============================
%==============================
\subsection{Trajectory estimation of dynamic targets}
\label{ch06:subsec:trajectory}

The coordination problem depends on forecasts of target motion rather than on instantaneous observations alone. The two parts below introduce the trajectory predictor and the conformal prediction regions used to quantify forecast uncertainty.

\subsubsection{Trajectory predictor}
\label{ch06:subsec:traj-predictor}

Given a task horizon $T_{\varphi}$ and the observation history
$Y_{0:t}$, an independent trajectory predictor is trained for each
target using the training dataset $D_{\texttt{tra}}$. For target
$m \in \mathcal{M}$, let
$D_{\texttt{tra}}^m \subset D_{\texttt{tra}}$ denote the subset of
training trajectories associated with target $m$. For each
$Y^{(i)} \in D_{\texttt{tra}}^m$, write
$Y_{0:T_{\varphi},m}^{(i)} \triangleq
(Y_{0,m}^{(i)},\cdots,Y_{t,m}^{(i)},Y_{t+1,m}^{(i)},\cdots,
Y_{T_{\varphi},m}^{(i)})$.
The predictor for target $m$ is a map
$\Upsilon_m : \mathbb{R}^{2(t+1)} \rightarrow
\mathbb{R}^{2(T_{\varphi}-t)}$ that estimates the future target states
over the horizon
$\mathcal{H}_t \triangleq \{t+1,\cdots,T_{\varphi}\}$. The predicted
trajectory of target $m$ is therefore
$\widehat{Y}_{\mathcal{H}_t,m} \triangleq \Upsilon_m(Y_{0:t,m})$,
with
$\Upsilon_m(Y_{0:t,m}) \triangleq
(\widehat{Y}_{t+1 \mid t,m},\cdots,\widehat{Y}_{T_{\varphi} \mid t,m})$.
Let
$\mathbf{\Upsilon} \triangleq (\Upsilon_1,\cdots,\Upsilon_M)$ denote the
collection of all target-wise predictors.
In principle, any suitable trajectory predictor may be used, including
long short-term memory networks, denoted by LSTM,
\cite{Hochreiter1997LSTM}, recurrent neural networks~\cite{rnn}, and
gated recurrent units~\cite{gru}. In the present setting, an LSTM
network is adopted for each target. The training objective is given by
\eqref{ch06:eq:traj-loss}:
\begin{equation}
\label{ch06:eq:traj-loss}
\min_{\Upsilon_m}
\frac{1}{|D_{\texttt{tra}}^m|}
\sum_{i=1}^{|D_{\texttt{tra}}^m|}
\left\|
Y_{\mathcal{H}_t,m}^{(i)} -
\Upsilon_m(Y_{0:t,m}^{(i)})
\right\|^2,
\end{equation}
where $Y_{\mathcal{H}_t,m}^{(i)}$ is the future segment of the
trajectory sample $Y_{0:T_{\varphi},m}^{(i)}$. The resulting
target-wise predictions are stacked as
$\widehat{Y}_{\mathcal{H}_t} \triangleq
(\widehat{Y}_{\mathcal{H}_t,1},\cdots,\widehat{Y}_{\mathcal{H}_t,M})$.
Based on the measured target velocities $V_{0:t}$, define
$v_m^{\ast}(t) \triangleq \max \{|V_{\varsigma,m}| : \varsigma \leq t\}$
as the maximum recorded speed of target $m$ up to time $t$. This
quantity is used later as a velocity bound in the planning stage.

\subsubsection{Conformal prediction regions}
\label{ch06:subsec:cpr}

Conformal prediction is used to construct uncertainty regions around the
nominal trajectory forecasts. These regions provide finite-sample
coverage guarantees for the true target trajectories; see
\cite{Angelopoulos2021CP} for background. Following
\cite{Lars2023CP}, the prediction regions are constructed
independently for each target.
Given observations $Y_{0:t}$ at time $t$, the predictors
$\mathbf{\Upsilon}$ generate the nominal forecasts
$\widehat{Y}_{\mathcal{H}_t}$. For each target
$m \in \mathcal{M}$ and each failure probability
$\delta \in (0,1)$, define
$G_{\mathcal{H}_t,m} \triangleq
(G_{t+1 \mid t,m},\cdots,G_{T_{\varphi} \mid t,m})$. The
coverage guarantee is given by \eqref{ch06:eq:cpg}:
\begin{equation}
\label{ch06:eq:cpg}
\Pr\Big(
\|Y_{h,m} - \widehat{Y}_{h \mid t,m}\|
\leq G_{h \mid t,m},
\ \forall h \in \mathcal{H}_t
\Big) \geq 1 - \delta,
\end{equation}
where $G_{h \mid t,m}$ is the conformal prediction radius associated
with target $m$ at step $h$.
To compute these radii, define the nonconformity score by
\eqref{ch06:eq:ncs}:
\begin{equation}
\label{ch06:eq:ncs}
R_{h \mid t,m}^{(i)} \triangleq
\left\|
Y_{h,m}^{(i)} - \widehat{Y}_{h \mid t,m}^{(i)}
\right\|,
\end{equation}
where $Y^{(i)} \in D_{\texttt{cal}}^m$ is a calibration trajectory and
$D_{\texttt{cal}}^m$ is the calibration dataset associated with target
$m$. For every $Y^{(i)} \in D_{\texttt{cal}}^m$, the predictor produces
$\widehat{Y}_{h \mid t,m}^{(i)}$, and the score
$R_{h \mid t,m}^{(i)}$ is then evaluated using
\eqref{ch06:eq:ncs}. Let
$\bar{\delta} \triangleq \delta/(T_{\varphi}-t)$, and augment the score
set by
$R_{h \mid t,m}^{(|D_{\texttt{cal}}^m|+1)} \triangleq \infty$.
After sorting the resulting scores in non-decreasing order, let
$p \triangleq
\left\lceil
(|D_{\texttt{cal}}^m|+1)(1-\bar{\delta})
\right\rceil$.
The bound $G_{h \mid t,m}$ is then chosen as the $p$-th smallest score.
Finally, the prediction regions for all targets are collected as
$G_{\mathcal{H}_t} \triangleq
(G_{\mathcal{H}_t,1},\cdots,G_{\mathcal{H}_t,M})$.

%=======================================
\setlength{\textfloatsep}{5pt}
\begin{algorithm}[t]
    \caption{Offline construction of conformal prediction regions}
    \label{ch06:alg:cp}
    \SetKwInOut{Input}{Input}
    \SetKwInOut{Output}{Output}
    \Input{Failure probability $\delta$, task horizon $T_{\varphi}$,
      calibration dataset $D_{\texttt{cal}}^m$.}
    \Output{Prediction radii $G_{h \mid t,m}$.}
    \For{$t$ from $0$ to $T_{\varphi}-1$}{
        $\bar{\delta} \gets \delta/(T_{\varphi}-t)$;\\
        $p \gets
        \left\lceil
        (|D_{\texttt{cal}}^m|+1)(1-\bar{\delta})
        \right\rceil$;\\
        \For{$h$ from $t+1$ to $T_{\varphi}$}{
            Obtain $\widehat{Y}_{h \mid t,m}^{(i)}$ for each
            $Y^{(i)} \in D_{\texttt{cal}}^m$;\\
            Compute $R_{h \mid t,m}^{(i)}$ by
            \eqref{ch06:eq:ncs};\\
            $R_{h \mid t,m}^{(|D_{\texttt{cal}}^m|+1)}
            \gets \infty$;\\
            Sort $R_{h \mid t,m}^{(i)}$ in non-decreasing order;\\
            $G_{h \mid t,m} \gets R_{h \mid t,m}^{(p)}$;\\
        }
    }
\end{algorithm}
%=======================================





















%==============================
%==============================
%==============================
%==============================
\subsection{R-posets for task formulas}
\label{ch06:subsec:poset}

The notion of an R-poset was introduced in
Chapter~\ref{chap:top-down}, especially in
Section~\ref{ch04:subsec:task-decompose-posets}, as a compact abstraction
of the temporal constraints induced by a team-level temporal task. That
construction is reused here. Accordingly, only the ingredients needed
for the present section are recalled.
Given the sc-LTL formula $\varphi$, first translate it into the
Nondeterministic B\"uchi Automaton
$\mathcal{A}_{\varphi} \triangleq (S,\Sigma,\delta,S_0,S_F)$, where
$S$ is the state set, $\Sigma \triangleq 2^{AP}$ is the alphabet,
$\delta : S \times \Sigma \rightarrow 2^{S}$ is the transition
relation, and $S_0, S_F \subseteq S$ are the initial and accepting
state sets. Along any satisfying run of $\mathcal{A}_{\varphi}$, the
decomposition procedure in Definition~\ref{ch04:def:subtasks} extracts a
finite set of indexed subtasks $\Omega_{\varphi}$ together with the
relevant temporal relations. The subsequent relaxation step retains only
those constraints that are essential for correctness, rather than the
full sequential order of the automaton run.
Recall that an R-poset associated with $\varphi$ is the triple
\begin{equation}
\label{ch06:eq:r-poset}
P_{\varphi} \triangleq
(\Omega_{\varphi},\preceq_{\varphi},\neq_{\varphi}),
\end{equation}
where $\Omega_{\varphi}$ is the subtask set,
$\preceq_{\varphi} \subseteq
\Omega_{\varphi} \times \Omega_{\varphi}$ is the precedence relation,
and $\neq_{\varphi} \subseteq 2^{\Omega_{\varphi}}$ is the
mutual-exclusion relation. The relation
$(\omega_1,\omega_2) \in \preceq_{\varphi}$ means that
$\omega_2$ cannot start before $\omega_1$ starts. The relation
$\widehat{\Omega} \in \neq_{\varphi}$ for some
$\widehat{\Omega} \subseteq \Omega_{\varphi}$ means that the subtasks in
$\widehat{\Omega}$ cannot be executed simultaneously.


The definition above for $P_{\varphi}$ is the present-section specialization
of Definition~\ref{ch04:def:partial},
Definition~\ref{ch04:def:poset}, and \eqref{ch04:eq:poset} to the task
model in Section~\ref{ch06:sec:problem}. The abstraction in
\eqref{ch06:eq:r-poset} is less restrictive than the full automaton
run, since only the essential temporal dependencies are retained. At
the same time, it remains task-consistent. In Chapter~\ref{chap:top-down},
Definition~\ref{ch04:def:language-poset} and
\eqref{ch04:eq:poset-language} define the language $L(P_{\varphi})$
induced by an R-poset, and Lemma~\ref{ch04:lemma:accepting-poset}
shows that, for every accepting R-poset returned by
Algorithm~\ref{ch04:alg:compute-poset}, one has
$L(P_{\varphi}) \subseteq \mathcal{L}_{\varphi}$. Therefore, any plan
synthesized to satisfy the relations encoded by $P_{\varphi}$ is
consistent with the original task formula $\varphi$.
Different accepting R-posets may correspond to the same task formula,
since the retained temporal relations need not be unique. In practice,
the R-poset in \eqref{ch06:eq:r-poset} is computed by the anytime
procedure $\texttt{compute\_poset}(\cdot)$ in
Algorithm~\ref{ch04:alg:compute-poset}. The resulting structure is an R-poset that retains the essential precedence and synchronization relations needed for assignment and execution.














%==============================
%==============================
%==============================
%==============================
%==============================
%==============================
%==============================
%==============================
\subsection{Uncertainty-aware task assignment}
\label{ch06:subsec:assignment}

Given the target-motion estimates
$(\widehat{Y}_{\mathcal{H}_t},G_{\mathcal{H}_t})$ and the R-poset
$P_{\varphi} \triangleq
(\Omega_{\varphi},\preceq_{\varphi},\neq_{\varphi})$, the objective is
to assign all subtasks in $\Omega_{\varphi}$ to the robot team
$\mathcal{N}$ so that the temporal relations in
$\preceq_{\varphi}$ and $\neq_{\varphi}$ are respected and the average
makespan $\overline{T}_{\varphi}$ is minimized. This subsection extends
the assignment perspective of Chapter~\ref{chap:top-down} to the present
setting, where completion times depend on predicted target motion and its
associated uncertainty.

\subsubsection{CP-based Monte Carlo tree search}
\label{ch06:sec:cp-mcts}

A centralized task-assignment procedure based on Monte Carlo Tree Search
is adopted to address the resulting combinatorial search problem. In
contrast to the deterministic assignment procedure in
Chapter~\ref{chap:top-down}, the present method evaluates partial plans
through predicted target trajectories together with their conformal
uncertainty regions. The resulting scheme is referred to as CP-based
Monte Carlo Tree Search, abbreviated as CP-MCTS. As summarized in
Algorithm~\ref{ch06:alg:cp-mcts}, the search proceeds through four
stages, namely selection, expansion, simulation, and backpropagation,
until the prescribed time budget is exhausted. A principal difficulty is
that the number of child nodes generated during expansion grows rapidly
with the numbers of robots and subtasks. If every expanded node were
forwarded to simulation, the search would become excessively
breadth-oriented and would devote insufficient effort to deeper and more
informative branches. To mitigate this effect, a prediction-based
screening metric is introduced after expansion so that only the most
promising nodes are passed to the simulation stage.

(I) \emph{Selection and Expansion}.
Each node of the search tree represents a partial assignment of
subtasks, written as
$\nu \triangleq (\tau_1,\cdots,\tau_N)$, where $\tau_n$ is the ordered
sequence of subtasks currently assigned to robot $n \in \mathcal{N}$.
During selection, the Upper Confidence Bound applied to Trees
criterion~\cite{2006Bandit} is used. The UCT score of the $i$-th child
is
$\overline{\xi}_i + Q\sqrt{\ln B/b_i}$, where $\overline{\xi}_i$ is the
estimated value of that child, $b_i$ is its visit count, $B$ is the
visit count of the current node, and $Q > 0$ balances exploration and
exploitation. Starting from the root, the child with the largest UCT
score is selected recursively until a leaf node is reached.

During expansion, unless a terminal node with a complete assignment has
been reached, child nodes are generated by assigning one admissible
next subtask. Let
$\Omega_{\nu} \triangleq
\bigcup_{n \in \mathcal{N}}\{\omega \mid \omega \in \tau_n\}$
denote the set of subtasks already assigned at node $\nu$, and let
$\Omega_{\nu}^{-} \triangleq \Omega_{\varphi}\setminus\Omega_{\nu}$
denote the set of unassigned subtasks. The admissible next-subtask set
is defined by
\begin{equation}
\label{ch06:eq:admissible}
\Omega_{\nu}^{\texttt{a}} \triangleq
\{\omega \in \Omega_{\nu}^{-} \mid
\operatorname{Pre}(\omega)\subseteq \Omega_{\nu}\},
\end{equation}
where $\operatorname{Pre}(\omega)$ is the set of predecessors of
$\omega$ in the R-poset. Equation~\eqref{ch06:eq:admissible} is the
same structural feasibility condition used in Chapter~\ref{chap:top-down}.
Once a subtask $\omega \in \Omega_{\nu}^{\texttt{a}}$ is selected, a
child node $\nu^{+}$ is created by assigning $\omega$ to a feasible
robot group $\mathcal{I}_{\omega}$ and appending it to the local plan
of each robot in that group. For the resulting child node, define the
set of key subtasks by
\begin{equation}
\label{ch06:eq:key-subtasks}
\Omega_{\nu^{+}}^{\texttt{s}} \triangleq
\Bigl\{
\omega \in \Omega_{\nu^{+}}
\;\Bigm|\;
\exists n \in \mathcal{N} \text{ such that } \omega=\tau_n[0],
\ \omega \notin \tau_{n'}[1:] \text{ for all } n' \in \mathcal{N}
\Bigr\},
\end{equation}
where $\Omega_{\nu^{+}}^{\texttt{s}}$ contains the subtasks predicted to
be executed first across the robots. Stepwise simulation based on the
nominal trajectories $\widehat{Y}_{\mathcal{H}_t}$ then yields a
predicted completion time $\widehat{T}_{\omega}$ for each assigned
subtask $\omega \in \Omega_{\nu^{+}}$.

%=======================================
\setlength{\textfloatsep}{5pt}
\begin{algorithm}[!t]
    \caption{CP-based MCTS for uncertainty-aware task assignment}
    \label{ch06:alg:cp-mcts}
    \SetKwInOut{Input}{Input}
    \SetKwInOut{Output}{Output}
    \Input{Robots $\mathcal{N}$, current team state $x_t$, R-poset
    $P_{\varphi}$, duration map $\rho$, time budget $t_b$, target
    estimates $\widehat{Y}_{\mathcal{H}_t}, G_{\mathcal{H}_t}$.}
    \Output{Plan $\Pi_{\texttt{c}}^\star$ and risk value
    $\eta_{\texttt{c}}^\star$.}
    Initialize the root node $\nu_0$ and set
    $\eta_{\texttt{c}}^\star \leftarrow \infty$\;
    \While{$\mathrm{time} < t_b$}{
        $\nu \leftarrow \texttt{Selection}(\nu_0)$\;
        $\{\nu_{+}\} \leftarrow
        \texttt{Expansion}(\nu,P_{\varphi})$\;
        Filter the candidate nodes by $\zeta$ in
        \eqref{ch06:eq:metric} and obtain
        $\{\nu_{\texttt{s}}\}$\;
        \ForEach{$\nu_{\texttt{s}} \in \{\nu_{\texttt{s}}\}$}{
            $\Pi_{\texttt{c}} \leftarrow
            \texttt{Rollout}(\nu_{\texttt{s}},P_{\varphi},\rho,
            \widehat{Y}_{\mathcal{H}_t},x_t)$\;
            $\eta_{\texttt{c}} \leftarrow
            \texttt{Eval}(\Pi_{\texttt{c}},P_{\varphi},\rho,
            \widehat{Y}_{\mathcal{H}_t},G_{\mathcal{H}_t},x_t)$\;
            \If{$\eta_{\texttt{c}} < \eta_{\texttt{c}}^\star$}{
                $\eta_{\texttt{c}}^\star \leftarrow
                \eta_{\texttt{c}}$\;
                $\Pi_{\texttt{c}}^\star \leftarrow
                \Pi_{\texttt{c}}$\;
            }
            $\xi_{\texttt{c}} \leftarrow
            2-\eta_{\texttt{c}}/\eta_{\texttt{c}}^\star$\;
            \texttt{Backpropagate}($\nu_{\texttt{s}},\xi_{\texttt{c}}$)\;
        }
    }
\end{algorithm}
%=======================================


(II) \emph{Simulation}.
The simulation stage consists of three components. First, a rollout
policy completes the current partial assignment. Second, the resulting
complete plan is evaluated under samples drawn from the prediction
regions. Third, a risk metric is computed from the induced
average-makespan distribution. Starting from the selected child node,
rollout is applied recursively until all subtasks are assigned. At each
iteration, the next subtask is chosen from the admissible set in
\eqref{ch06:eq:admissible} and is assigned to a feasible robot group.
To preserve rollout diversity, an $\epsilon$-randomized rule is used.
With probability $\epsilon$, a feasible robot group is selected
uniformly at random. With probability $1-\epsilon$, the group predicted
to start the subtask earliest is selected. Once a complete plan
$\Pi_{\texttt{c}}$ has been obtained, a sampling-based procedure
approximates its average-makespan distribution
$\mathbf{\overline{T}}_{\texttt{c}}$ through stepwise simulation under
$z \in \mathbb{N}$ samples drawn from the prediction regions
$(\widehat{Y}_{\mathcal{H}_t},G_{\mathcal{H}_t})$. The corresponding
risk assessment is then performed through the value-at-risk and
conditional value-at-risk criteria.

For a complete plan $\Pi_{\texttt{c}}$ and a risk level
$\alpha \in (0,1]$, the value at risk is defined by
$\mathrm{VaR}_{\alpha}(\Pi_{\texttt{c}}) \triangleq
\inf\bigl\{
\rho \in \mathbb{R}
\bigm|
\Pr(\mathbf{\overline{T}}_{\texttt{c}} \geq \rho) \geq \alpha
\bigr\}$,
and the associated conditional value at risk is defined by
$\mathrm{CVaR}_{\alpha}(\Pi_{\texttt{c}}) \triangleq
\mathbb{E}\bigl[
\mathbf{\overline{T}}_{\texttt{c}}
\bigm|
\mathbf{\overline{T}}_{\texttt{c}} \geq
\mathrm{VaR}_{\alpha}(\Pi_{\texttt{c}})
\bigr]$.
Here, $\mathrm{VaR}_{\alpha}$ identifies a tail threshold, whereas
$\mathrm{CVaR}_{\alpha}$ measures the expected cost beyond that
threshold and is therefore more informative for risk-sensitive
assignment. Accordingly, the evaluation metric of a complete plan is
defined by
$\eta_{\texttt{c}} \triangleq
\mathrm{CVaR}_{\alpha}(\Pi_{\texttt{c}})$.
The procedure that generates the sample distribution and evaluates this
criterion is denoted by $\texttt{Eval}(\cdot)$. During the search, the
best plan $\Pi_{\texttt{c}}^\star$ and its minimal risk value
$\eta_{\texttt{c}}^\star$ are maintained and updated whenever a
candidate plan yields a smaller value of $\eta_{\texttt{c}}$.

(III) \emph{Backpropagation}.
To support the tree search, define the normalized performance score
$\xi_{\texttt{c}} \triangleq
2-\frac{\eta_{\texttt{c}}}{\eta_{\texttt{c}}^\star}$,
which permits comparison of nodes across different branches. During
backpropagation, the value estimates and visit counts are propagated
from the selected node back to the root. The screening metric used
after expansion is defined below:
\begin{equation}
\label{ch06:eq:metric}
\zeta_{\nu^{+}} \triangleq
\frac{
\sum_{i=1}^{|\Omega_{\nu^{+}}|}\widehat{T}_{\omega_i}
+
\sum_{j=1}^{|\Omega_{\nu^{+}}^{\texttt{s}}|}
\widehat{T}_{\omega_j}^{\texttt{e}}
}{
|\Omega_{\nu^{+}}|
},
\end{equation}
where
$\widehat{T}_{\omega_j}^{\texttt{e}} \triangleq
G_{\widehat{T}_{\omega_j},m}/
(\min_{n \in \mathcal{I}_{\omega_j}} \{v_n - v_m^{\ast}\})$ is the
uncertainty-induced delay estimate. Equation~\eqref{ch06:eq:metric}
approximates the average completion time of the currently assigned
subtasks after uncertainty adjustment. After expansion, the child node
with smallest $\zeta_{\nu^{+}}$ for each candidate next subtask is sent
to simulation, and the retained nodes form the filtered set
$\{\nu_{\texttt{s}}\}$. If
$\min_{n \in \mathcal{I}_{\omega}} v_n \leq v_m^{\ast}$
for some key subtask $\omega \in \Omega_{\nu^{+}}^{\texttt{s}}$, the
assigned robots may fail to close the distance to the target. In that
case, the node is treated as infeasible by setting
$\widehat{T}_{\omega} = \infty$, which yields
$\zeta_{\nu^{+}} = \infty$. Such nodes are excluded before simulation.
For comparison, two simpler evaluation schemes are also considered. The
first ignores target motion and treats all targets as stationary, which
yields the no-trajectory-prediction baseline NTP. The second uses the
nominal predictions $\widehat{Y}_{\mathcal{H}_t}$ but ignores the
uncertainty regions $G_{\mathcal{H}_t}$, which yields the
no-uncertainty baseline NU. Relative to these baselines, the
risk-sensitive CVaR criterion captures both motion forecast and
uncertainty growth in the assignment process.


\begin{lemma}
\label{ch06:lemma:key-bound}
Given an expanded node $\nu^{+}$, the completion times of the key
subtasks in $\Omega_{\nu^{+}}^{\texttt{s}}$ satisfy:
\begin{equation}
\label{ch06:eq:key-bound}
\Pr\Bigl(
T_{\omega} \leq \widehat{T}_{\omega} +
\frac{G_{\widehat{T}_{\omega},m}}
{\min_{n \in \mathcal{I}_{\omega}} v_n - v_m^{\ast}},
\ \forall \omega \in \Omega_{\nu^{+}}^{\texttt{s}}
\Bigr)
\geq
(1-\delta)^{|\Omega_{\nu^{+}}^{\texttt{s}}|},
\end{equation}
where $T_{\omega}$ is the actual completion time of subtask
$\omega \in \Omega_{\nu^{+}}^{\texttt{s}}$, $m$ is the associated
target, $\mathcal{I}_{\omega}$ is the robot group assigned to
$\omega$, and $v_m^{\ast}$ is the target-speed bound defined in
Section~\ref{ch06:subsec:traj-predictor}.
\end{lemma}

\begin{proof}
Fix any $\omega \in \Omega_{\nu^{+}}^{\texttt{s}}$, and let $m$ be its
associated target. By construction of the stepwise simulation, all
robots in $\mathcal{I}_{\omega}$ are predicted to reach the nominal
target position $\widehat{Y}_{\widehat{T}_{\omega},m}$ by time
$\widehat{T}_{\omega}$. By \eqref{ch06:eq:cpg}, the actual target
position at that time satisfies
$
\left\|
Y_{\widehat{T}_{\omega},m}
-
\widehat{Y}_{\widehat{T}_{\omega},m}
\right\|
\leq
G_{\widehat{T}_{\omega},m}$,
with probability at least $1-\delta$. Applying the same bound to all
$\omega \in \Omega_{\nu^{+}}^{\texttt{s}}$ gives the probability factor
$(1-\delta)^{|\Omega_{\nu^{+}}^{\texttt{s}}|}$.
Hence, on that event, the remaining distance to the actual target
position is at most $G_{\widehat{T}_{\omega},m}$. Since the target speed
is bounded by $v_m^{\ast}$ and the slowest assigned robot has speed
$\min_{n \in \mathcal{I}_{\omega}} v_n$, the residual distance is closed
with relative speed at least
$\min_{n \in \mathcal{I}_{\omega}} v_n - v_m^{\ast}$. Therefore, the
additional delay is bounded above by
$G_{\widehat{T}_{\omega},m}/
(\min_{n \in \mathcal{I}_{\omega}} v_n - v_m^{\ast})$, which yields
\eqref{ch06:eq:key-bound}.
\end{proof}






















%=======================================
\setlength{\textfloatsep}{5pt}
\begin{algorithm}[!t]
    \caption{Online reactive and dynamic task assignment}
    \label{ch06:alg:ta}
    \SetKwInOut{Input}{Input}
    \SetKwInOut{Output}{Output}
    \Input{Robots $\mathcal{N}$, task formula $\varphi(t)$, duration
    map $\rho$, trajectory predictors $\mathbf{\Upsilon}$, prediction
    regions $G_{\mathcal{H}_t}$, initial team state $x_0$, and
    observations $\{Y_{0:t},V_{0:t}\}$.}
    \Output{Assignment $\Pi^{\star}$.}
    Initialize $t \leftarrow 0$,
    $P_{\varphi} \leftarrow \texttt{compute\_poset}(\varphi(0))$,
    and $\Omega_{\iota} \leftarrow \emptyset$\;
    Initialize
    $\Pi^{\star},\eta^{\star} \leftarrow
    \texttt{CP-MCTS}(P_{\varphi},\rho,
    \widehat{Y}_{\mathcal{H}_0},G_{\mathcal{H}_0},x_0)$\;
    \While{the execution is not terminated}{
        Each robot $n \in \mathcal{N}$ applies its local plan $\tau_n$\;
        Sense the team state $x_t$ and the target observation $Y_t$\;
        Obtain the updated prediction
        $\widehat{Y}_{\mathcal{H}_t}$\;
        Update $\varphi(t)$ by \eqref{ch06:eq:task-update}\;
        Update $P_{\varphi}$ by
        $\texttt{compute\_poset}(\varphi(t))$\;
        \If{$\varphi_{\texttt{r}}(t) \neq \emptyset$}{
            $\Pi^{\star},\eta^{\star} \leftarrow
            \texttt{CP-MCTS}(P_{\varphi},\rho,
            \widehat{Y}_{\mathcal{H}_t},G_{\mathcal{H}_t},x_t)$\;
        }
        \Else{
            $\eta^{\star} \leftarrow \eta^{\star} - 1$\;
            \If{$\Omega_{\texttt{c}}(t) \neq \emptyset$}{
                Update $\eta^{\star}$ by
                \eqref{ch06:eq:update_eta}\;
                $\Omega_{\iota} \leftarrow
                \Omega_{\iota}\cup\Omega_{\texttt{c}}(t)$\;
            }
            $\eta_t^{\star} \leftarrow
            \texttt{Eval}(\Pi^{\star},P_{\varphi},\rho,
            \widehat{Y}_{\mathcal{H}_t},G_{\mathcal{H}_t},x_t)$\;
            \If{$\eta_t^{\star} > (1+\gamma)\eta^{\star}$}{
                $\widehat{\Pi}_t,\widehat{\eta}_t \leftarrow
                \texttt{CP-MCTS}(P_{\varphi},\rho,
                \widehat{Y}_{\mathcal{H}_t},G_{\mathcal{H}_t},x_t)$\;
                \If{$\widehat{\eta}_t < \eta_t^{\star}$}{
                    $\Pi^{\star} \leftarrow \widehat{\Pi}_t$\;
                    $\eta^{\star} \leftarrow \widehat{\eta}_t$\;
                }
            }
        }
        $t \leftarrow t+1$\;
    }
\end{algorithm}
%=======================================

%==============================
%==============================
%==============================
\subsection{Overall analyses}
\label{ch06:subsec:execution}

The preceding assignment method is intended for repeated use inside a receding-horizon loop. The two parts below summarize the resulting online execution scheme and the corresponding complexity and generalization properties.

\subsubsection{Online execution and adaptation}

The assignment method developed above is embedded into an online
execution scheme with receding-horizon replanning. As in the online
adaptation framework of Chapter~\ref{chap:top-down}, the current plan is
executed non-preemptively, while replanning is triggered only when new
task information becomes available or when the quality of the current
assignment deteriorates sufficiently.
As summarized in
Algorithm~\ref{ch06:alg:ta}, the execution logic has two stages. The
first is initial planning. The second is online adaptation. After the
initial assignment has been computed, each robot
$n \in \mathcal{N}$ executes its local plan $\tau_n$ by moving toward
the targets associated with its assigned subtasks and then performing
the required actions. If the current action belongs to a multi-robot
collaboration $C_k$ with $|C_k|>1$, the participating robot must wait
for the other collaborators before execution. This partial
synchronization accommodates uncertainty in travel time and action
duration. If the action is not collaborative, the robot proceeds
independently.
At each time step $t>0$, the team state $x_t$ and the target
observations $Y_t$ are updated, and new target predictions
$\widehat{Y}_{\mathcal{H}_t}$ are generated from $Y_{0:t}$ by the
predictors $\mathbf{\Upsilon}$. The task formula is updated whenever new
reactive tasks are released:
\begin{equation}
\label{ch06:eq:task-update}
\varphi(t) \triangleq \varphi(t^{-}) \wedge \varphi_{\texttt{r}}(t),
\end{equation}
where $\varphi(t^{-})$ is the formula immediately before time $t$, and
$\varphi_{\texttt{r}}(t) \triangleq
\bigwedge_{\bar{e}\in\bar{E}}
\Diamond \varphi_{\texttt{rep}}^{\bar{e}}(t)$ is the conjunction of the
reactive tasks released at time $t$. The R-poset $P_{\varphi}$ is then
updated accordingly.

Replanning is triggered in either of two situations. The first is the
release of a new reactive task, namely
$\varphi_{\texttt{r}}(t) \neq \emptyset$. The second is a significant
degradation in the quality of the current plan, namely
$\eta_t^{\star} > (1+\gamma)\eta^{\star}$, where
$\gamma \in (0,1)$ is the replanning threshold. The quantity
$\eta_t^{\star}$ is obtained from $\texttt{Eval}(\cdot)$ using the
latest state and target information. For completed tasks, the expected
performance value is updated by
\begin{equation}
\label{ch06:eq:update_eta}
\eta_{+}^{\star} \triangleq
\frac{
|\varphi(t)|-|\Omega_{\iota}|+|\Omega_{\texttt{c}}(t)|
}{
|\varphi(t)|-|\Omega_{\iota}|
}
\eta_{-}^{\star},
\end{equation}
where $\eta_{+}^{\star}$ and $\eta_{-}^{\star}$ are the updated and
previous values, $|\varphi(t)|$ is the total number of tasks,
$|\Omega_{\iota}|$ is the number of completed tasks, and
$|\Omega_{\texttt{c}}(t)|$ is the number of tasks completed at time $t$.
Whenever replanning is activated, CP-MCTS produces a candidate plan
$\widehat{\Pi}_t$ with associated metric $\widehat{\eta}_t$. The new
plan is adopted only if it improves upon the current one, that is, if
$\widehat{\eta}_t < \eta_t^{\star}$. Otherwise, the current plan
$\Pi^{\star}$ is retained, and the value $\eta^{\star}$ is decreased
monotonically to reflect the progress already achieved. This procedure
is repeated until the mission terminates.

\subsubsection{Complexity and generalization}

The complexity of Algorithm~\ref{ch06:alg:ta} follows from the
individual components introduced in the preceding subsections. The
construction of an R-poset has worst-case complexity
$\mathcal{O}(J^{2})$, where $J$ is the number of subtasks and is
bounded by the number of edges in the underlying automaton. For the
assignment stage, the worst-case search space is
$\mathcal{O}(J!\cdot N^{J})$, since the subtask orderings are
combinatorial and the robot-group assignments grow exponentially with
the number of robots $N$. By contrast, the rollout stage in CP-MCTS has
complexity $\mathcal{O}(J\cdot N)$, since subtasks are assigned
sequentially through a random or greedy rule. The complexity of the
stepwise simulation is
$\mathcal{O}\bigl(zT(J+N)/\Delta t\bigr)$, where $z$ is the number of
samples, $T$ is the makespan, and $\Delta t$ is the simulation step
size.

Two immediate extensions of the execution scheme are worth noting. The
first concerns robot failures. If one or more robots fail during
execution, a modified CP-MCTS procedure may be used to generate a new
assignment. The root node is updated by removing the failed robots and
retaining the subtasks that have already been completed. Any subtask
interrupted by a failure is returned to the set of pending tasks, after
which the standard search procedure is applied to compute a revised
assignment.
The second extension concerns task priorities. If the tasks are
associated with weights $\alpha_{\ell}$, the optimization criterion can
be changed from the average makespan $\overline{T}_{\varphi}$ to the
weighted makespan
$\widetilde{T}_{\varphi} \triangleq
\sum_{\ell=1}^{|\varphi(t)|} \alpha_{\ell} T_{\varphi_{\ell}}$,
which places greater emphasis on high-priority tasks.
