 \section{Task Relaxation and Adaptation in Unknown Workspaces}
\label{ch03:sec:task_relaxation_adaptation}

The discrete planning framework developed in the previous chapter assumes that
the workspace abstraction is available before execution and remains fixed during
plan synthesis and implementation. This assumption is restrictive in partially
known environments. Blocked passages may be discovered only after motion begins,
semantic labels may be revised by sensing or communication, and a local temporal
task may appear infeasible on the current abstraction even though it is
achievable in the actual workspace. In such settings, the planning problem is
not only to generate a discrete plan, but also to determine how the task should
be relaxed, updated, or reinterpreted as new workspace information becomes
available.

This section studies that problem from the viewpoint of local temporal tasks.
It first considers a single robot and develops a progression from task
relaxation under infeasibility, to hard and soft specifications that preserve
safety, and then to online plan adaptation under workspace updates. It then
extends the same perspective to a multi-robot setting in which each robot is
still assigned its own local temporal task, while coordination is introduced
through collaborative knowledge transfer rather than through a centralized team
specification. In this way, the section establishes the logical basis for
bottom-up coordination under incomplete workspace knowledge.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Infeasible tasks}\label{ch03:sec:infeasible}

The notation of Chapter~\ref{chap:preliminaries} is adopted throughout this
subsection. Consider a robot with local task specification~$\varphi$. By
Definition~\ref{ch02:def:feasible}, $\varphi$ may be infeasible over the
current workspace abstraction~$\mathcal{T}_c$. This situation may arise because
the task is intrinsically infeasible, or because the task is feasible in the
actual workspace while the initial model~$\mathcal{T}_c$ is incomplete or
incorrect. In either case, the nominal synthesis procedure of
Section~\ref{ch02:sec:solution} returns failure. The relevant question is then
how the specification should be relaxed, and how a discrete plan should be
constructed so that the original task is violated as little as possible.

\begin{problem}
Assume that the task specification is infeasible. How should the specification
be relaxed, and how can a discrete plan be synthesized so that it satisfies the
original task as much as possible?
\end{problem}

A partial answer was given in~\cite{kim2012approximate}, where a relaxed
specification automaton~$\mathcal{A}_{\varphi}'$ is generated so that it
remains close to~$\mathcal{A}_{\varphi}$ and becomes feasible over
$\mathcal{T}_c$. Once such an automaton is obtained, a plan can be synthesized
by the same procedure as in Section~\ref{ch02:sec:solution}. The present
treatment adopts a more discriminating objective. Even when several accepting
runs exist in $\mathcal{T}_c \otimes \mathcal{A}_{\varphi}'$, they may satisfy
the original task to different extents. The goal is therefore to identify a
discrete plan that preserves the original specification as much as possible
according to an explicit criterion, and then infer from that plan the
associated relaxation of the specification automaton.

\subsubsection{Relaxed product automaton}
\label{ch03:subsec:relaxed-prod}

By Definition~\ref{ch02:def:feasible}, infeasibility of~$\varphi$ means that
the standard product automaton~$\mathcal{A}_p$ admits no accepting run. The
idea is therefore to relax the automaton-side constraints while keeping
the workspace unchanged. This leads to the relaxed product
B\"uchi automaton below.

\begin{definition}\label{ch03:def:relaxpro}
The relaxed product B\"uchi automaton is defined as:
\begin{equation}
\mathcal{A}_r \triangleq \mathcal{T}_c \times \mathcal{A}_{\varphi}
= (Q',\, 2^{AP},\, \delta',\, Q_0',\, \mathcal{F}',\, W_r),
\label{ch03:eq:relaxed-pba}
\end{equation}
where $Q' \triangleq \Pi \times Q$,
$AP \triangleq \{a_1,\, a_2,\, \cdots,\, a_K\}$,
$Q_0' \triangleq \Pi_0 \times Q_0$, and
$\mathcal{F}' \triangleq \Pi \times \mathcal{F}$. The transition relation
$\delta' : Q' \rightarrow 2^{Q'}$ satisfies
$\langle \pi_j,\, q_n \rangle \in \delta'(\langle \pi_i,\, q_m \rangle)$ if and
only if $(\pi_i,\, \pi_j) \in \longrightarrow_c$ and there exists
$l \in 2^{AP}$ such that $q_n \in \delta(q_m,\, l)$. The function
$W_r : Q' \times Q' \rightarrow \mathbb{R}^{+}$ is defined below.
\end{definition}

Compared with the standard product automaton in
Definition~\ref{ch02:def:product}, the construction in
Definition~\ref{ch03:def:relaxpro} differs in two respects. First, the usual
constraint $q_n \in \delta(q_m,\, L_c(\pi_i))$ is replaced by the weaker
condition that there exists some alphabet $l \in 2^{AP}$ for which
$q_n \in \delta(q_m,\, l)$. Second, the transition weight is no longer given
by the nominal product weight, but by a penalty-augmented cost that measures
how far the observed label is from the labels that would enable the automaton
transition.

To define that penalty, introduce the evaluation map
$\texttt{Eval} : 2^{AP} \rightarrow \{0,\, 1\}^{K}$ such that
$\texttt{Eval}(l) = \boldsymbol{\nu}$ if and only if
$\nu_i = 1$ when $a_i \in l$ and $\nu_i = 0$ otherwise, for
$i = 1,\, \cdots,\, K$. This map assigns to each alphabet a Boolean vector of
dimension~$K$. Based on this representation, the distance between two alphabets
$l,\, l' \in 2^{AP}$ is
$\rho(l,\, l') \triangleq \|\boldsymbol{\nu} - \boldsymbol{\nu}'\|_{1}
= \sum_{i=1}^{K} |\nu_i - \nu_i'|$, where
$\boldsymbol{\nu} = \texttt{Eval}(l)$ and
$\boldsymbol{\nu}' = \texttt{Eval}(l')$. This quantity is the Hamming
distance between the two Boolean encodings.
Next, for any nonempty set $\chi \subseteq 2^{AP}$, define the distance from
$l \in 2^{AP}$ to~$\chi$ as:
\begin{equation}
\texttt{Dist}(l,\, \chi) \triangleq
\begin{cases}
0, & \text{if } l \in \chi, \\
\min_{l' \in \chi} \rho(l,\, l'), & \text{otherwise},
\end{cases}
\label{ch03:eq:setdist}
\end{equation}
which is well defined only when $\chi \neq \emptyset$. An illustration of
\eqref{ch03:eq:setdist} is shown in Figure~\ref{ch03:distexam}.
Then, the transition weight of the relaxed product automaton is
\begin{equation}
\begin{aligned}
W_r(\langle \pi_i,\, q_m \rangle,\,
\langle \pi_j,\, q_n \rangle)
\triangleq {} & w_c(\pi_i,\, \pi_j) \\
& {} + \alpha \cdot
\texttt{Dist}(L_c(\pi_i),\, \chi(q_m,\, q_n)),
\end{aligned}
\label{ch03:eq:WR}
\end{equation}
where $\langle \pi_j,\, q_n \rangle \in
\delta'(\langle \pi_i,\, q_m \rangle)$ and $\alpha \geq 0$. Here,
$\chi(q_m,\, q_n) \triangleq
\{\, l \in 2^{AP} \mid q_n \in \delta(q_m,\, l) \,\}$ collects all alphabets
that enable the transition from $q_m$ to $q_n$ in
$\mathcal{A}_{\varphi}$. By Definition~\ref{ch03:def:relaxpro}, the existence
of such an alphabet is guaranteed whenever the relaxed transition is present,
so $\chi(q_m,\, q_n) \neq \emptyset$.
Equation~\eqref{ch03:eq:WR} has a direct interpretation. The term
$w_c(\pi_i,\, \pi_j)$ is the implementation cost of the transition in
$\mathcal{T}_c$, whereas
$\texttt{Dist}(L_c(\pi_i),\, \chi(q_m,\, q_n))$ quantifies the degree to which
the transition violates the original automaton constraint. A larger value of
$\alpha$ imposes a stronger penalty on such violations yielding
a stronger preference for preserving the original task.

\begin{example}\label{ch03:examp:dist}
Consider the NBA in Figure~\ref{ch02:fig:Buchi} and the transition from $q_1$
to $q_3$. In this case,
$\chi(q_1,\, q_3)
= \{\{a_2\},\, \{a_1,\, a_2\},\, \{a_2,\, a_3\},\,
\{a_1,\, a_2,\, a_3\}\}$. Therefore,
$\textup{\texttt{Dist}}(\{a_1\},\, \chi(q_1,\, q_3))
= \rho(\{a_1\},\, \{a_1,\, a_2\}) = 1$,
$\textup{\texttt{Dist}}(\{a_2\},\, \chi(q_1,\, q_3))
= \rho(\{a_2\},\, \{a_2\}) = 0$, and
$\textup{\texttt{Dist}}(\{a_2,\, a_3\},\, \chi(q_1,\, q_3))
= \rho(\{a_2,\, a_3\},\, \{a_2,\, a_3\}) = 0$.
\end{example}

\subsubsection{Balanced accepting run}\label{ch03:subsec:probStatement}

Since $\mathcal{A}_p$ has no accepting run, the search is carried out on the
relaxed automaton~$\mathcal{A}_r$. Mere existence of an accepting run is not
sufficient. Different accepting runs may induce different motion costs, and
they may deviate from the original task by different amounts. For that reason,
the same prefix-suffix structure as in the accepting runs introduced in
Chapter~\ref{chap:preliminaries} is retained:
\[
R \triangleq q_0' q_1' \cdots q_{f-1}'
[\, q_f' q_{f+1}' \cdots q_n' \,]^{\omega}
= \langle \pi_0,\, q_0 \rangle \cdots
\langle \pi_{f-1},\, q_{f-1} \rangle
[\, \langle \pi_f,\, q_f \rangle
\langle \pi_{f+1},\, q_{f+1} \rangle \cdots
\langle \pi_n,\, q_n \rangle \,]^{\omega},
\]
where $q_0' = \langle \pi_0,\, q_0 \rangle \in Q_0'$ and
$q_f' = \langle \pi_f,\, q_f \rangle \in \mathcal{F}'$.
The total cost of~$R$ in the relaxed product automaton is defined as:
\begin{equation}
\texttt{Cost}(R,\, \mathcal{A}_r)
\triangleq \texttt{cost}_{\tau}
+ \alpha \cdot \texttt{dist}_{\varphi},
\label{ch03:eq:prod-cost2}
\end{equation}
where $\tau \triangleq R|_{\Pi}$ is the induced motion plan. The implementation
cost in \eqref{ch03:eq:prod-cost2} is
\begin{equation}
\texttt{cost}_{\tau} \triangleq
\Big( \sum_{i=0}^{f-1} + \gamma \sum_{i=f}^{n-1} \Big)
w_c(\pi_i,\, \pi_{i+1}),
\label{ch03:eq:cost}
\end{equation}
and the accumulated specification distance is
\begin{equation}
\texttt{dist}_{\varphi} \triangleq
\Big( \sum_{i=0}^{f-1} + \gamma \sum_{i=f}^{n-1} \Big)
\texttt{Dist}(L_c(\pi_i),\, \chi(q_i,\, q_{i+1})),
\label{ch03:eq:dist}
\end{equation}
where $\gamma \geq 0$ weights the relative importance of the transient prefix
and the recurrent suffix.

\begin{problem}\label{ch03:pro:balanced}
Find an accepting run of $\mathcal{A}_r$ that minimizes the cost in
\eqref{ch03:eq:prod-cost2}.
\end{problem}

The solution to Problem~\ref{ch03:pro:balanced} is called the balanced
accepting run of~$\mathcal{A}_r$ and is denoted by $R_{\textup{bal}}$. The
induced balanced plan is
$\tau_{\textup{bal}} \triangleq R_{\textup{bal}}|_{\Pi}$.
As noted in the discussion of NBA representations in
Chapter~\ref{chap:preliminaries}, each transition of the NBA is encoded by a
Boolean formula that characterizes all admissible alphabets. Such a formula can
be represented by a binary decision diagram, so the distance in
\eqref{ch03:eq:setdist} can be evaluated without enumerating all alphabets in
$2^{AP}$. In particular, the operator ``$\vee$'' selects the smaller of the
two branch distances, whereas ``$\wedge$'' accumulates them.


\subsubsection{Balanced plan synthesis}\label{ch03:subsec:singleFeedback}

For fixed design parameters $\alpha$ and $\gamma$, the relaxed automaton
$\mathcal{A}_r$ may be built either explicitly from
Definition~\ref{ch03:def:relaxpro} or on the fly, exactly as in
Chapter~\ref{chap:preliminaries}. The only change is that the transition weight
is evaluated by \eqref{ch03:eq:WR}, which depends on the set distance in
\eqref{ch03:eq:setdist}. Under either construction,
Algorithm~\ref{ch02:alg:optimal} can then be applied to compute the balanced
accepting run~$R_{\textup{bal}}$, and hence the balanced plan
$\tau_{\textup{bal}} = R_{\textup{bal}}|_{\Pi}$.
Although $\mathcal{A}_r$ admits more automaton transitions than
$\mathcal{A}_p$, it never relaxes the transition relation of~$\mathcal{T}_c$.
Therefore, every balanced plan remains a valid path in~$\mathcal{T}_c$ and is
always implementable.


Once $R_{\textup{bal}}$ is known, the quantities
$\texttt{cost}_{\tau}$ and $\texttt{dist}_{\varphi}$ can be computed directly
from \eqref{ch03:eq:cost} and \eqref{ch03:eq:dist}. At the same time, a relaxed
specification automaton~$\mathcal{A}_{\varphi}'$ can be constructed by adding,
along the transitions of $R_{\textup{bal}}$, the automaton edges required to
support the observed labels. The resulting automaton is a valid relaxation of
$\mathcal{A}_{\varphi}$ in the sense of~\cite{kim2012approximate}. Hence each
balanced run determines simultaneously a balanced motion plan
$\tau_{\textup{bal}}$ and an associated relaxed specification automaton
$\mathcal{A}_{\varphi}'$.

\begin{lemma}\label{ch03:lem:distanceProof}
If $\textup{\texttt{dist}}_{\varphi} = 0$, then
$\tau_{\textup{bal}} \models \varphi$.
\end{lemma}

\begin{proof}
By \eqref{ch03:eq:setdist}, the function $\texttt{Dist}(\cdot,\, \cdot)$ is
nonnegative. Therefore, $\texttt{dist}_{\varphi} = 0$ implies that
$q_n \in \delta(q_m,\, L_c(\pi_i))$ for every transition
$(\langle \pi_i,\, q_m \rangle,\,
\langle \pi_j,\, q_n \rangle)$ along $R_{\textup{bal}}$. Hence every
transition of $R_{\textup{bal}}$ also satisfies the transition condition of the
standard product automaton $\mathcal{A}_p$ from
Definition~\ref{ch02:def:product}. Since $\mathcal{A}_p$ and $\mathcal{A}_r$
share the same state set, the same initial states, and the same accepting
states, $R_{\textup{bal}}$ is also an accepting run of~$\mathcal{A}_p$.
Lemma~\ref{ch02:lem:satisfy} then yields
$\tau_{\textup{bal}} \models \varphi$.
\end{proof}

Selecting the parameter~$\alpha$ is not always straightforward, since it
governs the trade-off between implementation cost and specification violation.
A simple extension is to run Algorithm~\ref{ch02:alg:optimal} for several
values of $\alpha$, retain the distinct balanced runs, and compare them through
the pairs $(\texttt{cost}_{\tau},\, \texttt{dist}_{\varphi})$. The preferred
plan $\tau_{\textup{opt}}$ can then be executed offline by the hybrid
controller in Algorithm~\ref{ch02:alg:hybrid}.

\begin{theorem}\label{ch03:thm:balanced-feasible}
If $\varphi$ is feasible over $\mathcal{T}_c$, then the balanced plan
$\tau_{\textup{bal}}$ satisfies $\varphi$ whenever
$\alpha > \underline{\alpha}$, where $\underline{\alpha}$ is given by
\eqref{ch03:underalpha}.
\end{theorem}

\begin{proof}
Suppose that $\varphi$ is feasible over $\mathcal{T}_c$. Then
Algorithm~\ref{ch02:alg:optimal} returns the optimal accepting run
$R_{\textup{opt}}$ of $\mathcal{A}_p$, whose total cost is
\begin{equation}
\texttt{Cost}(R_{\textup{opt}},\, \mathcal{A}_p)
\triangleq \underline{\alpha},
\label{ch03:underalpha}
\end{equation}
under the same value of $\gamma$ as in \eqref{ch03:eq:prod-cost2}. Since every
transition of $\mathcal{A}_p$ is also a transition of $\mathcal{A}_r$,
$R_{\textup{opt}}$ is an accepting run of $\mathcal{A}_r$ as well. Moreover,
because $R_{\textup{opt}}$ satisfies the original automaton constraints
exactly, its distance term vanishes, and thus
$\texttt{Cost}(R_{\textup{opt}},\, \mathcal{A}_r)
= \texttt{Cost}(R_{\textup{opt}},\, \mathcal{A}_p)
= \underline{\alpha}$.
Assume, for contradiction, that $\tau_{\textup{bal}} \not\models \varphi$.
Then \eqref{ch03:eq:dist} implies $\texttt{dist}_{\varphi} \geq 1$. Therefore,
by \eqref{ch03:eq:prod-cost2},
$\texttt{Cost}(R_{\textup{bal}},\, \mathcal{A}_r)
= \texttt{cost}_{\tau} + \alpha \cdot \texttt{dist}_{\varphi}
\geq \alpha$. If $\alpha > \underline{\alpha}$, then
$\texttt{Cost}(R_{\textup{bal}},\, \mathcal{A}_r)
> \underline{\alpha}
= \texttt{Cost}(R_{\textup{opt}},\, \mathcal{A}_r)$, which contradicts the
optimality of $R_{\textup{bal}}$ for Problem~\ref{ch03:pro:balanced}. Hence
$\tau_{\textup{bal}} \models \varphi$ holds.
\end{proof}



\begin{example}\label{ch03:examp:singleRS}
Consider the example in Figure~\ref{ch03:fig:toyExamSingle}. The robot starts
from region~$\pi_0$ and should eventually reach~$\pi_3$ and remain there, while
avoiding every region that satisfies $a_2$ or $a_3$. By varying $\alpha$ with
$\gamma = 5$, three distinct balanced plans are obtained:

(I) When the penalty on violating~$\varphi$ is small,
$\mathcal{A}_{\varphi}$ is revised by adding $q_1$ to
$\delta(q_0,\, \emptyset)$ and $q_1$ to $\delta(q_1,\, \emptyset)$. The
balanced plan is then $[\pi_0]^{\omega}$, indicated by the black hexagram, with
$\textup{\texttt{cost}}_{\tau} = 30$ and
$\textup{\texttt{dist}}_{\varphi} = 6$.
(II) When the penalty is increased, $\mathcal{A}_{\varphi}$ is revised by adding
$q_1$ to $\delta(q_0,\, \{a_2,\, a_3\})$. The resulting balanced plan is
$\pi_0\, \pi_1\, [\pi_3]^{\omega}$, indicated by the blue square, with
$\textup{\texttt{cost}}_{\tau} = 65$ and
$\textup{\texttt{dist}}_{\varphi} = 2$.
(III) When the penalty becomes sufficiently large,
$\mathcal{A}_{\varphi}$ is revised by adding $q_1$ to
$\delta(q_0,\, \{a_2\})$. The balanced plan becomes
$\pi_0\, \pi_2\, [\pi_3]^{\omega}$, indicated by the cyan triangle, with
$\textup{\texttt{cost}}_{\tau} = 85$ and
$\textup{\texttt{dist}}_{\varphi} = 1$. This case shows that the method
prefers to pass through~$\pi_2$, which satisfies only~$a_2$, rather than
through $\pi_1$, which satisfies both~$a_2$ and~$a_3$.
\end{example}












%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Soft and hard specifications}\label{ch03:sec:soft-hard}

The previous subsection addressed the synthesis of a balanced motion and task
plan when the specification may be infeasible. In many applications, however,
a temporal specification contains two qualitatively different components. One
component encodes safety-critical requirements that must never be
violated. The other encodes performance or achievement objectives that should
be satisfied as much as possible, but may need to be relaxed when the workspace
is only partially known. This subsection develops a synthesis framework for
that setting.

\subsubsection{Problem formulation}\label{ch03:subsec:formulation}

The robot transition system is still denoted by $\mathcal{T}_c$, as in
Definition~\ref{ch02:def:FTS}. The task specification remains an LTL formula
over $AP$, but now has the form
\begin{equation}
\varphi \triangleq \varphi^{\textup{soft}} \wedge \varphi^{\textup{hard}},
\label{ch03:eq:spec}
\end{equation}
where $\varphi^{\textup{hard}}$ and $\varphi^{\textup{soft}}$ are the hard and
soft components, respectively. Typical hard requirements include safety
constraints such as obstacle avoidance or recurrent charging, whereas soft
requirements encode additional objectives whose feasibility may depend on
workspace information that is initially incomplete. This decomposition is
motivated by the same partially known setting considered in
Section~\ref{ch03:sec:infeasible}: some task components may require relaxation,
but safety-critical constraints must remain inviolable.

\begin{problem}
Given the specification in \eqref{ch03:eq:spec}, synthesize a motion and task
plan such that $\varphi^{\textup{hard}}$ is satisfied fully, while
$\varphi^{\textup{soft}}$ is satisfied as much as possible.
\end{problem}

\subsubsection{Safety-ensured product automaton}
\label{ch03:subsec:intersection}

The decomposition in \eqref{ch03:eq:spec} leads naturally to three notions of
correctness.

\begin{definition}\label{ch03:def:vs}
A finite or infinite path $\tau$ of $\mathcal{T}_c$ is called \emph{valid}
if every consecutive pair of states along $\tau$ belongs to
$\longrightarrow_c$. An infinite valid path $\tau$ is called \emph{safe} if
$\tau \models \varphi^{\textup{hard}}$, and \emph{satisfying} if
$\tau \models \varphi$.
\end{definition}

Let
$\mathcal{A}^{\textup{hard}}
= (Q_1,\, 2^{AP},\, \delta_1,\, Q_{1,0},\, \mathcal{F}_1)$
and
$\mathcal{A}^{\textup{soft}}
= (Q_2,\, 2^{AP},\, \delta_2,\, Q_{2,0},\, \mathcal{F}_2)$
be the NBAs associated with $\varphi^{\textup{hard}}$ and
$\varphi^{\textup{soft}}$, respectively. Their construction was recalled in
Section~\ref{ch02:subsec:NBA}. For later use, define
$\chi_1(q_1,\, \check{q}_1) \triangleq
\{\, l \in 2^{AP} \mid \check{q}_1 \in \delta_1(q_1,\, l) \,\}$ and
$\chi_2(q_2,\, \check{q}_2) \triangleq
\{\, l \in 2^{AP} \mid \check{q}_2 \in \delta_2(q_2,\, l) \,\}$.
These sets play the same role as \eqref{ch03:eq:WR}, but are defined for the
hard and soft automata separately.
The hard component should be enforced exactly, whereas the soft component may
be relaxed. This leads to the following relaxed intersection automaton.

\begin{definition}\label{ch03:def:conj}
The relaxed intersection of $\mathcal{A}^{\textup{hard}}$ and
$\mathcal{A}^{\textup{soft}}$ is
\begin{equation}
\tilde{\mathcal{A}}_{\varphi}
\triangleq
(Q,\, 2^{AP},\, \delta,\, Q_0,\, \mathcal{F}),
\label{ch03:eq:intersect}
\end{equation}
where
$Q \triangleq Q_1 \times Q_2 \times \{1,\, 2\}$,
$Q_0 \triangleq Q_{1,0} \times Q_{2,0} \times \{1\}$, and
$\mathcal{F} \triangleq \mathcal{F}_1 \times Q_2 \times \{1\}$. The transition
relation $\delta : Q \times 2^{AP} \rightarrow 2^Q$ satisfies
$\langle \check{q}_1,\, \check{q}_2,\, \check{t} \rangle \in
\delta(\langle q_1,\, q_2,\, t \rangle,\, l)$ if and only if
$l \in \chi_1(q_1,\, \check{q}_1)$,
$\chi_2(q_2,\, \check{q}_2) \neq \emptyset$, and either
$q_t \notin \mathcal{F}_t$ with $\check{t} = t$, or
$q_t \in \mathcal{F}_t$ with
$\check{t} = \textup{\texttt{mod}}(t,\, 2) + 1$.
\end{definition}

Definition~\ref{ch03:def:conj} relaxes only the soft component. The actual
alphabet $l$ must satisfy the hard transition through
$l \in \chi_1(q_1,\, \check{q}_1)$, whereas the soft component requires only
that the transition from $q_2$ to $\check{q}_2$ be enabled by at least one
alphabet. Thus the soft component is relaxed structurally, while the hard
component remains exact. The auxiliary variable $t \in \{1,\, 2\}$ plays the
same role as in the standard NBA intersection~\cite{baier2008principles}: it
alternates the acceptance obligation between the two component automata.
Denote by $r|_{Q_1}$ and $r|_{Q_2}$ the projections of a run~$r$ of
$\tilde{\mathcal{A}}_{\varphi}$ onto the state sets of
$\mathcal{A}^{\textup{hard}}$ and $\mathcal{A}^{\textup{soft}}$,
respectively.

\begin{algorithm}[t]
\DontPrintSemicolon
\caption{Construction of the relaxed intersection, \texttt{RelaxInt}()}
\label{ch03:alg:relaxedint}
\KwIn{$\mathcal{A}^{\textup{soft}}$, $\mathcal{A}^{\textup{hard}}$}
\KwOut{$\tilde{\mathcal{A}}_{\varphi}$}

\ForEach{$q_1 \in Q_1$, $q_2 \in Q_2$, $t \in \{1,\,2\}$}{
  $q_m = \langle q_1,\, q_2,\, t \rangle \in Q$\;

  \If{$q_1 \in Q_{1,0}$, $q_2 \in Q_{2,0}$, and $t = 1$}{
    $q_m \in Q_0$\;
  }

  \If{$q_1 \in \mathcal{F}_1$ and $t = 1$}{
    $q_m \in \mathcal{F}$\;
  }

  \ForEach{$\check{q}_1 \in \textup{\texttt{Post}}(q_1)$,
           $\check{q}_2 \in \textup{\texttt{Post}}(q_2)$,
           $\check{t} \in \{1,\,2\}$}{
    $q_n = \langle \check{q}_1,\, \check{q}_2,\, \check{t} \rangle \in Q$\;

    \If{$(q_1 \notin \mathcal{F}_1$ and $t = 1$ and $\check{t} = 1)$ or
        $(q_2 \notin \mathcal{F}_2$ and $t = 2$ and $\check{t} = 2)$ or
        $(q_1 \in \mathcal{F}_1$ and $t = 1$ and $\check{t} = 2)$ or
        $(q_2 \in \mathcal{F}_2$ and $t = 2$ and $\check{t} = 1)$}{
      $q_n \in \delta(q_m,\, l)$, for all $l \in \chi_1(q_1,\, \check{q}_1)$\;
    }
  }
}
\textbf{return} $\tilde{\mathcal{A}}_{\varphi}$
\end{algorithm}

\begin{theorem}\label{ch03:theo:saferun}
Given an accepting run $r$ of $\tilde{\mathcal{A}}_{\varphi}$, the projection
$r|_{Q_1}$ is an accepting run of $\mathcal{A}^{\textup{hard}}$. Moreover,
$\mathcal{L}_{\omega}(\tilde{\mathcal{A}}_{\varphi}) \subseteq
\mathcal{L}_{\omega}(\mathcal{A}^{\textup{hard}})$.
\end{theorem}

\begin{proof}
Since $r$ is accepting, it visits the set
$\mathcal{F} = \mathcal{F}_1 \times Q_2 \times \{1\}$ infinitely often.
Therefore, the projection $r|_{Q_1}$ visits $\mathcal{F}_1$ infinitely often.
In addition, every transition of $r$ satisfies
$l \in \chi_1(q_1,\, \check{q}_1)$ by
Definition~\ref{ch03:def:conj}. Hence every projected transition of $r|_{Q_1}$
is a valid transition of $\mathcal{A}^{\textup{hard}}$. It follows that
$r|_{Q_1}$ is an accepting run of $\mathcal{A}^{\textup{hard}}$.
For the language inclusion, let
$\sigma \in \mathcal{L}_{\omega}(\tilde{\mathcal{A}}_{\varphi})$. Then
$\sigma$ induces an accepting run $r_{\sigma}$ of
$\tilde{\mathcal{A}}_{\varphi}$. By the first part of the proof,
$r_{\sigma}|_{Q_1}$ is an accepting run of
$\mathcal{A}^{\textup{hard}}$. Therefore,
$\sigma \in \mathcal{L}_{\omega}(\mathcal{A}^{\textup{hard}})$, which proves
the desired inclusion.
\end{proof}

To preserve the hard specification while still optimizing the soft one, combine
$\tilde{\mathcal{A}}_{\varphi}$ with the workspace transition system in a
relaxed product automaton.

\begin{definition}\label{ch03:def:safe}
The safety-ensured relaxed product B\"uchi automaton is defined as
\begin{equation}
\tilde{\mathcal{A}}_{r}
\triangleq
\mathcal{T}_c \times \tilde{\mathcal{A}}_{\varphi}
=
(Q',\, 2^{AP},\, \delta',\, Q_0',\, \mathcal{F}',\, W_r),
\label{ch03:eq:safe-product}
\end{equation}
where $Q' \triangleq \Pi \times Q$,
$Q_0' \triangleq \Pi_0 \times Q_0$, and
$\mathcal{F}' \triangleq \Pi \times \mathcal{F}$. The transition relation
$\delta' : Q' \rightarrow 2^{Q'}$ satisfies
$\langle \pi_j,\, q_n \rangle \in \delta'(\langle \pi_i,\, q_m \rangle)$ if and
only if $(\pi_i,\, \pi_j) \in \longrightarrow_c$ and
$q_n \in \delta(q_m,\, L_c(\pi_i))$. The weight function
$W_r : Q' \times Q' \rightarrow \mathbb{R}^{+}$ is given by:
\begin{equation}
\begin{aligned}
W_r(\langle \pi_i,\, q_m \rangle,\,
\langle \pi_j,\, q_n \rangle)
\triangleq {} & w_c(\pi_i,\, \pi_j) \\
& {} + \alpha \cdot
\textup{\texttt{Dist}}(L_c(\pi_i),\, \chi_2(q_2,\, \check{q}_2)),
\end{aligned}
\label{ch03:eq:weight-soft}
\end{equation}
where $q_m = \langle q_1,\, q_2,\, t \rangle$,
$q_n = \langle \check{q}_1,\, \check{q}_2,\, \check{t} \rangle$, and
$\alpha \geq 0$.
\end{definition}

The weight in \eqref{ch03:eq:weight-soft} has two components. The term
$w_c(\pi_i,\, \pi_j)$ is the implementation cost of the transition in
$\mathcal{T}_c$, whereas
$\texttt{Dist}(L_c(\pi_i),\, \chi_2(q_2,\, \check{q}_2))$ measures the degree
to which that transition violates the soft specification. The hard component
does not appear in the weight because it is enforced directly by the transition
relation of $\tilde{\mathcal{A}}_{\varphi}$.

\begin{theorem}\label{ch03:theo:safety}
Assume that $R$ is an accepting run of $\tilde{\mathcal{A}}_{r}$. Then its
projection $\tau \triangleq R|_{\Pi}$ is both valid and safe in the sense of
Definition~\ref{ch03:def:vs}.
\end{theorem}

\begin{proof}
The path $\tau$ is valid because every transition in $\delta'$ projects onto a
transition in $\longrightarrow_c$. Since $R$ is an accepting run of
$\tilde{\mathcal{A}}_{r}
= \mathcal{T}_c \times \tilde{\mathcal{A}}_{\varphi}$, the trace of $\tau$ is
accepted by $\tilde{\mathcal{A}}_{\varphi}$. By
Theorem~\ref{ch03:theo:saferun},
$\texttt{trace}(\tau) \in
\mathcal{L}_{\omega}(\mathcal{A}^{\textup{hard}})$. Then, by
Lemma~\ref{ch02:lem:LTL-BA},
$\texttt{trace}(\tau) \in \texttt{Words}(\varphi^{\textup{hard}})$, which is
equivalent to $\tau \models \varphi^{\textup{hard}}$. Hence, the resulting plan $\tau$ is safe.
\end{proof}

As in Section~\ref{ch03:sec:infeasible}, accepting runs are evaluated in the
prefix-suffix form of \eqref{ch02:eq:prefix-suffix}. Their total cost is
defined by
\begin{equation}
\begin{aligned}
\texttt{Cost}(R,\, \tilde{\mathcal{A}}_{r})
&\triangleq
\texttt{cost}_{\tau}
+ \alpha \cdot \texttt{dist}_{\varphi^{\textup{soft}}}, \\
\texttt{dist}_{\varphi^{\textup{soft}}}
&\triangleq
\Big( \sum_{i=0}^{f-1} + \gamma \sum_{i=f}^{n-1} \Big)
\texttt{Dist}(L_c(\pi_i),\,
\chi_2(q_i'|_{Q_2},\, q_{i+1}'|_{Q_2})),
\end{aligned}
\label{ch03:eq:prod-cost3}
\end{equation}
where $\tau \triangleq R|_{\Pi}$, $\texttt{cost}_{\tau}$ is given by
\eqref{ch03:eq:cost}, and $q_i'|_{Q_2}$ denotes the projection of the product
state $q_i'$ onto the soft-automaton state set $Q_2$.

\begin{problem}\label{ch03:pro:safe}
Find an accepting run of $\tilde{\mathcal{A}}_{r}$ that minimizes the cost in
\eqref{ch03:eq:prod-cost3}.
\end{problem}

\subsubsection{Safe plan synthesis}\label{ch03:subsec:safe-plan-synthesis}

The solution to Problem~\ref{ch03:pro:safe} is called the \emph{safe}
accepting run of $\tilde{\mathcal{A}}_{r}$ and is denoted by
$R_{\textup{safe}}$. Its induced plan is
$\tau_{\textup{safe}} \triangleq R_{\textup{safe}}|_{\Pi}$.
For fixed values of $\alpha$ and $\gamma$, the automaton
$\tilde{\mathcal{A}}_{r}$ may be constructed explicitly from
Definition~\ref{ch03:def:safe} or generated on the fly, exactly as in the
preceding subsection. The only substantive difference is that the transition
structure enforces the hard specification through
$\tilde{\mathcal{A}}_{\varphi}$, while the weight in
\eqref{ch03:eq:weight-soft} penalizes only deviation from the soft component.
Therefore, Algorithm~\ref{ch02:alg:optimal} can be applied directly to
$\tilde{\mathcal{A}}_{r}$ in order to compute the safe accepting run
$R_{\textup{safe}}$.
By Theorem~\ref{ch03:theo:safety}, the corresponding plan
$\tau_{\textup{safe}}$ is always valid and safe, regardless of the values of
$\alpha$ and $\gamma$. As in Section~\ref{ch03:sec:infeasible}, the parameter
$\alpha$ may be varied to generate several candidate runs, which can then be
compared through the pair
$(\texttt{cost}_{\tau},\, \texttt{dist}_{\varphi^{\textup{soft}}})$. Once the
preferred run has been selected, the associated plan
$\tau_{\textup{safe}}$ can be implemented by the hybrid control strategy of
Algorithm~\ref{ch02:alg:hybrid}.

\begin{lemma}\label{ch03:lem:distanceProofsafe}
If $\textup{\texttt{dist}}_{\varphi^{\textup{soft}}} = 0$, then
$\tau_{\textup{safe}} \models \varphi$.
\end{lemma}

\begin{proof}
Let $\mathcal{A}_{\varphi}$ denote the standard, unrelaxed intersection of
$\mathcal{A}^{\textup{hard}}$ and $\mathcal{A}^{\textup{soft}}$. Since
$\texttt{dist}_{\varphi^{\textup{soft}}} = 0$, every transition along
$R_{\textup{safe}}$ satisfies the soft component exactly. The hard component is
already enforced by the transition relation of
$\tilde{\mathcal{A}}_{\varphi}$. Therefore, $R_{\textup{safe}}$ is also an
accepting run of the standard product automaton
$\mathcal{T}_c \times \mathcal{A}_{\varphi}$. By
Lemma~\ref{ch02:lem:satisfy}, the plan
$\tau_{\textup{safe}}$ satisfies $\varphi$.
\end{proof}














%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{On-line sensing and plan adaptation}
\label{ch03:sec:partial}

In practice, the transition-system model is rarely fully consistent with the
actual workspace or the robot dynamics throughout execution. As a result, a
motion and task plan synthesized off-line may later become invalid, unsafe, or
suboptimal once new information is acquired. This motivates an on-line planning
framework in which execution and planning are interleaved, as illustrated in
Figure~\ref{ch03:fig:online}. The essential requirement is that the robot
should monitor plan execution, update its workspace model from real-time
observations, and revise the current plan whenever necessary while preserving
validity and safety.
Let
\begin{equation}\label{ch03:eq:t-FTS}
\mathcal{T}_c^t\triangleq (\Pi,\, \longrightarrow_c^t,\, \Pi_0,\,AP,\, L_c^t,\, w_c^t),
\end{equation}
denote the transition system available at time $t \geq 0$. The region set
$\Pi$, the initial-state set $\Pi_0$, and the proposition set $AP$ are taken
to be fixed. Only the transition relation, the labeling function, and the
transition weights are updated over time. The task specification is also fixed
during execution and retains the decomposition
$\varphi \triangleq
\varphi^{\textup{soft}} \wedge \varphi^{\textup{hard}}$
introduced in \eqref{ch03:eq:spec}. Let
$\tilde{\mathcal{A}}_{\varphi}$ denote the relaxed intersection automaton from
Definition~\ref{ch03:def:conj}.

\begin{problem}
Assume that the workspace is only partially known. The following questions
arise: (I) how should the robot's sensing information be modeled;
(II) how should the transition system be updated accordingly;
and (III) how can the motion and task plan be kept valid and safe throughout execution?
\end{problem}

\subsubsection{Initial plan synthesis}

At time $t = 0$, the initial motion and task plan is synthesized by the method
of Section~\ref{ch03:sec:soft-hard}, using the initial transition system
$\mathcal{T}_c^0$ and the fixed specification~$\varphi$. For each time
$t \geq 0$, let $R^t$ denote the current safe accepting run,
$\tau^t \triangleq R^t|_{\Pi}$ the corresponding plan, and
$\tilde{\mathcal{A}}_r^t$ the associated safety-ensured relaxed product
automaton. In this subsection, $\tilde{\mathcal{A}}_r^t$ is assumed to be
constructed on the fly, and $R^t$ is obtained by
Algorithm~\ref{ch02:alg:optimal} applied to $\tilde{\mathcal{A}}_r^t$.
Even if the soft component is initially infeasible,
Theorem~\ref{ch03:theo:safety} guarantees that $\tau^0$ is valid and safe with
respect to $\mathcal{T}_c^0$.

\subsubsection{Real-time knowledge update}
\label{ch03:subsec:knowledge-update}

The robot is assumed to acquire new workspace information through sensing during
execution. In the present subsection, only locally observed information is used
for the update. Knowledge obtained from other agents or external sources will
be introduced later in the collaborative knowledge-transfer setting.
Let $\textbf{Sense}^t$ denote the sensing information obtained at time
$t \geq 0$. This information may be generated either when the robot reaches a
region or while it traverses a transition. It is represented by
\begin{equation}
\textbf{Sense}^t \triangleq
\{\, ((\pi,\, S,\, S_{\neg}),\, E,\, E_{\neg}) \,\},
\label{ch03:eq:sense-info}
\end{equation}
where $\pi \in \Pi$ is an observed region, $S \subseteq AP$ is the set of
propositions confirmed to hold at~$\pi$, and
$S_{\neg} \subseteq AP$ is the set of propositions confirmed not to hold
there. Moreover, $(\pi_i,\, \pi_j,\, w) \in E$ means that the transition from
$\pi_i$ to $\pi_j$ must be added to $\longrightarrow_c^t$ with weight~$w$, or
that its weight must be updated to~$w$, while
$(\pi_i,\, \pi_j) \in E_{\neg}$ means that the transition from $\pi_i$ to
$\pi_j$ must be removed. Thus $\textbf{Sense}^t$ encodes the portion of the
workspace revealed at time~$t$.
This sensing mechanism can be modeled, for instance, by a sensing radius
$h > 0$ such that every region intersecting the ball
$\{y \in \mathbb{R}^n \mid \|y - x(t)\| \leq h\}$ is observable, where
$x(t) \in \mathbb{R}^n$ is the robot position at time~$t$. The specific
post-processing used to extract the symbolic information in
\eqref{ch03:eq:sense-info} from raw sensor data depends on the sensing modality
and is not needed here.

\subsubsection{Transition system update}\label{ch03:subsubsec:updateT}

At time~$t$, the robot may receive new information through
$\textbf{Sense}^t$. This information is used to update the current
transition-system model. Let $\mathcal{T}_c^{t^-}$ and
$\mathcal{T}_c^{t^+}$ denote the transition systems immediately before and
after the update at time~$t$, respectively.
If $(\pi,\, S,\, S_{\neg})$ is obtained from \eqref{ch03:eq:sense-info}, then
the labeling function is updated according to
$L_c^{t^+}(\pi) \triangleq
\big(L_c^{t^-}(\pi) \cup S\big) \setminus S_{\neg}$.
Similarly, each element of~$E$ adds a transition or updates its weight, whereas
each element of~$E_{\neg}$ removes a transition. For convenience,
$\widetilde{\Pi} \subseteq \Pi$ denotes the set of regions whose labeling
function changes during the update, and
\begin{equation}
\widehat{\Pi} \triangleq
\widetilde{\Pi} \cup
\{\, \pi_i \mid (\pi_i,\, \pi_j,\, w) \in E
\text{ or } (\pi_i,\, \pi_j) \in E_{\neg} \,\},
\label{ch03:eq:widehat}
\end{equation}
denotes the set of regions whose outgoing adjacency information must be
reconstructed. If $\textbf{Sense}^t$ is empty, then
$\mathcal{T}_c^{t^+} = \mathcal{T}_c^{t^-}$.

\subsubsection{Safety-ensured plan revision}\label{ch03:subsec:realtime}

Because $\mathcal{T}_c^t$ may change during execution, the current motion and
task plan must be re-evaluated with respect to its validity, its safety, and
its optimality. The first step is to determine how the product automaton should
be updated, and then to decide whether the current accepting run can be kept,
patched locally, or must be replaced.

(I) \emph{Product automaton update}.
Let $\tilde{\mathcal{A}}_r^{t^-}$ and $\tilde{\mathcal{A}}_r^{t^+}$ denote the
relaxed product automata associated with $\mathcal{T}_c^{t^-}$ and
$\mathcal{T}_c^{t^+}$, respectively. A straightforward update strategy would be
to reconstruct $\tilde{\mathcal{A}}_r^{t^+}$ entirely from scratch using
Definition~\ref{ch03:def:safe}, or to re-evaluate every transition potentially
affected by the new information, as in~\cite{guo2013revising}. Both approaches
have complexity proportional to the size of the product automaton, even though
most of the updated transitions may never be examined during the subsequent
revision process.
A more economical strategy is to incorporate the update information into the
adjacency routine of $\tilde{\mathcal{A}}_r^{t}$. Whenever a product state
$q_s'$ satisfies $q_s'|_{\Pi} \in \widehat{\Pi}$, its outgoing transitions are
marked for reconstruction. They are then rebuilt only when that state is
actually revisited by the revision procedure. In this way, the product
automaton is updated lazily rather than globally.
The updated transition system $\mathcal{T}_c^{t^+}$ and the current plan
$\tau^t$ then raise two questions. First, does $\tau^t$ remain valid and safe?
Second, if not, how should it be modified so that the revised plan becomes
valid and safe for $\mathcal{T}_c^{t^+}$ and~$\varphi$?

(II) \emph{Verifying validity and safety}.
Recall that the current accepting run $R^t$ contains only finitely many
distinct transitions in its prefix-suffix representation. Let
$\texttt{Edge}(R^t)$ denote the set of transitions appearing in the prefix and
suffix of $R^t$. By examining these transitions, it is possible to determine
whether the current plan remains valid and safe.

\begin{definition}\label{ch03:def:vsTS}
Given the updated transition system $\mathcal{T}_c^{t^+}$, a transition
$(\langle \pi_i,\, q_m \rangle,\,
\langle \pi_j,\, q_n \rangle) \in \textup{\texttt{Edge}}(R^t)$ is called
\emph{invalid} if $(\pi_i,\, \pi_j) \notin \longrightarrow_c^{t^+}$, and
\emph{unsafe} if
$L_c^{t^+}(\pi_i) \notin \chi_1(q_m|_{Q_1},\, q_n|_{Q_1})$.
\end{definition}

Let $\Xi$ and $\aleph$ denote the sets of invalid and unsafe transitions of
$R^t$, respectively. These sets are obtained by scanning the transitions in
$\texttt{Edge}(R^t)$ and checking whether a transition has been removed from
$\mathcal{T}_c^{t^+}$ or whether an updated label violates the hard-automaton
constraint.

\begin{theorem}\label{ch03:theo:validity}
Assume that $R^t$ is an accepting run of $\tilde{\mathcal{A}}_r^{t^-}$ and that
$\mathcal{T}_c^{t^-}$ is updated to $\mathcal{T}_c^{t^+}$. Let $\Xi$ and
$\aleph$ be the invalid and unsafe transitions above. Then the
following statements hold:
(I) $\tau^t$ remains valid if and only if $\Xi = \emptyset$;
(II) $\tau^t$ remains safe if $\Xi = \emptyset$ and $\aleph = \emptyset$.
\end{theorem}

\begin{proof}
Since $R^t$ is an accepting run of $\tilde{\mathcal{A}}_r^{t^-}$,
Theorem~\ref{ch03:theo:safety} implies that $\tau^t$ is valid and safe for
$\mathcal{T}_c^{t^-}$. Statement~(i) follows directly from the definition of
$\Xi$: the plan remains valid precisely when none of its transitions has been
removed from the updated transition relation.
For statement~(ii), assume that $\Xi = \emptyset$ and $\aleph = \emptyset$.
Then every transition of $R^t$ remains valid in $\mathcal{T}_c^{t^+}$ and
still satisfies the hard-automaton constraint under the updated labeling
function. Hence $R^t$ remains an accepting run of
$\tilde{\mathcal{A}}_r^{t^+}$, and Theorem~\ref{ch03:theo:safety} yields that
$\tau^t$ remains safe.
\end{proof}

(III) \emph{Plan revision}:
If either $\Xi$ or $\aleph$ is nonempty, then the current plan $\tau^t$ may no
longer be executable or safe. A revision mechanism is therefore needed. A
naive approach is to call Algorithm~\ref{ch02:alg:optimal} again on
$\mathcal{T}_c^{t^+}$ and the current task automaton, using the robot's
current region $\pi_{\textup{cur}}$ as the new initial region. Let the
resulting plan be $\tau_{\textup{new}}$. Although such a plan is valid and safe
from the current region onward, it need not be compatible with the robot's
actual past trajectory.
Let $\tau_{\textup{past}}$ denote the executed finite trajectory up to the
revision instant, and let $R_{\textup{past}}$ be the corresponding finite
product run recorded by Algorithm~\ref{ch02:alg:hybrid}. The complete
trajectory after replanning is
$\tau_{\textup{comp}} \triangleq
\tau_{\textup{past}} + \tau_{\textup{new}}$.
The key difficulty is that the current physical region
$\pi_{\textup{cur}}$ does not determine a unique current product state in
$\tilde{\mathcal{A}}_r^{t^+}$. Owing to the nondeterminism of the B\"uchi
automaton, several product states may be consistent with the observed past
trajectory. Therefore, the first state of a newly synthesized run need not
match the last product state reached by the executed past, even when both
project to the same workspace region.
This leads to the following problem.

\begin{problem}\label{ch03:prob:concatenate}
Assume that $\Xi$ or $\aleph$ is nonempty. Given the executed past trajectory
$\tau_{\textup{past}}$, find a new plan $\tau_{\textup{new}}$ such that the
complete trajectory
$\tau_{\textup{comp}} = \tau_{\textup{past}} + \tau_{\textup{new}}$
is valid and safe.
\end{problem}

To formalize the compatibility constraint, let
$\bar{\tau}_{\textup{past}} \triangleq
\tau_{\textup{past}}[1:(|\tau_{\textup{past}}|-1)]$, that is, the executed
trajectory without its last state. Because $\tilde{\mathcal{A}}_{\varphi}$ is
nondeterministic, the trace of $\bar{\tau}_{\textup{past}}$ may induce several
finite runs in $\tilde{\mathcal{A}}_{\varphi}$. Denote this set by
$r_{\tau_{\textup{past}}} \triangleq
\{\, r_{\bar{\sigma}}
\mid \bar{\sigma} = \texttt{trace}(\bar{\tau}_{\textup{past}}) \,\}$.
The corresponding set of finite runs in the updated product
automaton is
$R_{\tau_{\textup{past}}} \triangleq
\{\, R \mid R|_{\Pi} = \tau_{\textup{past}},\;
R|_{Q} \in r_{\tau_{\textup{past}}} \,\}$.
Accordingly, the product states consistent with the executed past is
\begin{equation}
Q'_{\tau_{\textup{past}}} \triangleq
\{\, \texttt{last}(R) \mid R \in R_{\tau_{\textup{past}}} \,\},
\label{ch03:eq:set-past}
\end{equation}
where $\texttt{last}(R)$ is the last state of the finite run~$R$.
By construction, $R_{\textup{past}} \in R_{\tau_{\textup{past}}}$ and the
current product state recorded during execution belongs to
$Q'_{\tau_{\textup{past}}}$.
Figure~\ref{ch03:fig:tworevise} illustrates the local revision idea. Rather
than recomputing a new run globally at every update, the algorithm first tries
to repair the current run by inserting a bridging segment around each invalid
or unsafe transition. This idea is close in spirit to the revision scheme
in~\cite{guo2013revising}, but the fallback global search is constrained to
start from the set $Q'_{\tau_{\textup{past}}}$ so that the new run remains
consistent with the already executed trajectory.
Let $R^{t^-}$ and $R^{t^+}$ denote the accepting runs before and after
revision, with corresponding plans $\tau^{t^-}$ and $\tau^{t^+}$.
Moreover, let $R_{\textup{pre}}^{t^-}$ and $R_{\textup{suf}}^{t^-}$ denote the
prefix and suffix of $R^{t^-}$, respectively. Given an invalid or unsafe
transition $(q_s',\, q_{s+1}')$ in $R^{t^-}$,
Algorithm~\ref{ch03:alg:revise} first searches for a bridging segment that
reconnects the run to a later state in the same prefix or suffix. If no such
bridge exists, the algorithm invokes \texttt{OptRun}$(\cdot)$ with the
initial-state set $Q'_{\tau_{\textup{past}}}$, rather than the default product
initial set $Q_0'$. This ensures that any newly synthesized accepting run is
compatible with the executed past.

\begin{algorithm}[t]
\caption{Revise the current plan, \texttt{Revise}()}
\label{ch03:alg:revise}
\DontPrintSemicolon
\KwIn{$\Xi$, $\aleph$, $R^{t^-}$, $\tilde{\mathcal{A}}_{r}^{t^+}$,
      $Q'_{\tau_{\textup{past}}}$}
\KwOut{$R^{t^+}$}

\ForAll{$(q'_s,\, q'_{s+1}) \in (\Xi \cup \aleph)$}{
  \If{$(q'_s,\, q'_{s+1}) \in R^{t^-}_{\textup{pre}}$}{
    $bridge = \texttt{DijksTarget}(
    \tilde{\mathcal{A}}_{r}^{t^+},\, q'_{s-1},\,
    \texttt{tail}(q'_s,\, R^{t^-}_{\textup{pre}}))$\;

    \If{$bridge \neq \emptyset$}{
      $R^{t^-}_{\textup{pre}} =
      \texttt{head}(q'_{s-1},\, R^{t^-}_{\textup{pre}})
      + bridge
      + \texttt{tail}(\texttt{last}(bridge),\, R^{t^-}_{\textup{pre}})$\;
    }
    \Else{
      $R^{t^+} =
      \texttt{OptRun}(\tilde{\mathcal{A}}_{r}^{t^+},
      Q'_{\tau_{\textup{past}}})$\;
      \textbf{return} $R^{t^+}$\;
    }
  }

  \If{$(q'_s,\, q'_{s+1}) \in R^{t^-}_{\textup{suf}}$}{
    Apply the same procedure to $R^{t^-}_{\textup{suf}}$\;
  }
}
\textbf{return}
$R^{t^+} = \langle R^{t^-}_{\textup{pre}},\, R^{t^-}_{\textup{suf}} \rangle$\;
\end{algorithm}

\begin{theorem}\label{ch03:theo:new-safe}
Assume that there exists an accepting run of
$\tilde{\mathcal{A}}_{r}^{t^+}$ whose projection onto $\Pi$ extends the
executed past trajectory $\tau_{\textup{past}}$. Then
Algorithm~\ref{ch03:alg:revise} returns a continuation
$\tau_{\textup{new}}$ such that
$\tau_{\textup{comp}} = \tau_{\textup{past}} + \tau_{\textup{new}}$
is valid and safe.
\end{theorem}

\begin{proof}
If a compatible accepting continuation exists, then there is an accepting run
of $\tilde{\mathcal{A}}_{r}^{t^+}$ whose initial state belongs to
$Q'_{\tau_{\textup{past}}}$. Algorithm~\ref{ch03:alg:revise} first attempts to
repair the current run locally by inserting bridging segments around the
transitions in $\Xi \cup \aleph$. Whenever such a bridge is found, the modified
run remains consistent with the unchanged portion of the original accepting
run.
If no bridge can be found for some transition, the algorithm calls
\texttt{OptRun}$(\cdot)$ with initial-state set
$Q'_{\tau_{\textup{past}}}$. By assumption, a compatible accepting run exists,
so this search succeeds and returns one such run. In either case, the resulting
run is an accepting run of $\tilde{\mathcal{A}}_{r}^{t^+}$ whose projection
extends $\tau_{\textup{past}}$. Theorem~\ref{ch03:theo:safety} then implies
that the complete trajectory $\tau_{\textup{comp}}$ is valid and safe.
\end{proof}

The overall on-line planning architecture is completed later by
Algorithm~\ref{ch03:alg:scheme}, in which the current accepting run $R^t$ is
updated through both the revision procedure of
Algorithm~\ref{ch03:alg:revise} and the optimal search routine of
Algorithm~\ref{ch02:alg:optimal}. Additional implementation details are given
in~\cite{guo2014cooperative}. The next target region supplied to the hybrid
controller is updated accordingly.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%











\subsection{Collaborative knowledge transfer}
\label{ch03:sec:multi-partial}

The previous subsection considered on-line sensing and plan adaptation for a
single robot. In a multi-robot setting, each robot may still maintain its own
local task and synthesize its own local plan, yet information gathered by one
robot may improve the workspace model used by others. When the local tasks are
independent at the specification level, there is no need to construct a
centralized team product automaton. Nevertheless, exchanging task-relevant
workspace knowledge can improve both the quality and the robustness of the
local plans. This subsection develops such a collaborative knowledge-transfer
mechanism.

Consider a team of autonomous robots indexed by
$i \in \mathcal{N} \triangleq \{1,\, 2,\, \cdots,\, N\}$, all operating in the
same partially known workspace~$\mathcal{W}$. The motion of robot~$i$ is
abstracted by the weighted transition system defined as:
\begin{equation}
\mathcal{T}_i^t \triangleq
(\Pi_i,\, \longrightarrow_i^t,\, \Pi_{i,0},\, AP_i,\, L_i^t,\, W_i^t),
\label{ch03:eq:Tk}
\end{equation}
which is defined in the same way as \eqref{ch03:eq:t-FTS}. The superscript
$t \geq 0$ indicates that the model may be updated during execution. Each
robot~$i$ is assigned a local task of the form
$\varphi_i \triangleq
\varphi_i^{\textup{soft}} \wedge \varphi_i^{\textup{hard}}$,
where the hard and soft components have the same interpretation as in
\eqref{ch03:eq:spec}. The local tasks are assumed to be independent in the
sense that $\varphi_i$ depends only on propositions in~$AP_i$.

\subsubsection{Initial synthesis}

At time $t = 0$, each robot~$i \in \mathcal{N}$ synthesizes its own initial
plan by the framework of Section~\ref{ch03:sec:soft-hard}, using
$\mathcal{T}_i^0$ and~$\varphi_i$. Let
$\tilde{\mathcal{A}}_{\varphi_i}$ denote the relaxed intersection automaton
associated with~$\varphi_i$, and let
$\tilde{\mathcal{A}}_{r,i}^t$ denote the corresponding safety-ensured relaxed
product automaton at time~$t$. The resulting safe accepting run and induced
safe plan are denoted by $R_i^t$ and $\tau_i^t \triangleq R_i^t|_{\Pi_i}$,
respectively. Even when the soft component of $\varphi_i$ is initially
infeasible, Theorem~\ref{ch03:theo:safety} guarantees that $\tau_i^0$ is valid
and safe with respect to $\mathcal{T}_i^0$.

\subsubsection{Knowledge transfer protocol}
\label{ch03:subsec:know-transfer}

Knowledge transfer is carried out through the communication network of the robot
team. This network describes how information can flow among the robots. Two
standard communication models are relevant here. Under fixed-topology
communication, each robot~$i$ has a prescribed neighbor set
$\mathcal{N}_i \subseteq \mathcal{N}$ and the effective communication set is
constant, namely $\mathcal{N}_i^t \triangleq \mathcal{N}_i$. Under
range-limited communication, robot~$i$ has communication radius $C_i \geq 0$
and can send messages to robot~$j$ only when
$\|x_j(t) - x_i(t)\| \leq C_i$, where
$x_i(t),\, x_j(t) \in \mathbb{R}^n$ are the robot positions. In that case,
$\mathcal{N}_i^t \triangleq
\{\, j \in \mathcal{N} \mid \|x_j(t) - x_i(t)\| \leq C_i \,\}$.
The notation $\mathcal{N}_i^t$ therefore covers both models.

Robot~$i$ is interested only in propositions that appear in its own local task,
namely the set $\varphi_i|_{AP_i}$. To reduce communication load, a
subscriber-publisher mechanism is adopted. When robot~$j$ communicates with
robot~$i \in \mathcal{N}_j^t$ for the first time, robot~$j$ subscribes to
task-relevant information held by robot~$i$ by sending the request:
\begin{equation}
\textbf{Request}_{j,i}^t \triangleq
(j,\, \varphi_j|_{AP_j},\, i),
\label{ch03:eq:robot-request}
\end{equation}
which records the identity of the subscriber, the relevant proposition set, and
the intended publisher. Each robot maintains a subscriber list consisting of
the request messages it has received. It also records which neighbors have
already been contacted, so that each request is sent at most once.
The local sensing update of robot~$i$ is denoted by $\textbf{Sense}_i^t$ and
has the structure defined in \eqref{ch03:eq:sense-info}. Whenever a tuple
$((\pi,\, S,\, S_{\neg}),\, E,\, E_{\neg}) \in \textbf{Sense}_i^t$ is
obtained, robot~$i$ checks whether the newly observed regional information is
relevant to any subscriber. If robot~$j$ is currently reachable and at least
one proposition in $S \cup S_{\neg}$ belongs to $\varphi_j|_{AP_j}$, then
robot~$i$ sends the reply:
\begin{equation}
\textbf{Reply}_{i,j}^t \triangleq
\{\, (\pi,\, S',\, S_{\neg}') \,\},
\label{ch03:eq:robot-reply}
\end{equation}
where
$S' \triangleq S \cap (\varphi_j|_{AP_j})$ and
$S_{\neg}' \triangleq S_{\neg} \cap (\varphi_j|_{AP_j})$.
Only task-relevant propositions are included in the reply, so every
message carries information that may affect robot~$j$'s local plan.

\begin{algorithm}[t]
\caption{Transfer knowledge to subscribed robots, \texttt{TranKnow}()}
\label{ch03:alg:transferAlgo}
\DontPrintSemicolon
\KwIn{$\textbf{Sense}_i^t$, subscriber list $\mathcal{S}_i^t$}
\KwOut{$\textbf{Reply}_{i,j}^t$ for reachable subscribers $j$}

Initialize $\textbf{Reply}_{i,j}^t = \emptyset$ for all $j \in \mathcal{N}_i^t$\;

\ForAll{$(j,\, \varphi_j|_{AP_j},\, i) \in \mathcal{S}_i^t$}{
  \If{$j \in \mathcal{N}_i^t$}{
    \ForAll{$((\pi,\, S,\, S_{\neg}),\, E,\, E_{\neg})
            \in \textbf{Sense}_i^t$}{
      $S' = S \cap (\varphi_j|_{AP_j})$\;
      $S_{\neg}' = S_{\neg} \cap (\varphi_j|_{AP_j})$\;
      \If{$S' \neq \emptyset$ or $S_{\neg}' \neq \emptyset$}{
        add $(\pi,\, S',\, S_{\neg}')$ to $\textbf{Reply}_{i,j}^t$\;
      }
    }
  }
}
\textbf{return} $\textbf{Reply}_{i,j}^t$ for all reachable subscribers $j$\;
\end{algorithm}

Algorithm~\ref{ch03:alg:transferAlgo} summarizes the protocol. The request
phase is subscription-based, whereas the reply phase is event-driven: only new
sensing information that is relevant to a subscriber's task is transmitted.
This avoids repeated request messages and reduces unnecessary traffic. The same
mechanism can be implemented over either multicast or unicast wireless
communication.

\subsubsection{On-line plan verification and adaptation}
\label{ch03:subsec:multi-revise}

Upon receiving local sensing information $\textbf{Sense}_i^t$ and reply
messages $\textbf{Reply}_{j,i}^t$ from neighboring robots
$j \in \mathcal{N}_i^t$, robot~$i$ updates its transition system
$\mathcal{T}_i^t$ by the same update logic as in
Section~\ref{ch03:sec:partial}. More precisely, local sensing may update both
labels and adjacency information, whereas communicated replies update only the
labeling function, since they provide proposition-level knowledge but do not
certify the traversability or cost of robot~$i$'s local transitions.

The subsequent verification and revision of the current plan follow the same
logic as in Section~\ref{ch03:sec:partial}. In particular,
$\widetilde{\Pi}_i^t$ denotes the set of regions with updated labels,
$\Xi_i^t$ the set of invalid transitions in $R_i^t$, and
$\aleph_i^t$ the set of unsafe transitions. If either $\Xi_i^t$ or
$\aleph_i^t$ is nonempty, Algorithm~\ref{ch03:alg:revise} is called to repair
the current run so that the resulting continuation remains valid and safe.
Local revision preserves validity and safety, but it does not generally recover
the safe accepting run that minimizes the current cost. A separate mechanism is
therefore needed to re-establish global optimality of the local plan from time
to time. This issue is related to the dynamic shortest-path problem, in which
edges may be inserted, deleted, or reweighted during execution
\cite{chan2009shortest,misra2005dynamic}. Recomputing the shortest path from
scratch after every update is generally inefficient, whereas maintaining exact
optimality continuously may be unnecessarily expensive.
To balance these considerations, an event-triggered criterion is introduced.
Let $\Upsilon_i^t$ denote the accumulated number of model changes observed by
robot~$i$ up to time~$t$, with initialization $\Upsilon_i^0 = 0$ and update
rule $\Upsilon_i^{t^+} \triangleq
\Upsilon_i^{t^-} + |\widehat{\Pi}_i^t|$.
Let $T_i$ denote the last time instant at which
Algorithm~\ref{ch02:alg:optimal} was called for robot~$i$. Choose thresholds
$N_i^{\textup{call}} \geq 0$ and $T_i^{\textup{call}} \geq 0$. Whenever either
$\Upsilon_i^t \geq N_i^{\textup{call}}$ or
$t - T_i \geq T_i^{\textup{call}}$ holds,
Algorithm~\ref{ch02:alg:optimal} is called on the updated product automaton
$\tilde{\mathcal{A}}_{r,i}^t$, using the set
$Q'_{i,\tau_{\textup{past}}}$ of product states consistent with the executed
past as the initial states. After that call,
$\Upsilon_i^t$ is reset to zero and $T_i$ is set to the current time.

\subsubsection{Overall structure}\label{ch03:subsec:structure}

The overall collaborative planning architecture is summarized in
Algorithm~\ref{ch03:alg:scheme}. When execution starts, the robot synthesizes
its initial safe accepting run. During execution, it sends request messages to
newly encountered neighbors, receives request and reply messages, transfers
task-relevant sensing information, updates its transition system, and then
decides whether local revision or full optimal synthesis should be performed.
The next goal region changes  whenever the current accepting run is
updated or the current goal region is reached.

\begin{algorithm}[t]
\caption{Cooperative on-line planning for each robot $i \in \mathcal{N}$}
\label{ch03:alg:scheme}
\DontPrintSemicolon
\KwIn{$\mathcal{T}_i^0$, $\tilde{\mathcal{A}}_{\varphi_i}$,
      $\tilde{\mathcal{A}}_{r,i}^{0}$, $x_i(t)$}
\KwOut{$R_i^t$, $\tau_i^t$, $\Upsilon_i^t$, $T_i$}

$R_i^0 = \texttt{OptRun}(\tilde{\mathcal{A}}_{r,i}^{0})$\;
$q'_{i,\textup{cur}} = q'_{i,\textup{next}} = R_{i,[\textup{pre},1]}^0$\;
$\pi_{i,\textup{next}} = q'_{i,\textup{next}}|_{\Pi_i}$\;
$R_{i,\textup{past}} = [\,]$, $\tau_{i,\textup{past}} = [\,]$\;
$\Upsilon_i^0 = 0$, $T_i = 0$\;

\While{\textup{\texttt{True}}}{
  send $\textbf{Request}_{i,g}^t$ to newly encountered neighbors
  $g \in \mathcal{N}_i^t$\;

  read $\textbf{Sense}_i^t$, incoming
  $\textbf{Request}_{j,i}^{t_0}$, and incoming
  $\textbf{Reply}_{h,i}^t$\;

  $\textbf{Reply}_{i,j}^t =
  \texttt{TranKnow}(\textbf{Sense}_i^t,\, \mathcal{S}_i^t)$\;

  send $\textbf{Reply}_{i,j}^t$ to reachable subscribers $j$\;

  $(\mathcal{T}_i^{t^+},\, \widetilde{\Pi}_i^t,\, \widehat{\Pi}_i^t) =
  \texttt{UpdateT}(\mathcal{T}_i^t,\,
  \textbf{Sense}_i^t,\, \textbf{Reply}_{h,i}^t)$\;

  $\tilde{\mathcal{A}}_{r,i}^{t^+} =
  \texttt{AdjProd}(\mathcal{T}_i^{t^+},\, \widehat{\Pi}_i^t,\,
  \tilde{\mathcal{A}}_{\varphi_i})$\;

  $(\aleph_i^t,\, \Xi_i^t) =
  \texttt{ValidRun}(\tilde{\mathcal{A}}_{r,i}^{t^+},\, R_i^t,\,
  \widetilde{\Pi}_i^t)$\;

  $\Upsilon_i^t = \Upsilon_i^t + |\widehat{\Pi}_i^t|$\;

  $Q'_{i,\tau_{\textup{past}}} =
  \texttt{CorProd}(\tilde{\mathcal{A}}_{r,i}^{t^+},\,
  \tau_{i,\textup{past}})$\;

  \If{$\Upsilon_i^t \geq N_i^{\textup{call}}$ or
      $t - T_i \geq T_i^{\textup{call}}$}{
    $R_i^{t^+} =
    \texttt{OptRun}(\tilde{\mathcal{A}}_{r,i}^{t^+},\,
    Q'_{i,\tau_{\textup{past}}})$\;
    $\Upsilon_i^t = 0$, $T_i = t$\;
  }
  \ElseIf{$\aleph_i^t \neq \emptyset$ or $\Xi_i^t \neq \emptyset$}{
    $R_i^{t^+} =
    \texttt{Revise}(\Xi_i^t,\, \aleph_i^t,\, R_i^t,\,
    \tilde{\mathcal{A}}_{r,i}^{t^+},\, Q'_{i,\tau_{\textup{past}}})$\;
  }
  \Else{
    $R_i^{t^+} = R_i^t$\;
  }

  \If{$x_i(t) \in q'_{i,\textup{next}}|_{\Pi_i}$ is confirmed}{
    $q'_{i,\textup{cur}} = q'_{i,\textup{next}}$\;
    $\tau_{i,\textup{past}} =
    \tau_{i,\textup{past}} + q'_{i,\textup{next}}|_{\Pi_i}$\;
    $R_{i,\textup{past}} =
    R_{i,\textup{past}} + q'_{i,\textup{next}}$\;
    $q'_{i,\textup{next}} =
    \texttt{NextGoal}(q'_{i,\textup{cur}},\, R_i^{t^+})$\;
  }

  $R_i^t = R_i^{t^+}$, $\tau_i^t = R_i^t|_{\Pi_i}$\;
  $\pi_{i,\textup{cur}} = q'_{i,\textup{cur}}|_{\Pi_i}$\;
  $\pi_{i,\textup{next}} = q'_{i,\textup{next}}|_{\Pi_i}$\;
  $u_i = \mathcal{U}(x_i(t),\, \pi_{i,\textup{cur}},\,
  \pi_{i,\textup{next}})$\;
}
\end{algorithm}

\begin{theorem}\label{ch03:theo:improvement}
For each robot $i \in \mathcal{N}$ and each time $t \geq 0$, the plan
$\tau_i^t$ produced by Algorithm~\ref{ch03:alg:scheme} is valid and safe with
respect to the current transition system $\mathcal{T}_i^t$ and the local hard
specification $\varphi_i^{\textup{hard}}$. Moreover, for any $t' \geq 0$,
there exists some $t \in [t',\, t' + T_i^{\textup{call}}]$ such that
$\tau_i^t$ is induced by a safe accepting run of
$\tilde{\mathcal{A}}_{r,i}^t$.
\end{theorem}

\begin{proof}
The first part follows from the update and revision mechanisms established in
Section~\ref{ch03:sec:partial}. If no invalid or unsafe transition is detected,
the current plan is retained and remains valid and safe by
Theorem~\ref{ch03:theo:validity}. If invalid or unsafe transitions are
detected, Algorithm~\ref{ch03:alg:revise} returns a compatible continuation,
and Theorem~\ref{ch03:theo:new-safe} guarantees validity and safety of the
resulting complete trajectory. Whenever
Algorithm~\ref{ch02:alg:optimal} is called, it computes a safe accepting run of
$\tilde{\mathcal{A}}_{r,i}^t$ from the compatible initial-state set
$Q'_{i,\tau_{\textup{past}}}$, so the induced plan is again valid and safe.
For the second part, the time-trigger condition
$t - T_i \geq T_i^{\textup{call}}$ guarantees that
Algorithm~\ref{ch02:alg:optimal} is invoked at least once within any interval
of length $T_i^{\textup{call}}$. At each such invocation, the resulting plan is
induced by a safe accepting run of the current product automaton. This proves
the claimed recurrence property.
\end{proof}
