%==============================
\section{Case study}
\label{sec:ch03_case_study}

The bottom-up coordination framework is illustrated next on a
representative multi-robot example involving local temporal tasks,
execution-time dependence, and collaborative actions. The case study is
intended to show how local plan synthesis, request-reply exchange, and
online adaptation interact in a concrete setting.

Consider a team in which each robot is assigned a local temporal task over its
own workspace abstraction, while several task clauses require assistance from
other robots through collaborative actions. Planning begins independently at
each robot, using the single-robot synthesis machinery of
Chapter~\ref{chap:preliminaries}. As execution unfolds, however, two forms of
coupling arise. First, previously unknown workspace information may invalidate a
locally feasible transition or reveal that part of a task cannot be completed
without revision. Second, the realization of a local task may require a support
action, shared resource, or synchronization event supplied by another robot.
The bottom-up framework addresses these difficulties without replacing the local
specifications by a monolithic team formula.

A representative execution therefore proceeds in three stages. The robots first
synthesize local plans and begin execution under the nominal assumptions of
known reachability and available assistance. When a robot detects a blocked
transition, missing proposition, or newly relevant environmental fact, the
local abstraction is updated and the current plan is re-verified against the
task logic. If the original task becomes infeasible, the relaxation and
safety-aware revision mechanisms of
Section~\ref{ch03:sec:task_relaxation_adaptation} produce a minimally modified
plan that preserves the hard part of the specification. When the revised local
plan requires collaboration, the request-reply protocol of
Section~\ref{ch03:sec:dep-local-collab} is activated, so that neighboring
robots confirm, reject, or reschedule the requested support actions according
to their own evolving commitments. In this way, the team maintains correctness
through incremental negotiation rather than through centralized resynthesis.

This integrated view clarifies the role of the bottom-up perspective. Task
adaptation remains local whenever the logical disturbance is local, knowledge is
shared only when it is relevant to feasibility, and execution-time dependence is
resolved through explicit coordination messages tied to concrete collaborative
actions. The resulting behavior is therefore neither fully independent nor
fully centralized. Instead, it is a layered coordination process in which local
temporal tasks remain the primary planning objects, while collaboration and
adaptation are introduced only where execution reveals that they are needed.

%==============================
\section{Summary}
\label{sec:ch03_summary}

This chapter developed a bottom-up view of multi-robot coordination
under local temporal tasks. Section~\ref{ch03:sec:task_relaxation_adaptation}
addressed task relaxation, safety-aware plan revision, online sensing,
and collaborative knowledge transfer in partially known workspaces.
Section~\ref{ch03:sec:dep-local-collab} then introduced dependent local
tasks with collaborative actions and a distributed execution-time
coordination mechanism for request, reply, assignment, and adaptation.
Taken together, these developments show how correctness can be
maintained without a monolithic global mission, by coordinating locally
specified tasks whose interactions arise during execution.
