\section{Dependent Local Tasks with Collaborative Actions}
\label{ch03:sec:dep-local-collab}

Robots are deployed to move around within the workspace and more importantly to perform various actions to interact with it.
Thus in this chapter, we first introduce a new method to construct the model of both robot motion and actions.
Based on this model, we can specify a local temporal task as the desired robot motion and actions.
Then a planning scheme for both robot motion and actions is proposed as an extension to the nominal solution presented in Chapter~\ref{chap:preliminaries}.
Furthermore, inter-robot collaborations are essential for many multi-robot applications to improve system capability, efficiency and robustness.
Thus we start from defining a dependency relation among the robots using collaborative and assisting actions. Then a distributed and on-line coordination scheme is proposed where each robot sends requests to its neighbors for collaborative tasks, confirms others' requests   and adapts its local plan to fulfill the confirmed collaboration. All decisions are made locally
by each robot based on local computation and communication between neighboring robots. This scheme is scalable and resilient to robot failures as the dependency relation is formed and removed dynamically according to the plan execution status and robot capabilities, instead of pre-assigned robot identities.
Numerical simulations are provided in the end to validate the proposed scheme.





%=============================================================================================

\subsection{Motion and action planning}
\label{ch03:sec:action-plan}

The task specifications considered so far describe only sequences of regions to
be visited. In many robotic applications, however, reaching a region serves the
purpose of enabling one or more operations at that location. The planning
problem therefore involves both motion and action. These two aspects should not
be treated independently, because action feasibility depends on the current
region and its properties, while executed actions may modify internal robot
states that affect later decisions.
A convenient modeling principle is to separate workspace-dependent information
from workspace independent action knowledge. The former concerns the workspace
structure and the robot's mobility. The latter concerns the conditions under
which actions are enabled and the resulting state updates. This separation
improves modularity and facilitates reuse when the workspace or the task
specification changes \cite{van2008handbook}.

(I) \emph{Mobility abstraction}.
The robot mobility is modeled by the weighted finite transition system below:
\begin{equation}
\label{ch03:eq:M}
\mathcal{M} \triangleq
(\Pi_{\mathcal{M}}, \Sigma_{\mathcal{M}}, \longrightarrow_{\mathcal{M}},
 \Pi_{\mathcal{M},0}, \Psi_{\mathcal{M}}, L_{\mathcal{M}},
 W_{\mathcal{M}}),
\end{equation}
where $\Sigma_{\mathcal{M}} \triangleq \{\sigma_{\mathcal{M}}\}$.
The structure in \eqref{ch03:eq:M} follows
Definition~\ref{ch02:def:FTS}. The set $\Pi_{\mathcal{M}}$ contains the
abstract workspace regions, and $\Pi_{\mathcal{M},0}$ denotes the initial
region set. The proposition set is
$\Psi_{\mathcal{M}} \triangleq \Psi_r \cup \Psi_p$, where
$\Psi_r \triangleq \{\Psi_{r,1}, \cdots, \Psi_{r,N}\}$ collects region names,
and $\Psi_p \triangleq \{\Psi_{p,1}, \cdots, \Psi_{p,I}\}$ collects workspace
properties relevant to task execution. For instance, the proposition
``the room contains product \textsf{A}'' is relevant to the action
``pickup \textsf{A}''. The action $\sigma_{\mathcal{M}}$ is the symbolic
counterpart of the continuous navigation controller in
\eqref{ch02:eq:control}.

(II) \emph{Action model}.
Action description languages provide a natural way to encode the preconditions
and effects of task-level actions \cite{gelfond1998action}. Classical
formalisms such as STRIPS and ADL offer concise models for such action
descriptions \cite{fikes1972strips,pednault1989adl}. Suppose that the robot can
perform $K$ actions
$\{\sigma_{\mathcal{B},1}, \cdots, \sigma_{\mathcal{B},K}\}$ through the
low-level controllers $\{\mathcal{K}_1, \cdots, \mathcal{K}_K\}$. Define
$\Sigma_{\mathcal{B}} \triangleq
\{\sigma_{\mathcal{B},0}, \sigma_{\mathcal{B},1}, \cdots,
  \sigma_{\mathcal{B},K}\}$,
where $\sigma_{\mathcal{B},0} \triangleq \texttt{None}$ indicates that no
task-level action is currently executed.
Two additional proposition sets are introduced. The set
$\Psi_s \triangleq \{\Psi_{s,1}, \cdots, \Psi_{s,J}\}$ represents internal
robot states, such as ``the robot holds product \textsf{A}''. The set
$\Psi_b \triangleq \{\Psi_{b,0}, \Psi_{b,1}, \cdots, \Psi_{b,K}\}$ records the
currently executed action. More precisely, $\Psi_{b,k} = \top$ means that
$\sigma_{\mathcal{B},k}$ is active. Only one proposition in $\Psi_b$ can be
true at any time, reflecting the non-concurrency assumption.
The enabledness of actions is characterized by the precondition function:
\begin{equation}
\label{ch03:eq:cond}
\texttt{Cond} :
\Sigma_{\mathcal{B}} \times 2^{\Psi_p} \times 2^{\Psi_s}
\longrightarrow \{\top,\bot\},
\end{equation}
where the value of \eqref{ch03:eq:cond} indicates whether a given action is
admissible under the current workspace properties and internal robot state.
For example, ``pickup \textsf{A}'' may be enabled only when the current region
contains product \textsf{A}. By definition,
$\texttt{Cond}(\sigma_{\mathcal{B},0}, w_p, w_s)=\top$ for all
$w_p \subseteq \Psi_p$ and $w_s \subseteq \Psi_s$.
The state update induced by an action is captured by the effect function:
\begin{equation}
\label{ch03:eq:eff}
\texttt{Eff} :
\Sigma_{\mathcal{B}} \times 2^{\Psi_s} \times \Psi_b
\longrightarrow 2^{\Psi_s} \times \Psi_b,
\end{equation}
where \eqref{ch03:eq:eff} modifies the internal state valuation and the active
action indicator. In particular,
$\texttt{Eff}(\sigma_{\mathcal{B},0}, w_s, \Psi_{b,\ell})
= (w_s, \Psi_{b,0})$ for every $\Psi_{b,\ell} \in \Psi_b$, so that the special
action $\sigma_{\mathcal{B},0}$ preserves the internal state and resets the
action status to \texttt{None}. For $k \geq 1$, the expression
$\texttt{Eff}(\sigma_{\mathcal{B},k}, w_s, \Psi_{b,\ell})
= (w_s', \Psi_{b,k})$ specifies the updated internal state $w_s'$ resulting
from the execution of $\sigma_{\mathcal{B},k}$. The effect function does not
modify workspace properties, which are treated as external inputs.

\begin{definition}
Given $\Psi_p$, $\Psi_s$, $\Psi_b$, $\Sigma_{\mathcal{B}}$,
$\texttt{Cond}$ from \eqref{ch03:eq:cond}, and $\texttt{Eff}$ from
\eqref{ch03:eq:eff}, define the action map by:
\begin{equation}
\label{ch03:eq:B}
\mathcal{B} \triangleq
(\Pi_{\mathcal{B}}, \Sigma_{\mathcal{B}}, \Psi_p,
 \hookrightarrow_{\mathcal{B}}, \Pi_{\mathcal{B},0}, \Psi_{\mathcal{B}},
 L_{\mathcal{B}}, W_{\mathcal{B}}),
\end{equation}
where the individual components are specified below.
The state set is
$\Pi_{\mathcal{B}} \triangleq 2^{\Psi_s} \times \Psi_b$, and the initial state
set satisfies
$\Pi_{\mathcal{B},0} \subseteq 2^{\Psi_s} \times \{\Psi_{b,0}\}$. Moreover,
$\Psi_{\mathcal{B}} \triangleq \Psi_s \cup \Psi_b$. The propositions in
$\Psi_p$ play the role of external inputs, so the transition relation is
conditional. For
$\pi_{\mathcal{B}} = (w_s,\Psi_{b,\ell})$ and
$\pi_{\mathcal{B}}' = (w_s',\Psi_{b,k})$, the tuple
$(\pi_{\mathcal{B}}, \alpha_{\mathcal{B}}, w_p, \pi_{\mathcal{B}}')$ belongs
to $\hookrightarrow_{\mathcal{B}}$ if and only if
$\alpha_{\mathcal{B}} \in \Sigma_{\mathcal{B}}$,
$\texttt{Cond}(\alpha_{\mathcal{B}}, w_p, w_s)=\top$, and
$\texttt{Eff}(\alpha_{\mathcal{B}}, w_s, \Psi_{b,\ell})
= (w_s', \Psi_{b,k})$. The labeling function is given by
$L_{\mathcal{B}}(w_s,\Psi_{b,k}) \triangleq w_s \cup \{\Psi_{b,k}\}$, and
$W_{\mathcal{B}}$ assigns a positive weight to each enabled transition
according to the cost of the corresponding action.
\end{definition}


Note that the state space of $\mathcal{B}$ is defined as
$2^{\Psi_s} \times \Psi_b$, rather than $2^{\Psi_s} \times 2^{\Psi_b}$,
because at most one proposition in $\Psi_b$ can be true at any time.
The action map is nondeterministic in the following sense. At a given state,
any action whose precondition evaluates to true may be selected. Once a
specific valuation of the workspace properties is fixed, all conditional
transitions in $\hookrightarrow_{\mathcal{B}}$ become either enabled or
disabled, and the resulting structure is a standard weighted finite transition
system \cite{baier2008principles}. An important feature is that
$\mathcal{B}$ is defined independently of the workspace geometry and can
therefore be reused across different environments.

(III) \emph{Complete functionality model}.
The mobility abstraction in \eqref{ch03:eq:M} and the action map in
\eqref{ch03:eq:B} describe complementary aspects of robot behavior. Richer task
specifications involving both navigation and action execution require a unified
model.

\begin{definition}
The complete robot functionality is defined by the composition:
\begin{equation}
\label{ch03:eq:R}
\mathcal{R} \triangleq
(\Pi_{\mathcal{R}}, \Sigma_{\mathcal{R}}, \longrightarrow_{\mathcal{R}},
 \Pi_{\mathcal{R},0}, \Psi_{\mathcal{R}}, L_{\mathcal{R}},
 W_{\mathcal{R}}),
\end{equation}
with the components specified as follows.
The state set is
$\Pi_{\mathcal{R}} \triangleq \Pi_{\mathcal{M}} \times \Pi_{\mathcal{B}}$, and
the action set is
$\Sigma_{\mathcal{R}} \triangleq
\Sigma_{\mathcal{M}} \cup \Sigma_{\mathcal{B}}$.
The initial state set is
$\Pi_{\mathcal{R},0} \triangleq
\Pi_{\mathcal{M},0} \times \Pi_{\mathcal{B},0}$, and the proposition set is
$\Psi_{\mathcal{R}} \triangleq
\Psi_{\mathcal{M}} \cup \Psi_{\mathcal{B}}$.
The transition relation in \eqref{ch03:eq:R} has two types of transitions.
First, for a mobility action $\sigma_{\mathcal{M}}$, one has
$\langle \pi_{\mathcal{M}}, \pi_{\mathcal{B}} \rangle
\xrightarrow{\sigma_{\mathcal{M}}}_{\mathcal{R}}
\langle \pi_{\mathcal{M}}', \pi_{\mathcal{B}}' \rangle$
if
$\pi_{\mathcal{M}} \xrightarrow{\sigma_{\mathcal{M}}}_{\mathcal{M}}
\pi_{\mathcal{M}}'$
and
$\pi_{\mathcal{B}}' =
\texttt{Eff}(\sigma_{\mathcal{B},0}, w_s, \Psi_{b,\ell})$ for
$\pi_{\mathcal{B}} = (w_s,\Psi_{b,\ell})$. Second, for a task action
$\alpha_{\mathcal{B}} \in \Sigma_{\mathcal{B}}$, one has
$\langle \pi_{\mathcal{M}}, \pi_{\mathcal{B}} \rangle
\xrightarrow{\alpha_{\mathcal{B}}}_{\mathcal{R}}
\langle \pi_{\mathcal{M}}, \pi_{\mathcal{B}}' \rangle$
if
$(\pi_{\mathcal{B}}, \alpha_{\mathcal{B}},
  L_{\mathcal{M}}(\pi_{\mathcal{M}}) \cap \Psi_p,
  \pi_{\mathcal{B}}') \in \hookrightarrow_{\mathcal{B}}$.
Thus, a mobility transition changes the region, whereas a task transition keeps
the region fixed and updates the internal action state.
The labeling function is defined by
$L_{\mathcal{R}}(\langle \pi_{\mathcal{M}}, \pi_{\mathcal{B}} \rangle)
\triangleq
L_{\mathcal{M}}(\pi_{\mathcal{M}}) \cup L_{\mathcal{B}}(\pi_{\mathcal{B}})$.
The weight function combines the weights inherited from the two component
models. Mobility transitions use the weight assigned by $W_{\mathcal{M}}$,
while task transitions use the corresponding weight from $W_{\mathcal{B}}$.
\end{definition}


Whenever a mobility transition is executed, the auxiliary action
$\sigma_{\mathcal{B},0}$ is applied automatically in the action component.
This represents the fact that no task-level action is active during motion and
is consistent with the non-concurrency assumption.
Figure~\ref{ch03:fig:composed} illustrates the composition underlying
\eqref{ch03:eq:R}. Informally, each region of the mobility model is paired with
a copy of the action map. The workspace properties associated with that region
then determine which conditional action transitions are enabled.

(IV) \emph{Local task over motion and action}.
The composed model $\mathcal{R}$ is a weighted finite transition system over
$\Psi_{\mathcal{R}} = \Psi_r \cup \Psi_p \cup \Psi_s \cup \Psi_b$. The sets
$\Psi_r$ and $\Psi_p$ describe the robot location and workspace properties.
The additional sets $\Psi_s$ and $\Psi_b$ allow task formulas to refer directly
to internal robot states and executed actions. This supports richer
specifications concerning both action ordering and the locations where actions
should occur \cite{bhatia2010sampling,kloetzer2010automatic,smith2011optimal}.
For example, the requirement that the robot should execute
``drop \textsf{A}'' at region \textsf{1} infinitely often can be expressed as
$\varphi = \square \Diamond (\Psi_{r,1} \wedge \Psi_{b,2})$. Here
$\Psi_{r,1}$ denotes ``the robot is in region \textsf{1}'', and
$\Psi_{b,2}$ denotes ``the action \texttt{drop \textsf{A}} is executed''.
The location where \textsf{A} is picked up need not be specified explicitly,
because this dependency is already enforced by the action preconditions.

(V) \emph{Planning for hybrid control strategy}.
The planning problem addressed in this subsection is stated as follows.

\begin{problem}
\label{ch03:pro:full}
Given the complete functionality model $\mathcal{R}$ and a task formula
$\varphi$, determine a motion-and-action plan that minimizes the cost in
\eqref{ch02:eq:prod-cost}, and construct a hybrid control strategy that
executes this plan.
\end{problem}

Since $\mathcal{R}$ in \eqref{ch03:eq:R} is a standard weighted finite
transition system, the synthesis procedure developed in
Chapter~\ref{chap:preliminaries} applies directly. Let $R_{\text{opt}}$ denote
the optimal accepting run of the product automaton
$\mathcal{R} \otimes \mathcal{A}_{\varphi}$ computed by
Algorithm~\ref{ch02:alg:optimal}. Its projection onto $\Pi_{\mathcal{R}}$
yields the optimal motion-and-action plan
$\tau_{\text{opt}} \triangleq R_{\text{opt}}|_{\Pi_{\mathcal{R}}}$.
For every consecutive pair of states
$(\pi_{\mathcal{R},s}, \pi_{\mathcal{R},s+1})$ in $\tau_{\text{opt}}$, there
exists an action $\alpha_{\mathcal{R}} \in \Sigma_{\mathcal{R}}$ such that
$(\pi_{\mathcal{R},s}, \alpha_{\mathcal{R}}, \pi_{\mathcal{R},s+1})
\in \longrightarrow_{\mathcal{R}}$. The hybrid control strategy is then
obtained by sequentially executing the corresponding low-level controller along
$\tau_{\text{opt}}$, in the same spirit as
Algorithm~\ref{ch02:alg:hybrid}. If
$\alpha_{\mathcal{R}} = \sigma_{\mathcal{B},k}$, the controller
$\mathcal{K}_k$ associated with the action $\sigma_{\mathcal{B},k}$ is
activated. If
$\alpha_{\mathcal{R}} = \sigma_{\mathcal{M}}$, the navigation controller in
\eqref{ch02:eq:control} is used to drive the robot from the current region to
the next one.
The resulting framework has four stages. First, construct the mobility model
$\mathcal{M}$ and the action map $\mathcal{B}$. Second, compose them into the
complete functionality model $\mathcal{R}$. Third, synthesize the optimal run
and project it to $\tau_{\text{opt}}$. Fourth, execute the plan by the hybrid
control strategy. The modularity of this structure is useful in practice. Once
$\mathcal{B}$ has been defined for a given robot, it can be reused across
different workspaces. When the workspace changes, only the mobility abstraction
$\mathcal{M}$ needs to be updated. When the task specification changes,
$\mathcal{R}$ remains unchanged and only the synthesis step must be repeated.






















%==============================
\subsection{Multi-robot systems with dependent local tasks}
\label{ch03:sec:collaborative}

The previous subsection considered the case in which each robot can fulfill its
local task independently. A more involved setting arises when the completion of
one robot's task requires assistance from other robots. This subsection treats
that case for a team of $N$ heterogeneous robots indexed by
$i \in \mathcal{N} \triangleq \{1,2,\cdots,N\}$, as introduced in
Section~\ref{ch03:sec:multi-partial}.

(I) \emph{Mobility abstraction}.
The workspace is partitioned into $M$ regions of interest,
$\Pi \triangleq \{\pi_1,\pi_2,\cdots,\pi_M\}$. These symbols are assumed to be
fixed a priori and known to all robots. The motion of robot $i$ is modeled by
the weighted finite transition system:
\begin{equation}
\label{ch03:M}
\mathcal{M}_i \triangleq
(\Pi, \longrightarrow_{\mathcal{M}}^i, \Pi_{\mathcal{M},0}^i,
 \Psi_{\mathcal{M}}^i, L_{\mathcal{M}}^i, T_{\mathcal{M}}^i),
\end{equation}
where $\longrightarrow_{\mathcal{M}}^i \subseteq \Pi \times \Pi$ is the motion
transition relation, $\Pi_{\mathcal{M},0}^i \subseteq \Pi$ is the initial
region set, $L_{\mathcal{M}}^i : \Pi \rightarrow 2^{\Psi_{\mathcal{M}}^i}$ is
the labeling function, and
$T_{\mathcal{M}}^i : \longrightarrow_{\mathcal{M}}^i \rightarrow \mathbb{R}^+$
assigns a positive duration to each transition. A finite path of
$\mathcal{M}_i$ is a sequence $\pi_0 \pi_1 \cdots \pi_P$ such that
$(\pi_n,\pi_{n+1}) \in \longrightarrow_{\mathcal{M}}^i$ for all
$n = 0,1,\cdots,P-1$. Because of heterogeneity, the models $\mathcal{M}_i$ may
differ across the robots.
Each robot $i$ is also associated with a set of neighboring robots
$\mathcal{N}_i \subseteq \mathcal{N}$. Messages can be exchanged directly
between robot $i$ and any robot $j \in \mathcal{N}_i$.

(II) \emph{Action model}.
In addition to motion, robot $i$ can execute a set of actions
$\Sigma_i \triangleq \Sigma_l^i \cup \Sigma_c^i \cup \Sigma_h^i$, where
$\Sigma_l^i$ contains local actions that robot $i$ can complete on its own,
$\Sigma_c^i$ contains collaborative actions initiated by robot $i$ but
requiring assistance, and $\Sigma_h^i$ contains assisting actions that
robot $i$ can provide to other robots. The active actions that can be initiated
by robot $i$ are
$\Sigma_a^i \triangleq \Sigma_l^i \cup \Sigma_c^i$. By convention,
$\sigma_0 \triangleq \texttt{None} \in \Sigma_l^i$ denotes the absence of an
active task-level action.
Let $\Sigma_h^{\sim i}$ denote the set of external assisting actions on which
robot $i$ may depend. This set satisfies
$\Sigma_h^{\sim i} \subseteq \bigcup_{j \in \mathcal{N}_i}\Sigma_h^j$.
The action model of robot $i$ is given by:
\begin{equation}
\label{ch03:A}
\mathcal{B}_i \triangleq
(\Sigma_i, \Psi_{\Sigma}^i, L_{\Sigma}^i, \texttt{Cond}_i,
 \texttt{Dura}_i, \texttt{Depd}_i),
\end{equation}
where $\Psi_{\Sigma}^i$ is the set of atomic propositions associated with the
active actions of robot $i$, and
$L_{\Sigma}^i : \Sigma_i \rightarrow 2^{\Psi_{\Sigma}^i}$ is the action
labeling function. For assisting actions,
$L_{\Sigma}^i(\sigma_h)=\emptyset$ for all $\sigma_h \in \Sigma_h^i$, while
$L_{\Sigma}^i(\sigma_a)\subseteq \Psi_{\Sigma}^i$ for all
$\sigma_a \in \Sigma_a^i$.
The function
$\texttt{Cond}_i : \Sigma_i \times 2^{\Psi_{\mathcal{M}}^i}
\rightarrow \{\top,\bot\}$ specifies the region properties required for an
action to be enabled. The function
$\texttt{Dura}_i : \Sigma_i \rightarrow \mathbb{R}^+$ assigns an execution
duration to each action, with $\texttt{Dura}_i(\sigma_0)=T_0>0$ as a design
parameter. The dependence function
$\texttt{Depd}_i : \Sigma_i \rightarrow 2^{\Sigma_h^{\sim i}}$
describes the assisting actions required by each action of robot $i$. It
satisfies $\texttt{Depd}_i(\sigma_s)=\emptyset$ for all
$\sigma_s \in \Sigma_l^i \cup \Sigma_h^i$, while
$\texttt{Depd}_i(\sigma_c)\subseteq \Sigma_h^{\sim i}$ for all
$\sigma_c \in \Sigma_c^i$. Defining dependency at the level of actions, rather
than directly at the level of robot identities, preserves flexibility because
the assisting robot need not be fixed in advance.

\begin{definition}
\label{ch03:done}
Let $\pi_v \in \Pi$. A local or assisting action
$\sigma_s \in \Sigma_l^i \cup \Sigma_h^i$ is said to be done at region $\pi_v$
if
$\texttt{Cond}_i(\sigma_s,L_{\mathcal{M}}^i(\pi_v))=\top$ and $\sigma_s$ is
activated for the duration $\texttt{Dura}_i(\sigma_s)$. A collaborative action
$\sigma_c \in \Sigma_c^i$ is said to be done at region $\pi_v$ if the same two
conditions hold and, in addition, every assisting action in
$\texttt{Depd}_i(\sigma_c)$ is done by other robots at the same region $\pi_v$.
\end{definition}

(III) \emph{Complete robot model}.
The complete model of robot $i$ combines its motion abstraction and its action
model into a single finite transition system.

\begin{definition}
\label{ch03:full-model}
Given $\mathcal{M}_i$ and $\mathcal{B}_i$, define the complete robot model by
\begin{equation}
\label{ch03:eq:Ri}
\mathcal{R}_i \triangleq
(\Pi_{\mathcal{R}}^i, \longrightarrow_{\mathcal{R}}^i,
 \Pi_{\mathcal{R},0}^i, \Psi_{\mathcal{R}}^i,
 L_{\mathcal{R}}^i, T_{\mathcal{R}}^i),
\end{equation}
where the components are specified below.
The state set is
$\Pi_{\mathcal{R}}^i \triangleq \Pi \times \Sigma_i$, and each state is written
as $\pi_{\mathcal{R},s}=\langle \pi_t,\sigma_n \rangle$ with
$\pi_t \in \Pi$ and $\sigma_n \in \Sigma_i$. The initial state set is
$\Pi_{\mathcal{R},0}^i \triangleq
\{\langle \pi_{\mathcal{M},0}^i,\sigma_0 \rangle\}$, where
$\Pi_{\mathcal{M},0}^i=\{\pi_{\mathcal{M},0}^i\}$. The proposition set is
$\Psi_{\mathcal{R}}^i \triangleq \Psi_{\mathcal{M}}^i \cup \Psi_{\Sigma}^i$.
The transition relation
$\longrightarrow_{\mathcal{R}}^i \subseteq
\Pi_{\mathcal{R}}^i \times \Pi_{\mathcal{R}}^i$
is defined as follows. A pair
$\big(\langle \pi_s,\sigma_m \rangle,\langle \pi_t,\sigma_n \rangle\big)$
belongs to $\longrightarrow_{\mathcal{R}}^i$ if one of the following cases
holds. First, $\sigma_m=\sigma_n=\sigma_0$ and
$\pi_s \longrightarrow_{\mathcal{M}}^i \pi_t$, which corresponds to a motion
transition. Second, $\sigma_m=\sigma_0$, $\sigma_n \neq \sigma_0$,
$\pi_s=\pi_t$, and
$\texttt{Cond}_i(\sigma_n,L_{\mathcal{M}}^i(\pi_s))=\top$, which corresponds to
the start of an enabled action. Third,
$\sigma_n=\sigma_0$ and
$\sigma_m \in \Sigma_i \setminus \{\sigma_0\}$ with $\pi_s=\pi_t$, which
is the completion of the action.
The labeling function
$L_{\mathcal{R}}^i : \Pi_{\mathcal{R}}^i \rightarrow 2^{\Psi_{\mathcal{R}}^i}$
is defined by
$L_{\mathcal{R}}^i(\langle \pi_s,\sigma_m \rangle) \triangleq
L_{\mathcal{M}}^i(\pi_s) \cup L_{\Sigma}^i(\sigma_m)$. The transition duration
$T_{\mathcal{R}}^i : \longrightarrow_{\mathcal{R}}^i \rightarrow \mathbb{R}^+$
is given by the corresponding motion or action duration. Motion transitions
inherit the value of $T_{\mathcal{M}}^i$, action-start transitions use
$\texttt{Dura}_i(\sigma_n)$, and action-completion transitions use the fixed
duration $T_0$.
\end{definition}

Note that the model $\mathcal{R}_i$ is therefore a standard finite
transition system \cite{baier2008principles}.
A finite path of $\mathcal{R}_i$ is written as
$\tau_i=\pi_{\mathcal{R},0}\pi_{\mathcal{R},1}\cdots\pi_{\mathcal{R},P}$,
where $\pi_{\mathcal{R},0}\in \Pi_{\mathcal{R},0}^i$ and
$(\pi_{\mathcal{R},s},\pi_{\mathcal{R},s+1})
\in \longrightarrow_{\mathcal{R}}^i$ for all $s=0,\cdots,P-1$. Its trace is
$\texttt{trace}(\tau_i)\triangleq
L_{\mathcal{R}}^i(\pi_{\mathcal{R},0})
L_{\mathcal{R}}^i(\pi_{\mathcal{R},1}) \cdots
L_{\mathcal{R}}^i(\pi_{\mathcal{R},P})$.
The local task of robot $i$, denoted by $\varphi_i$, is an sc-safe LTL formula
over the proposition set $\Psi_{\mathcal{R}}^i$ from
Definition~\ref{ch03:full-model}. Hence $\varphi_i$ can impose requirements on
motion, local actions, and collaborative actions. The syntax and semantics of
sc-LTL were introduced in Section~\ref{ch02:sec:LTL}. Since an sc-safe formula
admits satisfaction by a finite prefix, a finite path $\tau_i$ of
$\mathcal{R}_i$ fulfills $\varphi_i$ if
$\texttt{trace}(\tau_i)\models \varphi_i$. The special case
$\varphi_i \triangleq \top$ corresponds to a robot without an individual task,
which then acts purely as an assisting robot.

\begin{problem}
\label{ch03:big-problem}
Given $\mathcal{R}_i$ and the locally assigned task $\varphi_i$ of each robot
$i \in \mathcal{N}$, design a distributed control and coordination scheme such
that all tasks $\varphi_i$ are fulfilled, even when some local tasks require
collaboration from other robots.
\end{problem}

\subsection{Distributed task coordination}
\label{ch03:sec:coordination}

This subsection develops a bottom-up coordination scheme for multi-robot systems
with dependent local tasks. The scheme consists of off-line initial plan
synthesis, an on-line request-and-reply protocol, real-time plan adaptation,
and a recovery mechanism for robot failures.

(I) \emph{Initial plan synthesis}.
Given the complete model $\mathcal{R}_i$ and the local task $\varphi_i$, an
initial motion-and-action plan can first be synthesized off-line. This plan
serves as the starting point for the on-line coordination mechanism described
later.
Let
$\mathcal{A}_{\varphi_i} \triangleq
(Q_i, 2^{\Psi_{\mathcal{R}}^i}, \delta_i, Q_{i,0}, \mathcal{F}_i)$
be the NBA associated with $\varphi_i$, as in
Section~\ref{ch02:sec:LTL}. The product automaton between $\mathcal{R}_i$ and
$\mathcal{A}_{\varphi_i}$ is given by:
\begin{equation}
\label{ch03:Ap}
\mathcal{A}_{p,i} \triangleq
\mathcal{R}_i \otimes \mathcal{A}_{\varphi_i}
= (Q_{p,i}, \delta_{p,i}, Q_{p,i,0}, \mathcal{F}_{p,i}, W_{p,i}),
\end{equation}
where
$Q_{p,i} \triangleq \Pi_{\mathcal{R}}^i \times Q_i$,
$Q_{p,i,0} \triangleq \Pi_{\mathcal{R},0}^i \times Q_{i,0}$, and
$\mathcal{F}_{p,i} \triangleq \Pi_{\mathcal{R}}^i \times \mathcal{F}_i$.
Moreover,
$\big(\langle \pi_{\mathcal{R},s},q_m \rangle,
\langle \pi_{\mathcal{R},k},q_n \rangle\big) \in \delta_{p,i}$
if
$\pi_{\mathcal{R},s} \longrightarrow_{\mathcal{R}}^i \pi_{\mathcal{R},k}$ and
$(q_m,L_{\mathcal{R}}^i(\pi_{\mathcal{R},s}),q_n) \in \delta_i$. The weight
function $W_{p,i} : \delta_{p,i} \rightarrow \mathbb{R}^+$ is defined by:
\begin{equation}
\label{ch03:eq:wp}
W_{p,i}\big(
\langle \pi_{\mathcal{R},s},q_m \rangle,
\langle \pi_{\mathcal{R},k},q_n \rangle
\big)
\triangleq
T_{\mathcal{R}}^i(\pi_{\mathcal{R},s},\pi_{\mathcal{R},k}),
\end{equation}
where the right-hand side uses the transition duration in $\mathcal{R}_i$.
There exists a finite path of $\mathcal{R}_i$ satisfying $\varphi_i$ if and
only if $\mathcal{A}_{p,i}$ contains a finite path from an initial state to an
accepting state. Let
${R}_{p}^i = q_{p,0}^i q_{p,1}^i \cdots q_{p,P}^i$
be such a finite path, where
$q_{p,0}^i \in Q_{p,i,0}$,
$q_{p,P}^i \in \mathcal{F}_{p,i}$, and
$(q_{p,\ell}^i,q_{p,\ell+1}^i) \in \delta_{p,i}$ for all
$\ell=0,1,\cdots,P-1$. Its cost is defined by
\begin{equation}
\label{ch03:eq:init-cost}
\texttt{Cost}(R_p^i,\mathcal{A}_{p,i}) \triangleq
\sum_{\ell=0}^{P-1} W_{p,i}(q_{p,\ell}^i,q_{p,\ell+1}^i),
\end{equation}
which is the total weight accumulated along $R_p^i$. The notation
$R_p^i[\ell]$ refers to the $\ell$th element of the path, and
$R_p^i[\ell\!:\!k]$ denotes the segment from the $\ell$th element to the $k$th
element.

\begin{problem}
\label{ch03:prob-accept}
Find a finite path $R_p^i$ of $\mathcal{A}_{p,i}$ that starts from an initial
state, ends at an accepting state, and minimizes the cost in
\eqref{ch03:eq:init-cost}.
\end{problem}

Denote by $R_{p,\textup{init}}^i$ the solution of
Problem~\ref{ch03:prob-accept}. Algorithm~\ref{ch02:alg:optimal} yields this
path by a shortest-path search on the product automaton
\cite{guo2013motion,lavalle2006planning,korf1987planning}. Its worst-case
complexity is
$\mathcal{O}(|\delta_{p,i}| \cdot \log |Q_{p,i}| \cdot |Q_{p,i,0}|)$.
Projecting $R_{p,\textup{init}}^i$ onto $\Pi_{\mathcal{R}}^i$ produces the
initial motion-and-action plan
$\tau_{\mathcal{R},\textup{init}}^i \triangleq
R_{p,\textup{init}}^i|_{\Pi_{\mathcal{R}}^i}$, which fulfills $\varphi_i$.
The initial plans are synthesized locally by the robots, rather than by a
central coordinator or a cluster-level planner
\cite{chen2012formal,guo2012infeasible}. The plan
$\tau_{\mathcal{R},\textup{init}}^i$ can then be executed by following its
motion and action states in sequence. If the plan contains collaborative
actions from $\Sigma_c^i$, however, successful execution depends on assistance
from other robots, which is not guaranteed by the local off-line synthesis
alone. This motivates the on-line coordination mechanism described next.

(II) \emph{Plan in a finite horizon and request generation}.
Let $\pi_{\mathcal{R},t}^i \in \Pi_{\mathcal{R}}^i$ denote the current state of
robot $i$ at time $t$, and suppose that
$\pi_{\mathcal{R},t}^i=\tau_{\mathcal{R},\textup{init}}^i[\ell]$ for some
index $\ell$. Each robot is assigned a finite planning horizon
$0 < H_i < \infty$, which specifies how far ahead the robot examines its plan.
The segment of the initial plan expected to be reached within this horizon is
denoted by $\tau_{\mathcal{R},H}^i$ and is defined by
$\tau_{\mathcal{R},H}^i \triangleq
\tau_{\mathcal{R},\textup{init}}^i[\ell\!:\!k]$,
where $k \ge \ell$ is the smallest index satisfying
$
\sum_{s=\ell}^{k-1}
T_{\mathcal{R}}^i\!\left(
\tau_{\mathcal{R},\textup{init}}^i[s],
\tau_{\mathcal{R},\textup{init}}^i[s+1]
\right)
\ge H_i$.
If no such index exists, then $k$ is set to the final index of the remaining
plan. This horizon prevents coordination attempts for collaborative actions
that lie too far in the future.

The robot then checks whether $\tau_{\mathcal{R},H}^i$ contains a collaborative
action. Let $\langle \pi_v,\sigma_m \rangle$ be the first state in
$\tau_{\mathcal{R},H}^i$ such that $\sigma_m \in \Sigma_c^i$. In that case,
robot $i$ broadcasts a request to its neighbors. The request message is given by:
\begin{equation}
\label{ch03:request}
\textbf{Request}^i \triangleq
\{(\sigma_d,\pi_v,T_m) \mid
  \sigma_d \in \texttt{Depd}_i(\sigma_m)\},
\end{equation}
where $\pi_v$ is the region at which the collaborative action will be executed,
and $T_m \ge 0$ is the planned time from now until the execution of
$\sigma_m$. If $\langle \pi_v,\sigma_m \rangle$ is the $k$th state of
$\tau_{\mathcal{R},H}^i$, then
$T_m \triangleq
\sum_{s=\ell}^{k-1}
T_{\mathcal{R}}^i\!\left(
\tau_{\mathcal{R},H}^i[s],
\tau_{\mathcal{R},H}^i[s+1]
\right)$,
which is the elapsed time from the current state to the requested
collaboration. Each element $(\sigma_d,\pi_v,T_m)$ in \eqref{ch03:request}
means that robot $i$ requests the assisting action $\sigma_d$ at region
$\pi_v$ at time $T_m$ from the current moment. The message sent to each
neighbor $j \in \mathcal{N}_i$ is identical and is denoted by
$\textbf{Request}_j^i$.
Only the first collaborative action within the horizon is considered in
Algorithm~\ref{ch03:alg:request}. The outcome may
alter the rest of the plan and affect all later collaborative
actions within the same horizon.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{algorithm}[t]
\caption{Plan in horizon and request, \texttt{Request}()}
\label{ch03:alg:request}
\DontPrintSemicolon

\KwIn{$\tau_{\mathcal{R},\textup{init}}^i$, $\pi_{\mathcal{R},t}^i$, $H_i$}
\KwOut{$\tau_{\mathcal{R},H}^i$, $\textbf{Request}^i$}

Find $\ell$ such that
$\tau_{\mathcal{R},\textup{init}}^i[\ell]=\pi_{\mathcal{R},t}^i$\;
$k=\ell$, $T_m=0$, $\textbf{Request}^i=\emptyset$\;

\While{$(T_m < H_i)$ and $(k+1 < |\tau_{\mathcal{R},\textup{init}}^i|)$}{
  $k=k+1$\;
  $T_m=T_m+
  T_{\mathcal{R}}^i\!\left(
  \tau_{\mathcal{R},\textup{init}}^i[k-1],
  \tau_{\mathcal{R},\textup{init}}^i[k]\right)$\;
  $\langle \pi_v,\sigma_m \rangle=
  \tau_{\mathcal{R},\textup{init}}^i[k]$\;

  \If{$\textup{\textbf{Request}}^i=\emptyset$ and $\sigma_m \in \Sigma_c^i$}{
    \ForAll{$\sigma_d \in \textup{\texttt{Depd}}_i(\sigma_m)$}{
      add $(\sigma_d,\pi_v,T_m)$ to $\textbf{Request}^i$\;
    }
  }
}

$\tau_{\mathcal{R},H}^i=
\tau_{\mathcal{R},\textup{init}}^i[\ell\!:\!k]$\;

\Return{$\tau_{\mathcal{R},H}^i$, $\textup{\textbf{Request}}^i$}\;
\end{algorithm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

(III) \emph{Request evaluation and reply}.
After receiving a request, a neighbor
$j \in \mathcal{N}_i$ evaluates it in terms of feasibility and timing. The
reply message from robot $j$ to robot $i$ is given by:
\begin{equation}
\label{ch03:eq:reply}
\textbf{Reply}_i^j \triangleq
\{(\sigma_d,b_d^j,t_d^j) \mid
  (\sigma_d,\pi_v,T_m) \in \textbf{Request}_j^i\},
\end{equation}
where $b_d^j \in \{\top,\bot\}$ indicates whether robot $j$ can provide the
assisting action $\sigma_d$ at region $\pi_v$, and $t_d^j \ge 0$ is the time
at which the assistance can be provided.
Let $\overline{T}_j \ge 0$ denote the finishing time of the collaboration in
which robot $j$ is currently engaged. Initially $\overline{T}_j=0$. If
$\overline{T}_j>0$, then robot $j$ is unavailable and returns the rejection
reply $\{(\sigma_d,\bot,0)\}$ for all requested actions. If
$\overline{T}_j=0$, then robot $j$ is available and evaluates each request.
Let $\tau_{\mathcal{R},t^-}^j$ denote the current plan of robot $j$ before any
potential revision, and let $R_{p,t^-}^j$ be its corresponding accepting run in
$\mathcal{A}_{p,j}$. Suppose the current product state $q_{p,t}^j$ is the
$\ell$th state of $R_{p,t^-}^j$, and the accepting state $q_{p,f}^j$ is its
last state. The remaining segment is
$\widehat{R}_{p,-}^j \triangleq R_{p,t^-}^j[\ell\!:\!f]$. The objective is to
find an alternative segment $\widehat{R}_{p,+}^j$ from $q_{p,t}^j$ to
$q_{p,f}^j$ such that robot $j$ reaches the state
$\langle \pi_v,\sigma_d \rangle$, does so at a time close to $T_m$, and incurs
only a small additional cost.
Let
$S_d \triangleq
\{q_p \in Q_{p,j} \mid
  q_p|_{\Pi_{\mathcal{R}}^j}=\langle \pi_v,\sigma_d \rangle\}$,
which is the set of product states associated with the requested assisting
action. A candidate revised run has the structure below:
\begin{equation}
\label{ch03:segment}
\widehat{R}_{p,+}^j \triangleq
q_{p,t}^j \cdots q_{p,r}^j \cdots q_{p,f}^j,
\end{equation}
where $q_{p,r}^j \in S_d$. The balanced cost of
$\widehat{R}_{p,+}^j$ is defined by:
\begin{equation}
\label{ch03:balanced}
\begin{aligned}
\texttt{BalCost}(\widehat{R}_{p,+}^j,T_m,\mathcal{A}_{p,j})
\triangleq\;&
\left|
\sum_{s=1}^{r-t} W_{p,j}\big(\widehat{R}_{p,+}^j[s],\widehat{R}_{p,+}^j[s+1]\big)
- T_m
\right| \\
& + \alpha_j \big(
\texttt{Cost}(\widehat{R}_{p,+}^j,\mathcal{A}_{p,j})
-\texttt{Cost}(\widehat{R}_{p,-}^j,\mathcal{A}_{p,j})
\big),
\end{aligned}
\end{equation}
where the first term measures the timing mismatch relative to the request and
the second term measures the additional cost of the revised run. The parameter
$\alpha_j>0$ is a design weight.
The corresponding optimization problem is as follows.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{algorithm}[t]
\caption{Reply to request by robot $j$, \texttt{Reply}()}
\label{ch03:alg:reply}
\DontPrintSemicolon

\KwIn{$\textbf{Request}_j^i$, $\widehat{R}_{p,-}^j$, $\mathcal{A}_{p,j}$,
$\overline{T}_j$}
\KwOut{$\textbf{Reply}_i^j$, $\widehat{P}$}

$\textbf{Reply}_i^j=\emptyset$, $\widehat{P}=\emptyset$\;

\If{$\overline{T}_j>0$}{
  \ForAll{$(\sigma_d,\pi_v,T_m) \in \textbf{Request}_j^i$}{
    add $(\sigma_d,\bot,0)$ to $\textbf{Reply}_i^j$\;
  }
  \Return{$\textbf{Reply}_i^j$, $\widehat{P}$}\;
}

\ForAll{$(\sigma_d,\pi_v,T_m) \in \textup{\textbf{Request}}_j^i$}{
  $(\widehat{R}_{p,+}^j,b_d^j,t_d^j)=
  \texttt{EvalReq}(
  \widehat{R}_{p,-}^j,(\pi_v,\sigma_d,T_m),\mathcal{A}_{p,j})$\;

  \If{$b_d^j=\top$}{
    $\widehat{P}(\sigma_d)=\widehat{R}_{p,+}^j$\;
    add $(\sigma_d,\top,t_d^j)$ to $\textbf{Reply}_i^j$\;
  }
  \Else{
    add $(\sigma_d,\bot,0)$ to $\textbf{Reply}_i^j$\;
  }
}

\Return{$\textup{\textbf{Reply}}_i^j$, $\widehat{P}$}\;
\end{algorithm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{problem}
Given $\widehat{R}_{p,-}^j$, $S_d$, and $\mathcal{A}_{p,j}$, find a path
segment $\widehat{R}_{p,+}^j$ with the form \eqref{ch03:segment} that
minimizes \eqref{ch03:balanced}.
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{algorithm}[t]
\caption{Evaluate the request, \texttt{EvalReq}($\cdot$)}
\label{ch03:adapt}
\DontPrintSemicolon

\KwIn{$\widehat{R}_{p,-}^j$, $(\pi_v,\sigma_d,T_m)$, $\mathcal{A}_{p,j}$}
\KwOut{$\widehat{R}_{p,+}^j$, $b_d^j$, $t_d^j$}

$q_{p,t}^j=\widehat{R}_{p,-}^j[1]$,
$q_{p,f}^j=\widehat{R}_{p,-}^j[\text{-1}]$\;

Compute $S_d$ and $S_c$\;

$\bar{c}=\texttt{Cost}(\widehat{R}_{p,-}^j,\mathcal{A}_{p,j})$\;

$(P_1,C_1)=\texttt{DijksTA}
(\mathcal{A}_{p,j},q_{p,t}^j,S_d,S_c)$\;

$(P_2,C_2)=\texttt{DijksTA}
(\texttt{Reverse}(\mathcal{A}_{p,j}),q_{p,f}^j,S_d,\emptyset)$\;

\ForAll{$q_{p,r}^j \in S_d$}{
  \If{$P_1(q_{p,r}^j)$ and $P_2(q_{p,r}^j)$ exist}{
    $C_3(q_{p,r}^j)=
    |C_1(q_{p,r}^j)-T_m|
    +\alpha_j(C_1(q_{p,r}^j)+C_2(q_{p,r}^j)-\bar{c})$\;
  }
}

Find $q_{p,r}^{j,\star} \in S_d$ that minimizes $C_3(q_{p,r}^j)$\;

\If{$q_{p,r}^{j,\star} \neq \emptyset$}{
  $P=P_1(q_{p,r}^{j,\star})+
  \texttt{Reverse}(P_2(q_{p,r}^{j,\star}))$\;
  \Return{$P$, $\top$, $C_1(q_{p,r}^{j,\star})$}\;
}

\Return{$\emptyset$, $\bot$, $0$}\;
\end{algorithm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Algorithm~\ref{ch03:adapt} solves the problem by a bidirectional Dijkstra
search. The routine \texttt{DijksTA}($\cdot$) computes shortest paths from one
source state to all states in a target set while avoiding a specified set of
states. Let
\begin{equation}
\label{ch03:eq:Sc}
S_c \triangleq
\{q_p \in Q_{p,j} \mid
  q_p|_{\Pi_{\mathcal{R}}^j}
  = \langle \pi_{\mathcal{R}}^j,\sigma_n \rangle,\,
  \sigma_n \in \Sigma_c^j \cup \Sigma_h^j\},
\end{equation}
which excludes collaborative and assisting actions other than the requested one
from the forward search. Then
\texttt{DijksTA}$(\mathcal{A}_{p,j},q_{p,t}^j,S_d,S_c)$ computes the shortest
paths from $q_{p,t}^j$ to all states in $S_d$ while avoiding the states in
$S_c$, and
\texttt{DijksTA}$(\texttt{Reverse}(\mathcal{A}_{p,j}),q_{p,f}^j,S_d,\emptyset)$
computes shortest paths in the reversed graph from $q_{p,f}^j$ to all states
in $S_d$. The state that minimizes the balanced cost in
\eqref{ch03:balanced} determines the revised run $\widehat{R}_{p,+}^j$. If no
such state exists, then robot $j$ cannot provide the requested assistance.
The complexity of \texttt{DijksTA} over $\mathcal{A}_{p,j}$ is
$\mathcal{O}(|\delta_{p,j}| \cdot \log |Q_{p,j}|)$, while constructing the
reversed automaton is linear in $|\delta_{p,j}|$
\cite{baier2008principles,lavalle2006planning,korf1987planning}. The run
$\widehat{R}_{p,+}^j$ is only a potential revision. Robot $j$ does not
immediately change its current plan. Instead, it stores the candidate revised
run in $\widehat{P}$ and waits for a confirmation message from robot $i$.
If robot $i$ receives requests from several robots at the same time, it replies
to one request first and waits for the corresponding confirmation before
processing the next request.

\begin{lemma}
\label{ch03:td}
If Algorithm~\ref{ch03:adapt} returns $b_d^j=\top$, then action $\sigma_d$ can
be done by robot $j$ at time $t_d^j$ by following the run
$\widehat{R}_{p,+}^j$.
\end{lemma}

\begin{proof}
The first segment of $\widehat{R}_{p,+}^j$, from $q_{p,t}^j$ to the selected
state $q_{p,r}^{j,\star}$, is produced by \texttt{DijksTA} while avoiding all
states in $S_c$. Hence this segment contains no collaborative or assisting
actions other than the requested action $\sigma_d$. It therefore consists only
of motions and local actions that robot $j$ can execute independently. The time
required to reach $q_{p,r}^{j,\star}$ is exactly $t_d^j$.
\end{proof}


(IV) \emph{Confirmation and assignment}.
Based on the replies from the neighbors, robot $i$ sends confirmation messages
of the form:
\begin{equation}
\label{ch03:eq:confirm}
\textbf{Confirm}_j^i \triangleq
\{(\sigma_d,c_d^j,f_m) \mid
  \sigma_d \in \texttt{Depd}_i(\sigma_m)\},
\end{equation}
where $c_d^j \in \{\top,\bot\}$ indicates whether robot $j$ is confirmed to
provide the assisting action $\sigma_d$, and $f_m$ is the time at which the
collaborative action $\sigma_m$ is scheduled to finish.
The confirmation decision must satisfy two conditions. First, each required
assisting action must be assigned to exactly one robot. Second, each neighbor
may be assigned to at most one assisting action for the current collaboration.
The finishing time should be as early as possible.
Let $|\mathcal{N}_i|=N_1$ and
$|\texttt{Depd}_i(\sigma_m)|=N_2$. Without loss of generality, write
$\mathcal{N}_i=\{1,\cdots,N_1\}$ and
$\texttt{Depd}_i(\sigma_m)=\{\sigma_1,\cdots,\sigma_{N_2}\}$. The assignment
problem can then be written as the integer program below:
\begin{equation}
\label{ch03:integer}
\begin{aligned}
\textbf{minimize} \quad & f_m, \\
\textbf{subject to} \quad
& f_m \ge T_m, \\
& f_m \ge c_d^j t_d^j,
  \quad \forall j \in \{1,\cdots,N_1\},\;
  \forall d \in \{1,\cdots,N_2\}, \\
& \sum_{d=1}^{N_2} b_d^j c_d^j \le 1,
  \quad \forall j \in \{1,\cdots,N_1\}, \\
& \sum_{j=1}^{N_1} b_d^j c_d^j = 1,
  \quad \forall d \in \{1,\cdots,N_2\}, \\
& c_d^j \in \{0,1\},
  \quad \forall j \in \{1,\cdots,N_1\},\;
  \forall d \in \{1,\cdots,N_2\},
\end{aligned}
\end{equation}
where $(\sigma_d,b_d^j,t_d^j) \in \textbf{Reply}_i^j$ from
\eqref{ch03:eq:reply}. Any standard integer-programming solver can be used once
\eqref{ch03:integer} has been formed.
If \eqref{ch03:integer} is feasible, then $f_m$ and the decision variables
$c_d^j$ are obtained. In that case, robot $i$ sends
$(\sigma_d,\top,f_m)$ to the robot selected for action $\sigma_d$, and sends
$(\sigma_d,\bot,0)$ to the remaining robots. If \eqref{ch03:integer} is
infeasible, robot $i$ sends $(\sigma_d,\bot,0)$ to all neighbors for all
requested assisting actions. This indicates that $\sigma_m$ cannot be executed
as currently planned and must be delayed. This integer optimization is
solved locally by robot $i$ for the collaborative action $\sigma_m$ and
contains
$|\mathcal{N}_i| \cdot |\texttt{Depd}_i(\sigma_m)|$
Boolean decision variables.

(V) \emph{Plan adaptation}.
After sending the confirmation messages, robot $i$ proceeds as follows. If
\eqref{ch03:integer} is feasible, then the collaborative action $\sigma_m$ can
be executed as planned, the current accepting run of robot $i$ remains
unchanged, and $\overline{T}_i$ is set to $f_m$ to indicate that robot $i$ is
engaged in the collaboration until time $f_m$. If \eqref{ch03:integer} is
infeasible, then the collaborative action cannot be executed at the originally
planned time and the run of robot $i$ must be revised by delaying $\sigma_m$.
The delayed run is computed by Algorithm~\ref{ch03:alg:delay}. The idea is to
find a new segment from the current state to the accepting state that reaches
the collaborative action state $\langle \pi_v,\sigma_m \rangle$ no earlier than
$T_m + D_i$, where $D_i > 0$ is a design parameter. The existence of such a
path is guaranteed by the availability of the idle action $\sigma_0$, whose
duration $T_0$ allows waiting to be inserted as needed.
When a neighbor $j \in \mathcal{N}_i$ receives a confirmation message, it also
updates its plan. If $c_d^j=\top$, then robot $j$ is confirmed to provide the
assisting action $\sigma_d$ and replaces the current segment by the stored
candidate run from $\widehat{P}$. In this case, $\overline{T}_j$ is set to
$f_m$. If $c_d^j=\bot$ for all requested actions, then robot $j$ keeps its
current plan and resets $\overline{T}_j$ to $0$.
Each robot has a finite motion-and-action plan. Once a robot finishes its own
task, it becomes an assisting robot by setting $\varphi_i \triangleq \top$ and
remains at its current region unless it is later confirmed to collaborate with
another robot.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{algorithm}[t!]
\caption{Delay collaboration, \texttt{DelayCol}()}
\label{ch03:alg:delay}
\DontPrintSemicolon

\KwIn{$\widehat{R}_{p,-}^i$, $(\pi_v,\sigma_m,T_m)$, $D_i$, $\mathcal{A}_{p,i}$}
\KwOut{$\widehat{R}_{p,+}^i$}

$q_{p,t}^i=\widehat{R}_{p,-}^i[1]$,
$q_{p,f}^i=\widehat{R}_{p,-}^i[\text{-1}]$\;

Compute
$S_d=
\{q_p \in Q_{p,i} \mid
  q_p|_{\Pi_{\mathcal{R}}^i}
  = \langle \pi_v,\sigma_m \rangle\}$ and $S_c$\;

$(P_1,C_1)=\texttt{DijksTA}
(\mathcal{A}_{p,i},q_{p,t}^i,S_d,S_c)$\;

$(P_2,C_2)=\texttt{DijksTA}
(\texttt{Reverse}(\mathcal{A}_{p,i}),q_{p,f}^i,S_d,\emptyset)$\;

\ForAll{$q_{p,d}^i \in S_d$}{
  \If{$C_1(q_{p,d}^i)>T_m+D_i$ and $P_2(q_{p,d}^i)$ exists}{
    $\widehat{R}_{p,+}^i=
    P_1(q_{p,d}^i)+
    \texttt{Reverse}(P_2(q_{p,d}^i))$\;
    \Return{$\widehat{R}_{p,+}^i$}\;
  }
}
\end{algorithm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{theorem}
\label{ch03:tm}
If \eqref{ch03:integer} is feasible, the collaborative action $\sigma_m$
is accomplished at time $f_m$.
\end{theorem}

\begin{proof}
For each assisting action
$\sigma_d \in \texttt{Depd}_i(\sigma_m)$, feasibility of
\eqref{ch03:integer} implies that exactly one robot
$j \in \mathcal{N}_i$ is assigned to $\sigma_d$. By
Lemma~\ref{ch03:td}, that robot can complete $\sigma_d$ at time $t_d^j$. By
the constraints in \eqref{ch03:integer}, $f_m$ is at least the scheduled time
of every confirmed assisting action and is no earlier than $T_m$. Hence all
assisting actions required by $\sigma_m$ are completed by time $f_m$, and
$\sigma_m$ is accomplished at time $f_m$.
\end{proof}


(VI) \emph{Failure detection and recovery}.
Physical robots may fail at any time, which makes failure detection especially
important in collaborative tasks. Consider robot $i$, which has confirmed the
collaborative action $\sigma_m$. Until $\sigma_m$ is completed, robot $i$
periodically monitors the status of each confirmed collaborator associated with
an assisting action in $\texttt{Depd}_i(\sigma_m)$. If a confirmed collaborator
$h \in \mathcal{N}_i$ fails to acknowledge the inquiry within a bounded time,
robot $i$ treats robot $h$ as failed.
Robot $i$ then reassigns the affected assisting action $\sigma_d$ as follows.
First, it sends a new request $\{(\sigma_d,\pi_v,T_m)\}$ to its neighbors.
Second, based on the replies
$\{(\sigma_d,b_d^j,t_d^j)\mid j \in \mathcal{N}_i\}$, it selects the new
collaborator by
\begin{equation}
\label{ch03:eq:reassign}
\hat{h} \triangleq
\arg\min_{j \in \mathcal{N}_i}
|f_m - t_d^j b_d^j|,
\end{equation}
where $f_m$ is the previously confirmed finishing time. Third, it sends the
confirmation $\{(\sigma_d,\top,f_m)\}$ to robot $\hat{h}$. Finally, robot
$\hat{h}$ adapts its plan according to the nominal procedure.


\subsubsection{Overall structure}
\label{ch03:structure}

The proposed coordination scheme is intended for loosely coupled multi-robot
systems, in which collaborations are local and relatively sparse compared with
the total number of robot activities. This property is formalized by the
following assumption.

\begin{assumption}
\label{ch03:assump:loose}
There exists a finite constant ${T}>0$ such that, for every robot
$i \in \mathcal{N}$ and every collaborative action $\sigma_m$ requested by
robot $i$ at time $t_m>0$, Problem~\eqref{ch03:integer} becomes feasible no
later than time $t_m+{T}$.
\end{assumption}

Assumption~\ref{ch03:assump:loose} states that every collaborative action needed
for the fulfillment of a local task can eventually be supported, in finite
time, by the requesting robot and its neighbors.
During execution, each robot first checks whether it has received a request. If
so, it evaluates the request using Algorithm~\ref{ch03:alg:reply}, waits for a
confirmation, and updates its plan if necessary. Otherwise, it examines its own
plan by Algorithm~\ref{ch03:alg:request}, sends any required request, collects
the replies, solves the assignment optimization in~\eqref{ch03:integer}, and, if
needed, delays the collaborative action by
Algorithm~\ref{ch03:alg:delay}. The correctness of the overall scheme is stated
next.

\begin{theorem}
\label{ch03:all}
Under Assumption~\ref{ch03:assump:loose}, the proposed coordination scheme
solves Problem~\ref{ch03:big-problem}. That is, all local tasks $\varphi_i$ are
fulfilled in finite time for all $i \in \mathcal{N}$.
\end{theorem}

\begin{proof}
Consider an arbitrary robot $i \in \mathcal{N}$ and its initial plan
$\tau_{\mathcal{R},\textup{init}}^i$. All motion transitions and local actions
in this plan can be executed autonomously by robot $i$. It remains to show that
each collaborative action required by the plan is also completed in finite
time.
If the plan of robot $i$ remains unchanged, then every collaborative action
$\sigma_m$ appearing in the plan is associated with a feasible instance of
\eqref{ch03:integer}. By Theorem~\ref{ch03:tm}, each such action is completed
at a finite time $f_m$. Since the plan is finite, the entire plan is executed
in finite time, and the corresponding trace satisfies $\varphi_i$.
If the plan of robot $i$ is revised during execution, two cases may occur.
First, robot $i$ may be confirmed to assist a neighboring robot. In that case,
Algorithm~\ref{ch03:adapt} constructs a revised accepting run that still
reaches an accepting state, so the updated plan continues to satisfy
$\varphi_i$ in finite time. Second, robot $i$ may request assistance for a
collaborative action $\sigma_m$ and receive an infeasible outcome for
\eqref{ch03:integer}, which triggers Algorithm~\ref{ch03:alg:delay}. By
Assumption~\ref{ch03:assump:loose}, the delayed coordination problem becomes
feasible within at most ${T}$ additional time, so the postponed
collaborative action is still completed in finite time.
Therefore, every robot fulfills its local task in finite time.
\end{proof}
