\chapter{Preliminaries for Single-Robot Temporal Task Planning}
\label{chap:preliminaries}

This chapter presents the nominal framework for integrated motion planning and
task planning for a mobile robot. The robot motion in the workspace is first
represented by a finite-state transition system, which provides a discrete
abstraction of the underlying continuous dynamics. On the basis of this
abstraction, the mission specification is described by temporal logic formulas
defined over the set of abstract states. A model-checking-based synthesis
procedure is then employed to compute a discrete plan that satisfies the given
task while optimizing the associated execution cost. Finally, this discrete
plan is refined into a hybrid control strategy that steers the robot in the
continuous workspace and ensures satisfaction of the prescribed task
specification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Abstraction of Robot Motion}
\label{ch02:sec:system_abstraction_task_specification}

This section develops a discrete abstraction of robot motion in the workspace.
The objective is to derive a symbolic model that is sufficiently expressive for
task specification and formally correct plan synthesis, while remaining close to
the underlying continuous dynamics. The construction proceeds in three steps.
First, the workspace is partitioned into finitely many regions. Second, the
robot dynamics and the available navigation laws are used to determine which
inter-region motions are actually feasible. Third, the resulting motion
capability is encoded as a weighted finite-state transition system. This model
serves as the nominal discrete substrate for the temporal-logic framework used
in the remainder of the chapter.

\subsection{Workspace model}
\label{ch02:subsec:workspacemodel}

Consider a bounded workspace
$\mathcal{W}_0 \subset \mathbb{R}^{n}$.
Let
$\Pi \triangleq \{\pi_1,\cdots,\pi_N\}$
be a finite collection of regions of interest, where each
$\pi_i \subset \mathcal{W}_0$.
The set $\Pi$ is assumed to form a partition of the workspace. Hence, the
regions jointly cover the entire workspace and do not overlap pairwise. This
assumption is important because the abstraction should assign each robot
position to one and only one symbolic state.
Atomic propositions are introduced to describe properties of the robot and of
the workspace that are relevant to the planning task. The first class of
propositions records the robot location. Define
$AP_r \triangleq \{a_{r,1},\cdots,a_{r,N}\}$,
where each proposition $a_{r,i}$ indicates whether the robot is located in
region $\pi_i$. Its truth value is given by:
\begin{equation}
a_{r,i} \triangleq
\begin{cases}
\top, & \text{if the robot is located in } \pi_i, \\
\bot, & \text{otherwise},
\end{cases}
\label{ch02:eq:region_proposition}
\end{equation}
where the truth value in \eqref{ch02:eq:region_proposition} can be evaluated
from the output of a localization or positioning module.

Because $\Pi$ is a partition of $\mathcal{W}_0$, exactly one proposition from
$AP_r$ is true at every instant. This property is essential for obtaining an
unambiguous symbolic description of the robot motion. In addition to the
location propositions, task-dependent properties of the workspace are encoded by
a second set
$AP_p \triangleq \{a_{p,1},\cdots,a_{p,M}\}$.
These propositions may express, for example, the presence of an object, the
availability of a resource, or the designation of a service location. The full
set of atomic propositions is then defined by
$AP \triangleq AP_r \cup AP_p$.

\begin{definition}
\label{ch02:def:labeling}
A labeling function is a map
$L : \Pi \to 2^{AP}$
that assigns to each region $\pi_i \in \Pi$ the set of atomic propositions that
hold in that region. In particular,
$a_{r,i} \in L(\pi_i)$
for every $i \in \{1,\cdots,N\}$.
\end{definition}

Definition~\ref{ch02:def:labeling} associates semantic information with each
symbolic state. Once the labeling has been fixed, a motion of the robot through
the partition induces a corresponding sequence of truth assignments over the
proposition set. This sequence will later be interpreted against temporal logic
formulas.
The labeling is assumed to be uniform within each region. Therefore, partial
satisfaction of a proposition inside a single region is not permitted. If a
property holds only on a subset of a region, that region should be refined into
smaller regions so that the truth assignment becomes uniform on each resulting
cell. This requirement is standard in symbolic abstraction, since the discrete
model reasons at the level of regions rather than individual points.
A practical abstraction also requires each region to admit a compact geometric
representation. Rectangular regions may be described by center, width, and
height. Circular regions may be described by center and radius. Polygonal
regions may be encoded through their vertices. Standard geometric procedures,
such as Delaunay triangulation~\citep{lee1980two} and Voronoi-based
decomposition~\citep{lavalle2006planning}, can also be used to generate such a
partition. At this stage, however, the partition is purely geometric and does
not yet incorporate the robot dynamics or the controller structure.


\subsection{Robot dynamics and navigation controller}
\label{ch02:subsec:robotdynamics}

The robot is modeled by the continuous-time control system below:
\begin{equation}
\dot{x}(t) = f\bigl(x(t),u(t)\bigr),
\label{ch02:eq:dynamic}
\end{equation}
where $x(t) \in \mathbb{R}^{n}$ is the system state,
$u(t) \in \mathbb{R}^{m}$ is the control input, and
$f : \mathbb{R}^{n} \times \mathbb{R}^{m} \to \mathbb{R}^{n}$
is assumed to be Lipschitz continuous~\citep{khalil2002nonlinear}.
The regularity assumption on $f$ ensures that the system in
\eqref{ch02:eq:dynamic} is deterministic in the standard sense. For every
admissible initial condition and every admissible control signal, there exists a
unique state trajectory. The subsequent construction is stated for a general
nonlinear model. The examples in this chapter, however, primarily involve
single-integrator and unicycle dynamics.
Given the partition introduced in
Section~\ref{ch02:subsec:workspacemodel}, the next question is whether the robot
can be steered from one region to another under the dynamics
\eqref{ch02:eq:dynamic}. This notion is stronger than geometric adjacency. Two
regions may share a boundary and yet fail to admit a feasible transition under
the available control law. For that reason, the transition relation is defined
in a control-driven manner.

\begin{definition}
\label{ch02:def:transition}
Let $\pi_i,\pi_j \in \Pi$. A control-driven transition from $\pi_i$ to $\pi_j$
exists if there is an admissible navigation controller
$\mathcal{U} : \mathbb{R}^{n} \times \Pi \times \Pi \to \mathbb{R}^{m}$ such
that the control input is given by:
\begin{equation}
u(t) \triangleq \mathcal{U}\bigl(x(t),\pi_i,\pi_j\bigr),
\quad t \in [t_s,t_f],
\label{ch02:eq:control}
\end{equation}
where the corresponding closed-loop trajectory of
\eqref{ch02:eq:dynamic} satisfies
$x(t_s) \in \pi_i$,
$x(t_f) \in \pi_j$,
and
$x(t) \in \pi_i \cup \pi_j$
for all $t \in [t_s,t_f]$.
\end{definition}

Definition~\ref{ch02:def:transition} imposes two requirements. First, the
transition must be realizable under an admissible controller. Second, the
motion should remain within the union of the source and target regions during
its execution. The second requirement prevents the symbolic transition from
implicitly relying on intermediate regions that do not appear in the discrete
plan.
The above definition is closely connected to the hybrid execution mechanism that
will be introduced later in the chapter. A symbolic transition is included in
the abstraction only if a concrete controller exists that can implement it in
the continuous domain. In this sense, the abstraction is not merely geometric.
It is explicitly shaped by the dynamical capabilities of the robot and by the
available control architecture.

The existence of a suitable navigation law depends on both the robot model and
the geometry of the workspace decomposition. One representative approach is the
potential-field construction in~\citep{lindemann2006real}, where a
differential-drive robot is guided from any point in a region to an adjacent
region through a prescribed facet. This construction is naturally associated
with triangular decompositions derived from generalized Voronoi structures.
Another representative approach is the navigation-function framework
in~\citep{koditschek1990robot}, which provides provably correct point-to-point
navigation in sphere worlds with obstacles. Variants of this idea have been
used in both single-robot and multi-robot settings
\citep{loizou2008density,loizou2002closed,dimarogonas2007decentralized,
dimarogonas2005decentralized}.
A navigation method is rarely compatible with an arbitrary geometric partition.
Consequently, the preliminary partition may need to be refined or reshaped after
the dynamics and controller structure have been taken into account. Such a
refinement may produce either an over-approximation or an
under-approximation~\citep{saidi1999abstract} of the original workspace model.
If no available navigation strategy is compatible with the geometry, an
additional controller design step becomes necessary before the abstraction can
be completed.


\subsection{Control-driven weighted transition system}
\label{ch02:subsec:FTS}

The above ingredients now yield a symbolic model of robot motion. Each region
$\pi_i \in \Pi$ is interpreted as a discrete state representing the event that
the robot is located in that region. The admissible state changes are exactly
those induced by Definition~\ref{ch02:def:transition}. This leads to the
following control-driven transition system.

\begin{definition}
\label{ch02:def:FTS}
The control-driven weighted finite-state transition system is the tuple below:
\begin{equation}
\mathcal{T}_c \triangleq
\bigl(\Pi,\longrightarrow_c,\Pi_0,AP,L_c,w_c\bigr),
\label{ch02:eq:fts}
\end{equation}
where $\Pi$ is the finite state set induced by the workspace partition,
$\longrightarrow_c \subseteq \Pi \times \Pi$ is the control-driven transition
relation,
$\Pi_0 \subseteq \Pi$ is the set of initial states,
$AP$ is the set of atomic propositions,
$L_c : \Pi \to 2^{AP}$ is the labeling function inherited from
Definition~\ref{ch02:def:labeling}, and
$w_c : \longrightarrow_c \to \mathbb{R}_{\geq 0}$ assigns a nonnegative cost to
each enabled transition.
\end{definition}

The weighted transition system in \eqref{ch02:eq:fts} is the discrete model on
which the robot task will be interpreted. The function $w_c$ quantifies the
cost of executing a feasible symbolic transition. Depending on the application,
this quantity may represent travel time, path length, control effort, energy
consumption, or another performance metric. For later synthesis, the inclusion
of transition weights makes it possible to reason not only about task
satisfaction but also about cost-optimal execution.
The model is assumed to have no terminal states. This assumption is natural for
the infinite-horizon temporal specifications considered later. For each state
$\pi_i \in \Pi$, define the successor set
$\texttt{Post}(\pi_i) \triangleq
\{\pi_j \in \Pi \mid \pi_i \longrightarrow_c \pi_j\}$.
An infinite path of $\mathcal{T}_c$ is then an infinite sequence
$\tau = \pi_0\pi_1\pi_2\cdots$
such that
$\pi_0 \in \Pi_0$
and
$\pi_{k+1} \in \texttt{Post}(\pi_k)$
for all $k \geq 0$.

Every path induces a corresponding sequence of labels. More precisely, the trace
generated by a path $\tau$ is the infinite word
$\texttt{trace}(\tau) \triangleq
L_c(\pi_0)L_c(\pi_1)L_c(\pi_2)\cdots$.
Accordingly, the trace language of the transition system is
$\texttt{Trace}(\mathcal{T}_c) \triangleq
\{\texttt{trace}(\tau) \mid
\tau \text{ is an infinite path of } \mathcal{T}_c\}$.
This notion will be central later, since temporal logic formulas are evaluated
over such infinite words rather than directly over the continuous state
trajectories.
Ultimately periodic paths may be represented compactly by the
$\omega$-operator~\citep{baier2008principles}. Finally, the transition system
$\mathcal{T}_c$ is said to be fully known if it faithfully captures the
workspace model and the robot motion capabilities. It is said to be static if
it does not vary with time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Task Specification as LTL Formulas}
\label{ch02:sec:LTL}

A formal language is required to describe robotic tasks without ambiguity.
Such a language should express logical conditions, temporal ordering, and
recurrent mission requirements in a mathematically precise manner. In this
chapter, the specification language is chosen as linear-time temporal logic,
abbreviated as LTL. The reason is twofold. First, LTL is expressive enough to
capture a broad range of motion and surveillance tasks. Second, its semantics
is defined over infinite words, which matches the infinite executions generated
by the weighted transition system $\mathcal{T}_c$ introduced in
Section~\ref{ch02:subsec:FTS}.

\subsection{Syntax and semantics}
\label{ch02:subsec:syntax}

The syntax of LTL is defined over the atomic proposition set $AP$. The formulas
are constructed recursively from atomic propositions and temporal operators.

\begin{definition}
\label{ch02:def:ltl_syntax}
An LTL formula over $AP$ is generated by the grammar
\begin{equation}
\varphi ::= \top \mid a \mid \neg \varphi \mid \varphi_1 \vee \varphi_2
\mid \bigcirc \varphi \mid \varphi_1 \,\mathsf{U}\, \varphi_2,
\label{ch02:eq:LTLsyntax}
\end{equation}
where $a \in AP$, $\neg$ for negation,
$\vee$ for disjunction, $\bigcirc$ for the next operator, and
$\mathsf{U}$ for the until operator.
\end{definition}

Other operators used frequently in task
specification are derived from these primitives in the standard way. In
particular, conjunction, implication, always, and eventually may be introduced
as abbreviations. Standard derivations are given in Chapter~5 of
\citep{baier2008principles}.
The semantics of LTL is defined over infinite words on the alphabet $2^{AP}$.
Each symbol in such a word records the set of atomic propositions that are true
at a given discrete instant.

\begin{definition}
\label{ch02:def:word_semantics}
An infinite word over $2^{AP}$ is a sequence
$\sigma \triangleq S_0 S_1 S_2 \cdots$, where
$S_k \in 2^{AP}$ for every $k \geq 0$. The satisfaction relation
$(\sigma,k) \models \varphi$ is defined recursively by
\begin{equation}
\begin{aligned}
(\sigma,k) \models a
&\iff a \in S_k, \\
(\sigma,k) \models \neg \varphi
&\iff (\sigma,k) \not\models \varphi, \\
(\sigma,k) \models \bigcirc \varphi
&\iff (\sigma,k+1) \models \varphi, \\
(\sigma,k) \models \varphi_1 \vee \varphi_2
&\iff (\sigma,k) \models \varphi_1
\text{ or } (\sigma,k) \models \varphi_2, \\
(\sigma,k) \models \varphi_1 \,\mathsf{U}\, \varphi_2
&\iff \exists j \geq k \text{ such that }
(\sigma,j) \models \varphi_2 \\
&\quad \text{and }
(\sigma,\ell) \models \varphi_1
\text{ for every } \ell \in \{k,\cdots,j-1\},
\end{aligned}
\label{ch02:eq:ltl_semantics}
\end{equation}
where $a \in AP$, $k \in \mathbb{N}$, and $S_k$ is the set of propositions
satisfied at position $k$ of the word.
\end{definition}

According to \eqref{ch02:eq:ltl_semantics}, a formula $\varphi$ is satisfied by
a word $\sigma$ if $(\sigma,0)\models \varphi$. For brevity, this is written as
$\sigma \models \varphi$. The language generated by $\varphi$ is then defined as
$\texttt{Words}(\varphi) \triangleq
\{\sigma \in (2^{AP})^{\omega} \mid \sigma \models \varphi\}$.
This viewpoint is fundamental because temporal formulas are evaluated over
symbolic behaviors rather than directly over continuous state trajectories.
The connection with the transition system follows directly from the definitions
in Section~\ref{ch02:subsec:FTS}. Every infinite path $\tau$ of
$\mathcal{T}_c$ generates an infinite word
$\texttt{trace}(\tau)$ over the alphabet $2^{AP}$. Therefore, the
semantics in \eqref{ch02:eq:ltl_semantics} can be applied directly to traces of
the control-driven weighted transition system.

\begin{definition}
\label{ch02:def:path-satisfy}
Let $\tau$ be an infinite path of $\mathcal{T}_c$, and let $\varphi$ be an LTL
formula over $AP$. The path $\tau$ is said to satisfy $\varphi$ if
$\texttt{trace}(\tau) \models \varphi$. In that case, one writes
$\tau \models \varphi$. A satisfying path is also called a plan for
$\varphi$.
\end{definition}

LTL can express many classes of robotic tasks in a unified manner. Safety
requirements may be written as $\square \neg \psi$, which enforces permanent
avoidance of an undesired condition $\psi$. Ordered visiting requirements may
be expressed by formulas such as
$\Diamond(\psi_1 \wedge \Diamond(\psi_2 \wedge \Diamond \psi_3))$, which
require the propositions $\psi_1$, $\psi_2$, and $\psi_3$ to become true in
that order. Response requirements can be expressed by
$\square(\psi_1 \Rightarrow \Diamond \psi_2)$, meaning that whenever
$\psi_1$ holds, $\psi_2$ must hold at some future step. Persistent
surveillance tasks are captured by $\square \Diamond \psi$, which requires
$\psi$ to occur infinitely often.
A particularly important subclass in robotic planning is syntactically co-safe
LTL, abbreviated as sc-LTL~\citep{kupferman2001model}. Informally, an sc-LTL
formula has the property that satisfaction can be certified after observing a
finite prefix of an infinite word. Once such a satisfying prefix has been
generated, the remaining suffix no longer affects satisfaction. This feature is
especially useful for goal-reaching and sequencing tasks, where successful task
completion can be recognized in finite time. The syntactic characterization
used in this chapter follows the standard presentation in
\citep{kupferman2001model,fainekos2009robustness}.

\subsection{Problem formulation}
\label{ch02:sec:single-problem}

The semantic definitions above determine whether a given infinite path satisfies
a given task formula. For verification, it is often sufficient to identify one
counterexample in order to show that a specification is not universally
satisfied. The synthesis problem is different. In that case, a concrete path
must be computed and later executed through a continuous control law of the form
given in \eqref{ch02:eq:control}. For this reason, the desired path should
admit a finite and manipulable representation.
A standard representation is the prefix-suffix form, also known as the lasso
form~\citep{schuppan2005shortest}. In this representation, an infinite plan is
decomposed into a transient part and a recurrent part. The transient part is
executed once, while the recurrent part is repeated indefinitely. Formally, the
plan is written as:
\begin{equation}
\tau \triangleq \tau_{\textup{pre}}
[\tau_{\textup{suf}}]^{\omega},
\label{ch02:eq:prefix-suffix}
\end{equation}
where $\tau_{\textup{pre}}$ is a finite prefix executed once, and
$\tau_{\textup{suf}}$ is a finite suffix repeated infinitely.
The representation in \eqref{ch02:eq:prefix-suffix} is important for several
reasons. It provides a finite description of an infinite execution, which is
essential for computation. It also anticipates the accepting structure of the
product automaton introduced later in the chapter, where accepting runs admit
precisely this form. In addition, the decomposition separates transient
behavior from recurrent behavior, which is useful when analyzing feasibility,
execution cost, and control implementation.

\begin{definition}
\label{ch02:def:feasible}
An LTL formula $\varphi$ is said to be feasible on $\mathcal{T}_c$ if there
exists an infinite path $\tau$ of $\mathcal{T}_c$ such that
$\tau \models \varphi$.
\end{definition}

Definition~\ref{ch02:def:feasible} characterizes the existence of a symbolic
solution at the level of the transition system. If $\varphi$ is infeasible,
then no trace generated by $\mathcal{T}_c$ belongs to
$\texttt{Words}(\varphi)$. If $\varphi$ is feasible, the remaining
questions are computational: a satisfying path must be synthesized, and that
path must then be implemented in the continuous workspace through hybrid
control.
The nominal planning problem addressed in this chapter can now be stated as
follows.

\begin{problem}
\label{ch02:prob:nominal}
Given the control-driven weighted transition system $\mathcal{T}_c$ and an LTL
formula $\varphi$ over $AP$, determine a plan of the form
\eqref{ch02:eq:prefix-suffix} such that $\tau \models \varphi$, and construct a
hybrid control strategy based on \eqref{ch02:eq:control} that executes the
resulting symbolic plan in the continuous workspace.
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Hybrid Controller Synthesis}
\label{ch02:sec:solution}

This section develops the nominal solution to
Problem~\ref{ch02:prob:nominal}. The construction follows the standard
automata-theoretic synthesis pipeline, specialized to the control-driven
weighted transition system $\mathcal{T}_c$ introduced in
Section~\ref{ch02:subsec:FTS}. The development proceeds in four steps. First,
the task formula is translated into a Nondeterministic B\"uchi automaton.
Second, this automaton is composed with $\mathcal{T}_c$ to obtain a weighted
product automaton. Third, an accepting run with prefix-suffix structure is
computed by graph search. Finally, the resulting symbolic plan is executed by a
hybrid controller in the continuous workspace.

\subsection{B\"uchi automaton associated with the task formula}
\label{ch02:subsec:NBA}

Given an LTL formula $\varphi$ over the proposition set $AP$, there exists a
Nondeterministic B\"uchi automaton that accepts exactly the infinite words that
satisfy $\varphi$. This automaton provides the link between temporal logic and
algorithmic synthesis.

\begin{definition}
\label{ch02:def:NBA}
The Nondeterministic B\"uchi automaton associated with $\varphi$ is denoted by
$\mathcal{A}_{\varphi}$ and defined by:
\begin{equation}
\mathcal{A}_{\varphi} \triangleq
\left(Q,\,2^{AP},\,\delta,\,Q_0,\,\mathcal{F}\right),
\label{ch02:eq:NBA}
\end{equation}
where $Q$ is a finite set of automaton states, $2^{AP}$ is the input alphabet,
$\delta : Q \times 2^{AP} \to 2^{Q}$ is the transition map,
$Q_0 \subseteq Q$ is the set of initial states, and
$\mathcal{F} \subseteq Q$ is the set of accepting states.
\end{definition}

The automaton in \eqref{ch02:eq:NBA} evolves over infinite words on the
alphabet $2^{AP}$. Each symbol of such a word records the set of atomic
propositions satisfied at one discrete step. Acceptance is defined by infinite
repetition of designated states, which is the standard acceptance mechanism for
B\"uchi automata.

\begin{definition}
\label{ch02:def:buchi_run}
Let
$\sigma \triangleq S_0S_1S_2\cdots \in (2^{AP})^{\omega}$.
A resulting run of $\sigma$ in $\mathcal{A}_{\varphi}$ is an infinite sequence
$r_{\sigma} \triangleq q_0q_1q_2\cdots$ such that
$q_0 \in Q_0$
and
$q_{k+1} \in \delta(q_k,S_k)$
for every $k \geq 0$. The run is accepting if
$\texttt{Inf}(r_{\sigma}) \cap \mathcal{F} \neq \emptyset$,
where $\texttt{Inf}(r_{\sigma})$ denotes the set of states that appear
infinitely often in $r_{\sigma}$.
\end{definition}

Because $\mathcal{A}_{\varphi}$ is nondeterministic, a single word may induce
several resulting runs. The accepted language is therefore defined by the
existence of at least one accepting run.

\begin{definition}
\label{ch02:def:buchi_language}
The accepted language of $\mathcal{A}_{\varphi}$ is
$\mathcal{L}_{\omega}(\mathcal{A}_{\varphi})
\triangleq
\{
\sigma \in (2^{AP})^{\omega}
\,|\,
\text{some resulting run}$
$\text{ of $\sigma$ in $\mathcal{A}_{\varphi}$ is accepting}\}$.
\end{definition}


\begin{lemma}
\label{ch02:lem:LTL-BA}
The accepted language of $\mathcal{A}_{\varphi}$ coincides with the language of
$\varphi$, namely
$\mathcal{L}_{\omega}(\mathcal{A}_{\varphi})
= \texttt{Words}(\varphi)$.
\end{lemma}

\begin{proof}
The statement follows from the standard translation theorem from LTL to
B\"uchi automata. A proof is given in Theorem~5.37 of
\citep{baier2008principles}.
\end{proof}

The translation from LTL to B\"uchi automata has worst-case exponential
complexity in the size of the formula~\citep{baier2008principles}. In practice,
optimized translation procedures often generate substantially smaller automata.
A widely used construction is given in~\citep{gastin2001fast}. In such
implementations, transitions are typically labeled by Boolean formulas over
$AP$, rather than by explicitly listing all symbols in $2^{AP}$. This symbolic
encoding is important when the proposition set is nontrivial. Binary decision
diagrams provide an efficient representation for such Boolean
labels~\citep{akers1978binary}.


\begin{example}
\label{ch02:examp:NBA}
Consider the specification
$\varphi = (\square\Diamond a_1) \wedge (\square\Diamond a_2)
\wedge (\square\Diamond a_3) \wedge (\square \neg a_4)$.
A corresponding automaton is shown in Figure~\ref{ch02:fig:Buchi}. If an edge
from $q_1$ to $q_3$ is labeled by the Boolean expression
$a_2 \wedge \neg a_4$, then every symbol in $2^{AP}$ satisfying that
expression enables the same transition. Hence a single symbolic label may
represent several concrete alphabet symbols.
\end{example}

\subsection{Weighted product automaton}
\label{ch02:subsec:product}

The transition system $\mathcal{T}_c$ captures feasible robot motion in the
workspace, while the automaton $\mathcal{A}_{\varphi}$ captures the symbolic
task language. Their product yields a single model whose accepting runs encode
robot behaviors that satisfy the specification.

\begin{definition}
\label{ch02:def:product}
The weighted product B\"uchi automaton of $\mathcal{T}_c$ and
$\mathcal{A}_{\varphi}$ is denoted by
$\mathcal{A}_{p} \triangleq \mathcal{T}_c \otimes \mathcal{A}_{\varphi}$ and
defined by:
\begin{equation}
\mathcal{A}_{p}
\triangleq
\left(Q',\,\delta',\,Q_0',\,\mathcal{F}',\,w_p\right),
\label{ch02:eq:product}
\end{equation}
where
$Q' \triangleq \Pi \times Q$,
$Q_0' \triangleq \Pi_0 \times Q_0$,
$\mathcal{F}' \triangleq \Pi \times \mathcal{F}$, and
$\delta' : Q' \to 2^{Q'}$ satisfies
\begin{equation}
\langle \pi_j,q_n \rangle \in \delta'(\langle \pi_i,q_m \rangle)
\iff
\pi_i \longrightarrow_c \pi_j
\text{ and }
q_n \in \delta\!\left(q_m,L_c(\pi_i)\right),
\label{ch02:eq:product_transition}
\end{equation}
where the transition weight is defined by
\begin{equation}
w_p\bigl(\langle \pi_i,q_m \rangle,\langle \pi_j,q_n \rangle\bigr)
\triangleq
w_c(\pi_i,\pi_j),
\label{ch02:eq:product_weight}
\end{equation}
for every enabled transition in \eqref{ch02:eq:product_transition}.
\end{definition}

Equation~\eqref{ch02:eq:product_transition} shows that a product transition is
enabled only if two conditions hold simultaneously. The first condition is that
the robot can move from $\pi_i$ to $\pi_j$ in the transition system. The
second condition is that the task automaton can evolve from $q_m$ to $q_n$
under the label produced at $\pi_i$. The weight assignment in
\eqref{ch02:eq:product_weight} preserves the motion cost from $\mathcal{T}_c$,
which allows optimality to be analyzed directly at the product level.

\begin{definition}
\label{ch02:def:projection}
Let $q' = \langle \pi,q \rangle \in Q'$. The projection of $q'$ onto the
transition-system component is defined by
$q'|_{\Pi} \triangleq \pi$, and the projection onto the automaton component is
defined by
$q'|_{Q} \triangleq q$.
Given an infinite run
$R = q_0'q_1'q_2'\cdots$ of $\mathcal{A}_p$, its projected sequences are
denoted by
$R|_{\Pi} \triangleq q_0'|_{\Pi}q_1'|_{\Pi}q_2'|_{\Pi}\cdots$
and
$R|_{Q} \triangleq q_0'|_{Q}q_1'|_{Q}q_2'|_{Q}\cdots$.
\end{definition}

The product construction is the key step in synthesis. It converts the original
problem into a graph-search problem over $\mathcal{A}_p$. An accepting run in
the product automaton projects to a plan of $\mathcal{T}_c$ that satisfies the
task formula.

\begin{lemma}
\label{ch02:lem:satisfy}
If $R$ is an accepting run of $\mathcal{A}_p$, then
$R|_{\Pi} \models \varphi$.
\end{lemma}

\begin{proof}
Since $R$ is accepting, some state in $\mathcal{F}'$ appears infinitely often
in $R$. By Definition~\ref{ch02:def:projection}, the projection of that state
onto $Q$ belongs to $\mathcal{F}$. Hence $R|_{Q}$ is an accepting run of
$\mathcal{A}_{\varphi}$. By \eqref{ch02:eq:product_transition}, the run
$R|_{Q}$ is generated by the word
$\texttt{trace}(R|_{\Pi})$. Therefore,
$\texttt{trace}(R|_{\Pi})
\in \mathcal{L}_{\omega}(\mathcal{A}_{\varphi})$.
Applying Lemma~\ref{ch02:lem:LTL-BA} yields
$\texttt{trace}(R|_{\Pi}) \models \varphi$, which implies
$R|_{\Pi} \models \varphi$ by
Definition~\ref{ch02:def:path-satisfy}.
\end{proof}

Conversely, if there exists a path of $\mathcal{T}_c$ whose trace satisfies
$\varphi$, then the product automaton is nonempty in the accepting sense. This
fact underlies automata-based synthesis, but nonemptiness alone is not
sufficient here. An accepting run with suitable structure must also be
computed.

\subsection{Optimal accepting run search}
\label{ch02:subsec:optimality}

The plan in Problem~\ref{ch02:prob:nominal} is required to have prefix-suffix
structure. Since the product automaton is finite, every accepting behavior can
be represented by an ultimately periodic accepting run. Hence the search may be
restricted to lasso-shaped runs.

\begin{lemma}
\label{ch02:lem:prefix-suffix}
If there exists an infinite path $\tau$ of $\mathcal{T}_c$ with prefix-suffix
structure such that $\tau \models \varphi$, then there exists an accepting run
of $\mathcal{A}_p$ with prefix-suffix structure.
\end{lemma}

\begin{proof}
Since $\tau \models \varphi$, the word
$\sigma \triangleq \texttt{trace}(\tau)$ belongs to
$\texttt{Words}(\varphi)$. By Lemma~\ref{ch02:lem:LTL-BA},
$\sigma \in \mathcal{L}_{\omega}(\mathcal{A}_{\varphi})$. Therefore,
$\sigma$ induces an accepting run of $\mathcal{A}_{\varphi}$. Pairing the path
$\tau$ with such a resulting automaton run through
\eqref{ch02:eq:product_transition} yields a run of $\mathcal{A}_p$. Since both
components are ultimately periodic, the product run is also ultimately
periodic. Because the run is accepting, some accepting product state is visited
infinitely often and can be selected as the root of the suffix cycle.
\end{proof}

Accordingly, the search is restricted to accepting runs of the form below:
\begin{equation}
R \triangleq
q_0' q_1' \cdots q_{f-1}'
\left[q_f' q_{f+1}' \cdots q_n'\right]^{\omega},
\label{ch02:eq:prod-Apath}
\end{equation}
where $q_0' \in Q_0'$,
$q_f' \in \mathcal{F}'$, and the transition from $q_n'$ to $q_f'$ is enabled.
The representation in \eqref{ch02:eq:prod-Apath} provides a finite description
of an infinite accepting run. The prefix from $q_0'$ to $q_f'$ is executed
once. The suffix beginning at $q_f'$ is then repeated indefinitely. This is
precisely the form needed for finite optimization and subsequent hybrid
execution.

\begin{definition}
\label{ch02:def:run_cost}
Let $R$ be an accepting run of the form
\eqref{ch02:eq:prod-Apath}. Its total cost is defined by:
\begin{equation}
J(R)
\triangleq
\sum_{i=0}^{f-1} w_p(q_i',q_{i+1}')
+
\gamma
\left(
\sum_{i=f}^{n-1} w_p(q_i',q_{i+1}')
+
w_p(q_n',q_f')
\right),
\label{ch02:eq:prod-cost}
\end{equation}
where $\gamma \geq 0$ is a design parameter that specifies the relative weight
of the suffix cost with respect to the prefix cost.
\end{definition}

The first term in \eqref{ch02:eq:prod-cost} measures the transient cost of
reaching an accepting state. The second term measures the cost of one full
traversal of the recurrent suffix cycle. Consequently, larger values of
$\gamma$ place greater emphasis on long-term recurring performance.

\begin{problem}
\label{ch02:pro:opt-search}
Given the product automaton $\mathcal{A}_p$, determine an accepting run of the
form \eqref{ch02:eq:prod-Apath} that minimizes the cost in
\eqref{ch02:eq:prod-cost}.
\end{problem}

The search problem above can be decomposed into two finite graph problems. The
prefix corresponds to a shortest path from an initial product state to an
accepting product state. The suffix corresponds to a shortest cycle rooted at
that accepting state. Since all transition weights are nonnegative,
Dijkstra-type search can be used. Two construction strategies are relevant for
the product automaton. Under full construction, all states and transitions of
$\mathcal{A}_p$ are generated in advance by
Definition~\ref{ch02:def:product}. Under on-the-fly construction, the adjacency
relation of a product state is generated only when that state is visited during
search. The latter strategy is particularly useful when large parts of the
product graph are irrelevant to the optimal accepting run.
Let $S' \subseteq Q_0'$ denote the set of initial product states considered
by the search. For each starting state $q_s' \in S'$ and each accepting state
$q_f' \in \mathcal{F}'$, let
$d_{\textup{pre}}(q_s',q_f')$
denote the cost of a shortest path from $q_s'$ to $q_f'$ in $\mathcal{A}_p$.
Let
$d_{\textup{suf}}(q_f')$
denote the cost of a shortest cycle rooted at $q_f'$. Let
$R_{\textup{pre}}(q_s',q_f')$
and
$R_{\textup{suf}}(q_f')$
denote the corresponding prefix path and suffix cycle, respectively. Then the
cost of the accepting run induced by the pair $(q_s',q_f')$ is given by:
\begin{equation}
J(q_s',q_f')
\triangleq
d_{\textup{pre}}(q_s',q_f')
+
\gamma\, d_{\textup{suf}}(q_f'),
\label{ch02:eq:run_pair_cost}
\end{equation}
where $\gamma \geq 0$ is the weighting parameter introduced in
\eqref{ch02:eq:prod-cost}. The search problem is therefore reduced to finding a
pair $(q_s',q_f')$ that minimizes \eqref{ch02:eq:run_pair_cost}.

%==============================
%==============================
\begin{algorithm}[t]
\caption{Optimal accepting run search, \texttt{OptRun}().}
\label{ch02:alg:optimal}
\DontPrintSemicolon
\KwIn{$\mathcal{A}_p$, $S' \subseteq Q_0'$, $\gamma \geq 0$}
\KwOut{$R_{\textup{opt}}$ or failure}

\If{$S'=\emptyset$ or $\mathcal{F}'=\emptyset$}{
  \Return{failure}\;
}

$J^\star \leftarrow +\infty$\;
$R_{\textup{opt}} \leftarrow \emptyset$\;

\ForEach{$q_f' \in \mathcal{F}'$}{
  $(d_{\textup{suf}}(q_f'),\, R_{\textup{suf}}(q_f'))
  \leftarrow
  \texttt{DijksCycle}(\mathcal{A}_p, q_f')$\;
}

\ForEach{$q_s' \in S'$}{
  $(d_{\textup{pre}}(q_s',\cdot),\, R_{\textup{pre}}(q_s',\cdot))
  \leftarrow
  \texttt{DijksTargets}(\mathcal{A}_p, q_s', \mathcal{F}')$\;
  \ForEach{$q_f' \in \mathcal{F}'$}{
    \If{$d_{\textup{{pre}}}(q_s',q_f') < +\infty$
    \textup{ and }
    $d_{\textup{{suf}}}(q_f') < +\infty$}{
      $J \leftarrow d_{\textup{pre}}(q_s',q_f')
      + \gamma\, d_{\textup{suf}}(q_f')$\;
      \If{$J < J^\star$}{
        $J^\star \leftarrow J$\;
        $R_{\textup{opt}} \leftarrow
        R_{\textup{pre}}(q_s',q_f')
        \bigl[R_{\textup{suf}}(q_f')\bigr]^{\omega}$\;
      }
    }
  }
}

\If{$R_{\textup{opt}}=\emptyset$}{
  \Return{failure}\;
}
\Return{$R_{\textup{opt}}$}\;
\end{algorithm}
%==============================
%==============================

Algorithm~\ref{ch02:alg:optimal} is illustrated in
Figure~\ref{ch02:fig:accPath}. The algorithm first computes the shortest suffix
cycle for each accepting state in $\mathcal{F}'$. It then computes, for each
$q_s' \in S'$, the shortest prefix paths from $q_s'$ to all reachable
accepting states. Each feasible pair $(q_s',q_f')$ defines a candidate lasso
run with cost given by \eqref{ch02:eq:run_pair_cost}. The optimal accepting run
$R_{\textup{opt}}$ is obtained by selecting the minimizing pair. The
corresponding symbolic plan is
$\tau_{\textup{opt}} \triangleq R_{\textup{opt}}|_{\Pi}$.


\begin{remark}
\label{ch02:rem:optimality_loss}
The projected plan $\tau_{\textup{opt}}$ is optimal with respect to the search
performed on $\mathcal{A}_p$, but it need not coincide with the globally
optimal satisfying plan in $\mathcal{T}_c$. This possible loss of optimality is
caused by the translation from LTL to a nondeterministic B\"uchi automaton.
This phenomenon is discussed in~\citep{schuppan2005shortest}. Tighter automaton
constructions may preserve optimality more faithfully, but such constructions
typically produce larger automata and hence higher computational cost.
\end{remark}


\subsection{Hybrid execution of the optimal run}
\label{ch02:sec:hybrid}

The output of Algorithm~\ref{ch02:alg:optimal} is a symbolic accepting run in
the product automaton. This run must then be realized in the continuous
workspace by the navigation controller introduced in
Definition~\ref{ch02:def:transition}. The role of the hybrid controller is to
activate the appropriate local controller for each consecutive pair of regions
in the projected plan.
Let the optimal accepting run be written as below:
\begin{equation}
R_{\textup{opt}}
\triangleq
R_{\textup{pre}}
\left[R_{\textup{suf}}\right]^{\omega},
\label{ch02:eq:final-path}
\end{equation}
where $R_{\textup{pre}}$ is the finite prefix and $R_{\textup{suf}}$ is the
finite suffix cycle.
The projected motion plan is
$\tau_{\textup{opt}} \triangleq R_{\textup{opt}}|_{\Pi}$. Execution begins with
the prefix segment. After the prefix is completed, the execution switches to
the suffix segment and then repeats the suffix indefinitely. Thus the hybrid
controller realizes exactly the prefix-suffix structure required in
\eqref{ch02:eq:prefix-suffix}.

For implementation, the current position within $R_{\textup{opt}}$ is indexed
by a segment variable and an integer index. A helper function
$\texttt{NextGoal}(\cdot)$ returns the next product state in the
accepting run according to this indexing rule. The next target region is then
obtained by projection onto $\Pi$, and the active continuous controller is
generated by \eqref{ch02:eq:control}.
Algorithm~\ref{ch02:alg:hybrid} runs indefinitely because the suffix of
$R_{\textup{opt}}$ is repeated infinitely often. The condition
$x(t) \in \pi_{\textup{next}}$ is evaluated by the same localization mechanism
used in the labeling function. Once the next region is reached, the controller
advances to the subsequent symbolic goal and activates the corresponding local
navigation controller.

%==============================
%==============================
\begin{algorithm}[t]
\caption{Plan execution by hybrid control, \texttt{HybCon}().}
\label{ch02:alg:hybrid}
\DontPrintSemicolon
\KwIn{$R_{\textup{opt}}$, current state measurement $x(t)$}
\KwOut{$u(t)$, $R_{\textup{past}}$, $\tau_{\textup{past}}$}
initialize
$q_{\textup{cur}}' = q_{\textup{next}}' = R_{\textup{pre},1}$,
$\pi_{\textup{cur}} = q_{\textup{cur}}'|_{\Pi}$,
$\pi_{\textup{next}} = q_{\textup{next}}'|_{\Pi}$,
$R_{\textup{past}} = [\;]$,
$\tau_{\textup{past}} = [\;]$\;
\While{\textup{\texttt{True}}}{
  \If{$x(t) \in \pi_{\textup{next}}$ is confirmed}{
    $q_{\textup{cur}}' = q_{\textup{next}}'$\;
    append $q_{\textup{cur}}'$ to $R_{\textup{past}}$\;
    append $\pi_{\textup{next}}$ to $\tau_{\textup{past}}$\;
    $q_{\textup{next}}' =
    \texttt{NextGoal}(q_{\textup{cur}}',R_{\textup{opt}})$\;
    $\pi_{\textup{cur}} = q_{\textup{cur}}'|_{\Pi}$\;
    $\pi_{\textup{next}} = q_{\textup{next}}'|_{\Pi}$\;
  }
  $u(t) =
  \mathcal{U}\bigl(x(t),\pi_{\textup{cur}},\pi_{\textup{next}}\bigr)$\;
}
\end{algorithm}
%==============================
%==============================


\subsection{Complexity discussion}
\label{ch02:sssec:complexity_on_the_fly}

The computational burden of the synthesis procedure is determined mainly by the
size of the task automaton and the size of the product automaton. Let
$n_{\varphi} \triangleq |\varphi|$
denote the size of the LTL formula. Standard translation bounds imply:
$|Q| = 2^{\mathcal{O}(n_{\varphi})}$ and $|\delta| = 2^{\mathcal{O}(n_{\varphi})}$,
where $Q$ is the automaton state set and $|\delta|$ denotes the number of
enabled automaton transitions in $\mathcal{A}_{\varphi}$ from
Definition~\ref{ch02:def:NBA}.
Once $\mathcal{T}_c$ and $\mathcal{A}_{\varphi}$ are fixed, the fully
constructed product automaton satisfies:
$|Q'| = |\Pi|\,|Q|$ and $|\delta'| = \mathcal{O}\bigl(|\longrightarrow_c|\,|\delta|\bigr)$,
where $|\Pi|$ and $|\longrightarrow_c|$ denote the numbers of states and
transitions in $\mathcal{T}_c$.

Under on-the-fly construction, only the portion of the product graph that is
actually visited during search is generated. Let
$Q'_{\textup{vis}} \subseteq Q'$
denote the set of visited product states, and let
$\delta'_{\textup{vis}} \subseteq \delta'$
denote the set of product transitions that are explicitly constructed. Then the
effective synthesis time is bounded by:
$
T_{\textup{otf}}
=
\mathcal{O}
\Bigl(
\bigl(|Q_0'| + |\mathcal{F}'_{\textup{vis}}|\bigr)
\,|\delta'_{\textup{vis}}|\,
\log |Q'_{\textup{vis}}|
\Bigr)$,
where
$\mathcal{F}'_{\textup{vis}} \triangleq \mathcal{F}' \cap Q'_{\textup{vis}}$
is the set of accepting product states that are actually generated.
The corresponding memory requirement is given by:
$
M_{\textup{otf}}
=\mathcal{O}\bigl(|Q'_{\textup{vis}}| + |\delta'_{\textup{vis}}|\bigr)$,
where the two terms account for the explicitly stored product states and
product transitions.
The worst-case asymptotic complexity of the on-the-fly method coincides with
that of full construction, because the search may in principle visit the entire
product graph. In practice, however, on-the-fly construction is often much more
efficient, since large portions of the product automaton may be unreachable or
irrelevant to the optimal accepting run. This distinction becomes especially
important in adaptive settings where the transition system must be updated
repeatedly.

\begin{example}
\label{ch02:exam:complexity}
Consider a family of grid-based transition systems with $n^2$ states. The
transition cost between adjacent regions is chosen uniformly from the interval
$[0,1]$. The task is to transport two objects to two designated destinations
and then return to a base region. The corresponding specification has the same
general form as the case-study formula introduced later in the chapter.
For each instance, three CPU times are recorded. The quantity
$t_{\textup{full}}$ denotes the time required to construct $\mathcal{A}_p$
fully. The quantity $t_{\textup{syn}}$ denotes the time required to search for
the optimal accepting run on the fully constructed product automaton. The
quantity $t_{\textup{fly}}$ denotes the time required when the product automaton
is generated on the fly during search.
The associated B\"uchi automaton has $75$ states and
$877$ transitions. Table~\ref{ch02:tab:complexity} indicates that, for this
task class, product construction dominates the total computation time, whereas
the graph search itself is relatively inexpensive. This observation explains
why symbolic and on-the-fly constructions are important in practice.
\end{example}


\subsection{Scalability in the multi-robot setting}
\label{ch02:sssec:multi_robot_scalability}

For a single robot, the above synthesis procedure remains tractable when the
workspace abstraction and the task automaton are moderate in size. The
situation changes significantly in the multi-robot setting if the same method
is applied in a centralized way.
Consider $N$ robots with local transition systems
$\mathcal{T}_i \triangleq
\left(\Pi_i,\longrightarrow_i,\Pi_{i,0},AP_i,L_i,w_i\right)$,
where $i \in \{1,\cdots,N\}$. A centralized team abstraction yields the joint
state space
$\Pi_{\textup{team}} \triangleq \prod_{i=1}^{N} \Pi_i$.
Hence,
$|\Pi_{\textup{team}}| = \prod_{i=1}^{N} |\Pi_i|$,
which grows multiplicatively with the number of robots.
If a team specification $\Phi$ is translated into a B\"uchi automaton
$\mathcal{A}_{\Phi}$ with state set $Q_{\Phi}$, then the corresponding
centralized product space has cardinality below:
\begin{equation}
|Q_{\textup{team}}'|
=
|Q_{\Phi}|
\prod_{i=1}^{N} |\Pi_i|\approx 2^{|\Phi|} M^N,
\label{ch02:eq:team-product-cardinality}
\end{equation}
where the multiplicative growth arises from both the joint motion state and the
task automaton;
and $M$ is the size of homogeneous $\mathcal{T}_i$.
Equation~\eqref{ch02:eq:team-product-cardinality} explains why a direct
centralized extension of the single-robot method quickly becomes
computationally prohibitive. Even when each local model is small, the combined
team model can be too large for explicit product construction and exhaustive
graph search. This observation motivates the decentralized and compositional
developments presented in the later chapters. Similar scalability issues have
been reported in the multi-robot temporal-logic literature
\citep{kloetzer2010automatic,chen2012formal,kloetzer2011multi,guo2015multi}.

%==============================
%==============================
%==============================
\section{Case Study}
\label{ch02:sec:case_study}


 %==============================
\section{Summary}
\label{ch02:sec:ch02_summary}
