\chapter{Bottom-Up Coordination under Local Temporal Tasks}
\label{chap:bottom-up}

The previous chapter showed that centralized automata-based planning for
multi-robot systems faces a basic scalability limitation. When a global
temporal specification is synchronized with the joint transition system of all
robots, the resulting product construction grows rapidly with both the team size
and the task complexity. This growth is not only a computational issue in
offline synthesis, but also a structural limitation when coordination must be
maintained under incomplete information, inter-robot dependence, and online
changes. For this reason, the present chapter adopts a bottom-up perspective on
multi-robot coordination, in which each robot is assigned a local temporal task
and coordination is established through relations among these local tasks rather
than through a single monolithic team specification.

From this viewpoint, the main question is how local temporal tasks can be
planned, coordinated, and revised while preserving the logical structure
required for correct team behavior. Section~\ref{ch03:sec:task_relaxation_adaptation}
first studies task relaxation and plan adaptation when workspace information is
incomplete and feasibility cannot be determined in advance. Then,
Section~\ref{ch03:sec:dep-local-collab} introduces dependent local tasks with
explicit collaborative actions, so that the execution of one robot may require
functional support from others. Together, these
developments extend the discrete planning framework from isolated local task
execution to structured multi-robot coordination under dynamic and uncertain
conditions.

%==============================
%==============================
\input{chapters/ch03/ch03_sec01}
%==============================
%==============================
\input{chapters/ch03/ch03_sec02}
%==============================
%==============================
%\input{chapters/ch03/ch03_sec03}
%==============================
%==============================
\input{chapters/ch03/ch03_sec04}
