\chapter{Top-Down Synthesis for Global Temporal Tasks}
\label{chap:top-down}

The previous Chapter~\ref{chap:bottom-up} considered a bottom-up coordination framework in which
temporal tasks are assigned locally and coordination emerges through
execution-time dependence among the robots. That perspective is natural when
task coupling is sparse and collaboration is triggered only when local plans
intersect through shared actions, resources, or newly acquired information.
In many applications, however, the mission is specified from the outset at the
level of the whole team. The central problem is then no longer how local tasks
can be coordinated after they are assigned, but how a global temporal
specification should be decomposed, distributed, and refined so that the
resulting robot behaviors remain jointly correct. This top-down setting is
conceptually complementary to the bottom-up perspective of the previous
chapter, but it raises a different technical difficulty. The logical structure
of the team task must be preserved during decomposition and assignment, while
the resulting planning problem must remain computationally manageable for
heterogeneous teams and large workspaces.

This chapter develops a top-down synthesis framework for such global temporal
tasks. Section~\ref{ch04:sec:po_decomp_assignment} begins with
partial-order-based task decomposition and assignment, where the accepted
behavior of the global task automaton is converted into subtask occurrences
together with a compact temporal-dependence structure, which then supports
efficient allocation over the robot team. Section~\ref{ch04:subsec:task}
extends this perspective to hierarchical coordination of large-scale fleets,
where abstraction across multiple planning layers is used to reduce
computational growth while preserving mission-level consistency.
Section~\ref{ch04:sec:online_continual_tasks} then addresses online adaptation
for continual tasks, in which new demands may arise during execution and the
team plan must be revised without losing the global temporal structure. The
case study and summary at the end illustrate how these
components fit together within a unified top-down coordination framework.


%==============================
%==============================
\input{chapters/ch04/ch04_sec01}
%==============================
%==============================
\input{chapters/ch04/ch04_sec02}
%==============================
%==============================
\input{chapters/ch04/ch04_sec03}
%==============================
%==============================
\input{chapters/ch04/ch04_sec04}
