\chapter{Partial Knowledge and Limited Communication}
\label{chap:communication}


The preceding chapters developed coordination frameworks for multi-robot
systems under local and global temporal tasks, under the assumption that the
information required for planning and coordination can be obtained with
sufficient accuracy and timeliness. In many deployments, however, this
assumption is difficult to maintain. Robots often operate with only partial
knowledge of the workspace, communication links may be range-limited or
intermittent, and task-relevant information may become available only during
execution. Under such conditions, temporal-task planning is no longer concerned
only with motion, action, and allocation. It must also account for how
information is acquired, propagated, and used across the team. Communication is
therefore not merely a supporting mechanism for coordination, but an integral
part of the planning problem itself.

This chapter examines multi-robot coordination under partial knowledge and
limited communication from that perspective. In
Section~\ref{ch05:sec:ch05_interm_team}, communication and collaboration are
planned jointly, so that the robot team coordinates not only task execution but
also the information exchange required to support it. In
Section~\ref{ch05:sec:intermittent_comm}, the discussion turns to exploration and
coordination under intermittent communication, where continuous connectivity
cannot be assumed and planning must proceed with delayed or sparse information
sharing. Section~\ref{ch05:sec:ch05_human} addresses online human interaction
and request fulfillment, where new requests may be introduced during execution
and must be incorporated without discarding the underlying formal task
structure. The chapter then presents a case study in
Section~\ref{sec:ch05_case_study} and concludes with a summary in
Section~\ref{sec:ch05_summary}.

%==============================
%==============================
\input{chapters/ch05/ch05_sec01}
%==============================
%==============================
\input{chapters/ch05/ch05_sec02}
%==============================
%==============================
\input{chapters/ch05/ch05_sec03}
%==============================
%==============================
\input{chapters/ch05/ch05_sec04}
