\chapter{Dynamic Uncertainty and Constraints}
\label{chap:uncertainty}

Chapters~\ref{chap:bottom-up}, \ref{chap:top-down}, and
\ref{chap:communication} established coordination frameworks for temporal-task
planning under local interactions, global specifications, and limited
communication. In many robotic deployments, however, the main difficulty is not
only combinatorial coordination, but also the fact that the environment is
partially uncertain and the operative constraints evolve during execution.
Relevant propositions may be stochastic, targets may move, feasible motions may
change with time, and safety may depend on the continued availability of
recovery or return actions rather than on simple state avoidance alone.
Chapter~\ref{chap:uncertainty} addresses this next layer of difficulty by
studying temporal-task planning and coordination when uncertainty, dynamic
constraints, and reactive adaptation must be incorporated directly into the
decision process.

The chapter is organized around three complementary problem classes.
Section~\ref{ch06:sec:ch06_safe_plan} considers probabilistic temporal tasks
with safe-return requirements, where task fulfillment must be balanced against
the ability to retreat to designated safe states. Section~\ref{ch06:sec:ch06_uncertain}
then studies reactive coordination under dynamic
temporal-logic tasks, with particular emphasis on prediction, online
reassignment, and adaptation to evolving targets and mission conditions.
Section~\ref{ch06:sec:ch06_stl} turns to continuous multi-robot motion planning
under signal temporal logic specifications and dynamic couplings, using
collaborative optimal transport to handle nonlinear trajectory constraints. The
resulting methods are illustrated through the case study in
Section~\ref{ch06:sec:case}.

%==============================
%==============================
\input{chapters/ch06/ch06_sec01}
%==============================
%==============================
\input{chapters/ch06/ch06_sec02}
%==============================
%==============================
\input{chapters/ch06/ch06_sec03}
%==============================
%==============================
\input{chapters/ch06/ch06_sec04}
