\chapter{Outlook and Open Problems}
\label{chap:outlook}

The preceding chapters developed a progression of temporal-task planning and
coordination frameworks, beginning with single-robot automata-based synthesis
in Chapter~\ref{chap:preliminaries}, extending to bottom-up and top-down
multi-robot coordination in Chapters~\ref{chap:bottom-up}
and~\ref{chap:top-down}, and then addressing partial knowledge, limited
communication, and dynamic uncertainty in Chapters~\ref{chap:communication}
and~\ref{chap:uncertainty}. Taken together, these developments establish a
coherent methodological foundation, but they also reveal several unresolved
difficulties that cannot be treated as isolated technical refinements. As task
models become richer, environments less predictable, and decision processes
more distributed, the main obstacles shift from nominal task satisfaction on
fixed abstractions to broader questions of scalability, adaptation, semantic
robustness, and system-level correctness.

Rather than revisiting the earlier chapters one by one, this chapter isolates
five lines of difficulty that cut across the entire monograph. Section
\ref{sec:ch07_scalability} examines scalability and compositional synthesis
beyond monolithic constructions. Section~\ref{sec:ch07_continual} turns to
continual and reactive coordination in nonstationary environments, where
replanning is intrinsic rather than exceptional. Section
\ref{sec:ch07_open_world} addresses uncertainty, partial knowledge, and
open-world semantics, with emphasis on the limits of fixed symbolic models.
Section~\ref{sec:ch07_learning} considers how learning-based components may be
incorporated without dissolving meaningful formal guarantees. Finally, Section
\ref{sec:ch07_deployment} discusses verification, human interaction, and
deployment-level evaluation, where the practical validity of temporal-task
planning frameworks is ultimately tested.

%=================================================
\section{Scalability and Compositional Synthesis}
\label{sec:ch07_scalability}

A recurring theme throughout this monograph is that the expressive strength of
formal temporal-task planning is inseparable from its computational burden. In
Chapter~\ref{chap:preliminaries}, the product construction between robot
dynamics and task automata already reveals a basic combinatorial growth, even
before multi-robot coupling is introduced. Chapters~\ref{chap:bottom-up}
and~\ref{chap:top-down} show that this difficulty does not disappear when the
problem is distributed across multiple robots. Instead, it reappears through
dependence among local tasks, through assignment and decomposition under
shared temporal constraints, and through the need to preserve consistency
between local decisions and global task semantics. Chapters
\ref{chap:communication} and~\ref{chap:uncertainty} further indicate that once
communication limits, online requests, uncertainty, and dynamic constraints
are incorporated, scalability is no longer a question of state count alone. It
also depends on how formal structure is preserved across multiple layers, time
scales, and information regimes.

\textbf{(I) Limits of monolithic synthesis.}
One unresolved issue concerns the precise boundary beyond which monolithic
planning architectures cease to be useful. Product automata, joint transition
systems, and centralized optimization remain conceptually clean, but their
growth becomes difficult to control as the team size, task horizon, or
workspace complexity increases. The natural response is to seek
decomposition, yet the central question is not decomposition in itself. The
question is whether a decomposition exposes enough structure to reduce
complexity without discarding the logical dependencies that determine
correctness. In some settings, temporal coupling is sparse only locally or
only over part of the execution horizon. In others, coupling is dense in
principle but weak in effect, in the sense that only a small subset of
interactions is critical for feasibility or performance. A mature theory of
scalable synthesis therefore requires more than asymptotic complexity bounds.
It requires structural criteria that identify when a temporal-task problem
admits a compact compositional representation and when such a representation
is fundamentally unavailable.

\textbf{(II) Correctness-preserving decomposition.}
A second difficulty lies in the certification of decomposition itself. The
methods in Chapters~\ref{chap:bottom-up} and~\ref{chap:top-down} suggest that
useful structure can be extracted through local tasks, dependency relations,
relaxed orderings, and hierarchical abstractions. Yet the general problem
remains unsettled. When a global task is decomposed into subtasks, which
logical properties must be preserved at each interface between subproblems?
Which forms of temporal dependence can be relaxed, approximated, or postponed
without altering the admissible team behaviors in a harmful way? Conversely,
which dependencies are non-separable and therefore require centralized
reasoning or explicit synchronization? These questions become sharper in
heterogeneous teams, where robots differ not only in motion capabilities but
also in sensing, communication range, endurance, and access to task-relevant
actions. What is still missing is a unified notion of compositional soundness
that applies across symbolic decomposition, task assignment,
communication-aware coordination, and continuous trajectory coupling.

\textbf{(III) Cross-layer abstraction and refinement.}
A related challenge concerns the interaction among planning layers. Several
chapters exploit abstraction as a practical means of controlling complexity,
but abstraction is rarely static in realistic deployments. Discrete task
planning, motion planning, communication planning, and uncertainty handling
often operate on different models and at different resolutions. The difficulty
is therefore not only how to abstract, but how to refine and revise
abstractions when the assumptions under which they were constructed cease to
hold. This raises a broader question: what kinds of interface variables or
contracts are sufficient to connect symbolic task logic with geometric motion,
resource constraints, and information flow in a way that remains tractable
under revision? A satisfactory answer would support cross-layer synthesis in
which each layer can evolve without forcing complete recomputation of the
entire coordination policy.

\textbf{(IV) Scalability under continual revision.}
The problem of scale also changes once planning is no longer a one-shot
operation. When tasks arrive online, targets move, constraints change, or
communication becomes intermittent, a coordination architecture may be
computationally acceptable in an offline sense and still be unusable in
practice. The relevant question is then no longer the cost of solving a
single planning problem, but the cost of maintaining logical consistency
across a sequence of related problems. This suggests the need for
incremental, warm-started, or anytime variants of formal synthesis and
coordination, together with principled criteria for what information can be
reused safely across replanning stages. In that sense, scalability should be
understood not as a narrow implementation concern, but as a structural
question about how logical, dynamical, and informational couplings interact
over time. For that reason, scalability and compositional synthesis remain one
of the central open problems for temporal-task planning as a whole.

%=================================================
\section{Continual and Reactive Coordination in Nonstationary Environments}
\label{sec:ch07_continual}

If Section~\ref{sec:ch07_scalability} concerns the growth of the planning
problem, the next difficulty concerns the persistence of change during
execution. Several developments in this monograph already point in that
direction. Chapter~\ref{chap:top-down} considers continual tasks whose
requirements may evolve during execution. Chapter~\ref{chap:communication}
shows that communication opportunities, information availability, and external
requests may all change online. Chapter~\ref{chap:uncertainty} further
demonstrates that targets, constraints, and safety conditions may vary in
ways that directly affect the feasibility and quality of a current plan.
These examples suggest that continual revision is not merely an implementation
detail added after synthesis. In many realistic deployments, it is the
dominant operating condition. The central challenge is therefore to
understand how formal temporal-task planning should be reformulated when task
evolution, environmental change, and online coordination are treated as
intrinsic features of the problem.

\textbf{(I) Task persistence and continual arrival.}
Many missions do not terminate cleanly after the satisfaction of a fixed
specification. New subtasks may arrive while earlier commitments are still
being executed, temporary priorities may need to be elevated, and part of the
team may already be engaged in actions whose interruption is costly or unsafe.
Under such conditions, the planning problem is no longer adequately captured
by a single synthesis stage followed by execution. Instead, coordination must
account for temporal overlap among completed, ongoing, and newly introduced
obligations. This creates a theoretical gap between classical task
satisfaction and continual task management. What is needed is a richer notion
of feasibility that accounts not only for whether a new task can eventually be
incorporated, but also for whether it can be incorporated without invalidating
the logical and operational commitments that are already in force. More
generally, the field still lacks a systematic framework for representing
temporal-task queues, persistent mission contexts, and controlled revision of
priorities at the level of formal specification.

\textbf{(II) Reactive replanning under changing targets and constraints.}
A second source of difficulty lies in the triggers for replanning. In
nonstationary environments, revision is driven not only by new task requests,
but also by changes in targets, workspace accessibility, communication
topology, resource availability, and predicted risk. A coordination strategy
that is correct for a nominal model may therefore become suboptimal, overly
conservative, or even infeasible after the environment changes. The issue is
not simply how to recompute a new plan after such a change is detected. The
deeper issue is how to represent environmental variation in a form that can be
incorporated into temporal-task reasoning without forcing complete
reconstruction of the planning problem at every update. This includes the
question of which changes should be handled by local repair, which require
global reassignment, and which invalidate the underlying symbolic abstraction
itself. Existing methods often exploit specific structures, such as moving
targets, dynamic couplings, or safe-return conditions, but a general theory
of task-level replanning under heterogeneous forms of nonstationarity is still
missing.

\textbf{(III) Correctness, stability, and performance under repeated adaptation.}
Once adaptation becomes continual, the relevant guarantees also change. In
one-shot planning, correctness is usually expressed in terms of task
satisfaction, optimality, or feasibility with respect to a fixed model. Under
repeated adaptation, these notions are no longer sufficient by themselves. A
sequence of individually valid replanning steps may still produce undesirable
collective behavior, such as oscillatory task reassignment, excessive
sensitivity to transient disturbances, or progressive degradation of
long-horizon mission quality. For this reason, the relevant guarantees must be
expanded to include stability of coordination, bounded disruption to ongoing
execution, and performance criteria that reflect cumulative behavior over
extended horizons. This raises several unresolved questions. How should
logical progress be measured when the specification or environment is itself
evolving? What notion of robustness is appropriate when the main failure mode
is not immediate infeasibility, but repeated loss of coherence across
replanning stages? Which performance objectives should take precedence when
adaptation improves short-term feasibility but degrades global mission
efficiency or interpretability? These questions indicate that the analysis of
reactive coordination must move beyond isolated replanning events toward a
longer-horizon view in which correctness, stability, and performance are
evaluated jointly.

%=================================================
\section{Uncertainty, Partial Knowledge, and Open-World Semantics}
\label{sec:ch07_open_world}

Nonstationarity concerns change over time; a different challenge concerns
incompleteness of the model itself. Much of the formal synthesis literature,
including several constructions used throughout this monograph, relies on a
closed-world view of the planning problem: the relevant states, propositions,
task predicates, and transition relations are assumed to be specified in
advance with sufficient precision to support symbolic reasoning. That
assumption is methodologically useful and often unavoidable, but its
limitations become increasingly visible as the discussion moves from Chapter
\ref{chap:preliminaries} toward Chapters~\ref{chap:communication}
and~\ref{chap:uncertainty}. In realistic deployments, the planner may operate
with incomplete maps, delayed observations, uncertain labels, evolving
constraints, or task-relevant entities that are not known at design time.
Under such conditions, uncertainty is not only a perturbation of a fixed
model. It becomes part of the semantics of the planning problem itself. The
resulting challenge is therefore deeper than robust optimization or
stochastic control in the usual sense. It is the challenge of preserving
meaningful temporal-task reasoning when the world description remains
incomplete, revisable, and only partially aligned with the symbolic
structures used by the planner.

\textbf{(I) Limits of fixed symbolic models under partial knowledge.}
One limitation of many formal planning frameworks is that the relevant
propositions are assumed to be evaluated over a model that already exists,
even if transition outcomes remain uncertain. In practice, however,
uncertainty often enters earlier in the pipeline. The workspace may be only
partially mapped, the accessibility of regions may become known only through
exploration, and the truth of task-relevant conditions may depend on
information that is distributed across the team or acquired only during
execution. The difficulty is then not merely that the planner lacks
confidence in a known quantity, but that the symbolic model itself is
incomplete. This creates a tension with classical automata-based synthesis,
which presupposes a sufficiently stable state and proposition structure before
planning begins. The underlying question is how symbolic planning models
should be represented and updated when knowledge is accumulated
incrementally and when parts of the abstraction must remain provisional. A
satisfactory treatment would require more than uncertainty annotations
attached to a fixed transition system. It would require formal mechanisms for
expansion, correction, and revision of the symbolic model while preserving
enough logical continuity for planning to remain interpretable and
correctness claims to remain meaningful.

\textbf{(II) Semantic uncertainty and evolving task meaning.}
Another difficulty is that uncertainty may affect not only the state of the
environment, but also the meaning of the propositions used to specify the
task. In many robotic settings, a proposition is not a primitive fact given
directly by the environment. It is a semantic assertion whose truth depends on
perception, classification, context, or operational convention. Conditions
such as whether a location is safe, whether an object is relevant, whether a
region is traversable, or whether a request has been fulfilled may admit
borderline or time-varying interpretations. The central difficulty is that
such propositions are often treated as stable atomic labels within temporal
logic, whereas their truth conditions may in fact change as information is
updated or as mission context evolves. This leads to a broader question: what
should temporal-task correctness mean when the semantics of the task
vocabulary is itself uncertain or revisable? Addressing that question may
require formal models in which propositions are associated not only with
states, but also with confidence, provenance, context, or levels of semantic
commitment. More fundamentally, it suggests that a significant part of future
research will concern the interface between formal task languages and the
perceptual or inferential processes by which task-relevant meaning is
constructed.

\textbf{(III) Open-world temporal-task planning.}
The broader issue is how temporal-task planning should be formulated in an
open-world setting. In such a setting, it cannot be assumed that all relevant
objects, constraints, agents, and modes of interaction are known in advance
or remain fixed throughout execution. New entities may appear, old
assumptions may become invalid, and parts of the task structure may need to
be introduced only after additional information is obtained. This is a more
radical departure from standard synthesis assumptions than ordinary model
uncertainty, because it affects the very boundary of what the planner is
expected to reason about. The central question is no longer only how to find
a correct strategy on a given symbolic model, but how to maintain a coherent
task-directed decision process while the model, the proposition set, and even
the relevant interaction structure continue to evolve. At present, there is
no general framework that fully resolves this difficulty. What is needed is a
theory of open-world temporal-task planning in which symbolic abstractions can be revised without dissolving task-level guarantees, and in which model growth
is treated as a first-class aspect of coordination rather than as an external
preprocessing step.

%=================================================
\section{Learning-Augmented Planning with Formal Guarantees}
\label{sec:ch07_learning}

Once the symbolic model, the environment, and even the task vocabulary become
less stable, the attraction of learning-based components becomes evident.
Their role, however, is not straightforward. The methods developed in this
monograph are largely grounded in explicit models, symbolic abstractions, and
formally specified task semantics. That emphasis is methodologically
important, because it is precisely such structure that makes correctness,
feasibility, and coordination properties amenable to analysis. At the same
time, many of the components required in realistic deployments are difficult
to specify accurately in advance. Perception pipelines must interpret noisy
and context-dependent observations, models of environmental evolution may be
only approximate, and predictions about moving targets, external agents, or
human requests may depend on data-driven inference rather than analytic
modeling. Future progress is therefore unlikely to come from preserving a
strict separation between formal planning and learning-based adaptation. The
deeper question is how learning should be incorporated so that the resulting
system retains the task-level discipline, interpretability, and reliability
that motivate formal methods in the first place.

\textbf{(I) Learning across planning layers.}
The first difficulty is that learning is not a single operation inserted
uniformly into the planning pipeline. Different components may be learned for
different reasons and with different implications for certification.
Perception modules may infer the truth of task-relevant propositions from raw
sensory data. Predictive models may estimate the motion of external agents,
the evolution of workspace occupancy, or the likelihood of request arrival.
Data-driven representations may also be used to construct abstractions,
estimate costs, or shape heuristics for planning and assignment. The issue is
therefore not simply whether learning should be used, but where it should
enter and what kind of uncertainty it is meant to reduce. Learning a motion
primitive, for example, is conceptually different from learning a semantic
label or a model of environmental change, because the effect on task-level
reasoning arises at a different layer of the synthesis process. A systematic
theory of learning-augmented temporal-task planning would need to distinguish
these roles carefully and to specify which learned quantities may be treated
as auxiliary guidance, which may enter directly into the symbolic model, and
which should remain outside the core correctness logic.

\textbf{(II) Integration of learned components with formal guarantees.}
Even when the role of learning is clear, integration remains difficult. A
learned predictor may improve responsiveness but also introduce systematic
bias. A perception module may enable richer task semantics but also produce
classification errors whose effect propagates through the symbolic planner. A
learned abstraction may reduce computational cost, yet its failure modes may
be difficult to characterize using the same formal language in which the task
is specified. The methodological question is therefore how interfaces between
learned and symbolic components should be designed. This includes the need
for confidence-aware planning, explicit monitoring of model validity, fallback
or recovery mechanisms when learned outputs become unreliable, and principled
criteria for when a planner may trust learned information strongly enough to
act upon it. More broadly, it raises the issue of modular certification:
under what assumptions can the correctness or safety properties of a planning
architecture be preserved when part of the model is produced statistically
rather than analytically? At present, answers to this question remain
fragmented and highly application-dependent.

\textbf{(III) Guarantee notions for data-driven temporal-task planning.}
At a deeper level, the notion of guarantee itself may need to be reformulated
once learning becomes a central component of the decision process. In
classical automata-based synthesis, correctness is typically expressed against
a fixed abstraction and a stable task semantics. In a hybrid
formal--learning architecture, however, model uncertainty, distribution shift,
and context dependence may make such guarantees either too strong to be
realistic or too weak to be operationally useful. The challenge is therefore
not merely to preserve nominal correctness under approximation, but to define
guarantees that remain meaningful when some task-relevant quantities are
inferred from data and remain only partially reliable. Possible directions
include guarantees conditioned on calibrated confidence, bounded classes of
model mismatch, runtime-detectable failure modes, or verified adaptation
within restricted regions of the state and task space. Yet even these
directions leave open a broader conceptual question: should the goal be to
recover classical formal guarantees as closely as possible, or to develop a
new class of guarantees tailored to hybrid symbolic-statistical systems? The
answer is likely to shape the next generation of temporal-task planning
methods.

%=================================================
\section{Verification, Human Interaction, and Deployment-Level Evaluation}
\label{sec:ch07_deployment}

The final set of open questions arises not at the level of an isolated
planning algorithm, but at the level of deployment. The preceding chapters
develop methods that offer formal structure, explicit task semantics, and in
several cases correctness or feasibility guarantees relative to the adopted
models. Those features remain essential, but they do not by themselves
resolve the broader question of system-level dependability. Once planning is
performed under partial knowledge, intermittent communication, dynamic
constraints, continual revision, or learned inference, the relevant issue is
no longer only whether a synthesized policy is correct with respect to a fixed
model. The issue is whether correctness, interpretability, and operational
reliability can still be maintained when the system is executed as part of a
larger decision architecture that includes uncertain perception, online
supervision, and real-time intervention.

\textbf{(I) Verification beyond offline correctness.}
Classical correctness claims are often established under assumptions that
become fragile once the system is deployed. A task policy may be proven
correct relative to an abstraction, yet the abstraction itself may be
approximate, partially outdated, or revised online in response to new
information. Communication assumptions may fail intermittently, sensor inputs
may not support exact proposition evaluation, and part of the decision
process may depend on prediction or learning modules whose outputs are not
fixed in advance. Under such conditions, verification can no longer be
treated as a single offline step performed before execution begins. What is
needed are verification frameworks that remain meaningful under revision,
adaptation, and model mismatch. This includes runtime monitoring of logical
progress, detection of assumption violations, certification of fallback
behaviors, and principled criteria for when a system should continue, repair,
or suspend a plan. More fundamentally, it raises the question of what form of
verification is appropriate for a planning architecture whose symbolic model
and operative assumptions may evolve during deployment.

\textbf{(II) Human interaction and supervisory intervention.}
A second difficulty appears once human operators are treated not merely as
task designers, but as continuing participants in the coordination loop.
Requests may arrive during execution, priorities may be revised in light of
external concerns, and autonomous decisions may be overridden for reasons that
lie outside the original symbolic specification. Chapter
\ref{chap:communication} already points toward this difficulty through online
request fulfillment, but the broader challenge is more general.
Temporal-task planning must increasingly account for supervisory authority,
shared autonomy, and the need to explain or justify decisions in a form that
remains legible to human users. This creates a tension between formal
consistency and operational flexibility. A human intervention may be
reasonable at the mission level while appearing disruptive from the viewpoint
of the current task automaton, and a planner that is logically correct may
still be difficult to trust if its decisions cannot be interpreted or modified
coherently by a supervisor. Future temporal-task frameworks will therefore
need to treat supervisory interaction as part of the formal problem
formulation rather than as an external disturbance applied after planning.

\textbf{(III) Evaluation and benchmarking for system-level temporal-task planning.}
The final question concerns how progress in the field should be assessed. The
current literature often reports correctness, task completion, or
computation time on problem instances that are informative but still limited
in scope. Such metrics remain useful, yet they do not fully capture the
demands of multi-robot temporal-task planning under uncertainty,
communication limits, online adaptation, and human interaction. What is still
missing are benchmark families and evaluation criteria that reflect the
integrated nature of these difficulties. A method may perform well under
nominal execution but degrade severely under repeated replanning; another may
scale well computationally but offer only weak interpretability or poor
recoverability after supervisory intervention. The field therefore needs
richer system-level metrics, including measures of coordination stability,
adaptation cost, robustness to model revision, communication burden,
explainability, and cumulative mission quality over long horizons. It also
needs reproducible benchmark suites that enable meaningful comparison across
symbolic, optimization-based, and learning-based approaches.

%==============================
\section*{Concluding Perspective}

The developments surveyed in this chapter indicate that the future of
temporal-task planning will not be determined by progress along a single
technical axis. The main challenge is broader: to preserve the logical
discipline of formal task specification while extending planning and
coordination methods to larger teams, richer task structures, weaker modeling
assumptions, and more persistent forms of online adaptation. Across
Chapters~\ref{chap:preliminaries} through~\ref{chap:uncertainty}, the methods
developed in this monograph move from nominal synthesis on fixed abstractions
to coordination under communication limits, dynamic constraints, and
uncertain environments. That progression also reveals the boundaries of the
current theory. Scalability remains constrained by the lack of general
compositional principles, reactive coordination remains difficult to analyze
over long horizons, open-world settings challenge the stability of symbolic
models, and the integration of learning raises unresolved questions about what
system-level guarantees should mean in the first place.

The outlook for the field is therefore best understood not as a departure from
formal methods, but as a demand for their extension. Future frameworks will
likely need to combine symbolic reasoning, adaptive abstraction, runtime
verification, learned inference, and human supervision within a unified
decision architecture. The long-term objective is not merely to satisfy
temporal specifications more efficiently, but to develop planning systems
whose correctness, robustness, and interpretability remain meaningful as the
world, the model, and the coordination structure all evolve during execution.
In that sense, temporal-task planning remains an open and expanding research
area, whose most important advances may come from clarifying how formal
guarantees can be preserved under increasingly realistic conditions of
deployment.
