%==============================
\section{Topology-adaptive Fulfillment of Temporal Requests}
\label{ch05:sec:ch05_human}


The previous section established latency-bounded exploration through an embedded
wheel backbone. Online temporal requests may, however, require more than local
plan updates: some requests modify priorities or timing within the current
backbone, whereas others require temporary topology reconfiguration for
communication chains, operator relocation, or external-demand exchange.

This section develops a topology-adaptive coordination layer. The wheel remains
the nominal structure and is modified only when the request cannot be fulfilled
within the current embedding. Then a subset of robots is detached, assigned to a
request-consistent topology, and merged back after completion. The section first
formulates single-request topology optimization, including resizing, role
assignment, and realization cost, and then extends the mechanism to multiple
online requests through admissibility, compatibility, and sequential execution.

%==============================
%==============================
%==============================
\subsection{Topology optimization for diverse temporal requests}
\label{ch05:subsec:requests}

This subsection develops the topology-optimization layer. Starting from the
current wheel backbone, the planner decides whether a request can be handled by
an embedding update or requires an explicit topology change. In the latter
case, it constructs a request-consistent topology, a resized backbone, a
feasible detachment schedule, and a role assignment. The subsection formulates
the instantiation cost, gives the optimization procedure, and separates
topology-preserving from topology-changing realization modes.

% =========================================================
\subsubsection{Topology match for temporal request}
\label{ch05:subsubsec:request_problem}

For each request \(\xi\), let
\(\widehat{\mathcal{G}}^{\xi}\) denote the family of candidate
embedded topologies that satisfy the requested robot-role structure
and the required communication pattern. Given the current wheel
backbone \(\mathcal{G}^{\texttt{w}}\), the task is to determine a
resized backbone \(\mathcal{G}^{\texttt{w}'}\), a selected target
topology \(\mathcal{G}^{\xi}\in\widehat{\mathcal{G}}^{\xi}\), a
detachment schedule \(\widehat{\pi}^{\xi}\), and a role assignment
\(\Pi^{\xi}\). The purpose of this topology-matching step is to convert
a temporal request into an executable topology transition whose cost
balances backbone reconfiguration and request-realization time. Define
\(\widehat{\pi}^{\xi}\triangleq
\{\pi_r^{\xi}\}_{r\in\mathcal{N}^{\xi}}\),
where \(\mathcal{N}^{\xi}\) is the set of robots detached from
\(\mathcal{G}^{\texttt{w}}\) to realize \(\mathcal{G}^{\xi}\),
and \(\pi_r^{\xi}\triangleq(r,t_r,p_r)\) is the detach event
assigned to robot \(r\). The realization cost is defined by:
\begin{equation}
\label{ch05:eq:match_objective}
J^{\xi}\triangleq
T^{\texttt{trs}}\!\left(
\mathcal{G}^{\texttt{w}},\mathcal{G}^{\texttt{w}'}
\right)
+
w_{\texttt{n}}
\max_{r\in\mathcal{N}^{\xi}}
\left\{
T_r^{\texttt{nav}}(\pi_r^{\xi},\Pi^{\xi})
\right\},
\end{equation}
where \(T^{\texttt{trs}}(
\mathcal{G}^{\texttt{w}},\mathcal{G}^{\texttt{w}'})\) is the time
required to resize the wheel backbone,
\(T_r^{\texttt{nav}}(\pi_r^{\xi},\Pi^{\xi})\) is the time needed by
robot \(r\) to execute its detach event and reach the assigned role,
and \(w_{\texttt{n}}>0\) weights backbone resizing against request
realization.
The topology-instantiation problem is therefore to minimize
\eqref{ch05:eq:match_objective} over all admissible candidates in
\(\widehat{\mathcal{G}}^{\xi}\).

\begin{problem}
\label{ch05:prob:topology_matching}
Given the current wheel backbone \(\mathcal{G}^{\texttt{w}}\) and
the candidate family \(\widehat{\mathcal{G}}^{\xi}\) associated
with request \(\xi\), determine the topology matching procedure:
\begin{equation}
\label{ch05:eq:single_topologymatch}
\texttt{TopologyMatch}\!\left(
\mathcal{G}^{\texttt{w}},\widehat{\mathcal{G}}^{\xi},\xi
\right)
\triangleq
\left(
\mathcal{G}^{\texttt{w}'},\mathcal{G}^{\xi},
\widehat{\pi}^{\xi},\Pi^{\xi}
\right),
\end{equation}
which returns a request-consistent topology transition that minimizes
\eqref{ch05:eq:match_objective}.
\end{problem}

%==============================
\begin{algorithm}[t!]
\caption{Request-consistent topology instantiation
\(\texttt{TopologyMatch}(\cdot)\)}
\label{ch05:alg:topologymatch}
\LinesNumbered
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}

\Input{\(\mathcal{G}^{\texttt{w}}\), request \(\xi\), candidate
family \(\widehat{\mathcal{G}}^{\xi}\), weight \(w_{\texttt{n}}\)}
\Output{\(\mathcal{G}^{\texttt{w}'}\), \(\mathcal{G}^{\xi}\),
\(\widehat{\pi}^{\xi}\), \(\Pi^{\xi}\)}

Extract \(\mathcal{C}^{+}\) from the local plans associated with
\(\mathcal{G}^{\texttt{w}}\)\;
Initialize \(J^\star\leftarrow +\infty\)\;
Choose an initial candidate
\(\mathcal{G}_{1}^{\xi}\in\widehat{\mathcal{G}}^{\xi}\)\;

\For{\(q=1,2,\cdots,|\widehat{\mathcal{G}}^{\xi}|\)}{
  Extract \(\mathcal{N}^{\xi}\) and \(\mathbf{p}_0^{\xi}\) from
  \(\mathcal{G}_{q}^{\xi}\)\;

  \(\mathcal{G}^{\texttt{w}'}\leftarrow
  \mathcal{G}^{\texttt{w}}[
  \mathcal{N}^{\texttt{w}}\setminus\mathcal{N}^{\xi}]\)\;

  \(\left(
  T^{\texttt{trs}}(
  \mathcal{G}^{\texttt{w}},\mathcal{G}^{\texttt{w}'}),
  \widehat{\pi}^{\xi}
  \right)
  \leftarrow
  \texttt{MultiDetach}(
  \mathcal{G}^{\texttt{w}},\mathcal{N}^{\xi})\) by
  \eqref{ch05:eq:topologymatch_trs}\;

  \If{a feasible detachment schedule is returned}{
    \(\Pi^{\xi}\leftarrow
    \texttt{LBAP}(
    \mathcal{N}^{\xi},\mathbf{p}_0^{\xi},\widehat{\pi}^{\xi})\) by
    \eqref{ch05:eq:role_assignment}\;

    Evaluate \(J^{\xi}\) by \eqref{ch05:eq:match_objective}\;

    \If{\(J^{\xi}<J^\star\)}{
      Store the current solution and set
      \(J^\star\leftarrow J^{\xi}\)\;
    }
  }

  \If{unevaluated candidates remain}{
    Select the next candidate by
    \eqref{ch05:eq:topologymatch_next_candidate}\;
  }
}
\Return the stored best solution\;
\end{algorithm}
%==============================


% =========================================================
\subsubsection{Optimization of topology instantiation}
\label{ch05:subsubsec:request_optimization}

For a candidate family \(\widehat{\mathcal{G}}^{\xi}\), the procedure evaluates
feasible backbone-resizing schedules, assigns detached robots to target roles,
and returns the candidate minimizing \eqref{ch05:eq:match_objective}.
Consider a candidate
\(\mathcal{G}^{\xi}\in\widehat{\mathcal{G}}^{\xi}\). Let
\(\mathcal{N}^{\xi}\) be the robots assigned to this target topology.
Removing \(\mathcal{N}^{\xi}\) from the current wheel backbone
\(\mathcal{G}^{\texttt{w}}\) yields the resized backbone
\(\mathcal{G}^{\texttt{w}'}\). The detachment routine in
Section~\ref{ch05:sec:detach-join} is then invoked to compute a feasible
timed schedule:
\begin{equation}
\label{ch05:eq:topologymatch_trs}
\left(
T^{\texttt{trs}}\!\left(
\mathcal{G}^{\texttt{w}},\mathcal{G}^{\texttt{w}'}
\right),
\widehat{\pi}^{\xi}
\right)
\triangleq
\texttt{MultiDetach}\!\left(
\mathcal{G}^{\texttt{w}},\mathcal{N}^{\xi}
\right),
\end{equation}
where each \(\pi_r^{\xi}\in\widehat{\pi}^{\xi}\) has the form
\(\pi_r^{\xi}\triangleq(r,t_r,p_r)\). The quantity
\(T^{\texttt{trs}}\) is the earliest time at which the resized
backbone \(\mathcal{G}^{\texttt{w}'}\) is fully formed and its
embeddings have been updated through \eqref{ch05:eq:embed-update}.
The detached robots are next assigned to the role locations of the
selected topology. Define
\(\mathbf{p}_0^{\xi}\triangleq
\{p_r^{\xi}\}_{r\in\mathcal{N}^{\xi}}\) as the set of initial role
locations in \(\mathcal{G}^{\xi}\). An assignment
\(\Pi^{\xi}:\mathcal{N}^{\xi}\to\mathbf{p}_0^{\xi}\) induces the
bottleneck realization time
\(\max_{r\in\mathcal{N}^{\xi}}
\{T_r^{\texttt{nav}}(\pi_r^{\xi},\Pi^{\xi})\}\).
Selecting \(\Pi^{\xi}\) to minimize this bottleneck yields the
unbalanced linear bottleneck assignment problem:
\begin{equation}
\label{ch05:eq:role_assignment}
\Pi^{\xi}\leftarrow
\texttt{LBAP}\!\left(
\mathcal{N}^{\xi},\mathbf{p}_0^{\xi},\widehat{\pi}^{\xi}
\right),
\end{equation}
where the LBAP formulation follows~\cite{burkard2012assignment}. A
practical implementation can be obtained from standard optimization
packages such as OR-Tools~\cite{gor}.
To reduce the search cost over
\(\widehat{\mathcal{G}}^{\xi}\), the candidates are ordered by a
best-first surrogate. After evaluating the current candidate
\(\mathcal{G}_{q}^{\xi}\), the next candidate is selected by:
\begin{equation}
\label{ch05:eq:topologymatch_next_candidate}
\mathcal{G}_{q+1}^{\xi}
=
\underset{
\mathcal{G}\in
\widehat{\mathcal{G}}^{\xi}\setminus\{\mathcal{G}_{q}^{\xi}\}
}{\textbf{argmin}}
\Big(
\widetilde{T}^{\texttt{trs}}(\mathcal{G},\mathcal{C}^{+})
+
w_{\texttt{n}}\,
\widetilde{T}^{\texttt{nav}}(\mathcal{G},\mathcal{C}^{+})
\Big),
\end{equation}
where \(\mathcal{C}^{+}\) is the set of future meeting events
implied by the local plans \(\{\Gamma_i\}\). The quantities
\(\widetilde{T}^{\texttt{trs}}(\mathcal{G},\mathcal{C}^{+})\) and
\(\widetilde{T}^{\texttt{nav}}(\mathcal{G},\mathcal{C}^{+})\) are
fast predictors of the backbone-transition time and the bottleneck
realization time for candidate \(\mathcal{G}\). Thus,
\eqref{ch05:eq:topologymatch_next_candidate} prioritizes candidates that
are likely to admit early feasible detachment and small worst-case
travel, without requiring full evaluation at each intermediate step.
The complete procedure is summarized in
Algorithm~\ref{ch05:alg:topologymatch}. Its role within the proposed
coordination method is to return an executable request-consistent
transition for a given candidate family, as stated below.

\begin{lemma}
\label{ch05:thm:topologymatch-exec-short}
For any request \(\xi\), suppose that there exists at least one
candidate \(\mathcal{G}^{\xi}\in\widehat{\mathcal{G}}^{\xi}\) for
which \eqref{ch05:eq:topologymatch_trs} returns a feasible detachment
schedule for the corresponding robot set \(\mathcal{N}^{\xi}\). Then
Algorithm~\ref{ch05:alg:topologymatch} terminates in finite time and returns
\((\mathcal{G}^{\texttt{w}'},\mathcal{G}^{\xi},
\widehat{\pi}^{\xi},\Pi^{\xi})\) such that the resulting transition is
executable under the detachment mechanism, and the objective in
\eqref{ch05:eq:match_objective} is minimized over all candidates in
\(\widehat{\mathcal{G}}^{\xi}\).
\end{lemma}
\begin{proof}
The proof follows directly from the finite candidate search in
Algorithm~\ref{ch05:alg:topologymatch} and from the construction steps in
\eqref{ch05:eq:topologymatch_trs}, \eqref{ch05:eq:role_assignment},
and \eqref{ch05:eq:match_objective}.
First, the candidate family \(\widehat{\mathcal{G}}^{\xi}\) is finite
by construction. Algorithm~\ref{ch05:alg:topologymatch} examines one
candidate at each iteration and never revisits an already evaluated
candidate, since the next candidate is selected from the unevaluated
set through the ordering rule associated with the search step. Hence
the outer loop can execute at most
\(|\widehat{\mathcal{G}}^{\xi}|\) times.

For a fixed candidate \(\mathcal{G}^{\xi}\), the corresponding robot
set \(\mathcal{N}^{\xi}\) and role locations
\(\mathbf{p}_0^{\xi}\) are uniquely extracted in finite time. The
resized backbone \(\mathcal{G}^{\texttt{w}'}\) is then obtained
directly from the induced subgraph operation on
\(\mathcal{G}^{\texttt{w}}\). Next,
\(\texttt{MultiDetach}(\mathcal{G}^{\texttt{w}},\mathcal{N}^{\xi})\)
in \eqref{ch05:eq:topologymatch_trs} either returns infeasibility or
returns a feasible detachment schedule
\(\widehat{\pi}^{\xi}\) together with the corresponding transition
time \(T^{\texttt{trs}}(
\mathcal{G}^{\texttt{w}},\mathcal{G}^{\texttt{w}'})\). By the
detachment mechanism developed in
Section~\ref{ch05:sec:detach-join}, feasibility of this subroutine is
equivalent to executability of the backbone-resizing transition.
Whenever a feasible detachment schedule is returned, the subsequent
assignment \(\Pi^{\xi}\) is obtained by solving the bottleneck
assignment problem in \eqref{ch05:eq:role_assignment}, which is a
finite combinatorial optimization over the finite sets
\(\mathcal{N}^{\xi}\) and \(\mathbf{p}_0^{\xi}\). Therefore each
candidate evaluation terminates in finite time.
By assumption, there exists at least one candidate in
\(\widehat{\mathcal{G}}^{\xi}\) for which
\eqref{ch05:eq:topologymatch_trs} is feasible. Hence, during the finite
search, the algorithm encounters at least one executable candidate and
updates the incumbent cost \(J^\star\). For every feasible candidate,
the value \(J^{\xi}\) is evaluated through
\eqref{ch05:eq:match_objective}, which combines the backbone
transition time and the bottleneck navigation time induced by
\(\Pi^{\xi}\). The algorithm stores the candidate if and only if its
objective value is smaller than the current incumbent. Since all
candidates in \(\widehat{\mathcal{G}}^{\xi}\) are examined in finite
time, the final stored solution attains the minimum of
\eqref{ch05:eq:match_objective} over all executable candidates.

Finally, the returned tuple
\((\mathcal{G}^{\texttt{w}'},\mathcal{G}^{\xi},
\widehat{\pi}^{\xi},\Pi^{\xi})\) is executable by construction:
\(\widehat{\pi}^{\xi}\) is a feasible detach schedule realizing the
transition from \(\mathcal{G}^{\texttt{w}}\) to
\(\mathcal{G}^{\texttt{w}'}\), and \(\Pi^{\xi}\) assigns each detached
robot to a valid role in the selected request-consistent topology
\(\mathcal{G}^{\xi}\). Therefore
Algorithm~\ref{ch05:alg:topologymatch} terminates in finite time and returns
an executable topology transition that minimizes
\eqref{ch05:eq:match_objective} over
\(\widehat{\mathcal{G}}^{\xi}\).
\end{proof}

% =========================================================
\subsubsection{Realization modes for temporal requests}
\label{ch05:subsubsec:request_modes}

Requests are realized either by preserving the wheel backbone and updating only
its embeddings, or by invoking Algorithm~\ref{ch05:alg:topologymatch} to
instantiate a request-consistent subtopology with a detach schedule and role
assignment. In both cases, the remaining robots keep the wheel active so that
latency-bounded intermittent communication is preserved.

(I) \emph{Intra-fleet communication with bounded latency.}
For the intra-fleet exploration requests \(\xi_1\), \(\xi_2\), and
\(\xi_5^i\), the wheel backbone \(\mathcal{G}^{\texttt{w}}\) is kept
and no topology instantiation is required. The base request \(\xi_1\) is
satisfied directly by the backbone itself, which guarantees
latency-bounded information flow. The requests \(\xi_2\) and
\(\xi_5^i\) are realized by updating the embeddings in
\eqref{ch05:eq:embed-update} while retaining the same topology.
For \(\xi_2\), frontiers in the prohibited region
\(\mathcal{P}^{-}\) are excluded, while the remaining frontiers are
prioritized toward \(\mathcal{P}^{+}\) through the evaluation:
\begin{equation}
\label{ch05:eq:frontier_weight}
J(f)\triangleq
\sum_{p\in\mathcal{P}^{+}}\|x(f)-p\|,
\qquad \forall f\in\mathcal{F},
\end{equation}
where larger values of \(J(f)\) are assigned lower priority. For
\(\xi_5^i\), an immediate return opportunity is introduced for timely
delivery of \(D_i\), and the relevant time stamps in the embeddings
are updated accordingly. In both cases, the backbone topology is
unchanged.

(II) \emph{External-demand exchange with bounded latency.}
For the external-demand request \(\xi_7^{\texttt{e}}\), the fleet
periodically exchanges compressed mission data with an external demand
under the latency bound \(T_{\texttt{e}}\). This request is realized
through a singleton messenger subtopology. A messenger robot detaches
from the backbone, reaches an external exchange event
\(\mathbf{c}_{\texttt{e}}\triangleq(p_{\texttt{d}},t_{\texttt{e}})\),
performs the data exchange, and then rejoins the backbone.
Consistency with the intra-fleet meeting sequence is enforced by
selecting the messenger from the most recent meeting pair and by
choosing the external exchange event so that:
\begin{equation}
\label{ch05:eq:interteam_req}
t_{ij}^{+}+T_{\ell}^{\texttt{nav}}
\le t_{\texttt{e}}, \, \forall \ell\in\{i,j\},
\end{equation}
where \(\mathbf{c}_{ij}^{+}\triangleq(p_{ij}^{+},t_{ij}^{+})\) is the
next intra-fleet meeting event. After the data exchange, the external
meeting time is updated to
\(t_{\texttt{e}}^{+}\triangleq t_{\texttt{e}}+T_{\texttt{e}}\),
after which the messenger rejoins the backbone.

(III) \emph{Operator relocation and group migration.}
Request \(\xi_3\) relocates the operator \(\texttt{h}\) to a
target position \(p_{\texttt{h}}^{+}\) while preserving latency
feasibility. If \(p_{\texttt{h}}^{+}\) satisfies:
\begin{equation}
\label{ch05:eq:base-location}
T_i^{\texttt{nav}}(p_{\texttt{h}}^{+},p_i^{\ell_i})
\le
T_i^{\texttt{nav}}(p_{\texttt{h}},p_i^{\ell_i}),
\qquad
\forall i\in\mathcal{N},\ \ell_i\le L_i,
\end{equation}
then the embeddings are updated in place and the backbone topology is
retained. Otherwise, the relocation is realized as a
connectivity-preserving migration. In that case,
\texttt{TopologyMatch} instantiates an embedded star subtopology centered at
\(\texttt{h}\), and assigns robots to relative placements that
preserve connectivity during motion. Once the migration is completed,
the wheel backbone is restored for continued exploration.

(IV) \emph{Consistent link through a communication-ensured chain.}
Requests \(\xi_4\) and \(\xi_6^i\) require a reliable and
consistent chain between the operator \(\texttt{h}\) and a
designated robot at \(p_i^{\star}\), as required by~\ref{ch05:eq:chain}.
These requests are therefore realized
through a communication-ensured chain by selecting anchor points
along the shortest path from \(p_{\texttt{h}}\) to \(p_i^{\star}\)
so that the hop-by-hop communication quality satisfies:
\begin{equation}
\label{ch05:eq:cec}
\mathbf{p}^{\xi}\triangleq p_0p_1\cdots p_C,
\qquad
\texttt{Qual}(p_c,p_{c+1},\mathcal{M}_{\texttt{h}})
\ge \underline{\delta}^{\texttt{c}},
\end{equation}
where \(c=0,\cdots,C-1\). Assigning robots to these anchor points
yields an embedded chain subtopology that enforces multi-hop
connectivity using the estimated communication map. Any robots not
allocated to the chain continue exploration within the resized wheel
backbone.































% ========================================
% ========================================
% ========================================
\subsection{Online topological coordination for multiple requests}
\label{ch05:subsec:online-requests}

This subsection handles multiple temporal requests. Several requests may arrive
close in time and compete for the same robots, so the online problem is to
incorporate them into the evolving embedded topology without violating the
remaining communication backbone. The wheel remains the nominal structure:
topology-preserving requests update embeddings, whereas topology-changing
requests instantiate subtopologies by detaching robot subsets. The full topology
is therefore the union of the residual wheel and active request subtopologies.

% =========================================================
\subsubsection{Compatibility and sequential request processing}
\label{ch05:subsubsec:multi_req}

At time \(t\), the embedded topology consists of two components: the
current wheel backbone that continues to support exploration and
latency-bounded intermittent communication, and the collection of
request-specific subtopologies that have already been instantiated but
have not yet completed their assigned tasks. This decomposition is
needed because newly arriving requests are evaluated relative to the
robots that remain available on the current backbone, rather than
relative to the full fleet. It also makes explicit that online request
handling takes place on an evolving topology whose structure depends on
both previously accepted requests and the residual communication
backbone that they leave behind. Accordingly, the embedded topology is
written as:
\begin{equation}
\label{ch05:eq:decomposition}
\mathcal{G}(t) \triangleq
\mathcal{G}^{\texttt{w}}
\cup
\bigcup_{j=1}^{L_{\texttt{a}}}\mathcal{G}^{j},
\end{equation}
where \(\mathcal{G}^{\texttt{w}}\) is the current wheel backbone used
for exploration, \(\mathcal{G}^{j}\) is the instantiated subtopology
associated with the active request \(\xi^{j}\), and
\(L_{\texttt{a}}\) is the number of active requests.
Suppose that a robot
\(\texttt{r}\in\mathcal{N}^{\texttt{w}}\) receives a bundle of
\(L>0\) new requests
\(\Xi \triangleq \{\xi^{\ell}\}_{\ell=1}^{L}\). The requests
\(\xi_1\), \(\xi_2\), and \(\xi_5\) preserve the wheel topology
and are handled directly through embedding updates, as described in
Section~\ref{ch05:subsec:requests}. The request \(\xi_3\) is
also topology-preserving when the in-place condition in
\eqref{ch05:eq:base-location} is satisfied. Otherwise, \(\xi_3\)
induces a topology-changing migration. By contrast, the requests
\(\xi_4\), \(\xi_6\), \(\xi_7^{\texttt{e}}\), and the migration case of
\(\xi_3\) require explicit topology instantiation through candidate
families \(\widehat{\mathcal{G}}^{\xi}\), constructed according to
\eqref{ch05:eq:base-location}, \eqref{ch05:eq:cec}, and
\eqref{ch05:eq:interteam_req}.

For the topology-changing requests in \(\Xi\), a subset
\(\Xi' \subseteq \Xi\) is called compatible on
\(\mathcal{G}^{\texttt{w}}\) if the current wheel backbone can supply
disjoint robot groups for all requests in \(\Xi'\), it holds that:
\begin{equation}
\label{ch05:eq:compat}
\mathcal{N}^{\xi}\cap\mathcal{N}^{\xi'}=\emptyset,
\quad \forall \xi\neq\xi'\in\Xi',
\qquad
\bigcup_{\xi\in\Xi'}\mathcal{N}^{\xi}
\subseteq
\mathcal{N}^{\texttt{w}},
\end{equation}
where \(\mathcal{N}^{\xi}\) denotes the set of robots allocated to
fulfill request \(\xi\). The migration case of \(\xi_3\) is more
restrictive. It is compatible only with the complete wheel backbone.
Once the fleet has already been split into active request subtopologies,
robots detached from the backbone cannot, in general, be guaranteed to
receive the updated operator location during migration. For this
reason, the migration case of \(\xi_3\) is deferred until all other
active requests have been completed and the full wheel backbone has
been restored.

Finding a maximum compatible subset of \(\Xi\) under \eqref{ch05:eq:compat} is
combinatorial. The method therefore uses a sequential policy: requests are
ordered by operator priority or first-in-first-out order and examined on the
current wheel backbone. A topology-preserving request is retained when its
embedding update is admissible; a topology-changing request is retained only if
it is admissible on the residual wheel and \(\texttt{TopologyMatch}(\cdot)\)
returns an executable instantiation.
Let \(\mathcal{G}^{\texttt{w},0}\) denote the wheel backbone before
the new request bundle is processed. For \(\ell=1,\cdots,L\), if the
request \(\xi^{\ell}\) requires topology instantiation and is admissible
on \(\mathcal{G}^{\texttt{w},\ell-1}\), then it follows that:
\begin{equation}
\label{ch05:eq:combinatorial}
\mathcal{G}^{\texttt{w},\ell},\,
\mathcal{G}^{\ell},\,
\widehat{\pi}^{\ell},\,
\Pi^{\ell}
\leftarrow
\texttt{TopologyMatch}\!\left(
\mathcal{G}^{\texttt{w},\ell-1},\,
\widehat{\mathcal{G}}^{\xi^{\ell}},\,
\xi^{\ell}
\right),
\end{equation}
where \(\mathcal{G}^{\ell}\) is the instantiated subtopology for
\(\xi^{\ell}\), \(\widehat{\pi}^{\ell}\) is the associated
detachment schedule, and \(\Pi^{\ell}\) is the resulting role
assignment, consistently with \eqref{ch05:eq:single_topologymatch}. If no
executable instantiation exists on the remaining wheel backbone, the
request is deferred and the procedure moves to the next request in the
priority order. After all new requests have been processed, the
decomposition in \eqref{ch05:eq:decomposition} is augmented by the
newly instantiated subtopologies and by the embedding updates that were
found admissible.

%==============================
\begin{algorithm}[t]
  \caption{Overall online execution of the proposed coordination
  method}
  \label{ch05:alg:moroco}
  \LinesNumbered
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}

  \Input{\(\mathcal{G}_{0}^{\texttt{w}}\), \(\{\Gamma_i\}\)}
  \Output{Local plans \(\{\Gamma_i\}\) and evolving topology
  \(\mathcal{G}\)}

  \While{the mission is not terminated}{

    \tcc{\textbf{Wheel-backbone execution}}
    Each robot \(i\in\mathcal{N}^{\texttt{w}}\) executes \(\Gamma_i\)
    in parallel\;

    \uIf{meeting event \(\mathbf{c}_{ij}\) occurs}{
      \(\Gamma_i,\Gamma_j,\mathbf{c}_{ij}^{+}
      \leftarrow
      \texttt{IntmCoord}(\Gamma_i,\Gamma_j,\chi_{ij})\)\;
      Update the embeddings of \(\mathcal{G}^{\texttt{w}}\)\;
    }

    \tcc{\textbf{Request processing}}
    \uIf{a request bundle \(\Xi\) is received}{
      Order the requests in \(\Xi\) by priority\;

      \ForEach{request \(\xi^{\ell}\in\Xi\)}{

        \uIf{\(\xi^{\ell}\) is topology-preserving and admissible}{
          Apply the corresponding embedding update on
          \(\mathcal{G}^{\texttt{w}}\)\;
        }

        \ElseIf{\(\xi^{\ell}\) is topology-changing and admissible on
        \(\mathcal{G}^{\texttt{w}}\)}{
          \(\widetilde{\mathcal{G}}^{\texttt{w}},\,
          \widetilde{\mathcal{G}}^{\ell},\,
          \widetilde{\pi}^{\ell},\,
          \widetilde{\Pi}^{\ell}
          \leftarrow
          \texttt{TopologyMatch}\!\left(
          \mathcal{G}^{\texttt{w}},\,
          \widehat{\mathcal{G}}^{\xi^{\ell}},\,
          \xi^{\ell}
          \right)\)\;

          \uIf{an executable instantiation is returned}{
            \(\mathcal{G}^{\texttt{w}}
            \leftarrow
            \widetilde{\mathcal{G}}^{\texttt{w}}\)\;
            Add \(\widetilde{\mathcal{G}}^{\ell}\) to
            \(\mathcal{G}\)\;
            Store \(\widetilde{\pi}^{\ell}\) for execution\;
          }
        }

        \Else{
          Defer \(\xi^{\ell}\)\;
        }
      }

      Execute the stored detachment schedules and update
      \(\mathcal{G}\)\;
    }

    \tcc{\textbf{Active-request execution}}
    \ForEach{active request \(\xi^{\ell}\) in parallel}{
      All robots in the associated subgroup perform \(\xi^{\ell}\)\;
      \uIf{\(\xi^{\ell}\) is fulfilled}{
        Trigger the rejoin procedure and update \(\mathcal{G}\)\;
      }
    }
  }
\end{algorithm}
%==============================

% =========================================================
\subsubsection{Online interaction and execution}
\label{ch05:subsubsec:interaction}

The interaction interface parses an operator command at time \(t\) into a
request bundle \(\Xi \triangleq \{\xi^{\ell}\}\) with parameters and transmits
it through intermittent communication. Execution remains distributed and is
driven by Algorithm~\ref{ch05:alg:coordination}. Initially,
\(\mathcal{G}=\mathcal{G}_{0}^{\texttt{w}}\) satisfies the intra-fleet latency
bound \(T_{\texttt{h}}\), while external-demand exchange under
\(\xi_7^{\texttt{e}}\) is subject to \(T_{\texttt{e}}\). At each meeting,
\(\texttt{IntmCoord}(\cdot)\) fuses embeddings and data, maintaining
latency-bounded propagation without centralized synchronization.

When a new request bundle \(\Xi\) is received, the online
coordination layer processes the requests sequentially under the
admissibility rule described above and within the runtime summarized
by Algorithm~\ref{ch05:alg:moroco}. As a result, the embedded topology
\(\mathcal{G}(t)\) evolves as the union of the current wheel
backbone and the instantiated subtopologies required by the active
topology-changing requests. Typical examples include a singleton
messenger subtopology for \(\xi_7^{\texttt{e}}\), a chain subtopology for
\(\xi_4\) and \(\xi_6\), and a star subtopology for the migration
case of \(\xi_3\). By contrast, topology-preserving requests modify
only the embeddings of the wheel backbone and do not alter the
decomposition in \eqref{ch05:eq:decomposition}. Robots assigned to
instantiated subtopologies detach and execute their requests, while the
remaining wheel backbone continues exploration and preserves
latency-bounded intermittent connectivity. Once a request is
completed, the corresponding subtopology is removed and the involved
robots rejoin the wheel backbone, thereby restoring the fleet
structure for continued exploration and later requests.
The following theorem states the framework-level guarantee induced by
the online execution policy in Algorithm~\ref{ch05:alg:moroco}. The proof
is deferred to the appendix.

\begin{theorem}
\label{ch05:thm:moroco-complete-short}
Consider any finite-time arrival sequence of request bundles.
Suppose that every topology-preserving request admitted on the current
wheel backbone can be executed through its associated embedding update,
and that whenever a topology-changing request \(\xi^{\ell}\) becomes
admissible under \eqref{ch05:eq:compat}, there exists at least one
candidate \(\mathcal{G}^{\xi^{\ell}}\in
\widehat{\mathcal{G}}^{\xi^{\ell}}\) such that
\eqref{ch05:eq:topologymatch_trs} returns a feasible detach-and-rejoin
schedule for \(\mathcal{N}^{\xi^{\ell}}\). Then the following
properties hold.
(I) After each request activation or completion, the embedded topology
\(\mathcal{G}(t)\) remains an executable decomposition of the form
\eqref{ch05:eq:decomposition}, consisting of a remaining wheel
backbone and finitely many request subtopologies.
(II) Every request admitted by Algorithm~\ref{ch05:alg:moroco} is fulfilled
in finite time.
(III) There exists a finite time \(t_f\) such that
\(\mathcal{G}(t)=\mathcal{G}^{\texttt{w}}\) for all \(t\ge t_f\),
that is, all request subtopologies have been removed and the complete
wheel backbone has been restored.
\end{theorem}

\begin{proof}
Property (I) follows by induction over request-activation and
request-completion events. Initially,
\(\mathcal{G}(t)=\mathcal{G}^{\texttt{w}}\), so
\eqref{ch05:eq:decomposition} holds trivially. If an admitted request
is topology-preserving, only the embeddings of the current backbone are
updated, so the decomposition is unchanged. If an admitted request is
topology-changing, admissibility under \eqref{ch05:eq:compat} ensures
that its robot set is disjoint from those of the active requests and is
drawn from the current backbone. By the assumption on
\eqref{ch05:eq:topologymatch_trs} and
Lemma~\ref{ch05:thm:topologymatch-exec-short},
Alg.~\ref{ch05:alg:moroco} returns an executable tuple
\((\mathcal{G}^{\texttt{w}'},\mathcal{G}^{\xi^{\ell}},
\widehat{\pi}^{\ell},\Pi^{\ell})\). Replacing the current backbone by
\(\mathcal{G}^{\texttt{w}'}\) and adding
\(\mathcal{G}^{\xi^{\ell}}\) preserves an executable decomposition of
the form \eqref{ch05:eq:decomposition}. Completion of a request removes
the corresponding subtopology and rejoins its robots to the backbone,
so the same form is preserved. Hence (I) holds.
For (II), every admitted topology-preserving request is executable by
assumption through its embedding update. Every admitted
topology-changing request is executable by
Lemma~\ref{ch05:thm:topologymatch-exec-short}, since it is realized by
a feasible detach-and-rejoin schedule together with a valid role
assignment. Each such execution consists of finitely many motions,
detachment events, task actions, and rejoin events, all with finite
duration. Therefore every admitted request is fulfilled in finite time.
For (III), the arrival sequence of request bundles is finite in time,
so there exists a last activation event. After that time, no new
requests are introduced. By (II), every request admitted before that
time is completed in finite time. Each completion removes one active
request subtopology from \eqref{ch05:eq:decomposition} and returns its
robots to the backbone. Since only finitely many requests are admitted,
there exists a finite time \(t_f\) after which no active request
subtopology remains. Therefore
\(\mathcal{G}(t)=\mathcal{G}^{\texttt{w}}\) for all \(t\ge t_f\).
\end{proof}
