%==============================
\section{Exploration and Coordination via Intermittent Communication}
\label{ch05:sec:intermittent_comm}

The previous section considered communication and collaboration as a
joint planning problem under online temporal requests. The present
section turns to a complementary setting in which exploration itself
is the primary mechanism for acquiring new information, while
communication among the operator and the robot fleet is available only
intermittently. The environment is initially unknown, so map
information, task-relevant observations, and request-dependent updates
must be generated and fused online under sparse communication events
rather than under persistent connectivity.
This setting introduces a distinct coordination challenge. Since
robots cannot rely on continuously shared global knowledge, coherent
exploration must be maintained despite delayed information exchange,
partial local awareness, and communication opportunities that depend
on the evolving team topology. At the same time, operator requests and
newly discovered workspace information must still propagate through
the fleet with bounded delay whenever possible. These requirements
motivate a framework in which exploration, intermittent
communication, and online coordination are treated within a unified
planning and execution architecture.
The development in this section therefore centers on an embedded
communication topology that supports frontier-based exploration while
preserving periodic information fusion between the operator and the
fleet. On top of this topology, local exploration plans, map updates,
and request-triggered coordination decisions are integrated so that
the team can continue to explore efficiently without assuming
continuous end-to-end communication.


%==============================
%==============================
%==============================

\subsection{Operator-fleet system under limited communication}
\label{ch05:sec:problem-exploration}

This subsection introduces the operator-fleet model used throughout the intermittent
communication framework. The discussion proceeds from the workspace and communication
assumptions to the request model and the resulting local planning objective.

\subsubsection{Workspace, robot model, and operator model}
\label{ch05:subsec:ws}

Consider a three-dimensional workspace
$\mathcal{A} \subset \mathbb{R}^3$, whose map, including the boundary, free
space, and obstacles, is initially unknown. To construct a complete map of the
workspace together with the relevant features, a single operator and a fleet of
autonomous robots are deployed in $\mathcal{A}$. Let
$\mathcal{N} \triangleq \{1,\cdots,N\}$ denote the robot index set, and let
$\mathcal{T} \triangleq \{\texttt{h}\} \cup \mathcal{N}$ denote the overall
operator-fleet system, where $\texttt{h}$ is the human operator. Each robot
$i \in \mathcal{N}$ is equipped with sensing devices such as IMU, LiDAR, and
RGB-D cameras, together with the actuators required for mobility in the
workspace. These capabilities support simultaneous localization and mapping, as
formalized below.

\begin{definition}
Each robot is capable of simultaneous localization and mapping with collision
avoidance, namely:
\begin{equation}
\bigl(M_i^+,\, p_i^+,\, \boldsymbol{\tau}_i\bigr) \triangleq
\texttt{SLAM}\bigl(p_i,\, p_{\texttt{g}},\, M_i,\, \mathcal{A}\bigr),
\label{ch05:eq:slam}
\end{equation}
where $\texttt{SLAM}(\cdot)$ is a generic SLAM module, $M_i(t) \subset
\mathcal{A}$ is the local map of robot $i$ at time $t \geq 0$,
$p_i(t) \in M_i(t)$ is the pose of robot $i$, $p_{\texttt{g}} \in M_i(t)$ is a
goal pose in the current local map, $\boldsymbol{\tau}_i \subset M_i(t)$ is a
collision-free path from $p_i(t)$ to $p_{\texttt{g}}$, $M_i^+ \subset
\mathcal{A}$ is the updated map after traversing the path at time $t' > t$, and
$p_i^+(t')$ is the updated pose, close to $p_{\texttt{g}}$ when navigation
succeeds.
\end{definition}

The exact map representation depends on the underlying SLAM algorithm and
software implementation, for instance an occupancy map
\cite{thrun2002probabilistic} or a semantic map \cite{tian2022kimera}. These
details are not essential to the coordination model here, provided that each
robot can update its local map and generate collision-free motion within the
known part of the workspace.
The operator $\texttt{h}$ does not construct a map directly, but can localize
itself with pose $p_{\texttt{h}}(t) \in \mathcal{A}$ and move within its
current local map $M_{\texttt{h}}(t) \subset \mathcal{A}$ for all $t \geq 0$.
Consequently, communication with the robotic fleet is required for the operator
to obtain updated local maps and maintain situational awareness during the
exploration process.

\subsubsection{Communication model under limited connectivity}
\label{ch05:subsec:com}

Each robot and the operator
are equipped with wireless communication devices for robot-robot and
robot-operator information exchange, such as ad-hoc networks, UWB, radio, or
5G-based communication \cite{klaesson2020planning,xia2023relink}. The same
communication model is used throughout the operator-fleet system.

\begin{definition}
Two communicating entities $x, y \in \mathcal{N} \cup \{\texttt{h}\}$ can
exchange data at time $t > 0$ if the \emph{communication quality exceeds a threshold}, i.e.,
\begin{equation}
\bigl(D_x^+,\, D_y^+\bigr) \triangleq \texttt{Com}(D_x, D_y),
\quad \text{if } \texttt{Qual}(p_x, p_y, \mathcal{A}) > \underline{\delta},
\label{ch05:eq:com}
\end{equation}
where $\texttt{Com}(\cdot)$ is the communication module, $D_x(t), D_y(t) \in
\mathcal{D}$ are the local data held by the two entities at time $t$,
$\mathcal{D}$ includes map data, video streams, coordination messages, and
online requests, $\texttt{Qual}(\cdot)$ evaluates the communication quality
from the positions and the workspace geometry, and $\underline{\delta} > 0$ is
the minimum admissible quality level.
\end{definition}

By default, the pair $(p_i, M_i)$ is contained in $D_i$ for every robot
$i \in \mathcal{N}$, and $(p_{\texttt{h}}, M_{\texttt{h}})$ is contained in the
operator data $D_{\texttt{h}}$. For any entity
$x \in \mathcal{N} \cup \{\texttt{h}\}$, let
$\mathcal{C}_x(t) \subseteq \mathcal{N} \cup \{\texttt{h}\}$ denote the set of
entities that can communicate with $x$ at time $t$ according to
\eqref{ch05:eq:com}. To maintain reliable access between the operator and a
remote robot over a nonzero time interval, a communication chain is introduced.

\begin{definition}
A \emph{communication chain} from robot $i_0 \in \mathcal{N}$ to the operator
$\texttt{h}$ at time $t \geq 0$ is a sequence below:
\begin{equation}
\overline{i\texttt{h}} \triangleq i_0 i_1 \cdots i_\ell \cdots i_L \texttt{h},
\label{ch05:eq:chain}
\end{equation}
where each neighboring pair in \eqref{ch05:eq:chain} satisfies the
communication condition in \eqref{ch05:eq:com}. Equivalently,
$i_{\ell+1} \in \mathcal{C}_{i_\ell}(t)$ for all
$\ell \in \{0,\cdots,L-1\}$, and
$\texttt{h} \in \mathcal{C}_{i_L}(t)$.
\end{definition}

The pairwise communication model in \eqref{ch05:eq:com} is suitable for
intermittent data exchange, whereas the communication chain in
\eqref{ch05:eq:chain} is needed when the operator must maintain reliable access
to a robot for some duration, for instance in remote supervision or
tele-operation.

\subsubsection{Online and contingent requests}
\label{ch05:subsec:robot-human}

Both the operator and the robots can send and receive online requests through
the communication model above. These requests are contingent in the sense that
they are not known a priori, but arise during exploration as the map evolves,
new features are revealed, and unforeseen operational needs appear. In the
single-operator setting considered here, six request types are modeled across
two directions of interaction.

(I) \emph{Operator-to-robot requests}.
The operator can issue four types of requests.
(i) The first request, denoted by $\xi_1$, is a latency constraint requiring the
operator to obtain the local maps of all robots with delay smaller than a
prescribed threshold $T_{\texttt{h}} > 0$ at all times. Formally,
\begin{equation}
\bigcup_{i \in \mathcal{N}} M_i(t)
\subseteq M_{\texttt{h}}(t + T_{\texttt{h}}),
\quad \forall t \geq 0,
\label{ch05:eq:latency}
\end{equation}
where \eqref{ch05:eq:latency} means that any map information explored by a
robot at time $t$ must be available to the operator no later than time
$t + T_{\texttt{h}}$. Once issued, this request remains active until it is
modified or removed;
(ii) The second request, denoted by $\xi_2$, specifies a prioritized or avoided
region in the workspace. Such a request indicates that a region should either be
explored with priority or avoided during exploration. It is fulfilled once the
corresponding exploration or avoidance requirement has been completed;
(iii) The third request, denoted by $\xi_3$, is an operator-movement request. Here the
operator requests to move to a desired location
$p_{\texttt{h}}^\star \in \mathcal{A}$. The request is fulfilled once the
operator reaches $p_{\texttt{h}}^\star$;
(iv) The fourth request, denoted by $\xi_4$, concerns remote access or
tele-operation. It is written as
$\xi_4 \triangleq (i^\star, p^\star, T_{i^\star})$, where
$i^\star \in \mathcal{N}$ is the target robot, $p^\star \in \mathcal{A}$ is the
required access location, and $T_{i^\star} > 0$ is the required duration. This
request is fulfilled if a communication chain of the form
\eqref{ch05:eq:chain} is established from the operator to robot $i^\star$ and
maintained without interruption for duration $T_{i^\star}$.

(II) \emph{Robot-to-operator requests}.
Any robot can issue two types of requests to the operator.
(i) The fifth request, denoted by $\xi_5^i$, is a confirmation or modification
request sent by robot $i \in \mathcal{N}$. It is written as
$\xi_5^i \triangleq D_i$ and requires the operator to inspect, confirm, or
modify the data held by robot $i$. The request is fulfilled if $D_i$ reaches
the operator and is returned as modified data $D_i^+$;
(ii) The sixth request, denoted by $\xi_6^i$, is a direct-assistance request sent by
robot $i$. It is written as $\xi_6^i \triangleq (p_i, T_i)$, where
$p_i \in M_i$ is the current robot location and $T_i > 0$ is the required
assistance duration. This request is fulfilled if a communication chain from the
operator to robot $i$ is established at location $p_i$ and maintained without
interruption for duration $T_i$.
Accordingly, the set of all requests received by time $t \geq 0$ is denoted by:
\begin{equation}
\Xi(t) \triangleq
\Big\{\xi_1, \xi_2, \xi_3, \xi_4\Big\}
\cup
\Big\{\xi_5^i, \xi_6^i \mid i \in \mathcal{N}\Big\},
\label{ch05:eq:requests}
\end{equation}
where the individual requests are not known beforehand and are removed from
$\Xi(t)$ once fulfilled.

\subsubsection{Local plans and mission objective}
\label{ch05:subsec:problem}

Consequently, the behavior of each robot and the operator can be represented by
a timed sequence of navigation and communication events.
The communication-aware local plan of robot $i \in \mathcal{N}$ is denoted by:
\begin{equation}
\Gamma_i \triangleq
\boldsymbol{\tau}_i^0 \mathbf{c}_i^0
\boldsymbol{\tau}_i^1 \mathbf{c}_i^1
\boldsymbol{\tau}_i^2 \mathbf{c}_i^2 \cdots,
\label{ch05:eq:plan}
\end{equation}
where $\mathbf{c}_i^\ell \triangleq
(p_i^\ell, \iota_i^\ell, D_i^\ell)$ is a communication event defined by the
communication location, the communication partner, and the exchanged data under
\eqref{ch05:eq:com}, $\boldsymbol{\tau}_i^\ell \subset \mathcal{A}$ is the
timed sequence of waypoints visited by robot $i$ between two consecutive
communication events, and $\ell \geq 0$ records the communication round.
The behavior of the operator is defined similarly and is denoted by
$\Gamma_{\texttt{h}}$. For brevity, denote by:
$\widehat{\Gamma} \triangleq
\{\Gamma_i\}_{i \in \mathcal{N}} \cup \{\Gamma_{\texttt{h}}\}$
as the joint communication-aware plan of the full operator-fleet system.
Write $\widehat{\Gamma} \models \Xi(\overline{T})$ if the behaviors of the
robots and the operator collectively fulfill all requests received up to the
terminal time $\overline{T}$.
The overall mission-level objective is then formulated as:
\begin{subequations}
\begin{align}
& \min_{\widehat{\Gamma},\, \overline{T}} \; \overline{T},
\label{ch05:eq:problem_obj}
\\
\text{s.t.} \quad
& \mathcal{A} \subseteq M_{\texttt{h}}(\overline{T}),
\label{ch05:subeq:terminal}
\\
& \eqref{ch05:eq:slam}\text{--}\eqref{ch05:eq:plan},
\quad \widehat{\Gamma} \models \Xi(\overline{T}),
\label{ch05:subeq:system_requests}
\end{align}
\label{ch05:eq:problem}
\end{subequations}
where $\widehat{\Gamma}$ is the joint plan of the operator and all robots, and
$\overline{T} > 0$ is the time at which the environment has been completely
mapped and the resulting map is available to the operator, as required by
\eqref{ch05:subeq:terminal}. Constraint
\eqref{ch05:subeq:system_requests} enforces the operational constraints
together with the fulfillment of all requests received up to time
$\overline{T}$.

Problem~\eqref{ch05:eq:problem} defines the overall mission objective. In the
online setting considered here, however, the request set is revealed only
incrementally during execution, and future operator motions, robot encounters,
and communication opportunities are not known a priori. For this reason, the
coordination framework does not solve \eqref{ch05:eq:problem} directly as a
globally optimal offline planning problem. Instead, it realizes this objective
through a sequence of online feasible coordination decisions over the current
communication backbone, as detailed later in
Section~\ref{ch05:subsec:online}.















%==============================
%==============================
%==============================
\subsection{Latency-bounded intermittent communication protocol with dynamic resizing}
\label{ch05:subsec:spread}

Timely information sharing between the operator and the fleet is essential for
handling online requests. The framework therefore first establishes a
latency-bounded intermittent communication backbone on an embedded joint-wheel
graph, and then augments it with a distributed detach-and-rejoin mechanism for
dynamic resizing. Together, these two components provide the communication
backbone for the request-instantiation and online coordination layers developed
later in this section.

\subsubsection{Latency-bounded intermittent communication protocol}
\label{ch05:subsubsec:spread-problem}

Consider a single operator-fleet system
$\mathcal{T} \triangleq \{\texttt{h}\} \cup \mathcal{N}$, where
$\texttt{h}$ is the operator and $\mathcal{N}$ is the robot fleet. The system
is assigned information-gathering objectives
$\mathcal{F} \triangleq \{f_m\}$, where each task $f_m$ specifies a location to
be visited for information acquisition, such as exploration. Information
collected by any robot $i$ at time $t$ must reach the operator within a
prescribed latency bound $T_{\texttt{h}} > 0$. Let
$\delta_{\texttt{h}}(t) > 0$ denote the maximum information latency over all
robots at time $t$.

\begin{problem}
\label{ch05:prob:latency}
Given the task set $\mathcal{F}$ and the latency bound $T_{\texttt{h}}$,
synthesize the communication-aware local plans $\{\Gamma_i\}_{i \in \mathcal{N}}$ so that the
completion time of $\mathcal{F}$ is minimized and the latency constraint
$\delta_{\texttt{h}}(t) \leq T_{\texttt{h}}$ holds for all $t > 0$.
\end{problem}

The latency bound is enforced through scheduled pairwise communication.
Robots exchange information at communication events and replan the next event
so that a timely return to the operator remains feasible within
$T_{\texttt{h}}$. This mechanism is represented through an embedded
joint-wheel coordination graph.
An embedded graph is a tuple
$\mathcal{G} \triangleq (\mathcal{V}, \mathcal{E}, \sigma, x)$, where
$\mathcal{V}$ is the vertex set, $\mathcal{E} \subseteq
\mathcal{V} \times \mathcal{V}$ is the edge set,
$\sigma \triangleq (\sigma_{\mathcal{V}}, \sigma_{\mathcal{E}})$ assigns vertex
and edge attributes in $(\Sigma_{\mathcal{V}}, \Sigma_{\mathcal{E}})$, and
$x : \mathcal{V} \rightarrow \boldsymbol{X}_{\mathcal{V}}$ is the geometric
embedding.
An embedded joint-wheel coordination graph is an embedded graph
$\mathcal{G}^{\texttt{w}} \triangleq
(\mathcal{V}, \mathcal{E}, \sigma, x)$ such that the vertex set is
$\mathcal{V} \triangleq \mathcal{N} \cup \{\texttt{h}\}$ and the edge set is
$\mathcal{E} \triangleq \mathcal{E}_{\texttt{r}} \cup
\mathcal{E}_{\texttt{h}}$, where $\mathcal{E}_{\texttt{r}}$ forms a ring over
$\mathcal{N}$, namely
$(k_\ell, k_{\ell+1}) \in \mathcal{E}_{\texttt{r}}$ for
$\ell = 1,\cdots,N$ with $k_{N+1} \triangleq k_1$, and
$\mathcal{E}_{\texttt{h}} \triangleq
\{(\texttt{h}, i) \mid i \in \mathcal{N}\}$.
Here, the term
\emph{joint-wheel} refers to joint connectivity over time rather than to a
static wheel graph. At any given time, the protocol activates at most one rim
edge in $\mathcal{E}_{\texttt{r}}$, so only one robot pair communicates.
However, over any time interval longer than $T_{\texttt{h}}$, the union of the
activated rim edges together with the hub edges in
$\mathcal{E}_{\texttt{h}}$ induces a wheel-like connectivity pattern. This is
different from the static line and star graphs used later.
The embedded graph $\mathcal{G}^{\texttt{w}}$ is constructed and maintained in
a fully distributed manner through local coordination at intermittent
communication events, as described next.

(I) \emph{Initialization}.
The vertex embedding records the local coordination state:
\begin{equation}
\sigma_{\mathcal{V}}(i) \triangleq (\Gamma_i, \chi_i), \qquad
\sigma_{\mathcal{V}}(\texttt{h}) \triangleq
(\widehat{p}_{\texttt{h}}, \chi_{\texttt{h}}),
\label{ch05:eq:def-vertex-embed}
\end{equation}
where the communication-aware local plan $\Gamma_i$ specifies the visiting locations and
communication events of robot $i$ as in \eqref{ch05:eq:plan},
$\chi_i \in \mathbb{R}^{N}$ estimates the minimum time stamp of information
from each robot that is guaranteed to reach the operator,
$\widehat{p}_{\texttt{h}} \in \mathbb{R}^{3}$ is the current target location of
the operator, and $\chi_{\texttt{h}} \in \mathbb{R}^{N}$ is the confirmed time
stamp vector available at the operator. The initialization sets
$\Gamma_i = \varnothing$ and $\chi_i = \mathbf{0}$ for all
$i \in \mathcal{N}$, together with
$\widehat{p}_{\texttt{h}} = \varnothing$ and
$\chi_{\texttt{h}} = \mathbf{0}$.
The edge embedding stores the scheduled communication events:
\begin{equation}
\sigma_{\mathcal{E}}(i, j) \triangleq (t_{ij}^{+}, p_{ij}^{+}),
\quad \forall (i, j) \in \mathcal{E}_{\texttt{r}},
\label{ch05:eq:def-edge-embed}
\end{equation}
where $\sigma_{\mathcal{E}}(\texttt{h}, i) \triangleq
(t_{\texttt{h}i}^{+}, p_{\texttt{h}i}^{+})$ is defined similarly for
each $i \in \mathcal{N}$. Initially,
$\sigma_{\mathcal{E}}(i, j) = (0, p_{\texttt{h}})$, since all robots start
near the operator and communicate immediately, whereas
$\sigma_{\mathcal{E}}(\texttt{h}, i) = \varnothing$ indicates that no return is
scheduled initially. Finally, the geometric embedding stores the physical state
and local data of each entity, namely
$x(i) \triangleq (p_i(t), D_i(t))$ and
$x(\texttt{h}) \triangleq (p_{\texttt{h}}(t), D_{\texttt{h}}(t))$.

(II) \emph{Pairwise coordination}.
At a communication event $\mathbf{c}_{ij}$, robots $i$ and $j$ read their
current geometric states $x(i)$ and $x(j)$, exchange the incident embeddings,
and compute the fused local map $\mathcal{M}_{ij}$. The resulting information
exchange yields the merged latency vector
\begin{equation}
\chi_{ij} \triangleq \max\{\chi_i, \chi_j\},
\label{ch05:eq:chi-merge}
\end{equation}
where the maximum is taken component-wise. The subsequent coordination
computation proceeds in three stages, summarized in
Algorithm~\ref{ch05:alg:coordination}.

\textbf{Stage-I: determine the return event}.
Define
$\underline{\chi}_{ij} \triangleq \min_{k \in \mathcal{N}} \chi_{ij}(k)$,
which is the minimum team time stamp guaranteed to reach the operator as
inferred at $\mathbf{c}_{ij}$. Let $t_{\ell}^{L_\ell}$ and
$p_{\ell}^{L_\ell}$ denote the terminal time and terminal location in the
communication-aware local plan $\Gamma_\ell$ of robot $\ell \in \{i, j\}$. The earliest return
time to the operator for a candidate return robot
$\ell \in \{i, j\}$ is estimated by
\begin{equation}
t_{\texttt{h}\ell}^{\star} \triangleq
t_{\ell}^{L_\ell} +
T_{\ell}^{\texttt{nav}}(p_{\ell}^{L_\ell}, \widehat{p}_{\texttt{h}}),
\label{ch05:eq:estret}
\end{equation}
where $T_{\ell}^{\texttt{nav}}(\cdot,\cdot)$ is the navigation-time model. A
return to the operator is unnecessary if
\begin{equation}
\min_{\ell \in \{i,j\}} t_{\texttt{h}\ell}^{\star}
\leq T_{\texttt{h}} + \underline{\chi}_{ij},
\label{ch05:eq:return_cond}
\end{equation}
and otherwise the return robot is selected as
$n \triangleq
\arg\min_{\ell \in \{i,j\}} t_{\texttt{h}\ell}^{\star}$.
Thus, robot $n$ is the robot that returns to the operator immediately after its
last currently planned event.

\textbf{Stage-II: optimize the task sequence and the next
communication event}.
Conditioned on the return decision, the optimization computes the tail plans
$\Gamma_i^{+} \triangleq \tau_i^{+}\mathbf{c}_{ij}^{+}$ and
$\Gamma_j^{+} \triangleq \tau_j^{+}\mathbf{c}_{ij}^{+}$, where
$\tau_i^{+}$ and $\tau_j^{+}$ are the optimized task sequences and
$\mathbf{c}_{ij}^{+}$ is the next communication event by the two robots.
The schedule of $\mathbf{c}_{ij}^{+}$ is denoted by
$(t_{ij}^{+}, p_{ij}^{+})$ and must satisfy the local latency condition:
\begin{equation}
t_{ij}^{+} +
T_{\ell}^{\texttt{nav}}(p_{ij}^{+}, \widehat{p}_{\texttt{h}})
\leq T_{\texttt{h}} + \chi^{\star},
\label{ch05:eq:pairing_cond}
\end{equation}
where $\chi^{\star} \triangleq \underline{\chi}_{ij}$ if no return is scheduled
in Stage-I, and $\chi^{\star}$ is computed from the scheduled return otherwise.
Feasible outputs $(\tau_i^{+}, \tau_j^{+}, \mathbf{c}_{ij}^{+})$ are obtained
by an iterative TSP-based procedure. At each iteration, a traveling-salesman
problem is constructed over the remaining task locations in $\mathcal{F}^{-}$
together with the current terminal locations of robots $i$ and $j$. The edge
weights are given by the corresponding navigation-time model. The TSP solution
yields a joint visiting order $\mathbf{p}_{ij}$, from which
$\tau_i^{+}$ and $\tau_j^{+}$ are extracted, and a candidate communication
event $\mathbf{c}_{ij}^{+}$ is selected along $\mathbf{p}_{ij}$. The event is
accepted only if the feasibility condition in
\eqref{ch05:eq:pairing_cond} holds. Otherwise, one task in
$\tau_i^{+}$ or $\tau_j^{+}$ is removed and the procedure is repeated.

\textbf{Stage-III: update the embeddings}.
After the next communication event has been determined, the current event
$\mathbf{c}_{ij}$ is removed from the plans to obtain the prefixes
$\Gamma_i^{-}$ and $\Gamma_j^{-}$. The updated plans satisfy
$\widehat{\Gamma}_i = \Gamma_i^{-}\Gamma_i^{+}$ and
$\widehat{\Gamma}_j = \Gamma_j^{-}\Gamma_j^{+}$. The latency vectors are
updated by information exchange as
$\widehat{\chi}_i \triangleq \widehat{\chi}_j \triangleq \chi_{ij}$, where
$\chi_{ij}$ is defined in \eqref{ch05:eq:chi-merge}. The local embedding update
is then:
\begin{equation}
\begin{aligned}
\sigma_{\mathcal{V}}(i) &\leftarrow
(\widehat{\Gamma}_i, \widehat{\chi}_i), \qquad
\sigma_{\mathcal{V}}(j) \leftarrow
(\widehat{\Gamma}_j, \widehat{\chi}_j), \\
\sigma_{\mathcal{E}}(i,j) &\leftarrow (t_{ij}^{+}, p_{ij}^{+}), \qquad
\sigma_{\mathcal{E}}(\texttt{h}, n) \leftarrow
(t_{\texttt{h}n}^{\star}, \widehat{p}_{\texttt{h}}),
\end{aligned}
\label{ch05:eq:embed-update}
\end{equation}
where the second line is applied only when Stage-I schedules a return. If no
return is scheduled, the corresponding hub-edge embedding remains
$\varnothing$.
Combined with the joint-wheel communication topology, this pairwise procedure
ensures exploration with bounded latency shown below.


\begin{lemma}
\label{ch05:prop:exist-solution}
The iterative process in Algorithm~\ref{ch05:alg:coordination} terminates in
finite steps and always returns a feasible plan satisfying the meeting-event
constraint \eqref{ch05:eq:pairing_cond} during each pairwise coordination.
\end{lemma}
\begin{proof}
The claim follows from the structure of
Algorithm~\ref{ch05:alg:coordination}. Stage-I is finite, since the
quantities \(t^{\star}_{\texttt{h}\ell}\) are computed directly from
\eqref{ch05:eq:estret} for \(\ell\in\{i,j\}\), and the return decision
is made by the feasibility test \eqref{ch05:eq:return_cond}.
Hence only finitely many operations are required before Stage-II.
It remains to consider the iterative loop in Stage-II. At each
iteration, the candidate next meeting event is accepted only if the
meeting-event constraint \eqref{ch05:eq:pairing_cond} holds.
Otherwise, one task \(f^{\star}\) is removed from the remaining set
\(F^{-}\), and the candidate path and communication event are
recomputed. Since \(F^{-}\) is finite and each unsuccessful iteration
strictly decreases its cardinality, the loop can be repeated only
finitely many times.

Moreover, feasibility is preserved when tasks are removed. If
\eqref{ch05:eq:return_cond} already holds in Stage-I, then the
current tails admit a feasible continuation without forcing an
immediate operator return. If \eqref{ch05:eq:return_cond} fails,
Stage-I explicitly appends a return event for the robot
\(n=\arg\min_{\ell\in\{i,j\}} t^{\star}_{\texttt{h}\ell}\), so that a
hub-reaching continuation is inserted before the latency bound is
violated. Therefore, after finitely many task removals, the remaining
tail contains only a feasible continuation consistent with the return
decision, and the selected event necessarily satisfies
\eqref{ch05:eq:pairing_cond}.
Once such an event is found, Stage-III updates the communication-aware local plans and edge
embeddings according to \eqref{ch05:eq:embed-update}. Hence the
algorithm terminates in finite steps and returns revised communication-aware local plans
whose next meeting event satisfies
\eqref{ch05:eq:pairing_cond}.
\end{proof}

\begin{theorem}
\label{ch05:prop:Ts-constraint}
Under the joint-wheel communication topology and the coordination protocol in
Algorithm~\ref{ch05:alg:coordination}, the operator-fleet latency constraint is
satisfied, namely
$\delta_{\texttt{h}}(t) \leq T_{\texttt{h}}$ for all $t > 0$.
\end{theorem}
\begin{proof}
The proof is by induction over the sequence of scheduled communication
events on the embedded joint-wheel graph.
At initialization, the embeddings are set by
\eqref{ch05:eq:def-vertex-embed}--\eqref{ch05:eq:def-edge-embed}. In
particular, all robots initially communicate near the operator, so the
operator possesses the initial information state and the latency bound
holds at \(t=0\).
Assume now that the bound is satisfied up to some communication event
\(c_{ij}\). At \(c_{ij}\), robots \(i\) and \(j\) exchange their
incident embeddings and fuse their information, producing the merged
latency vector \(\chi_{ij}\) in \eqref{ch05:eq:chi-merge}. Since the
maximum in \eqref{ch05:eq:chi-merge} is taken component-wise, no confirmed
time stamp is lost in this fusion step: every piece of information that
was guaranteed to reach the operator through either robot remains
guaranteed after the meeting.

Stage-I then checks whether a direct operator return is needed. If
\eqref{ch05:eq:return_cond} holds, one of the current tails
already implies operator delivery within the admissible horizon. If
\eqref{ch05:eq:return_cond} fails, the algorithm appends a return
event for the robot \(n\) with smallest \(t^{\star}_{\texttt{h}\ell}\),
computed through \eqref{ch05:eq:estret}. Hence, after Stage-I, the
information available at \(c_{ij}\) is guaranteed to reach the operator
no later than \(\underline{\chi}_{ij}+T_{\texttt{h}}\).
Stage-II then selects the next pairwise event
\(c_{ij}^{+}=(t_{ij}^{+},p_{ij}^{+})\) subject to the local feasibility
constraint \eqref{ch05:eq:pairing_cond}. This condition states exactly
that, from the next meeting location \(p_{ij}^{+}\), at least one of
the two robots can still return to the operator within the residual
latency budget determined by the current guaranteed time stamp. Thus
the pairwise schedule preserves the invariant that all information
known to the communicating pair is still deliverable to the operator
within \(T_{\texttt{h}}\) from its guaranteed time stamp. Stage-III
records this invariant in the updated embeddings through
\eqref{ch05:eq:embed-update}.
Because the communication topology is a joint wheel, every robot lies on
the ring \(E_{\texttt{r}}\), and over any interval longer than
\(T_{\texttt{h}}\) the activated rim edges together with the hub edges
\(E_{\texttt{h}}\) induce a time-respecting wheel-like communication
pattern. Consequently, the above invariant propagates from each local
pairwise coordination step to the whole operator-fleet system: any data
generated by any robot is forwarded through successive scheduled rim
meetings and, when required, a hub return, so that it reaches the
operator within time \(T_{\texttt{h}}\).
Therefore, for every \(t>0\), all information older than
\(t-T_{\texttt{h}}\) has already reached the operator. Equivalently,
the maximum information latency satisfies
\(\delta_{\texttt{h}}(t)\le T_{\texttt{h}}\) for all \(t>0\).
\end{proof}


%==============================
\begin{algorithm}[t]
  \caption{Latency-aware Intermittent Pairwise Coordination:
  $\texttt{IntmCoord}(\cdot)$}
  \label{ch05:alg:coordination}
  \LinesNumbered
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}
\Input{Local plans $\Gamma_i$, $\Gamma_j$, stamp $\chi_{ij}$, map
$\mathcal{M}_{ij}$, and unassigned tasks $\mathcal{F}^{-}$.}
\Output{Revised communication-aware local plans $\widehat{\Gamma}_i$, $\widehat{\Gamma}_j$, and next
event $\mathbf{c}_{ij}$.}

\tcc{\textbf{Determine Return Event}}
Compute $t_{\texttt{h}\ell}^{\star}$ by \eqref{ch05:eq:estret} for
$\ell \in \{i,j\}$\;
\If{condition \eqref{ch05:eq:return_cond} is not satisfied}{
  Determine the return robot by
  $n \triangleq
  \arg\min_{\ell \in \{i,j\}} t_{\texttt{h}\ell}^{\star}$\;
  Append a return event to $\Gamma_n$\;
}

\tcc{\textbf{Optimize Task Sequence and Next Pairing Event}}
$\mathbf{p}_{ij} \leftarrow
\texttt{TSP-Path}(p_i^{L_i}, \mathcal{F}^{-}, p_j^{L_j})$\;
$\mathbf{c}_{ij} \leftarrow
\texttt{SelComm}(\mathbf{c}_i^{L_i}, \mathbf{c}_j^{L_j}, \mathbf{p}_{ij})$\;
\While{condition \eqref{ch05:eq:pairing_cond} is not satisfied}{
  Remove $f^{\star}$ with maximum cost from $\mathcal{F}^{-}$\;
  Recompute $\mathbf{p}_{ij}$ and $\mathbf{c}_{ij}$\;
}
$\boldsymbol{\tau}_i^{+}, \boldsymbol{\tau}_j^{+} \leftarrow
\texttt{Split}(\mathbf{p}_{ij}, \mathbf{c}_{ij})$\;

\tcc{\textbf{Update Local Plans}}
$\Gamma_{\ell}^{-} \leftarrow
\texttt{Remove}(\Gamma_{\ell}, \mathbf{c}_{ij})$,
for $\ell \in \{i,j\}$\;
$\widehat{\Gamma}_{\ell} \leftarrow
\Gamma_{\ell}^{-} + \boldsymbol{\tau}_{\ell}^{+} + \mathbf{c}_{ij}$,
for $\ell \in \{i,j\}$\;
Update embeddings through \eqref{ch05:eq:embed-update}\;
\end{algorithm}
%==============================


\subsubsection{Distributed detach-and-rejoin mechanism}
\label{ch05:sec:detach-join}

Flexible resizing under intermittent communication is enabled through a
distributed detach-and-rejoin procedure that preserves the wheel topology and
respects the latency constraint. A key design choice is that the rewiring is
completed through two consecutive meeting events. In this way, the successor
robots receive consistent plan and adjacency updates before a robot leaves or
rejoins the wheel.

(I) \emph{Detaching from the wheel}.
Consider three consecutive robots $(i,j,k)$ along the wheel, where robot $j$
intends to detach. Let $\Gamma_{\ell}$ denote the nominal communication-aware local plan of robot
$\ell$, and let $\widehat{\Gamma}_{\ell}$ denote the optimized plan maintained
online. At the meeting between robots $i$ and $j$, robot $j$ does not depart
immediately. Instead, using robot $i$'s most recent knowledge of robot $k$'s
plan $\Gamma_k$, robot $i$ computes a bypass schedule that directly connects
$(i,k)$ through
$\widehat{\Gamma}_i,\ \widehat{\Gamma}_k,\ \mathbf{c}_{ik}
= \texttt{IntmCoord}(\Gamma_i,\Gamma_k)$,
and updates the edge embeddings by removing $(i,j)$ and inserting $(i,k)$, that
is,
$\sigma_{\mathcal{E}}(i,j) \leftarrow \emptyset,
\sigma_{\mathcal{E}}(i,k) \leftarrow (t_{ik}, p_{ik})$,
where $(t_{ik}, p_{ik})$ is the time-location pair of the next event
$\mathbf{c}_{ik}$. Robot $j$ then follows its original schedule to meet robot
$k$ and forwards the bypass message so that robot $k$ adopts
$\widehat{\Gamma}_k$ and records the updated predecessor implied by
$\mathbf{c}_{ik}$. Only after this second meeting does robot $j$ finalize the
detachment and depart to its remote location. The detach event is denoted by
$\pi_j \triangleq (j, t_{jk}, p_{jk})$, where $(t_{jk}, p_{jk})$ is the
confirmed meeting between robots $j$ and $k$, which serves as the time and
location at which robot $j$ leaves the current coordination graph.

(II) \emph{Rejoining the wheel}.
To rejoin, robot $j$ first returns to the operator $\texttt{h}$ and waits until
some team robot $m$ also arrives at the operator. Reinsertion is then completed
through two consecutive meetings. Let $\ell$ denote the current successor of
robot $m$ on the wheel. First, robots $m$ and $j$ execute pairwise coordination
at the operator location in order to synchronize plans and schedule the next
event as
$\widehat{\Gamma}_m,\ \widehat{\Gamma}_j,\ \mathbf{c}_{mj}
= \texttt{IntmCoord}(\Gamma_m,\Gamma_j)$.
The edge embeddings are then updated by replacing the successor edge $(m,\ell)$
with $(m,j)$, that is,
$\sigma_{\mathcal{E}}(m,\ell) \leftarrow \emptyset,
\sigma_{\mathcal{E}}(m,j) \leftarrow (t_{mj}, p_{mj})$,
where $(t_{mj}, p_{mj})$ is taken from the event $\mathbf{c}_{mj}$. Second,
robot $j$ meets robot $\ell$ in order to establish the new successor relation
through $\texttt{IntmCoord}(\Gamma_j,\Gamma_{\ell})$, and the edge embedding
$\sigma_{\mathcal{E}}(j,\ell) \leftarrow (t_{j\ell}, p_{j\ell})$ is added.
This two-step reinsertion prevents inconsistent topology updates and restores a
valid wheel graph.

When several robots detach or rejoin, the same logic extends by increasing the
number of required forwarding meetings. For detachment, nonconsecutive robots
may leave sequentially through repeated use of the same two-meeting handoff. If
a consecutive block detaches, a single bypass plan is computed by the
predecessor of the block and forwarded across the block until it reaches the
first non-detaching successor. Each intermediate robot forwards once and then
exits. For rejoining, a block waiting at the operator can be inserted between a
returning robot $m$ and its successor through consecutive pairwise coordination,
and the last inserted robot then performs the successor take-over meeting that
closes the wheel.
