\section{Online Adaptation for Continual Tasks}
\label{ch04:sec:online_continual_tasks}

The previous sections considered mission decomposition, team formation, and
local coordination for mission sets that are fixed during each planning stage.
Continual operation requires new mission formulas to be inserted after part of
the current plan has already been executed, while preserving that valid prefix.
Rebuilding the conjunction of all active formulas is generally expensive because
automaton construction, partial-order extraction, and team assignment grow with
the active mission set
\citep{schillinger2018simultaneous,kantaros2020stylus,luo2022temporal,
liu2024time,luo2025hulk}. This section therefore maintains the high-level task
structure online and updates only the portion affected by new missions, using
R-poset products, receding-horizon replanning, and request-driven updates.


%==============================
%==============================
\subsection{Online product of R-posets}
\label{ch04:subsec:online_poset_product}

An online product operation incorporates a new mission without rebuilding the
entire global task structure. The new formula is translated into a B\"uchi
automaton, converted into R-posets by
Section~\ref{ch04:sec:po_decomp_assignment}, and merged with the current global
R-poset so that the executed prefix is preserved and only the residual structure
is updated.
Let the current global R-poset be
$P_1 \triangleq (\Omega_1,\preceq_1,\neq_1)$, and let the R-poset induced by
the new mission be
$P_2 \triangleq (\Omega_2,\preceq_2,\neq_2)$. Let $w_0$ be the finite executed
word of the current plan, and let
$\Omega_1^{\mathrm{f}} \subseteq \Omega_1$ be the set of subtasks already
completed in $w_0$. The role of $w_0$ is only to fix the executed prefix of
$P_1$. It does not impose additional constraints on the future subtasks in
$P_2$. When the new mission is introduced before any execution starts, $w_0$ is
the empty word.

\begin{definition}
\label{ch04:def:product_of_r_posets}
Given a finite executed word $w_0$, the product of two R-posets
$P_1$ and $P_2$ is a set of R-posets defined as:
\begin{equation}
\label{ch04:eq:product_of_r_posets}
\mathcal{P}^{r} \triangleq P_1 \otimes_{w_0} P_2,
\end{equation}
such that the following two conditions hold:
(I) if $w_0w \in L(P_1)$ and $w \in L(P_2)$, then
$w_0w \in \bigcup_{P' \in \mathcal{P}^{r}} L(P')$;
(II) if $w_0w \in L(P')$ for some $P' \in \mathcal{P}^{r}$, then
$w_0w \in L(P_1)$ and $w \in L(P_2)$.
\end{definition}

Definition~\ref{ch04:def:product_of_r_posets} preserves exactly the
continuations compatible with both the current and newly introduced task
structures. The construction has two stages: subtask composition and partial-
relation calculation.

(I) \emph{Composition of subtasks}.
The first stage generates candidate subtask sets satisfying both $\Omega_1$ and
$\Omega_2$. For two subtasks
$\omega'=(j,\sigma',{\sigma'}^{s})$ and
$\omega=(i,\sigma,\sigma^{s})$, write $\omega' \models \omega$ if
$\sigma \subseteq \sigma'$ and $\sigma^{s}\subseteq {\sigma'}^{s}$. This means
that $\omega'$ satisfies all action and self-loop requirements of $\omega$.
The construction starts from $\Omega' \triangleq \Omega_1$. Then $\Omega'$
trivially satisfies $\Omega_1$. A partial mapping
$M_{\Omega} : \Omega_2 \rightharpoonup \Omega'$ is introduced to record which
subtasks in $\Omega'$ satisfy which subtasks in $\Omega_2$. Its domain and
range are denoted by $\operatorname{dom}(M_{\Omega})$ and
$\operatorname{ran}(M_{\Omega})$, respectively. Initially, $M_{\Omega}$ is
empty. Whenever $\omega'_j \models \omega_i^2$, the relation
$M_{\Omega}(\omega_i^2)=\omega'_j$ is recorded.
A depth-first search~\citep{kozen1992depth} is then used to process the subtasks
of $\Omega_2$ in order. The search state is a pair $(\Omega',M_{\Omega})$, and
the initial queue is
$que=[(\Omega'=\Omega_1,\,M_{\Omega}=\emptyset)]$. Suppose the current search
state is $(\Omega',M'_{\Omega})$, and let the next unmapped subtask of
$\Omega_2$ be $\omega_i^2$, where
$i=|\operatorname{dom}(M'_{\Omega})|+1$. If there exists a subtask
$\omega'_j=(j,\sigma'_j,{\sigma'}_j^{s})$ in
$\Omega'\setminus
(\operatorname{ran}(M'_{\Omega})\cup\Omega_1^{\mathrm{f}})$ such that
$\omega'_j \models \omega_i^2$ or $\omega_i^2 \models \omega'_j$, then a new
candidate composition is created by merging the two subtasks:
\begin{equation}
\label{ch04:eq:generate_new_subtask}
\hat{\omega}'_j \triangleq
\bigl(j,\sigma'_j\cup\sigma_i^2,{\sigma'}_j^{s}\cup{\sigma_i^{s}}^{2}\bigr),\
\hat{\Omega}' \triangleq
\bigl(\Omega' \setminus \{\omega'_j\}\bigr)\cup\{\hat{\omega}'_j\},\
\hat{M}'_{\Omega}(\omega_i^2) \triangleq \hat{\omega}'_j,
\end{equation}
where all remaining entries of $\hat{M}'_{\Omega}$ are inherited from
$M'_{\Omega}$.
If no suitable existing subtask is used, then $\omega_i^2$ is appended as a
new subtask:
\begin{equation}
\label{ch04:eq:generate_new_subtask2}
j \triangleq |\Omega'|+1,\
\hat{\omega}'_j \triangleq \omega_i^2,\
\hat{\Omega}' \triangleq \Omega' \cup \{\hat{\omega}'_j\},\
\hat{M}'_{\Omega}(\omega_i^2) \triangleq \hat{\omega}'_j,
\end{equation}
where all existing mapping entries are again preserved. Once
$|\operatorname{dom}(M_{\Omega})| = |\Omega_2|$, a candidate subtask set
$\Omega'$ satisfying both $\Omega_1$ and $\Omega_2$ has been obtained. The
second stage is then triggered.

(II) \emph{Calculation of partial relations}.
Given a candidate subtask set $\Omega'$ and the corresponding mapping
$M_{\Omega}$, the precedence relation is first initialized by inheriting the
precedence constraints from both input R-posets:
\begin{equation}
\label{ch04:eq:inherit_precedence}
\preceq \triangleq
\preceq_1 \cup
\Bigl\{
\bigl(M_{\Omega}(\omega_{\ell_1}^2),M_{\Omega}(\omega_{\ell_2}^2)\bigr)
\mid
(\omega_{\ell_1}^2,\omega_{\ell_2}^2)\in\preceq_2
\Bigr\},
\end{equation}
where the inherited precedence may introduce successor constraints that were not
present in one of the original R-posets. For each subtask
$\omega_i=(i,\sigma_i,\sigma_i^s)\in\Omega'$, let $\omega_i^1$ denote the
subtask of $\Omega_1$ represented by $\omega_i$ whenever such a source exists,
and let $\omega_i^2 \triangleq M_{\Omega}^{-1}(\omega_i)$ whenever
$\omega_i \in \texttt{ran}(M_{\Omega})$. The newly induced successors of
$\omega_i$ relative to $P_1$ and $P_2$ are then collected by:
\begin{equation}
\label{ch04:eq:new_partial_order}
\begin{aligned}
S_i^1 &\triangleq
\bigl\{
\omega_j \in \Omega'
\mid
(\omega_i,\omega_j)\in\preceq,\,
\omega_i^1,\omega_j^1 \text{ exist},\,
(\omega_i^1,\omega_j^1)\notin\preceq_1
\bigr\},\\
S_i^2 &\triangleq
\bigl\{
\omega_\ell^2 \in \Omega_2
\mid
\omega_i^2 \text{ exists},\,
(\omega_i,M_{\Omega}(\omega_\ell^2))\in\preceq,\,
(\omega_i^2,\omega_\ell^2)\notin\preceq_2
\bigr\}.
\end{aligned}
\end{equation}
To preserve the self-loop constraints of these newly induced successors, the
labels of $\omega_i$ are updated by:
\begin{equation}
\label{ch04:eq:omega_task_update}
\hat{\sigma}_i \triangleq
\bigcup_{\omega_\ell^1\in S_i^1}{\sigma_\ell^{s1}}
\cup
\bigcup_{\omega_\ell^2\in S_i^2}{\sigma_\ell^{s2}},\
\sigma_i \leftarrow \sigma_i \cup \hat{\sigma}_i,\
\sigma_i^s \leftarrow \sigma_i^s \cup \hat{\sigma}_i,
\end{equation}
which guarantees that the additional self-loop requirements inherited from
$P_1$ and $P_2$ are respected before executing $\omega_i$.
After this update, further precedence constraints may still be necessary. If
$\sigma_j$ conflicts with the self-loop requirement $\sigma_i^s$ of another
subtask $\omega_i$, and $(\omega_i,\omega_j)\notin\preceq$, then $\omega_i$
must precede $\omega_j$. This yields the augmentation below:
\begin{equation}
\label{ch04:eq:neq_update}
\preceq \leftarrow
\preceq \cup
\bigl\{
(\omega_i,\omega_j)\mid
(\omega_i,\omega_j)\notin\preceq,\,
\sigma_j\cap\sigma_i^s\neq\emptyset
\bigr\},
\end{equation}
after which \eqref{ch04:eq:new_partial_order} and
\eqref{ch04:eq:omega_task_update} are applied again if necessary.
For subtasks that remain unordered in $\preceq$, the opposition
relation is obtained by mapping the opposition sets of $P_2$ into $\Omega'$:
\begin{equation}
\label{ch04:eq:neq_calculate}
\neq \triangleq
\neq_1 \cup
\Bigl\{
\{M_{\Omega}(\omega)\mid \omega\in\widehat{\Omega}^2\}
\ \Big|\
\widehat{\Omega}^2\in\neq_2
\Bigr\},
\end{equation}
where the resulting product R-poset
$P' \triangleq (\Omega',\preceq,\neq)$ is then added to the product set
$\mathcal{P}^{r}$.

The two-stage construction is anytime: it may stop after the first feasible
product R-poset or continue until the queue or time budget is exhausted. Repeating
this update whenever a mission arrives maintains the R-poset family while
preserving the executed prefix.









%==============================
%==============================
\subsection{Online execution and receding-horizon adaptation}
\label{ch04:subsec:online_execution}

After decomposition, team instantiation, and local coordination, the resulting
plans are executed online. Let
$\overline{\mathcal{N}}=\{\mathcal{N}_1,\cdots,\mathcal{N}_{K^\star}\}$ be the
current team formation, and let
$\Gamma_k=(\omega_k^1,\omega_k^2,\cdots,\omega_k^{L_k})$ be the task sequence
assigned to team $\mathcal{N}_k$. For each team, the corresponding robot-level
task sequence is induced by \eqref{ch04:eq:robot-plan}. Online adaptation is
then organized around three components: execution of the current plan,
receding-horizon replanning, and status updates for the next planning
cycle.

\subsubsection{Online execution of the current team plans}
\label{ch04:subsubsec:online_exec_model}

A team alternates among navigation, synchronization, and action execution. The
robots move to $S_k^\ell$, synchronize when simultaneous actions or cross-task
waiting are required, and then execute the local subtasks in $\Xi_k^\ell$ using
the strategy of Section~\ref{ch04:subsec:local_coordination}. Execution is
concurrent across teams and parallel within teams when permitted by the temporal
structure. Tasks that have not started may be reassigned, whereas tasks already
under execution are not preempted; replanning therefore affects only the
unfinished and not-yet-started portion of the plan.
For instances, suppose the teams
$\mathcal{N}_1$ and $\mathcal{N}_2$ are formed from the corresponding capacity
profiles obtained in the team-formation stage. After task $\omega_2$, team
$\mathcal{N}_2$ must wait until team $\mathcal{N}_1$ reaches the required
coordination point for task $\omega_5$. This waiting interval is generated by
the synchronization state and enforces the temporal coupling between the two
tasks.

\subsubsection{Receding-horizon replanning upon new missions}
\label{ch04:subsubsec:receding_horizon_replan}

Let $H>0$ denote the receding-horizon length. At each planning cycle, only $H$
tasks are committed for detailed assignment and team formation; the remainder is
deferred, which keeps each update finite. Replanning is triggered by sufficient
progress, new missions, or infeasibility caused by failure or resource loss, and
uses the current system state. The candidate pool contains assigned-but-not-
started tasks, unassigned tasks, and newly introduced tasks; tasks already under
execution are excluded by non-preemption. Let $\Omega^{\mathrm{ns}}$,
$\Omega^{\mathrm{ua}}$, and $\Omega^{\mathrm{new}}$ denote these three sets. The
replanning pool is then given by:
\begin{equation}
\label{ch04:eq:replanning_pool}
\Omega^{\mathrm{rh}} \triangleq
\Omega^{\mathrm{ns}} \cup
\Omega^{\mathrm{ua}} \cup
\Omega^{\mathrm{new}},
\end{equation}
where only tasks in \eqref{ch04:eq:replanning_pool} are reconsidered in the
next assignment cycle.

\subsubsection{Rolling update of robot status}
\label{ch04:subsubsec:rolling_state_update}

The rolling state of each robot is updated from the current execution status.
Suppose the current team sequence is
$(\omega_k^0,\cdots,\omega_k^{L_k})$, and that $\omega_k^{L_k}$ is the task
currently in execution. Then, for each robot $i\in\mathcal{N}_k$, the rolling
availability time and rolling position are updated by:
\begin{equation}
\label{ch04:eq:rolling_state}
\widehat{t}_i \triangleq t_{\mathrm{e}}(\omega_k^{L_k}), \qquad
\widehat{x}_i \triangleq S_k^{L_k},
\end{equation}
where $t_{\mathrm{e}}(\omega_k^{L_k})$ is the estimated completion time of the
currently executing task. The state update in
\eqref{ch04:eq:rolling_state} ensures that the next team-formation and
assignment cycle remains consistent with the executed prefix of the current
plan.
The horizon length $H$, redundancy margin $\alpha_j$, and replanning threshold
control the tradeoff between look-ahead efficiency and responsiveness. The
mechanism preserves the executed prefix, avoids preemption of ongoing tasks, and
keeps each replanning step finite.




















%==============================
\subsection{Human requests as online task updates}
\label{ch04:subsec:human_requests}

During continual operation, a human operator may insert, remove, or modify
missions. These interventions are modeled as online task updates that change the
active mission set or admissible fleet allocation, allowing the same R-poset
product and receding-horizon mechanisms to be reused
\citep{ferreira2021survey,swamy2020scaledautonomy}. Let $\Phi_t$ denote the
active missions at time $t\geq0$; an operator request changes either $\Phi_t$ or
the constraints under which it is executed.

\subsubsection{Request model}
\label{ch04:subsubsec:request_model}


Let $\Phi_t$ be the set of missions that remain active at time $t$. An online
request at time $t$ is one of the following four types.
(I) A mission-insertion request is
$\kappa_1(t) \triangleq (\varphi_t,t)$, where $\varphi_t$ is a newly specified
mission formula of the form in \eqref{ch04:eq:mission_formula}.
(II) A mission-cancellation request is
$\kappa_2(t) \triangleq \{\varphi_\ell\}$, where
$\varphi_\ell \in \Phi_t$ is an active mission to be removed.
(III) A deadline or priority update is
$\kappa_3(t) \triangleq \{(\varphi_\ell,d_\ell^\star,\varpi_\ell^\star)\}$,
where $\varphi_\ell \in \Phi_t$, $d_\ell^\star > 0$ is the revised deadline,
and $\varpi_\ell^\star > 0$ is the revised priority weight.
(IV) A robot-reassignment request is
$\kappa_4(t) \triangleq \{(\mathcal{N}_\ell,\varphi_\ell)\}$, where
$\mathcal{N}_\ell \subseteq \mathcal{N}$ is the robot subset designated for
mission $\varphi_\ell \in \Phi_t$.
The set of active requests at time $t$ is denoted by:
\begin{equation}
\label{ch04:eq:request_set}
\mathcal{U}(t) \triangleq
\{\kappa(t) \mid \kappa(t) \text{ is active at time } t\},
\end{equation}
where each $\kappa(t)$ is of one of the four types above, and completed or
inactive requests are removed from $\mathcal{U}(t)$.
The request model reuses the uncertain-mission model of
Section~\ref{ch04:subsec:uncertain_mission}; newly inserted missions may still
have partially known subtasks, locations, or durations. The four request types
respectively add a temporal objective, remove an obsolete objective, change a
mission's urgency, or constrain the admissible use of fleet resources.

\subsubsection{Integration into online planning}
\label{ch04:subsubsec:request_integration}

Each request in \eqref{ch04:eq:request_set} is handled by the same online
pipeline. A new mission request $\kappa_1$ is translated into a B\"uchi
automaton, converted into R-posets by Section~\ref{ch04:sec:po_decomp_assignment},
and merged with the current global R-poset through
\eqref{ch04:eq:product_of_r_posets}, preserving the executed prefix.
(I) A cancellation request $\kappa_2$ removes the corresponding mission from the
active set $\Phi_t$ and deletes its contribution from the currently maintained
task structure. Since the execution process in
Section~\ref{ch04:subsec:online_execution} is non-preemptive, tasks that are
already under execution are not interrupted. The cancellation therefore affects
only the residual plan and the future replanning pool.
(II) A deadline or priority update $\kappa_3$ changes the urgency of an existing
mission without changing its logical specification. In the abstract assignment
stage, this modifies the mission weight in the search-value function
\eqref{ch04:eq:search_value}. More urgent missions are therefore favored in the
subsequent node selection, while lower-priority missions may be deferred to a
later planning cycle. In the receding-horizon layer, the updated mission is
reinserted into the replanning pool
\eqref{ch04:eq:replanning_pool} together with the other unfinished tasks.
(III) A robot-reassignment request $\kappa_4$ modifies the admissible team-formation
constraints. More precisely, it imposes additional restrictions on the binary
membership variables used in the team-formation model of
Subsubsection~\ref{ch04:subsubsec:team_formation}. These restrictions may lock
certain robots to a designated mission or exclude them from other teams. The
subsequent team-formation step is then recomputed from the rolling robot state
in \eqref{ch04:eq:rolling_state}, so that the revised assignment remains
consistent with the current execution status.
Conflicting requests may make the online problem infeasible, for example when a
high-priority mission is combined with a reassignment that removes the required
capabilities. Such infeasibility is detected at the online-product, abstract-
assignment, or concrete team-formation stage. The planner then preserves the
current non-preemptive prefix and reports that $\mathcal{U}(t)$ is inconsistent
with the active mission and capacity constraints.
