%==============================
\section{Hierarchical Coordination of Large-Scale Fleets}\label{ch04:subsec:task}

Temporal-task coordination for large heterogeneous fleets must address two
coupled difficulties. The first is logical: a mission typically imposes
ordering, eventual completion, concurrency, and collaborative-execution
requirements across multiple tasks. The second is combinatorial: concrete
robots with different capabilities, current locations, and availability times
must be grouped into teams and deployed over a shared workspace. Monolithic
methods that synchronize temporal automata with all robot models can preserve
formal correctness, but their size grows rapidly with the numbers of robots and
tasks
\cite{kantaros2020stylus,schillinger2018simultaneous,luo2022temporal,
leahy2021scalable}. Hierarchical decompositions are therefore essential for
large fleets
\cite{chen2025real,liu2024time,luo2025simultaneous,gosrich2025online,
luo2025hulk}.
This section develops a three-layer coordination architecture for that
setting. The upper layer performs temporal decomposition together with abstract
team assignment. The intermediate layer instantiates these abstract teams by
assigning concrete robots to capability profiles. The lower layer then realizes
each team plan through local coordination and motion generation. Human-centric
protocols are omitted here, and online adaptation is mentioned only when it
helps clarify the hierarchical construction. The objective of this section is
to formalize how mission-level temporal tasks are converted into executable
team-level and robot-level plans.

%==============================
%==============================
\subsection{Temporal Mission with Uncertain Subtasks}
\label{ch04:subsec:uncertain_mission}

In contrast to the previous
section, where the starting point was one fully specified global task, the
present setting considers a collection of mission formulas over collaborative
tasks whose internal subtask structures may not be completely known in
advance. The purpose of the formulation is therefore to describe the fleet, the
mission set, and the optimization objective in a form that is suitable for
hierarchical decomposition and team-level coordination.
Consider a fleet of $N$ robots operating in a shared workspace
$\mathcal{W}\subset\mathbb{R}^{3}$, and let
$\mathcal{N}\triangleq\{1,\cdots,N\}$ be the robot index set. Each robot
$i\in\mathcal{N}$ is characterized by its current position
$x_i\in\mathcal{W}$, a reference velocity $v_i\in\mathbb{R}^{3}$, and a set of
primitive actions $\mathcal{A}_i$. The local plan of robot $i$ is represented
by a timed sequence:
\begin{equation}
\label{ch04:eq:tau}
\tau_i \triangleq
(t_i^1,p_i^1,a_i^1)(t_i^2,p_i^2,a_i^2)\cdots,
\end{equation}
where $t_i^\ell\geq 0$ is the execution time, $p_i^\ell\in\mathcal{W}$ is the
target location, and $a_i^\ell\in\mathcal{A}_i$ is the action to be executed.
Thus, robot $i$ moves to $p_i^\ell$ and initiates action $a_i^\ell$ at time
$t_i^\ell$ for each $\ell\geq 1$.

Let $\Phi \triangleq \{\varphi_1,\cdots,\varphi_M\}$ be a finite set of
missions. Each mission $\varphi_m\in\Phi$ is an sc-LTL formula over a
collaborative-task alphabet
$\boldsymbol{\omega}_m \triangleq
\{\omega_{m,1},\cdots,\omega_{m,K_m}\}$, namely
\begin{equation}
\label{ch04:eq:mission_formula}
\varphi_m \in \mathrm{sc\mbox{-}LTL}(\boldsymbol{\omega}_m),
\end{equation}
where the syntax and semantics follow
Chapter~\ref{chap:preliminaries}. Each collaborative task
$\omega_{m,k}\in\boldsymbol{\omega}_m$ is characterized by an execution region
$S_{m,k}\subset\mathcal{W}$, a duration estimator
$\eta_{m,k}(\cdot)$, and a collection of required subtasks defined as:
\[
\Xi_{m,k}\triangleq
\{(n_{m,k}^{j},a_{m,k}^{j},s_{m,k}^{j})\}_{j=1}^{J_{m,k}},
\]
where each tuple $(n_{m,k}^{j},a_{m,k}^{j},s_{m,k}^{j})$ specifies that at
least $n_{m,k}^{j}$ robots are required to perform action
$a_{m,k}^{j}$ at location $s_{m,k}^{j}$. Accordingly,
$\omega_{m,k}$ is written as
$\omega_{m,k}\triangleq(S_{m,k},\eta_{m,k},\Xi_{m,k})$.
The uncertainty lies in the internal subtask structure of its
collaborative tasks. In particular, the number of required subtasks, their
precise execution locations, or their effective durations may not be fully
specified when the mission formula is first given. This differs from temporal
task models with fixed deterministic task alphabets
\cite{sahin2019multirobot}. The duration estimator
$\eta_{m,k}(n,a,\mathcal{N}')$ depends on the assigned robot subset
$\mathcal{N}'\subseteq\mathcal{N}$ and therefore captures the fact that task
duration varies with team composition and may exhibit saturation effects
\cite{ferreira2021survey,krizmancic2022cooperative_ag,apt2009generic,liu2024time}.

Let $t_0$ denote the initial planning time, and let $t_m^{\mathrm{f}}$ denote
the first time at which mission $\varphi_m$ is satisfied. The completion time
of mission $\varphi_m$ is defined by
$T_m \triangleq t_m^{\mathrm{f}}-t_0$.
Thus, the fleet objective is to synthesize collective plans
$\{\tau_i\}_{i\in\mathcal{N}}$ that satisfy all missions in $\Phi$ while
minimizing the average mission completion time, i.e.,
\begin{equation}
\label{ch04:eq:mission_set_objective}
\min_{\{\tau_i\}_{i\in\mathcal{N}}}
\frac{1}{M}\sum_{m=1}^{M} T_m.
\end{equation}

\begin{problem}
\label{ch04:prob:uncertain_mission_set}
Given the fleet model in \eqref{ch04:eq:tau} and the mission set
$\Phi=\{\varphi_1,\cdots,\varphi_M\}$ with mission formulas in
\eqref{ch04:eq:mission_formula}, synthesize collective plans
$\{\tau_i\}_{i\in\mathcal{N}}$ such that all missions in $\Phi$ are satisfied
and the average mission completion time in
\eqref{ch04:eq:mission_set_objective} is minimized.
\end{problem}

The formulation above combines temporal mission specifications, uncertain
collaborative subtasks, and fleet-level execution constraints in one planning
problem. Direct centralized formulations become rapidly intractable as the
numbers of missions, tasks, and robots increase. This motivates the
hierarchical treatment developed in the remainder of this section.
















%==============================
%==============================
\subsection{Simultaneous Task Decomposition and Team Assignment}
\label{ch04:subsubsec:simultaneous_assignment}

Consider the finite mission set
$\Phi \triangleq \{\varphi_1,\cdots,\varphi_M\}$, where each mission is given
by a temporal formula over collaborative tasks with uncertain internal
subtasks. The next step is to decompose these missions into executable task
symbols and assign them to abstract teams at the same time. Existing methods
for temporal-task coordination often rely on the synchronized product between
a mission automaton and a global fleet model obtained from all local robot
models
\cite{schillinger2018simultaneous,luo2021temporal}. Although this construction
preserves correctness, its size grows rapidly with both the number of robots
and the number of missions. Improved scalability has been pursued through
local coordination~\cite{lindemann2019coupled}, decision trees
\cite{chen2025real}, distributed sampling~\cite{kantaros2020stylus},
hierarchical decoupling~\cite{luo2025simultaneous}, and partial-order
reasoning~\cite{liu2024time}. The present subsection adopts a different
hierarchical viewpoint, in which mission decomposition and team
assignment are carried out simultaneously within one automaton-guided search.
For each mission $\varphi_m \in \Phi$, let
\[
\mathcal{B}_{\varphi_m} \triangleq
(Q^m,Q_0^m,\Sigma^m,\delta^m,Q_F^m)
\]
be the B\"uchi automaton associated with $\varphi_m$, defined as in
Definition~\ref{ch02:def:NBA}. Let
$\mathcal{M} \triangleq \{1,\cdots,M\}$ be the mission index set. The fleet is
partitioned into $K$ abstract teams indexed by
$\mathcal{K} \triangleq \{1,\cdots,K\}$, where the value of $K$ is not fixed in
advance. To avoid conflict with the collaborative behaviors $C_k$ introduced in
Section~\ref{ch04:sec:po_decomp_assignment}, the teams are denoted here by
$\mathcal{T}_k$. Each team is described by an abstract capability profile to be
instantiated later, rather than by a fixed set of robot identities. Let
$\Gamma_k$ denote the local plan of team $\mathcal{T}_k$, written as a sequence
of task symbols assigned to that team.

\begin{problem}
\label{ch04:prob:simultaneous_assignment}
Given the mission set $\Phi$, determine the number of teams $K$, the capability
profiles of the teams $\{\mathcal{T}_k\}_{k\in\mathcal{K}}$, and the local
plans $\{\Gamma_k\}_{k\in\mathcal{K}}$ such that all mission specifications are
satisfied under the temporal and fleet-capacity constraints, and the mission
execution cost is \emph{minimized}.
\end{problem}

The search is organized as a rooted tree
$\mathfrak{T} \triangleq (\mathcal{V},\rightarrow)$,
where $\mathcal{V}$ is the node set and
$\rightarrow \subseteq \mathcal{V}\times\mathcal{V}$ is the edge relation.
Each node
$\nu \triangleq
\bigl(\{\Gamma_k\}_{k\in\mathcal{K}},
\{\widehat{Q}_m\}_{m\in\mathcal{M}}\bigr)$
contains two components: the current local plans of all teams and the sets of
currently reachable automaton states for all missions. The root node is
$\nu_0 \triangleq
\bigl(\emptyset,\{\widehat{Q}_m^0\}_{m\in\mathcal{M}}\bigr)$,
where $\widehat{Q}_m^0 \triangleq Q_0^m$. As summarized in
Algorithm~\ref{ch04:alg:ss-TaskAssign} and illustrated in
Fig.~\ref{ch04:fig:Tree}, the procedure consists of four stages: selection,
expansion, bounding, and termination, as detailed below.


\emph{(I) Selection.}
At each iteration, a set of $P>0$ candidate nodes is selected for parallel
expansion. Denote this set by
$\mathcal{V}_P \triangleq \{\nu_1^\star,\cdots,\nu_P^\star\}$, where the nodes
are chosen from the current frontier in ascending order of a value function
$\chi:\mathcal{V}\rightarrow\mathbb{R}_{\ge 0}$. This value function is defined
by:
\begin{equation}
\label{ch04:eq:search_value}
\chi(\nu) \triangleq
\max_{k\in\mathcal{K}} T_k
+ \eta_1 \sum_{k\in\mathcal{K}} \kappa_k
+ \eta_2 \sum_{m\in\mathcal{M}}
\min_{q_m\in\widehat{Q}_m}\psi_m(q_m,Q_F^m),
\end{equation}
where $\eta_1,\eta_2>0$ are weighting parameters, $T_k$ is the ending time of
the local plan $\Gamma_k$, $\kappa_k$ is the estimated cost of $\Gamma_k$, and
$\psi_m(q_m,Q_F^m)$ is the length of the shortest path from state $q_m$ to the
accepting set $Q_F^m$ in $\mathcal{B}_{\varphi_m}$. The first term in
\eqref{ch04:eq:search_value} estimates the current makespan, the second
penalizes costly assignments, and the third favors nodes with greater mission
progress. The weights $\eta_1$ and $\eta_2$ regularize the search order and
mainly affect the convergence rate rather than feasibility.
For each team $\mathcal{T}_k$, the local plan $\Gamma_k$ induces an abstract
capacity profile below:
\begin{equation}
\label{ch04:eq:capacity-team}
\Lambda_k \triangleq
\{(\beta_k^j,a^j)\mid a^j\in\mathcal{A},\ \beta_k^j>0\},
\end{equation}
where $\mathcal{A} \triangleq \bigcup_{i\in\mathcal{N}}\mathcal{A}_i$ is the
global action set, and $\beta_k^j$ is the minimum number of robots capable of
action $a^j$ required by the tasks currently assigned to $\Gamma_k$. A node is
declared infeasible if the induced team requirements exceed the fleet
capability, that is, if
\begin{equation}
\label{ch04:eq:capacity-bound}
\sum_{k\in\mathcal{K}} \beta_k^j
\leq
\sum_{i\in\mathcal{N}} \mathds{1}(a^j\in\mathcal{A}_i),
\qquad \forall a^j\in\mathcal{A},
\end{equation}
is violated,
where the left-hand side is the required capacity and the right-hand side is
the available fleet capacity.

\emph{(II) Expansion.}
Each selected node $\nu_h^\star\in\mathcal{V}_P$ is expanded by assigning one
additional task symbol to one team. The candidate task set is given by:
\begin{equation}
\label{ch04:eq:task-candidate}
\Omega^-_{\nu_h^\star} \triangleq
\bigcup_{m\in\mathcal{M}}
\Bigl\{
\omega\in\Sigma^m \,\Big|\, \exists q_m\in\widehat{Q}_m :
\delta^m(q_m,\omega)\neq\emptyset
\Bigr\},
\end{equation}
where $\delta^m:Q^m\times\Sigma^m\to 2^{Q^m}$ is the transition function of
$\mathcal{B}_{\varphi_m}$. Therefore, a task symbol is eligible only if it
enables at least one valid automaton transition for some mission.
For each $\omega\in\Omega^-_{\nu_h^\star}$ and each team index
$k\in\mathcal{K}$, a child node is constructed as:
\begin{equation}
\label{ch04:eq:expand}
\nu^+ \triangleq
\Bigl(
\{\Gamma_1,\cdots,\Gamma_k^+,\cdots,\Gamma_K\},
\{\widehat{Q}_m^+\}_{m\in\mathcal{M}}
\Bigr),
\end{equation}
where $\Gamma_k^+$ is obtained by appending $\omega$ to $\Gamma_k$, and
$\widehat{Q}_m^+ \triangleq
\bigcup_{q_m\in\widehat{Q}_m}\delta^m(q_m,\omega)$
is the updated set of reachable states for mission $\varphi_m$. The edge
$(\nu_h^\star,\nu^+)$ is then added to the search tree $\mathfrak{T}$. If
$\Omega^-_{\nu_h^\star}=\emptyset$, the node admits no further expansion.

A new team can also be created during expansion. In that case, the index set is
enlarged to $\mathcal{K}\cup\{K+1\}$, an empty local plan is introduced for the
new team, and the selected task symbol is assigned to it. In this way, team
formation and task assignment emerge jointly from the search rather than being
fixed in advance.
After assigning a task $\omega_k^\ell$ to team $k$, the capacity profile is
updated by:
\begin{equation}
\label{ch04:eq:update-constraint}
\beta_k^j \triangleq
\max\{n_r \mid \exists \omega\in\Gamma_k
\text{ such that } (n_r,a^j,s_r)\in\omega\},
\qquad \forall a^j\in\mathcal{A},
\end{equation}
where $\beta_k^j$ is set to zero if no task in $\Gamma_k$ requires action
$a^j$. The corresponding estimated ending time is updated by:
\begin{equation}
\label{ch04:eq:estimated_time}
t_{\mathrm{e}}(\omega_k^\ell) \triangleq
\max_{\omega_j\in\texttt{Pre}(\omega_k^\ell)}
\bigl\{ t_{\mathrm{e}}(\omega_j) \bigr\}
+ T_{\mathrm{nav}}(S_k^{\ell-1},S_k^\ell)
+ T_{\mathrm{exec}}(\omega_k^\ell),
\end{equation}
where $\texttt{Pre}(\omega_k^\ell)$ is the set of preceding tasks of
$\omega_k^\ell$ that have already been assigned, and
$T_{\mathrm{nav}}(S_k^{\ell-1},S_k^\ell)$ is the estimated navigation time
between the execution regions of two consecutive tasks.

\begin{algorithm}[t!]
  \caption{Simultaneous task decomposition and team assignment}
  \label{ch04:alg:ss-TaskAssign}
  \SetAlgoLined
  \KwIn{$\Phi=\{\varphi_1,\cdots,\varphi_M\}$, automata
        $\{\mathcal{B}_{\varphi_m}\}$, fleet $\mathcal{N}$.}
  \KwOut{Number of teams $K$, capability profiles $\{\Lambda_k\}$, and local
         plans $\{\Gamma_k\}$.}

  Initialize $\mathfrak{T}=(\mathcal{V},\rightarrow)$ with
  $\mathcal{V}=\{\nu_0\}$.\;

  \While{the search is not terminated}{
    \tcc{Selection}
    Select $\mathcal{V}_P=\{\nu_1^\star,\cdots,\nu_P^\star\}$ by the value
    function in~\eqref{ch04:eq:search_value}.\;
    Compute $\Lambda_k$ for all $k\in\mathcal{K}$ by
    \eqref{ch04:eq:capacity-team}.\;

    \tcc{Expansion}
    \ForEach{$\nu_h^\star \in \mathcal{V}_P$}{
      Obtain $\Omega^-_{\nu_h^\star}$ by~\eqref{ch04:eq:task-candidate}.\;
      \ForEach{$\omega \in \Omega^-_{\nu_h^\star}$}{
        \If{a new team is created}{
          Create team $K+1$ and set $\Gamma_{K+1}=\emptyset$.\;
          Update the team index set $\mathcal{K}$.\;
        }
        \ForEach{$k \in \mathcal{K}$}{
          Generate $\nu^+$ by~\eqref{ch04:eq:expand}.\;
          Update $\widehat{Q}_m^+$ for all $m\in\mathcal{M}$.\;
          Update $\Lambda_k$ by~\eqref{ch04:eq:update-constraint}.\;
          Update $t_{\mathrm{e}}(\omega_k^\ell)$ by
          \eqref{ch04:eq:estimated_time}.\;
          \lIf{\eqref{ch04:eq:capacity-bound} is violated}
             {mark $\nu^+$ infeasible}
          Add $(\nu_h^\star,\nu^+)$ to $\mathfrak{T}$.\;
        }
      }
    }

    \tcc{Bounding}
    \ForEach{$\nu \in \mathcal{V}$}{
      Compute $\zeta(\nu)$ by~\eqref{ch04:eq:profile}.\;
      Prune dominated nodes by~\eqref{ch04:eq:dominance}.\;
    }
    Update $\overline{\mathcal{V}}$ by~\eqref{ch04:eq:frontier}.\;
  }

  \tcc{Termination}
  Compute $\overline{\mathcal{V}}^\star$ by~\eqref{ch04:eq:complete-nodes}.\;
  Select $\nu_{K^\star}^\star$ by~\eqref{ch04:eq:termination}.\;
  \Return $K^\star$, $\{\Lambda_k\}$, and $\{\Gamma_k\}$.\;
\end{algorithm}

\emph{(III) Bounding.}
To reduce unnecessary exploration, each node is evaluated through a performance
profile that aggregates the current team schedules, team costs, and mission
progress:
\begin{equation}
\label{ch04:eq:profile}
\zeta(\nu) \triangleq
\Bigl[
\{T_k\}_{k\in\mathcal{K}},
\{\kappa_k\}_{k\in\mathcal{K}},
\Bigl\{
\min_{q_m\in\widehat{Q}_m}\psi_m(q_m,Q_F^m)
\Bigr\}_{m\in\mathcal{M}}
\Bigr],
\end{equation}
where $T_k$ is the ending time of $\Gamma_k$, $\kappa_k$ is the estimated cost
of $\Gamma_k$, and the last term measures the minimum distance from the current
mission-progress set $\widehat{Q}_m$ to the accepting set $Q_F^m$. The profile
$\zeta(\nu)$ is therefore a nonnegative vector of dimension $(2K+M)$.
Given two nodes $\nu_1$ and $\nu_2$, node $\nu_1$ is said to dominate node
$\nu_2$ if
\begin{equation}
\label{ch04:eq:dominance}
\zeta(\nu_1)\leq \zeta(\nu_2), \qquad
\zeta(\nu_1)\neq \zeta(\nu_2),
\end{equation}
where the inequality is interpreted element-wise. In other words, $\nu_1$ has
no larger team-end time or team cost, and no worse mission progress, with
strict improvement in at least one component. The set of frontier nodes is then
defined by:
\begin{equation}
\label{ch04:eq:frontier}
\overline{\mathcal{V}} \triangleq
\bigl\{
\nu\in\mathcal{V} \,\big|\,
\nexists\,\nu'\in\mathcal{V} \text{ such that } \nu' \text{ dominates } \nu
\bigr\},
\end{equation}
where only nodes in $\overline{\mathcal{V}}$ are retained as candidates for
further expansion. This pruning rule discards strictly inferior branches while
preserving all non-dominated search directions.

\emph{(IV) Termination.}
The stages above are repeated until the computational budget is exhausted or no
new frontier nodes are generated. Among the frontier nodes, the set of complete
assignments is given by:
\begin{equation}
\label{ch04:eq:complete-nodes}
\overline{\mathcal{V}}^\star \triangleq
\bigl\{
\nu\in\overline{\mathcal{V}} \,\big|\,
\widehat{Q}_m\cap Q_F^m\neq\emptyset,\ \forall m\in\mathcal{M}
\bigr\},
\end{equation}
which means that every mission has reached an accepting state. The final node is
selected by:
\begin{equation}
\label{ch04:eq:termination}
\nu_{K^\star}^\star \triangleq
\texttt*{arg\,min}_{\nu\in\overline{\mathcal{V}}^\star}\chi(\nu),
\end{equation}
where $\chi(\nu)$ is the value function in
\eqref{ch04:eq:search_value}. The resulting capability profiles
$\{\Lambda_k\}$ and local plans $\{\Gamma_k\}$ define the abstract team
assignment.

\begin{example}
\label{ch04:example:task-decomp}
As illustrated in Fig.~\ref{ch04:fig:Tree}, suppose the mission automata
contain the task symbols
$\omega_1,\omega_2,\omega_3,\omega_4,\omega_5$. The optimal leaf node
$\nu_{K^\star}^\star$ assigns these five tasks to three teams, so that
$K^\star=3$. One feasible outcome is that the first team executes
$\omega_1$ followed by $\omega_3$, the second executes $\omega_2$ followed by
$\omega_4$, and the third executes $\omega_5$.
\end{example}

The construction above avoids the explicit formation of a monolithic product
between all mission automata and all robot models. It also allows the number of
teams to emerge from the search rather than being fixed a priori. In contrast to
the poset-based methods in~\cite{liu2024time,luo2025hulk}, each node tracks the
reachable-state sets $\widehat{Q}_m$ of all mission automata directly, so that
mission decomposition and abstract team assignment are handled in one search
process. This reduces the amount of preprocessing required before the
subsequent team-instantiation stage.
Correctness and completeness of the construction in the static case are stated
next. The mission set is treated as fixed throughout this subsection.

\begin{theorem}
\label{ch04:thm:correctness}
Consider an instance of Problem~\ref{ch04:prob:simultaneous_assignment} with a
fixed finite mission set $\Phi$. Suppose that the search is allowed to terminate
on the finite tree and that feasible team plans exist. Then
Algorithm~\ref{ch04:alg:ss-TaskAssign} returns local plans
$\{\Gamma_k\}_{k\in\mathcal{K}}$ such that:
\emph{(I)} all temporal constraints encoded by the automata
$\{\mathcal{B}_{\varphi_m}\}_{m\in\mathcal{M}}$ are satisfied;
\emph{(II)} all fleet-capacity constraints in
\eqref{ch04:eq:capacity-team}--\eqref{ch04:eq:capacity-bound} are respected;
\emph{(III)} among the complete frontier nodes, the returned node minimizes the
search value in \eqref{ch04:eq:search_value}.
\end{theorem}

\begin{proof}
Temporal correctness follows from the update of the reachable-state sets
$\widehat{Q}_m$ along enabled automaton transitions. Every root-to-leaf path
ending at a node in $\overline{\mathcal{V}}^\star$ satisfies
\eqref{ch04:eq:complete-nodes}, so for each mission index $m\in\mathcal{M}$
the corresponding task sequence induces an accepting run of
$\mathcal{B}_{\varphi_m}$. Hence the local plans
$\{\Gamma_k\}_{k\in\mathcal{K}}$ satisfy all temporal constraints encoded by
the mission automata.
Fleet-capacity feasibility follows from the expansion and pruning rules.
Whenever a task is added to a team plan, the induced capability profile is
updated by \eqref{ch04:eq:update-constraint}, and every node violating the fleet
bound \eqref{ch04:eq:capacity-bound} is marked infeasible and removed from
further consideration. Therefore, every node that remains on the frontier
satisfies the capacity constraints.
The search space is finite because the mission set $\Phi$, the automata
$\mathcal{B}_{\varphi_m}$, their state sets $Q^m$, the alphabets $\Sigma^m$,
and the admissible team-capability profiles are all finite. The frontier update
\eqref{ch04:eq:frontier} removes only dominated nodes and therefore preserves at
least one representative of each non-dominated feasible solution class. Hence,
if feasible plans exist, at least one complete frontier node is generated in
finite time. By \eqref{ch04:eq:termination}, the returned node is the minimizer
of the value function \eqref{ch04:eq:search_value} over
$\overline{\mathcal{V}}^\star$.
\end{proof}

















%======================================================================================
%======================================================================================
\subsection{Capacity-Based and Redundancy-Aware Team Formation}
\label{ch04:subsubsec:team_formation}

Given the abstract assignment $\nu_{K^\star}^\star$ returned by
Subsubsection~\ref{ch04:subsubsec:simultaneous_assignment}, the remaining task
is to instantiate each abstract team by concrete robots. The capability
profiles computed there specify only aggregate action requirements for each
team, namely the minimum numbers of robots needed for different capabilities.
They do not specify which robots should be assigned. This subsection therefore
determines the concrete team formation while enforcing disjoint membership and
allowing controlled redundancy.

\begin{problem}
\label{ch04:prob:team_formation}
Given the abstract assignment $\nu_{K^\star}^\star$ and the fleet
$\mathcal{N}$, determine the team formation as
$\overline{\mathcal{N}} \triangleq
\{\mathcal{N}_1,\cdots,\mathcal{N}_{K^\star}\}$,
where $\mathcal{N}_k$ is the set of robots assigned to the abstract team
$\mathcal{T}_k$. Each team must satisfy its required capacities, and the team
memberships must be disjoint, that is,
$\mathcal{N}_{k_1}\cap\mathcal{N}_{k_2}=\emptyset$ for all $k_1\neq k_2$.
The objective is to minimize the overall response time of the assigned
missions.
\end{problem}

The team-formation problem is cast as a redundancy-aware mixed-integer linear
program. Two components are central. The first is the capacity requirement
induced by the plans $\{\Gamma_k\}_{k\in\mathcal{K}}$. The second is a
min-max objective that minimizes the worst team response time after robots are assigned.
Let $b_{ik}\in\{0,1\}$ be the binary decision variable indicating whether robot
$i\in\mathcal{N}$ is assigned to team $\mathcal{T}_k$. The capacity profile
$\Lambda_k$ from \eqref{ch04:eq:capacity-team} specifies the minimum number
$\beta_k^j$ of robots capable of action $a^j$ required by team $k$. To allow
certain redundancy, introduce a margin $\alpha_j\geq 1$ for each action
$a^j\in\mathcal{A}$. Then the team-capacity bounds are written as:
\begin{equation}
\label{ch04:eq:bound_form}
\beta_k^j
\leq
\sum_{i\in\mathcal{N}} b_{ik}\,\mathds{1}(a^j\in\mathcal{A}_i)
\leq
\alpha_j\beta_k^j,
\qquad \forall a^j\in\mathcal{A},\ \forall k\in\mathcal{K},
\end{equation}
where the left inequality enforces the minimum required capacity and the right
inequality limits the amount of redundancy admitted for that action. In
addition, concrete team memberships must remain disjoint, which is imposed by:
\begin{equation}
\label{ch04:eq:disjoint_membership}
\sum_{k\in\mathcal{K}} b_{ik}\leq 1,
\qquad \forall i\in\mathcal{N},
\end{equation}
where each robot can belong to at most one team.
The margin $\alpha_j$ accounts for uncertainty in the actual workload of the
assigned tasks and for possible robot unavailability during execution. In
practice, $\alpha_j$ is chosen as a modest safety factor above one. Larger
values improve robustness but reserve more robots, whereas smaller values
increase utilization but provide less tolerance to uncertainty and failure.
The second component of the formulation is the response-time objective. Let
$S_k^1$ be the region of the first task in the local plan $\Gamma_k$. For each
robot $i$ and team $k$, define the expected arrival time by
$t_{ik}\triangleq
\widehat{t}_i+T_{\mathrm{nav}}(\widehat{x}_i,S_k^1)$,
where $\widehat{t}_i$ is the time at which robot $i$ becomes available and
$\widehat{x}_i$ is its position at that time. Once the team members are fixed,
the execution cost of team $\mathcal{N}_k$ is given by:
\begin{equation}
\label{ch04:eq:team_cost}
J_k \triangleq
\max_{i\in\mathcal{N}_k}\{t_{ik}\}
+\sum_{\omega\in\Gamma_k} T_{\mathrm{exec}}(\omega),
\end{equation}
where the first term is the synchronization delay of the slowest robot
in team $k$, and the second term is the total execution time of the
tasks in $\Gamma_k$. The overall objective is given by:
\begin{equation}
\label{ch04:eq:formation_objective}
\min \max_{k\in\mathcal{K}} J_k,
\end{equation}
which minimizes the worst response time across all teams.

The formulation defined by
\eqref{ch04:eq:bound_form}--\eqref{ch04:eq:formation_objective} is a mixed-
integer linear program once the standard linearization of the max operators is
introduced. Its binary variables are the membership indicators $\{b_{ik}\}$,
whose number is $\mathcal{O}(|\mathcal{N}|K^\star)$. The number of capacity
constraints is $\mathcal{O}(|\mathcal{A}|K^\star)$. Although the problem
remains NP-hard in general, the number of teams $K^\star$ is typically small
in the hierarchical construction here, so the solve time remains manageable.
This formulation is also substantially smaller than a direct robot-to-subtask
assignment model without hierarchical decomposition
\cite{sahin2019multirobot,kurtz2021more}.
Once the binary solution $\{b_{ik}\}$ is obtained, the concrete team formation
is recovered by
$\mathcal{N}_k \triangleq
\{i\in\mathcal{N}\mid b_{ik}=1\},
\forall k\in\mathcal{K}$.
The plan of each robot $i\in\mathcal{N}_k$ is then induced by the local plan
$\Gamma_k$ of its team, namely
\begin{equation}
\label{ch04:eq:robot-plan}
\xi_i \triangleq
(S_k^1,\omega_k^1)(S_k^2,\omega_k^2)\cdots(S_k^{L_k},\omega_k^{L_k}),
\qquad \forall i\in\mathcal{N}_k,
\end{equation}
where $\omega_k^\ell$ is the $\ell$th task in $\Gamma_k$, $S_k^\ell$ is its
associated execution region, and $L_k$ is the number of tasks assigned to team
$k$. Thus, every robot receives a concrete task sequence that is consistent
with the team assignment and the mission-level decomposition.

\begin{example}
\label{ch04:example:formation}
Figure~\ref{ch04:fig:formation} illustrates the team formation induced by the
optimal abstract assignment $\nu_{K^\star}^\star$. In that example, the team
$\mathcal{N}_1=\{2,4,6,7,9\}$ executes the sequence
$\omega_2\omega_3\omega_6\omega_8$, the team
$\mathcal{N}_2=\{3,8,13,14,15\}$ executes $\omega_1\omega_5$, and the team
$\mathcal{N}_{K^\star}=\{5,10,11,17,18,19\}$ executes
$\omega_9\omega_{11}\omega_{10}$.
\end{example}




















%==============================
%==============================
\subsection{Local Task and Trajectory Coordination within Teams}
\label{ch04:subsec:local_coordination}

Given the concrete team formation
$\overline{\mathcal{N}}=\{\mathcal{N}_1,\cdots,\mathcal{N}_{K^\star}\}$ from
Subsection~\ref{ch04:subsubsec:team_formation}, the local task sequence of each
robot is induced by \eqref{ch04:eq:robot-plan}. For each team
$k\in\{1,\cdots,K^\star\}$, let
$\Gamma_k \triangleq (\omega_k^1,\omega_k^2,\cdots,\omega_k^{L_k})$ be the
assigned sequence of collaborative tasks. Each task $\omega_k^\ell$ is
associated with an execution region $S_k^\ell$ and a set of required subtasks
$\Xi_k^\ell \triangleq \{(n_j,a_j,s_j)\}_{j=1}^{J_k^\ell}$,
where each tuple $(n_j,a_j,s_j)$ specifies that at least $n_j$ robots are
required to perform action $a_j$ at location $s_j$.
The symbolic team assignment alone is not sufficient for execution. Each robot
$i\in\mathcal{N}_k$ must still be given a timed local execution plan together
with a feasible trajectory. To distinguish this execution-level object from the
fleet-level plan in Subsection~\ref{ch04:subsec:uncertain_mission}, denote the
local execution plan of robot $i$ during task $\omega_k^\ell$ by:
\[
\widetilde{\tau}_i^\ell \triangleq
(t_i^1,p_i^1,a_i^1)(t_i^2,p_i^2,a_i^2)\cdots,
\]
where $t_i^r$ is the execution time, $p_i^r$ is the target position, and
$a_i^r$ is the action executed at that stage. The collective local plan for
task $\omega_k^\ell$ is therefore
$\widetilde{\tau}_k^\ell \triangleq
\{\widetilde{\tau}_i^\ell \mid i\in\mathcal{N}_k\}$, and the corresponding
robot trajectories are denoted by $\{x_i^\ell\}_{i\in\mathcal{N}_k}$.
The local coordination problem is coupled with trajectory generation. The time
required to execute a subtask depends not only on which robot performs it, but
also on how the robot reaches the corresponding location. Hence, local task
allocation and trajectory design must be optimized jointly. This subsection
addresses that problem under three classes of local tasks: static and known
subtasks, static and unknown subtasks, and dynamic and known subtasks.

\begin{problem}
\label{ch04:prob:local_coordination}
Given a task $\omega_k^\ell$ with region $S_k^\ell$, required subtask set
$\Xi_k^\ell$, and assigned team $\mathcal{N}_k$, determine the local execution
plans $\widetilde{\tau}_k^\ell$ and the trajectories
$\{x_i^\ell\}_{i\in\mathcal{N}_k}$ such that the task is executed feasibly and
the local makespan is minimized.
\end{problem}

The distinction among the three cases comes from the subtask set
$\Xi_k^\ell$. If the number of subtasks and their locations are both known and
fixed, the problem reduces to coordinated assignment and trajectory planning
over a static set. If the subtasks are fixed but initially unknown, the problem
must combine exploration with online insertion of newly discovered subtasks. If
the subtasks are known but dynamic, then continual reassignment and trajectory
updates are required during execution. These three cases motivate the
specialized coordination strategies developed below.

\subsubsection{Static and Known Subtasks}
\label{ch04:sec:static-known}

In the static and known case, both the number of subtasks $J_k^\ell$ and their
locations $\{s_j\}_{j=1}^{J_k^\ell}$ are predetermined. This situation arises
in tasks such as delivery and inspection, where all points of interest are
known before execution begins. The remaining difficulty is then to coordinate
the robots so that the subtasks are completed with minimum makespan under the
given motion model.
For holonomic robots, the problem reduces to a variant of the multi-vehicle
routing problem without return. Each robot travels along straight-line segments
between assigned subtask locations, and the coordination problem becomes the
selection of visit sequences that minimize the completion time of the slowest
robot. The resulting routing solution directly determines the local execution
plans $\widetilde{\tau}_k^\ell$, while the trajectories
$\{x_i^\ell\}_{i\in\mathcal{N}_k}$ are given by straight-line connections.
For non-holonomic robots, sequencing and motion feasibility cannot be separated.
Each subtask location must be augmented by a feasible orientation, and
transitions between subtasks must respect the robot dynamics. This leads to a
joint optimization over execution sequences and motion primitives. A convenient
formulation is
\begin{equation}
\label{ch04:eq:augmented-cost}
J_k^\ell \triangleq
\min_{\{\widetilde{\tau}_i^\ell,x_i^\ell\}}
\max_{i\in\mathcal{N}_k}
\Biggl\{
\sum_{(u,v)\in \widetilde{\tau}_i^\ell}
\min_{\kappa\in\mathbb{K}(x_u,x_v)}
\texttt{Cost}(\kappa)
\Biggr\},
\end{equation}
where $\mathbb{K}(x_u,x_v)$ is the set of feasible motion primitives
connecting the states $x_u$ and $x_v$. The cost in
\eqref{ch04:eq:augmented-cost} therefore combines subtask sequencing and
dynamically feasible motion. Practical implementations can employ motion
primitives such as Dubins curves~\cite{ny2011dubins,vavna2015dubins}, together
with hybrid-A$^\star$ type search~\cite{erke2020improved}, to obtain the plans
$\widetilde{\tau}_k^\ell$ and trajectories
$\{x_i^\ell\}_{i\in\mathcal{N}_k}$.

\begin{example}
\label{ch04:example:StaticKnown}
Figure~\ref{ch04:fig:StaticKnown} shows one instance of the static and known
case. A total of $14$ delivery subtasks are assigned to $5$ robots with
different dynamic constraints. The resulting assignments and trajectories differ
substantially between the holonomic and non-holonomic settings.
\end{example}

\subsubsection{Static and Unknown Subtasks}
\label{ch04:sec:staticunknown}

In the static and unknown case, the subtasks in $\Xi_k^\ell$ are initially
unknown, but once discovered they remain fixed. This setting is typical of
search-and-rescue tasks, where the numbers and locations of targets are
revealed only during execution. Since the task set is not fully available at
the outset, the static assignment strategy above cannot be applied directly. A
suitable approach combines exploration of the task region with online insertion
of newly discovered subtasks.
The first component is collaborative exploration of the region $S_k^\ell$. In
open spaces, the objective is efficient coverage of the region while accounting
for heterogeneous robot velocities, initial positions, and sensing footprints.
In cluttered or structured environments, obstacle avoidance must also be taken
into account, and frontier-based exploration is a standard mechanism
\cite{holz2010evaluating}. In either case, the exploration component generates
trajectories $\{x_i^\ell\}_{i\in\mathcal{N}_k}$ that reveal the unknown
subtasks while avoiding redundant search effort.
The second component is the online insertion of newly discovered subtasks into
the current local plans. The guiding principle is to minimize disruption to the
existing execution schedule. If a new subtask is discovered and is feasible for
some robot in $\mathcal{N}_k$, the corresponding local plans are updated by
solving
\begin{equation}
\label{ch04:eq:min-disruption}
\min_{\{\widetilde{\tau}_i^{\ell,+},x_i^{\ell,+}\}}
\sum_{i\in\mathcal{N}_k}
\Bigl(
\texttt{Cost}(\widetilde{\tau}_i^{\ell,+})
-
\texttt{Cost}(\widetilde{\tau}_i^\ell)
\Bigr),
\end{equation}
where $\widetilde{\tau}_i^{\ell,+}$ is the updated local plan obtained after
inserting the newly discovered subtasks. The objective in
\eqref{ch04:eq:min-disruption} measures the increase in execution cost caused
by the insertion. Online assignment then selects the robot that yields the
smallest incremental cost, and the associated trajectories are updated
accordingly. For non-holonomic robots, the same motion-primitive construction as
in the static and known case is used to ensure trajectory feasibility.

\begin{example}
\label{ch04:example:StaticUnknown}
Figure~\ref{ch04:fig:StaticUnknown} shows one instance of the static and
unknown case. Five robots are assigned to a task containing $8$ initially
unknown targets. The initial trajectories are chosen for coverage, and the
plans are then updated as targets are discovered. The resulting execution
simultaneously explores the task region and completes all revealed subtasks.
\end{example}

\subsubsection{Dynamic and Known Subtasks}
\label{ch04:sec:dynamicknown}

In the third case, the subtask set $\Xi_k^\ell$ and the initial locations of
its elements are known, but the subtasks move during execution. This setting
arises in tasks such as cooperative pursuit and capture, where the robots must
form coalitions to intercept moving targets. Since the target positions evolve
over time, static execution plans quickly become suboptimal or infeasible.
A distributed dynamic coalition formation mechanism is adopted here. At time
$t$, each subtask $j\in\Xi_k^\ell$ is associated with a coalition
$\mathcal{R}_j(t)\subseteq\mathcal{N}_k$, and the collection of all coalitions
is denoted by
$\overline{\mathcal{R}}_t \triangleq
\{\mathcal{R}_j(t)\mid j\in\Xi_k^\ell\}$.
The coalitions are disjoint and collectively exhaustive, namely
$\mathcal{R}_{j_1}(t)\cap\mathcal{R}_{j_2}(t)=\emptyset$ for
$j_1\neq j_2$, and
$\bigcup_{j\in\Xi_k^\ell}\mathcal{R}_j(t)=\mathcal{N}_k$. The current local
execution plan of robot $i$ is determined by its present coalition assignment,
and its trajectory follows the motion required by that coalition.
To evaluate coalition updates, each robot maintains local estimates of the
coalition costs $\chi(\mathcal{R}_j)$, which depend on robot-target distances
and relative velocities. The instantaneous coordination cost is given by:
\begin{equation}
\label{ch04:eq:coalition-cost}
\chi(\overline{\mathcal{R}}_t)\triangleq
\max_{j\in\Xi_k^\ell}\chi(\mathcal{R}_j)
+
\frac{1}{J_k^\ell}\sum_{j\in\Xi_k^\ell}\chi(\mathcal{R}_j),
\end{equation}
where the first term captures the current makespan across subtasks, and the
second promotes balanced workload among coalitions.

The local switch rule is based on whether moving one robot from its current
coalition to a neighboring coalition reduces the worst coalition cost. If robot
$i$ is currently assigned to subtask $j_i$, and $j'$ is a candidate neighboring
subtask, then the switch is accepted only if
\begin{equation}
\label{ch04:eq:switch-condition}
\max\Bigl\{
\chi(\mathcal{R}_{j'}\cup\{i\}),
\chi(\mathcal{R}_{j_i}\setminus\{i\})
\Bigr\}
<
\max\Bigl\{
\chi(\mathcal{R}_{j_i}),
\chi(\mathcal{R}_{j'})
\Bigr\},
\end{equation}
which is evaluated locally using the current coalition estimates. If
\eqref{ch04:eq:switch-condition} holds, a switch intent is announced to the
relevant neighboring robots, conflicts are resolved by a priority rule, and the
coalition update is synchronized. The local execution plans and trajectories are
then updated accordingly.
The distributed update requires only local information exchange and preserves
disjoint coalitions throughout execution. Its effectiveness depends on the
connectivity of the communication graph induced by the team. If the graph is
disconnected, coalition information cannot propagate reliably and coordinated
updates may fail. Delays or packet losses mainly slow information flow and may
cause temporary inconsistencies, but sufficiently frequent connectivity over
time windows is enough to recover a consistent coalition configuration.


\begin{example}
\label{ch04:example:DynamicKnown}
Figure~\ref{ch04:fig:DynamicKnown} shows one instance of the dynamic and known
case. Five robots are assigned to five moving targets through the distributed
coalition updates. Initially, each
robot is assigned to one target. Later, robots $2$ and $3$ exchange targets,
and once target $3$ is completed, robots $2$ and $4$ form a coalition for
target $2$, while robot $3$ proceeds to target $4$.
\end{example}

Note that the case of dynamic and unknown subtasks is not pursued here. Without knowing
the total number and locations of subtasks, it becomes difficult to determine
whether the current task has been completed, since new moving subtasks may
enter the explored region from previously unobserved areas. If such a setting
must be addressed, the appropriate strategy would combine the exploration-based
mechanism of the static and unknown case with the dynamic coalition updates of
the dynamic and known case.

\begin{theorem}
\label{ch04:thm:local-correctness}
For each task $\omega_k^\ell$ and assigned team $\mathcal{N}_k$, the
coordination strategies in Section~\ref{ch04:subsec:local_coordination}
produce feasible local execution plans and trajectories. In the static and
known case, the local makespan objective is optimized directly. In the static
and unknown case, feasibility is preserved while the increase in execution cost
caused by online insertions is minimized. In the dynamic and known case, the
distributed coalition updates decrease the local coalition cost until a locally
optimal coalition configuration is reached.
\end{theorem}

\begin{proof}
In the static and known case, the subtask set and the subtask locations are
fully specified. For holonomic robots, the coordination problem reduces to
routing with makespan minimization. For non-holonomic robots, the feasible
motion-primitive construction in \eqref{ch04:eq:augmented-cost} ensures that
the resulting trajectories satisfy the robot dynamics while the local objective
is minimized.
In the static and unknown case, exploration guarantees eventual discovery of
all fixed subtasks under complete coverage of the task region. Each discovered
subtask is then inserted by solving \eqref{ch04:eq:min-disruption}, which
preserves feasibility of the current plan while minimizing the incremental cost
of the update. Hence the resulting local execution remains feasible throughout
the exploration process.
In the dynamic and known case, the coalition update rule in
\eqref{ch04:eq:switch-condition} admits a switch only when the worst affected
coalition cost decreases. Therefore, the aggregate coalition cost in
\eqref{ch04:eq:coalition-cost} decreases monotonically under accepted updates.
Since the number of feasible coalition structures over the finite team
$\mathcal{N}_k$ is finite, the update process terminates at a locally optimal
coalition configuration. The associated local plans and trajectories are
therefore feasible and locally consistent with the moving subtasks.
\end{proof}

%==============================
\subsection{Complexity and Scalability Analysis}\label{ch04:subsubsec:complexity}

The computational complexity of the proposed method is analyzed as follows.
{In each iteration of Alg.~\ref{ch04:alg:ss-TaskAssign}, selecting a batch of~$P$
nodes and expanding up to~$B \triangleq (K{+}1)\,|\Omega^-_{\nu_h^\star}|$
children incurs $\mathcal{O}(PBK)$ for the automaton-state
updates~$\widehat{Q}^+_m$, together with the capacity and timing
updates in~\eqref{ch04:eq:update-constraint} and~\eqref{ch04:eq:estimated_time};
maintaining the non-dominated frontier based
on~\eqref{ch04:eq:profile}-\eqref{ch04:eq:frontier} requires
$\mathcal{O}(|\mathcal{V}|(2K{+}M))$ to reach
the completeness condition in \eqref{ch04:eq:complete-nodes}. The capacity-based
team-formation MILP introduces $\mathcal{O}(N K^\star)$ binary variables $b_{ik}$,
which is NP-hard in general but limited by the small number of teams.
For the case of static and known tasks, the
routing problem with subtour elimination is NP-hard,
and its number of variables scales with $|\mathcal{N}_k|(J_k^\ell)^2$.
For static and unknown tasks,
polygonal coverage generation and assignment are near-linear in the region
representation, while rolling insertion of newly discovered subtasks is
$\mathcal{O}(|\mathcal{N}_k|)$ per discovery. Lastly, for dynamic and known tasks, the
distributed coalition updates are linear in
team size and the number of active targets per iteration.
Although the
local MILPs and coalition updates are NP-hard in the worst case, the hierarchical
decomposition confines them to small teams, enabling event-based
replanning with standard solvers.
Detailed empirical runtime evaluations of each component and scaling benchmarks are provided in Sec.~\ref{sec:ch04_case_study}.}
