\section{Partial-Order-Based Task Decomposition and Assignment}
\label{ch04:sec:po_decomp_assignment}

This section develops the first component of the top-down synthesis framework.
It introduces the team model and global task specification, prunes the task
automaton, extracts a relaxed partial-order representation, and assigns the
resulting subtasks to feasible robots or coalitions. The outcome is an
assignment-oriented representation that preserves task correctness while
allowing concurrency whenever the temporal structure permits it.



%==================================================
%==================================================
%==================================================
%==================================================
\subsection{Global temporal task over collective behaviors}
\label{ch04:subsec:problem_statement}

This subsection introduces the team-level model, the global task specification,
and the optimization objective. Unlike the bottom-up setting of
Chapter~\ref{chap:bottom-up}, the task is specified directly at the team level.
The model therefore retains only the motion capabilities, action capabilities,
and collaborative requirements needed for global-task synthesis.

\subsubsection{Team model}
\label{ch04:subsubsec:team_model}

Consider a team of $N$ robots operating in a workspace
$\mathcal{W} \subset \mathbb{R}^{3}$. Let
$\mathbb{W} \triangleq \{W_{1}, W_{2}, \cdots, W_{M}\}$ be the finite set of
regions of interest. For each robot $n \in \mathcal{N} \triangleq
\{1, \cdots, N\}$, the admissible motion is represented by the weighted graph
$\mathcal{G}_{n} \triangleq (\mathbb{W}, \rightarrow_{n}, d_{n})$, where
$\rightarrow_{n} \subseteq \mathbb{W} \times \mathbb{W}$ is the transition
relation and $d_{n} : \rightarrow_{n} \to \mathbb{R}_{>0}$ assigns the travel
time of each feasible transition.
Each robot $n$ is equipped with an action set
$\mathcal{A}_{n} \triangleq \mathcal{A}^{\mathrm{l}}_{n} \cup
\mathcal{A}^{\mathrm{c}}_{n}$, where
$\mathcal{A}^{\mathrm{l}}_{n}$ contains actions that can be executed
independently and $\mathcal{A}^{\mathrm{c}}_{n}$ contains actions that require
cooperation with other robots. Let
$\mathcal{A}^{\mathrm{l}} \triangleq \bigcup_{n \in \mathcal{N}}
\mathcal{A}^{\mathrm{l}}_{n}$ and
$\mathcal{A}^{\mathrm{c}} \triangleq \bigcup_{n \in \mathcal{N}}
\mathcal{A}^{\mathrm{c}}_{n}$. Let
$\mathcal{C} \triangleq \{C_{1}, \cdots, C_{K}\}$ be the set of collaborative
behaviors available to the team. Each behavior
$C_{k} \triangleq \{a_{k,1}, a_{k,2}, \cdots, a_{k,\ell_{k}}\}$ is a finite set
of collaborative actions, where $\ell_{k} > 0$ and
$a_{k,j} \in \mathcal{A}^{\mathrm{c}}$ for all
$j \in \{1, \cdots, \ell_{k}\}$. The duration of behavior $C_{k}$ is denoted by
$d_{c}(C_{k}) \in \mathbb{R}_{>0}$.

A collaborative behavior is characterized by required action capabilities rather
than fixed robot identities, so any feasible robot or coalition can be selected
later. Representative examples include formation, herding, consensus, and
coverage behaviors
\cite{chen2005formation,pan2007multi,li2009consensus,mesbahi2010graph}.
Compared with Chapter~\ref{chap:bottom-up}, this abstraction omits the detailed
execution logic of local and assisting actions and describes collaboration only
through capability requirements and execution durations, which is sufficient for
global-task decomposition and assignment.

\subsubsection{Global task specification}
\label{ch04:subsubsec:global_task_spec}

The global mission is given by atomic propositions of three types. For each
region $W_{m} \in \mathbb{W}$, the proposition $p_{m}$ is true whenever at
least one robot is located in $W_{m}$. Let
$\mathbf{p} \triangleq \{p_{m} \mid W_{m} \in \mathbb{W}\}$. For each local
action $a \in \mathcal{A}^{\mathrm{l}}$ and each region
$W_{m} \in \mathbb{W}$, the proposition $a^{m}$ is true whenever action $a$ is
executed at $W_{m}$. Let
$\mathbf{a} \triangleq \{a^{m} \mid a \in \mathcal{A}^{\mathrm{l}},
\, W_{m} \in \mathbb{W}\}$. For each collaborative behavior
$C_{k} \in \mathcal{C}$ and each region $W_{m} \in \mathbb{W}$, the proposition
$c_{k}^{m}$ is true whenever behavior $C_{k}$ is completed at $W_{m}$. Let
$\mathbf{c} \triangleq \{c_{k}^{m} \mid C_{k} \in \mathcal{C},
\, W_{m} \in \mathbb{W}\}$.
The global task is given by:
\begin{equation}
\label{ch04:eq:global_task}
\varphi \in \mathrm{sc\mbox{-}LTL}(\mathbf{p} \cup \mathbf{a} \cup \mathbf{c}),
\end{equation}
where the syntax and semantics of sc-LTL follow
Chapter~\ref{chap:preliminaries}.
To state the synthesis problem precisely, the output object is introduced next.

\begin{definition}
\label{ch04:def:team_plan}
For each robot $n \in \mathcal{N}$, a finite motion-and-action plan is a pair
$(\rho_{n}, \mu_{n})$, where $\rho_{n} \triangleq
W_{n,0} W_{n,1} \cdots W_{n,L_{n}}$ is a finite path of $\mathcal{G}_{n}$, and
$\mu_{n}$ is a finite set of action-execution records of the form
$(a, W_{m}, t)$, where $a \in \mathcal{A}_{n}$,
$W_{m} \in \mathbb{W}$, and $t \in \mathbb{R}_{\ge 0}$. The collection
$\Pi_{\mathrm{team}} \triangleq
\{(\rho_{n}, \mu_{n}) \mid n \in \mathcal{N}\}$ is called a team plan.
\end{definition}

An action-execution record $(a, W_{m}, t) \in \mu_{n}$ means that robot $n$
executes action $a$ at region $W_{m}$ and time $t$. Let $t_{0}$ denote the
initial time, and let $t_{f}$ denote the first time when the team execution
induced by $\Pi_{\mathrm{team}}$ satisfies $\varphi$. The completion time of
the global task is defined by
\begin{equation}
\label{ch04:eq:completion_time}
T_{\varphi} \triangleq t_{f} - t_{0},
\end{equation}
where $T_{\varphi}$ is finite because sc-LTL formulas admit satisfaction by a
finite prefix. The objective is to minimize the team completion time in
\eqref{ch04:eq:completion_time}. This criterion differs from minimizing the sum
of all robot execution times, which does not capture the benefit of concurrent
task execution.

\begin{problem}
\label{ch04:prob:global_task_assignment}
Given the team model above and the global task formula
in \eqref{ch04:eq:global_task}, synthesize a team plan
$\Pi_{\mathrm{team}}$ in the sense of
Definition~\ref{ch04:def:team_plan} such that the induced team execution
satisfies $\varphi$ and the completion time $T_{\varphi}$ in
\eqref{ch04:eq:completion_time} is minimized.
\end{problem}

Problem~\ref{ch04:prob:global_task_assignment} is computationally difficult: it
already contains NP-hard makespan minimization for multi-stage scheduling
\cite{hochba1997approximation,bovet1994introduction}, with additional mobility,
temporal-logic, and collaboration constraints. Approximate multi-robot
allocation and scheduling methods address related aspects of this difficulty
\cite{khamis2015multi,hoos2004stochastic}.












%==================================================
%==================================================
%==================================================
%==================================================

%==============================
\subsection{Task decomposition based on posets}
\label{ch04:subsec:task-decompose-posets}

This subsection converts the compact global formula into explicit subtask
occurrences and temporal constraints. The decomposition must be sufficiently
expressive for assignment but less restrictive than a fully sequential run, so
that later allocation can exploit admissible parallelism.

\subsubsection{B\"{u}chi automaton pruning}
\label{ch04:subsubsec:NBA-pruning}

The construction starts from the NBA $\mathcal{B}_{\varphi}$ associated with
$\varphi$, obtained by standard translation tools such as~\cite{gastin2001fast}
and having the structure of Definition~\ref{ch02:def:NBA}. In the multi-robot
setting, this automaton may contain infeasible alphabets, states outside all
accepting runs, and simultaneity constraints that admit equivalent sequential
realizations. Pruning these elements reduces the automaton used for later
decomposition and assignment.
The pruning procedure contains three parts. First, a transition
$q_{j} \in \delta(q_{i}, \sigma)$ is removed if the alphabet $\sigma$ is
infeasible for the team. This can be checked by verifying whether the
propositions in $\sigma$ can be generated either by one robot that can reach
the required region and execute the required local action, or by a feasible
group of robots that can jointly realize the required collaborative action.
Second, a state is removed if it is unreachable from the initial set, or if
it cannot reach an accepting state that is itself reachable from that state.
Such a state cannot belong to any accepting run. Third, transitions that
impose exact simultaneity and admit equivalent decompositions through
intermediate states are removed. This last case is formalized next.

\begin{definition}
\label{ch04:def:decomposable-transition}
A transition from $q_{i}$ to $q_{j}$ in $\mathcal{B}_{\varphi}$ is called
\emph{decomposable} if there exist a state $q_{k}$ and alphabets
$\sigma_{ik}, \sigma_{kj} \subseteq \Sigma$ such that
\begin{equation}
\label{ch04:eq:decomposable_transition}
q_{k} \in \delta(q_{i}, \sigma_{ik}), \quad
q_{j} \in \delta(q_{k}, \sigma_{kj}), \quad
q_{j} \in \delta(q_{i}, \sigma_{ik} \cup \sigma_{kj}),
\end{equation}
where the direct transition and the two-step path connect the same pair of
states.
\end{definition}

Definition~\ref{ch04:def:decomposable-transition} identifies a transition whose
alphabet can be split without changing the logical effect of the run. Such a
transition imposes unnecessary simultaneity on distributed execution, so the
direct edge can be replaced by the corresponding two-step path whenever
\eqref{ch04:eq:decomposable_transition} holds. The pruned automaton is denoted
by $\mathcal{B}_{\varphi}^{-}$; it retains the accepting behaviors needed for
decomposition while reducing the search space.

\begin{lemma}
\label{ch04:lemma:pruned_nba_equiv}
If a feasible word $w$ is accepted by $\mathcal{B}_{\varphi}$, then there
exists an equivalent word $w'$ that is accepted by
$\mathcal{B}_{\varphi}^{-}$.
\end{lemma}

\begin{proof}
Let $w$ be a feasible accepting word of $\mathcal{B}_{\varphi}$, and let
$\rho = q_{0} q_{1} \cdots q_{L}$ be a corresponding accepting run. Since
$\rho$ is accepting, every state on $\rho$ is reachable from the initial set
and can reach an accepting state. Therefore, none of these states is removed
during the invalid-state pruning step. Since $w$ is feasible for the team,
none of the transitions used by $\rho$ is removed by the infeasibility test.
If every transition in $\rho$ is retained after pruning, then $\rho$ is also
an accepting run of $\mathcal{B}_{\varphi}^{-}$, and one may take
$w' = w$. Assume now that $\rho$ contains a transition from $q_{i}$ to
$q_{j}$ that is removed because it is decomposable. By
Definition~\ref{ch04:def:decomposable-transition}, there exist a state
$q_{k}$ and alphabets $\sigma_{ik}, \sigma_{kj}$ such that
$q_{k} \in \delta(q_{i}, \sigma_{ik})$ and
$q_{j} \in \delta(q_{k}, \sigma_{kj})$, while the direct transition from
$q_{i}$ to $q_{j}$ is induced by the combined alphabet
$\sigma_{ik} \cup \sigma_{kj}$. Hence, the segment $q_{i} q_{j}$ in the
accepting run can be replaced by the segment $q_{i} q_{k} q_{j}$. This
replacement yields a new accepting run in which the direct step is realized
by two consecutive steps.
If one of the newly introduced transitions is itself removed as decomposable,
the same replacement can be applied again. Since the automaton is finite,
this process terminates after finitely many steps. The result is an
accepting run of $\mathcal{B}_{\varphi}^{-}$ associated with a word $w'$
that is equivalent to $w$ for the purpose of task satisfaction.
\end{proof}

\begin{example}
\label{ch04:example:prune_nba}
For the task formula in~\ref{ch04:example:subtask}, the automaton
$\mathcal{B}_{\varphi}$ generated by~\cite{gastin2001fast} contains
$707$ states and $16044$ transitions, together with more than
$1.29 \times 10^{7}$ accepting words. After the pruning procedure above, the
automaton $\mathcal{B}_{\varphi}^{-}$ still contains $707$ states, but only
$2423$ transitions and $174469$ accepting words. Therefore, the pruning step
reduces the number of transitions by $84.9\%$ and the number of accepting
words by $98.6\%$.
\end{example}

\subsubsection{Task decomposition}
\label{ch04:subsubsec:task-decompose}

An intermediate decomposition is needed because temporal formulas impose
logical ordering constraints among subtask occurrences. For instance,
$\Diamond(\alpha_{1}\land\Diamond(\alpha_{2}\land\Diamond\alpha_{3}))$
requires sequential satisfaction of $\alpha_{1}$, $\alpha_{2}$, and
$\alpha_{3}$, whereas
$\Diamond\alpha_{1}\land\Diamond\alpha_{2}\land\Diamond\alpha_{3}$ does not.
The construction therefore extracts indexed subtasks from an accepting run of
$\mathcal{B}_{\varphi}^{-}$ and relaxes the induced total order into an
R-poset that keeps only the temporal constraints needed for correctness.

\begin{definition}
\label{ch04:def:subtasks}
Consider an accepting run
$\rho \triangleq q_{0} q_{1} \cdots q_{L}$ of
$\mathcal{B}_{\varphi}^{-}$, where $q_{0} \in Q_{0}^{-}$ and
$q_{L} \in Q_{F}^{-}$. A \emph{decomposition} of $\varphi$ induced by $\rho$
is the set
\begin{equation}
\label{ch04:eq:subtask}
\Omega_{\varphi} \triangleq
\bigl\{\omega_{\ell}=(\ell,\sigma_{\ell},\sigma_{\ell}^{s})
\mid \ell=1,\cdots,L\bigr\},
\end{equation}
where $\ell$ is the index of subtask $\omega_{\ell}$. For each
$\ell \in \{1,\cdots,L\}$, the label $\sigma_{\ell}\subseteq\Sigma$
satisfies $q_{\ell}\in\delta(q_{\ell-1},\sigma_{\ell})$ and
$q_{\ell}\notin\delta(q_{\ell-1},\sigma_{\ell}\setminus\{z\})$ for every
$z\in\sigma_{\ell}$. The set $\sigma_{\ell}^{s}\subseteq\Sigma$ is the
self-loop requirement at $q_{\ell-1}$ and satisfies the following
conditions: if $\sigma\cap\sigma_{\ell}^{s}\neq\emptyset$, then
$q_{\ell-1}\notin\delta(q_{\ell-1},\sigma)$; if
$\sigma\cap\sigma_{\ell}^{s}=\emptyset$, then
$q_{\ell-1}\in\delta(q_{\ell-1},\sigma)$.
\end{definition}

Each subtask records its index, transition label, and self-loop label. The
index distinguishes repeated occurrences of the same proposition set;
$\sigma_{\ell}$ specifies what must hold for the transition, and
$\sigma_{\ell}^{s}$ specifies what must remain inactive before it. The induced
order of the accepting run is correct but often too restrictive, so two partial
relations are introduced to recover admissible concurrency.

\begin{example}
\label{ch04:example:subtask}
Consider the accepting run
$\rho=q_{1}q_{2}q_{4}q_{5}$. The corresponding decomposition is
$\Omega_{\varphi}=
\bigl\{
(1,\{\texttt{sweep}_{\texttt{p}_{21}}\},\{\texttt{p}_{24}\}),
(2,\{\texttt{mow}_{\texttt{p}_{21}}\},\{\texttt{p}_{24}\}),
(3,\{\texttt{scan}_{\texttt{p}_{21}}\},\{\texttt{p}_{24}\})
\bigr\}$.
For the first subtask, one has
$\sigma_{1}=\{\texttt{sweep}_{\texttt{p}_{21}}\}$ and
$\sigma_{1}^{s}=\{\texttt{p}_{24}\}$. The reduced label
$\sigma_{1}\setminus\{\texttt{sweep}_{\texttt{p}_{21}}\}=\emptyset$ does not
enable the transition from $q_{1}$ to $q_{2}$, whereas $\sigma_{1}$ does.
Moreover, any alphabet that intersects $\sigma_{1}^{s}$ disables the
self-loop at $q_{1}$, whereas any alphabet that does not intersect
$\sigma_{1}^{s}$ keeps that self-loop enabled.
\end{example}

\begin{definition}
\label{ch04:def:partial}
Given subtasks $\omega_{h},\omega_{\ell}\in\Omega_{\varphi}$, define two
relations. The \emph{precedence} relation
$\preceq_{\varphi}\subseteq
\Omega_{\varphi}\times\Omega_{\varphi}$ satisfies
$\omega_{h}\preceq_{\varphi}\omega_{\ell}$ if
$\omega_{h}$ must start no later than $\omega_{\ell}$.
The \emph{opposition} relation
$\neq_{\varphi}\subseteq 2^{\Omega_{\varphi}}$ satisfies
$\widehat{\Omega}\in\neq_{\varphi}$ for some
$\widehat{\Omega}\subseteq\Omega_{\varphi}$ if the subtasks in
$\widehat{\Omega}$ cannot all be executed simultaneously.
\end{definition}

The pair $(\preceq_{\varphi},\neq_{\varphi})$ is needed because action
propositions remain true over nonzero intervals. Unlike models with
instantaneous action satisfaction, where a single precedence relation often
suffices
\cite{kantaros2020stylus,guo2015multi,tumova2016multi,
luo2021abstraction,luo2022temporal,sahin2019multirobot,
jones2019scratchs}, the present model must distinguish start-order constraints,
finish-before-start constraints, mutual exclusion, and full parallelism. The
construction contains the decomposable-state abstraction of
\cite{schillinger2018simultaneous} as a special case, while allowing independent
subtasks within the same accepting-run segment.

\begin{definition}
\label{ch04:def:poset}
A \emph{relaxed partially ordered set}, abbreviated as an \emph{R-poset},
over the decomposition $\Omega_{\varphi}$ is the triple
\begin{equation}
\label{ch04:eq:poset}
P_{\varphi}\triangleq
(\Omega_{\varphi},\preceq_{\varphi},\neq_{\varphi}),
\end{equation}
where the two relations are given by
Definition~\ref{ch04:def:partial}.
\end{definition}

The relation $\preceq_{\varphi}$ is irreflexive, asymmetric, and transitive,
whereas $\neq_{\varphi}$ is not transitive. The number of retained relations
determines the restrictiveness of execution: empty relations allow arbitrary
arrangement, while the full relation recovers the original sequential order.
For later use, let
\[
\mathbb{F}\triangleq
\{(i,j)\mid i,j\in\{1,\cdots,L\},\,i<j\},
\]
denote the full precedence relation. Fewer retained relations generally allow
more concurrency. The language of an R-poset is defined next as the set of
timed subtask words that satisfy these partial constraints.

\begin{definition}
\label{ch04:def:language-poset}
Given an R-poset
$P_{\varphi}=
(\Omega_{\varphi},\preceq_{\varphi},\neq_{\varphi})$,
its \emph{language} is the set of all finite timed words generated by the
subtasks in $\Omega_{\varphi}$ while satisfying the relations in
$P_{\varphi}$. More precisely,
\begin{equation}
\label{ch04:eq:poset-language}
L(P_{\varphi})\triangleq
\Bigl\{
w_{\varphi}=
(t_{1},\omega_{1})(t_{2},\omega_{2})\cdots(t_{L},\omega_{L})
\Bigr\},
\end{equation}
where $\omega_{\ell}\in\Omega_{\varphi}$ and $t_{\ell}$ is the starting time
of $\omega_{\ell}$. The timed word $w_{\varphi}$ must satisfy
$t_{\ell}\le t_{\ell'}$ for every
$(\omega_{\ell},\omega_{\ell'})\in\preceq_{\varphi}$. In addition, for every
$\widehat{\Omega}\in\neq_{\varphi}$, there must exist
$\omega_{\ell},\omega_{\ell'}\in\widehat{\Omega}$ such that
$t_{\ell}+d_{\ell}<t_{\ell'}$, where $d_{\ell}$ and $d_{\ell'}$ are the
durations of $\omega_{\ell}$ and $\omega_{\ell'}$, respectively.
\end{definition}

With a slight abuse of notation, $w_{\varphi}$ may also denote the symbolic
sequence $\omega_{1}\omega_{2}\cdots\omega_{L}$. An R-poset is accepting if
$L(P_{\varphi})\subseteq L_{\varphi}$, so synthesis can search for accepting
R-posets rather than single accepting words. Since enumerating all accepting
runs of $\mathcal{B}_{\varphi}^{-}$ has worst-case complexity
$\mathcal{O}(|Q^{-}|!)$, an anytime construction is used. For a subtask set
$\Omega$, let
$2_{\ge 2}^{\Omega}\triangleq \{\widehat{\Omega}\subseteq\Omega \mid |\widehat{\Omega}|\ge 2\}$,
which is the family of all subsets of $\Omega$ with cardinality at least
two. Also, for any pair of automaton states $q,q'$, let
$\delta^{-1}(q,q')\triangleq
\{\sigma\in\Sigma \mid q'\in\delta(q,\sigma)\}$,
which denotes the set of alphabets that induce the transition from $q$ to
$q'$.

\begin{algorithm}[t!]
\caption{Anytime computation of accepting R-posets
$\texttt{compute\_poset}(\cdot)$}
\label{ch04:alg:compute-poset}
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\Input{Pruned automaton $\mathcal{B}_{\varphi}^{-}$ and time budget $t_{0}$.}
\Output{R-posets $\mathcal{P}_{\varphi}$ and languages
$\mathcal{L}_{\varphi}$.}

Initialize $\mathcal{P}_{\varphi}=\mathcal{L}_{\varphi}=\emptyset$.\;
Begin a modified DFS that explores accepting runs over all pairs
$(q_{0},q_{f})\in Q_{0}^{-}\times Q_{F}^{-}$.\;

\While{$\mathrm{time}<t_{0}$}{
    Compute $\Omega$ and a word $w$ from $\rho$ according to
    Definition~\ref{ch04:def:subtasks}.\label{ch04:alg-line:decompose}\;

    \If{$w \notin \mathcal{L}_{\varphi}$\label{ch04:alg1:inlp}}{
        Initialize
        $P=(\Omega,\preceq_{\varphi}=\mathbb{F},
        \neq_{\varphi}=2_{\ge 2}^{\Omega})$.\;
        \label{ch04:alg-line:initial}
        Set $L(P)=\{w\}$, $Que=[w]$, and $I_{1}=I_{2}=\emptyset$.\;
        \label{ch04:alg-line:que}

        \While{$|Que|>0$ and $\mathrm{time}<t_{0}$
        \label{ch04:alg1:begin_que}}{
            Pop one word $w$ from $Que$.\;

            \For{$i=1,2,\cdots,|w|-1$}{
                Let $\omega_{1}=w[i]$ and $\omega_{2}=w[i+1]$.\;
                Obtain $w'$ by switching $\omega_{1}$ and $\omega_{2}$ in
                $w$.\label{ch04:alg1:switch}\;

                \eIf{$w'$ is accepting\label{ch04:alg1:waccepting}}{
                    Add $(\omega_{1},\omega_{2})$ to $I_{1}$.\;
                    Add $w'$ to $Que$ and $L(P)$ if $w' \notin L(P)$.\;
                    \label{ch04:alg1:add-w}
                }{
                    Add $(\omega_{1},\omega_{2})$ to $I_{2}$.\;
                    \label{ch04:alg1:end_que}
                }
            }

            Remove $I_{1}\setminus I_{2}$ from $\preceq_{\varphi}$.\;
            \label{ch04:alg-line:remove}
        }

        \ForEach{$\widehat{\omega}\in\neq_{\varphi}$
        \label{ch04:alg-line:calculate_neq}}{
            Obtain $w'$ by replacing every
            $\omega_{i}\in\widehat{\omega}$ in $w$ by
            $\bigcup_{\omega_{i}\in\widehat{\omega}}\omega_{i}$.\;
            \label{ch04:alg1:word-neq}

            \If{$w'$ is accepting\label{ch04:alg1:waccepting-neq}}{
                Remove $\widehat{\omega}$ from $\neq_{\varphi}$.\;
                \label{ch04:alg-line:remove_neq}
            }
        }

        \For{$w\in L(P)$}{
            Recover the associated path $\rho'$.\;
            \For{$i=1,2,\cdots,|\rho'|$}{
                Update $\sigma^{s}_{\ell_{i}}$ by intersecting it with
                $\delta^{-1}(\rho'[i-1],\rho'[i-1])$.\;
            }
        }
        \label{ch04:alg-line:self-loop}

        Add $P$ to $\mathcal{P}_{\varphi}$ and add $L(P)$ to
        $\mathcal{L}_{\varphi}$.\;
    }

    Continue the modified DFS and update $\rho$ and the elapsed time.\;
}

\Return{$\mathcal{P}_{\varphi},\mathcal{L}_{\varphi}$}.\;
\end{algorithm}

Algorithm~\ref{ch04:alg:compute-poset} uses a modified depth-first search with
local visited sets~\cite{sedgewick2001algorithms}. Each accepting run is
converted into a subtask set and seed word, from which a fully ordered R-poset
is initialized. Adjacent switches that preserve acceptance remove unnecessary
precedence constraints, while simultaneous-execution tests remove unnecessary
opposition constraints. The self-loop labels are then updated from feasible
words, producing one accepting R-poset; the search continues to return more
R-posets as time permits.

\begin{example}
\label{ch04:example:r_poset}
Continuing Example~\ref{ch04:example:subtask}, one obtains an R-poset with
$\preceq_{\varphi}=\{(1,2),(1,3)\}$ and
$\neq_{\varphi}=\{\{2,3\}\}$.
\end{example}

\begin{lemma}
\label{ch04:lemma:accepting-poset}
Any R-poset in $\mathcal{P}_{\varphi}$ produced by
Algorithm~\ref{ch04:alg:compute-poset} is accepting.
\end{lemma}

\begin{proof}
By definition, it is sufficient to show that the language $L(P)$ associated
with each returned R-poset $P$ is contained in $L_{\varphi}$. First, every
word added to $L(P)$ is explicitly checked for acceptance in
Line~\ref{ch04:alg1:add-w}. Hence, all stored words are accepting.
Now consider any word $w$ that satisfies the partial relations of $P$. For
the precedence relation, the queue process in
Line~\ref{ch04:alg1:begin_que}--\ref{ch04:alg1:end_que} explores every word
that can be generated from the seed word by a sequence of adjacent switches
while remaining accepting. Therefore, any precedence constraint removed from
$\preceq_{\varphi}$ is unnecessary for acceptance, while every constraint
that remains is necessary. It follows that every word satisfying the final
precedence relation can be generated through accepting switch operations and
belongs to $L(P)$.
For the opposition relation, every candidate relaxation is checked by
replacing the corresponding subtasks with their union and verifying whether
the resulting word is accepting. Therefore, every removal from
$\neq_{\varphi}$ preserves acceptance. Hence, any word satisfying the final
relations of $P$ is accepting, which proves $L(P)\subseteq L_{\varphi}$.
\end{proof}

Because Algorithm~\ref{ch04:alg:compute-poset} is anytime, a finite-time output
may cover only part of the accepting language. With unbounded time, however, the
complete family of accepting R-posets is obtained, and the union of their
languages is as expressive as the pruned automaton.

\begin{lemma}
\label{ch04:lemma:complete-poset}
The output of Algorithm~\ref{ch04:alg:compute-poset} satisfies
$L(P_{i})\cap L(P_{j})=\emptyset$ for all $i\neq j$, and
$L(P_{i})\subseteq L_{\varphi}$ for every
$P_{i}\in\mathcal{P}_{\varphi}$. Moreover, if
$t_{0}\rightarrow\infty$, then the complete set of R-posets is returned and
\begin{equation}
\label{ch04:eq:complete_poset_language}
L_{\varphi}=
\bigcup_{P_{i}\in\mathcal{P}_{\varphi}}L(P_{i}),
\end{equation}
so the union of the R-poset languages equals the accepting language of the
task.
\end{lemma}

\begin{proof}
The inclusion $L(P_{i})\subseteq L_{\varphi}$ follows directly from
Lemma~\ref{ch04:lemma:accepting-poset}. It remains to prove disjointness and
completeness.
For disjointness, assume that there exist two distinct R-posets
$P_{1},P_{2}\in\mathcal{P}_{\varphi}$ and a word
$w\in L(P_{1})\cap L(P_{2})$. Then the indexed subtask sets induced by
$P_{1}$ and $P_{2}$ must coincide. For a fixed subtask set, the queue phase
of Algorithm~\ref{ch04:alg:compute-poset} computes the maximal relaxation of
the partial relations supported by accepting words. Hence two different
R-posets cannot share the same accepting word. This contradicts the
assumption, and therefore $L(P_{1})\cap L(P_{2})=\emptyset$.
For completeness, observe that the underlying DFS explores all accepting
runs of $\mathcal{B}_{\varphi}^{-}$ when the time budget is unbounded. Each
accepting run induces a decomposition and, through the subsequent reduction
steps, an accepting R-poset. Therefore, every accepting word of the pruned
automaton is covered by at least one returned R-poset. Since the pruning
step preserves the accepting behaviors relevant here, the union in
\eqref{ch04:eq:complete_poset_language} equals $L_{\varphi}$.
\end{proof}

\begin{definition}
\label{ch04:def:poset-graph}
Let
$P_{\varphi}=
(\Omega_{\varphi},\preceq_{\varphi},\neq_{\varphi})$.
Its \emph{R-poset graph} is the mixed graph
$\mathcal{G}_{P_{\varphi}}=(\Omega,E,R,\Omega_{0})$,
where $\Omega=\Omega_{\varphi}$ is the node set,
$E\subseteq\Omega\times\Omega$ is the set of directed edges,
$R\subseteq 2^{\Omega}$ is the set of undirected hyperedges, and
$\Omega_{0}\subseteq\Omega$ is the set of root nodes. A directed edge
$(\omega_{1},\omega_{2})\in E$ exists if
$(\omega_{1},\omega_{2})\in\preceq_{\varphi}$ and there is no
$\omega_{3}\in\Omega$ such that
$\omega_{1}\preceq_{\varphi}\omega_{3}\preceq_{\varphi}\omega_{2}$. A
hyperedge $\widehat{\Omega}\in R$ exists if
$\widehat{\Omega}\in\neq_{\varphi}$. A node belongs to $\Omega_{0}$ if it
has no incoming directed edge in $E$.
\end{definition}

As in a Hasse diagram~\cite{simovici2008mathematical},
$\mathcal{G}_{P_{\varphi}}$ compactly represents the temporal structure of the
decomposition: directed edges encode precedence, hyperedges encode mutual
exclusion, and disconnected components indicate independent roots.


%========================================
%========================================
%========================================
%========================================















\subsubsection{Task assignment}
\label{ch04:subsubsec:task-assignment}

The next step assigns the subtasks of each accepting R-poset to the robot team.
For a fixed $P_{\varphi}\in\mathcal{P}_{\varphi}$, the objective is to respect
its temporal relations while minimizing completion time; candidate R-posets are
then compared through the resulting assignments.

\begin{problem}
\label{ch04:prob:task_assignment}
Given an R-poset
$P_{\varphi}=(\Omega_{\varphi},\preceq_{\varphi},\neq_{\varphi})$ with
$P_{\varphi}\in\mathcal{P}_{\varphi}$, determine an assignment of all subtasks
in $\Omega_{\varphi}$ to the robot team $\mathcal{N}$ such that the relations
$\preceq_{\varphi}$ and $\neq_{\varphi}$ are respected and the makespan of the
resulting team execution is minimized.
\end{problem}

Problem~\ref{ch04:prob:task_assignment} contains the multi-vehicle routing
problem~\cite{khamis2015multi} and job-shop scheduling
\cite{morrison2016branch} as special cases, and is therefore NP-hard. A direct
mixed-integer formulation is possible but unsuitable for large real-time
instances because useful intermediate assignments may not be available before
convergence. An anytime branch-and-bound procedure is therefore adopted; it
remains complete while returning progressively better feasible solutions. Its
main components are node expansion, branching, and upper and lower bounds.

(I) \emph{Node expansion.}
Each node of the search tree represents one partial assignment of subtasks,
written as:
\begin{equation}
\label{ch04:eq:node}
\nu \triangleq (\tau_{1},\tau_{2},\cdots,\tau_{N}),
\end{equation}
where $\tau_{n}$ is the ordered sequence of subtasks currently assigned to
robot $n\in\mathcal{N}$. For example, if
$\nu=((\omega_{1},\omega_{2}),(),())$ in a team of three robots, then
$\omega_{1}$ and $\omega_{2}$ are assigned to robot~$1$, while robots~$2$
and~$3$ are still unassigned. Let
\[
\Omega_{\nu} \triangleq \bigcup_{n\in\mathcal{N}}\{\omega \mid \omega\in\tau_n\},
\qquad
\Omega_{\nu}^{-} \triangleq \Omega_{\varphi}\setminus\Omega_{\nu},
\]
where $\Omega_{\nu}$ is the set of subtasks already assigned at node $\nu$ and
$\Omega_{\nu}^{-}$ is the set of remaining subtasks.
To avoid infeasible expansions, only eligible next subtasks are considered.
More precisely, a subtask $\omega$ can be chosen at node $\nu$ only if all of
its predecessors in the R-poset graph
$\mathcal{G}_{P_{\varphi}}$ from
Definition~\ref{ch04:def:poset-graph} have already been assigned, that is,
\begin{equation}
\label{ch04:eq:next-task}
\operatorname{Pre}(\omega)\subseteq\Omega_{\nu},
\end{equation}
where $\operatorname{Pre}(\omega)$ denotes the parent set of $\omega$ in
$\mathcal{G}_{P_{\varphi}}$. Once such a subtask is selected, a child node
$\nu^{+}$ is created by assigning it to a feasible robot set
$\mathcal{N}_{\omega}\subseteq\mathcal{N}$. If $\omega$ is a local subtask,
then $|\mathcal{N}_{\omega}|=1$. If $\omega$ is associated with a
collaborative behavior $C_{k}$, then $\mathcal{N}_{\omega}$ is admissible only
if the combined capabilities of the robots in $\mathcal{N}_{\omega}$ cover the
required actions in $C_{k}$. In this way, infeasible nodes that violate the
precedence structure are excluded from the search tree.

(II) \emph{Branching.}
Once a node has been expanded, the branching rule determines the order in which
its child nodes are visited. Standard choices such as breadth-first search,
depth-first search, and $A^{\star}$ search are all possible. Here,
$A^{\star}$-type branching is used, since its evaluation order is naturally
compatible with the lower bounds introduced below. The child nodes are therefore
prioritized according to the lower-bound estimate of the makespan that can
still be achieved from them.

(III) \emph{Upper and lower bounds.}
The upper bound is used to update the best feasible assignment found so far,
while the lower bound is used to determine whether a branch can still improve
the current incumbent. For a node $\nu$, the upper bound is computed by a greedy
completion policy:
\begin{equation}
\label{ch04:eq:upper-bound}
\overline{J}_{\nu},\,\overline{T}_{\nu},\,\overline{T}_{\Omega}
\triangleq
\texttt{upper\_bound}(\nu,P_{\varphi}),
\end{equation}
where $\overline{J}_{\nu}$ is a complete assignment extending the current node,
$\overline{T}_{\nu}$ is its makespan, and $\overline{T}_{\Omega}$ is the vector
of subtask starting times. Starting from $\nu$, each remaining subtask
$\omega\in\Omega_{\nu}^{-}$ is tentatively assigned to each admissible robot
set, thus generating a family of child nodes. For each such node, the
concurrency index is estimated by:
\begin{equation}
\label{ch04:eq:node-makespan}
T_{\nu}\triangleq\max_{n\in\mathcal{N}}T_{\tau_n}, \qquad
T_{\nu}^{\mathrm{s}}\triangleq
\sum_{\omega\in\Omega_{\nu}} D_{\omega}N_{\omega}, \qquad
\eta_{\nu}\triangleq \frac{T_{\nu}^{\mathrm{s}}}{T_{\nu}},
\end{equation}
where $T_{\tau_n}$ is the execution time of the sequence $\tau_n$ assigned to
robot~$n$, $D_{\omega}$ is the duration of subtask $\omega$, and
$N_{\omega}$ is the number of robots participating in $\omega$. The quantity
$T_{\nu}$ is computed under the precedence relation $\preceq_{\varphi}$, the
opposition relation $\neq_{\varphi}$, the motion constraints induced by the
graphs $\mathcal{G}_{n}$, and the synchronization requirements of
collaborative subtasks. The greedy upper-bound routine expands the child node
with the largest value of $\eta_{\nu}$ until all subtasks are assigned.
Moreover, the lower bound is estimated from two relaxations of the original problem.
The first retains only the temporal relations and relaxes the robot-capacity
constraints. The second retains the robot-capacity constraints and relaxes the
temporal relations. Accordingly, the lower bound is written as:
\begin{equation}
\label{ch04:eq:lower-bound}
\underline{T}_{\nu}\triangleq \texttt{lower\_bound}(\nu,P_{\varphi})
= \max\bigl\{\underline{T}_{\nu}^{\mathrm{ord}},
\underline{T}_{\nu}^{\mathrm{cap}}\bigr\},
\end{equation}
where $\underline{T}_{\nu}^{\mathrm{ord}}$ is the bound obtained from the
ordering relaxation and $\underline{T}_{\nu}^{\mathrm{cap}}$ is the bound
obtained from the capacity relaxation. The estimate in
\eqref{ch04:eq:lower-bound} is used in
Algorithm~\ref{ch04:alg:BnB} to prune any node that cannot improve the current
incumbent. Details of the two relaxed optimization problems can be found
in~\cite{liu2022arxiv}. Both the upper-bound and lower-bound routines are
designed to avoid integer optimization, which helps keep their runtime
predictable during the search.

Algorithm~\ref{ch04:alg:BnB} summarizes the resulting anytime search. The root
is the empty assignment, the queue stores unvisited nodes with lower-bound keys,
and each selected node is greedily completed to update the incumbent. Children
whose lower bounds exceed the current best makespan are discarded, and the
search stops when the queue is empty or the time budget is exhausted.

\begin{algorithm}[t]
\caption{Anytime branch-and-bound search for task assignment}
\label{ch04:alg:BnB}
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\Input{Robots $\mathcal{N}$, R-poset $P_{\varphi}$, and time budget $t_{0}$.}
\Output{Best assignment $J^{\star}$, makespan $T^{\star}$, and subtask starting
times $T_{\Omega}^{\star}$.}

Initialize the root node $\nu_{0}$ and queue
$Q=\{(\nu_{0},0)\}$.\label{ch04:algline:init-start}\;
Set $T^{\star}=\infty$, $J^{\star}=()$, and
$T_{\Omega}^{\star}=()$.\label{ch04:algline:init-end}\;

\While{$Q\neq\emptyset$ and $\mathrm{time}<t_{0}$}{
    Remove from $Q$ the node $\nu$ with smallest lower-bound key.\;

    $\overline{J}_{\nu},\,\overline{T}_{\nu},\,\overline{T}_{\Omega}
    = \texttt{upper\_bound}(\nu,P_{\varphi})$.\label{ch04:algline:upper}\;

    \If{$\overline{T}_{\nu}<T^{\star}$ \label{ch04:algline:update1}}{
        Set $T^{\star}=\overline{T}_{\nu}$,
        $J^{\star}=\overline{J}_{\nu}$, and
        $T_{\Omega}^{\star}=\overline{T}_{\Omega}$.\;
        \label{ch04:algline:update2}
    }

    Expand the child nodes $\{\nu^{+}\}$ of $\nu$.\;

    \ForEach{$\nu^{+}\in\{\nu^{+}\}$}{
        $\underline{T}_{\nu^{+}}
        = \texttt{lower\_bound}(\nu^{+},P_{\varphi})$.\;
        \label{ch04:algline:lower}

        \If{$\underline{T}_{\nu^{+}}\leq T^{\star}$
        \label{ch04:algline:store1}}{
            Add $(\nu^{+},\underline{T}_{\nu^{+}})$ to $Q$.\;
            \label{ch04:algline:store2}
        }
    }
}
\Return{$J^{\star},T^{\star},T_{\Omega}^{\star}$}.\;
\end{algorithm}

\begin{lemma}
\label{ch04:lemma:BnB-satisfying}
Any task assignment $J^{\star}$ returned by
Algorithm~\ref{ch04:alg:BnB} satisfies the temporal relations in
$P_{\varphi}$.
\end{lemma}

\begin{proof}
The assignment $J^{\star}$ is obtained from the upper-bound routine at some
node of the search tree. Therefore, it is sufficient to show that every
assignment returned by \eqref{ch04:eq:upper-bound} satisfies the temporal
relations in $P_{\varphi}$.
During node expansion, a subtask is selected only when all of its parent
subtasks in $\mathcal{G}_{P_{\varphi}}$ have already been assigned. By
\eqref{ch04:eq:next-task}, this guarantees that the precedence relation
$\preceq_{\varphi}$ is respected structurally by every partial assignment
generated in the search tree. Moreover, the routine
$\texttt{upper\_bound}(\nu,P_{\varphi})$ computes the makespan and the subtask
starting times under the precedence relation $\preceq_{\varphi}$, the
opposition relation $\neq_{\varphi}$, the robot-motion constraints, and the
synchronization requirements of collaborative behaviors. Hence the complete
assignment produced by the upper-bound routine satisfies both temporal
relations of the R-poset.
Since $J^{\star}$ is one such complete assignment, it satisfies the temporal
relations in $P_{\varphi}$.
\end{proof}
